### 2023-08-28 analysis.qmd
First commit

---
title: "Two Mode Networks and Relational Similarity"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
---

# Setup
```{r setup}
    knitr::opts_chunk$set(include = TRUE, 
                          echo = FALSE, 
                          warning = FALSE, 
                          message = FALSE)
    library(here)
    library(patchwork)
```

# Data
```{r Southern women data}
    A <- matrix(
         c(
          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0,

          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,

          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
          ncol = 14, byrow = TRUE)
     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
     e <- paste("E", c(1:14), sep = "")
     rownames(A) <- w
     colnames(A) <- e
```

# Analysis
```{r}
    source("Functions/TwoModeCA.R")
    tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 3, d.r = 12, d.c = 12)
    a <- tmCA.res$A.plot + theme(axis.text.x = element_text(angle = 0, 
                                                            size = 9,
                                                            vjust = -3))
    b <- tmCA.res$corr.plot
    c <- tmCA.res$sim.plot.r
    d <- tmCA.res$sim.plot.c + theme(axis.text.x = element_text(angle = 0))
    e <- tmCA.res$eigvec.plot.r
    f <- tmCA.res$eigvec.plot.c
    g <- tmCA.res$eigval.plot.r
    h <- tmCA.res$eigval.plot.c
    e <- e + inset_element(g, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
    f <- f + inset_element(h, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
    png(file = here("Plots", "main.png"), width=950, height=1200)
        (a | e | f) / (c | d) / b
    dev.off() 
```

```{r}
    source("Functions/TwoModeCA.R")
    res <- TwoModeCA(A)
```

### 2023-08-28 Functions/CA.clust.R

@@ -1,0 +1,31 @@
+CA.clust <- function(x) {
+     library(Matrix)
+     library(factoextra)
+
+     A <- x[ , colSums(x) != 0, drop = FALSE]
+     A <- Matrix(A, sparse = TRUE)
+     At <- t(A)
+     iDr <- matrix(0, nrow(A), nrow(A))
+     iDc <- matrix(0, ncol(A), ncol(A))
+     diag(iDr) <- 1/rowSums(A)
+     diag(iDc) <- 1/colSums(A)
+     eig.r <- svd(iDr %*% A %*% iDc %*% At)$u[, 2] #row canonical correlation
+     eig.c <- svd(iDc %*% At %*% iDr %*% A)$u[, 2] #column canonical correlation
+     res.km.r <- hkmeans(data.frame(eig.r), 2, hc.method = "ward.D2") #clustering rows
+     res.km.c <- hkmeans(data.frame(eig.c), 2, hc.method = "ward.D2") #clustering cols
+     clus.dat.r <- data.frame(actors = rownames(x),
+                              cluster = factor(res.km.r$cluster))
+     clus.dat.c <- data.frame(events = colnames(x),
+                              cluster = factor(res.km.c$cluster))   
+     actors.1 <- dplyr::filter(clus.dat.r, cluster == 1)$actors
+     actors.2 <- dplyr::filter(clus.dat.r, cluster == 2)$actors
+     events.1 <- dplyr::filter(clus.dat.r, cluster == 1)$events
+     events.2 <- dplyr::filter(clus.dat.r, cluster == 2)$events
+     a.clusters$1 <- list(actors.1 = actors.1, actors.2 = actors.2)
+     
+    return(list(
+         clus.dat.r = clus.dat.r,
+         clus.dat.c = clus.dat.c,
+         a.clusters = a.clusters)
+         )
+}

### 2023-08-28 Functions/TwoModeCA.R

@@ -1,0 +1,172 @@
+TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
+        library(factoextra)
+        library(ggcorrplot)
+        library(ggpubr)
+        library(Matrix)
+
+        # Plot Functions
+        eigval.scatter <- function(x) {
+                p <- ggscatter(x, 
+                               x = "k", 
+                               y = "value",
+                               size = 2,
+                               color = "red")
+                p <- p + labs(y = "", x = "k") 
+                p <- p + theme(axis.text.x = element_blank(),
+                               axis.text.y = element_text(size = 8),
+                               axis.line.y = element_blank(),
+                               axis.line.x = element_blank())
+                return(p)
+                }
+        eigvec.scatter <- function(x) {
+                p <- ggscatter(x, 
+                          y = "rank", 
+                          x = "value", 
+                          point = FALSE,
+                          font.label =  12,
+                          label = "lab", 
+                          color = "cluster",
+                          palette = "Dark2",
+                          repel = TRUE
+                          #ellipse = TRUE,
+                          #ellipse.type = "convex",
+                          #ellipse.alpha = 0.15,
+                          #ellipse.border.remove = TRUE
+                          )
+                p <- p + theme(axis.text.y = element_blank(),
+                               axis.line.x = element_blank(),
+                               axis.line.y = element_blank(),
+                               legend.position = "none")
+                p <- p + labs(y = "Rank", x = "First Axis")
+                return(p)
+                }
+        corr.scatter <- function(x) {
+                p <- ggscatter(data.frame(x), 
+                          x = "d1", 
+                          y = "d2", 
+                          point = FALSE,
+                          color = "cluster",
+                          palette = "Dark2",
+                          font.label =  10,
+                          label = rownames(x), 
+                          repel = TRUE)
[…]

### 2023-08-28 Functions/CA.clust.R
Updated function

@@ -1,31 +1,0 @@
-CA.clust <- function(x) {
-     library(Matrix)
-     library(factoextra)
-
-     A <- x[ , colSums(x) != 0, drop = FALSE]
-     A <- Matrix(A, sparse = TRUE)
-     At <- t(A)
-     iDr <- matrix(0, nrow(A), nrow(A))
-     iDc <- matrix(0, ncol(A), ncol(A))
-     diag(iDr) <- 1/rowSums(A)
-     diag(iDc) <- 1/colSums(A)
-     eig.r <- svd(iDr %*% A %*% iDc %*% At)$u[, 2] #row canonical correlation
-     eig.c <- svd(iDc %*% At %*% iDr %*% A)$u[, 2] #column canonical correlation
-     res.km.r <- hkmeans(data.frame(eig.r), 2, hc.method = "ward.D2") #clustering rows
-     res.km.c <- hkmeans(data.frame(eig.c), 2, hc.method = "ward.D2") #clustering cols
-     clus.dat.r <- data.frame(actors = rownames(x),
-                              cluster = factor(res.km.r$cluster))
-     clus.dat.c <- data.frame(events = colnames(x),
-                              cluster = factor(res.km.c$cluster))   
-     actors.1 <- dplyr::filter(clus.dat.r, cluster == 1)$actors
-     actors.2 <- dplyr::filter(clus.dat.r, cluster == 2)$actors
-     events.1 <- dplyr::filter(clus.dat.r, cluster == 1)$events
-     events.2 <- dplyr::filter(clus.dat.r, cluster == 2)$events
-     a.clusters$1 <- list(actors.1 = actors.1, actors.2 = actors.2)
-     
-    return(list(
-         clus.dat.r = clus.dat.r,
-         clus.dat.c = clus.dat.c,
-         a.clusters = a.clusters)
-         )
-}

### 2023-08-28 main.tex
First overleaf commit

%% 
%% Copyright 2019-2020 Elsevier Ltd
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-dc documentclass for 
%% double column output.

%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
\documentclass[a4paper,fleqn]{cas-sc}

%\usepackage[numbers]{natbib}
%\usepackage[authoryear]{natbib}
\usepackage[authoryear,longnamesfirst]{natbib}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}

%%%Author definitions
\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
\tsc{WGM}
\tsc{QE}
\tsc{EP}
\tsc{PMS}
\tsc{BEC}
\tsc{DE}
%%%

% Uncomment and use as if needed
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}[theorem]{Lemma}
%\newdefinition{rmk}{Remark}
%\newproof{pf}{Proof}
%\newproof{pot}{Proof of Theorem \ref{thm}}

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}

% Short title
\shorttitle{Correspondence Analysis of Two-Mode Networks}

% Short author
%\shortauthors{O. Lizardo}

% Main title of the paper
\title [mode = title]{The Correspondence Analysis of Two-Mode Networks: Applications to Centrality and Positional Analysis}                      
% Title footnote mark
% eg: \tnotemark[1]
%\tnotemark[1,2]

% Title footnote 1.
% eg: \tnotetext[1]{Title footnote text}
% \tnotetext[<tnote number>]{<tnote text>} 
%\tnotetext[1]{}

%\tnotetext[2]{.}

% First author
%
% Options: Use if required
% eg: \author[1,3]{Author Name}[type=editor,
%       style=chinese,
%       auid=000,
%       bioid=1,
%       prefix=Sir,
       %orcid=0000-0002-5405-3007,
%       facebook=<facebook id>,
%       twitter=<twitter id>,
%       linkedin=<linkedin id>,
%       gplus=<gplus id>]
\author[1]{Omar Lizardo}[orcid=0000-0001-7511-2910]

% Corresponding author indication
\cormark[1]

% Footnote of the first author
\fnmark[1]

% Email id of the first author
\ead{olizardo@soc.ucla.edu}

% URL of the first author
\ead[url]{http://olizardo.bol.ucla.edu/}

%  Credit authorship
\credit{Conceptualization of this study, Methodology, Software}

%Address/affiliation
\affiliation[1]{organization={University of California, Los Angeles},
    addressline={264 Haines Hall}, 
    city={Los Angeles},
    %citysep={},  Uncomment if no comma needed between city and postcode
     postcode={90095}, 
     state={CA},
     country={USA}}

 %Corresponding author text
\cortext[cor1]{Corresponding author}
%\cortext[cor2]{Principal corresponding author}

%Footnote text
\fntext[fn1]{This paper was written while the author was partially supported by National Science Foundation grant HNDS-R \#2214215.}
 % author as well.}
%\fntext[fn2]{Another author footnote, this is a very long footnote and
 % it should be a really long footnote. But this footnote is not yet
 % sufficiently long enough to make two lines of footnote text.}

% For a title note without a number/mark
%\nonumnote{This note has no numbers. In this work we demonstrate $a_b$
%  the formation Y\_1 of a new type of polariton on the interface
 % between a cuprous oxide slab and a polystyrene micro-sphere placed
 % on the slab.
 % }
% Here goes the abstract
\begin{abstract}
Lorem...
\end{abstract}

% Use if graphical abstract is present
% \begin{graphicalabstract}
% \includegraphics{figs/grabs.pdf}
% \end{graphicalabstract}

%Research highlights
\begin{highlights}
    \item 
    \item 
    \item 
\end{highlights}

% Keywords
% Each keyword is separated by \sep
\begin{keywords}
Generalized similarity \sep Duality \sep Two-mode networks \sep Correspondence Analysis
\end{keywords}

\maketitle
\newpage

\section{Introduction}
While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature. Initially dismissed by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic visualizations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}.

This paper contributes to the stream of previous work applying CA to analyze two-mode data. Still, it departs from the aforementioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been centered or treated in detail before.

Particularly, I show that CA can be thought of as computing a kind of \textit{centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}. I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent linear orderings.

I also show that CA can be interpreted as providing \textit{positional} information on both modes. First, I show that there is a clear link between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity" in two-mode networks \citep{kovacs2010generalized, lizardo2024two}. Second, drawing on recent work by \citet{van2021correspondence}, I show that the first two dimensions provide information on the degree-weighted similarity of nodes in the same mode with respect to their connections to nodes in the other mode. Thus, CA can be used as an approach to indirect ``clique'' identification in two-mode networks that connect to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method--- e.g., the discovery of maximally similar subsets of entities.

\subsection{Organization of the Paper}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via singular value decomposition (SVD) of the (inverse degree-weighted) one-mode projection of the affiliation matrix, which is computationally and mathematically equivalent to the CA of the affiliation matrix. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. I close by providing some pointers as to how CA can be used as a method of structural analysis (e.g., centrality and positional analysis), independently of the usual emphasis on visualization.

\section{Reflective Centralities in Two-Mode Networks}
In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach to iterated centralities. Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as ``economic complexity.'' Subsequent shows that there is no logical connection between the formal method and this particular application. As such, I introduce it here using the more intuitive---and classical---case of the duality of persons and groups \citep{breiger1974duality}.

If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}. A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $P$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.

\begin{equation}
    C^R_p(1) = \sum_g a_{pg}
    \label{eq:R1_p}
\end{equation}

And for groups:

\begin{equation}
   C^R_g(1) = \sum_p a_{pg}
  \label{eq:R1_g}
\end{equation}

The key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second order centralities" for both people and groups using the centralities of the other mode.

For people, this is:

\begin{equation}
    C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
    \label{eq:R2_p}
\end{equation}

And for groups:

\begin{equation}
   C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
    \label{eq:R2_g}
\end{equation}

Equation \ref{eq:R1_p} says ``people are more central when the average sum of the memberships of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number). Equation \ref{eq:R2_g} says ``groups are more central when the average sum of the number of memberships of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).

Of course, we can keep on going and define third-order reflections. For the people these are:

\begin{equation}
   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
   \label{eq:R3_p}
\end{equation}

And for groups:

\begin{equation}
   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
   \label{eq:R3_g}
\end{equation}

Equation \ref{eq:R3_p} says something like ``people are more central when the average of the average number of memberships of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number). Equation \ref{eq:R3_g} says something like ``groups are more central when the average of the average number of members of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).

Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, their degree centrality). On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to. In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their degree centrality). On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{groups their members belong to}.

More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases):

\begin{equation}   
    C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
   \label{eq:Rq_p}
\end{equation}

\begin{equation}
   C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
   \label{eq:Rq_g}
\end{equation}

\section{Correspondence Analysis and Dual Centralities}

%% Loading bibliography style file
%\bibliographystyle{model1-num-names}
\bibliographystyle{cas-model2-names}
% Loading bibliography database
\bibliography{ca.bib}
\end{document}

The Elsevier cas-dc class is based on the
standard article class and supports almost all of the functionality of
that class. In addition, it features commands and options to format the
\begin{itemize} \item document style \item baselineskip \item front
matter \item keywords and MSC codes \item theorems, definitions and
proofs \item labels of enumerations \item citation style and labeling.
\end{itemize}

This class depends on the following packages
for its proper functioning:

\begin{enumerate}
\itemsep=0pt
\item {natbib.sty} for citation processing;
\item {geometry.sty} for margin settings;
\item {fleqn.clo} for left aligned equations;
\item {graphicx.sty} for graphics inclusion;
\item {hyperref.sty} optional packages if hyperlinking is
  required in the document;
\end{enumerate}

All the above packages are part of any
standard \LaTeX{} installation.
Therefore, the users need not be
bothered about downloading any extra packages.

\section{Installation}

The package is available at author resources page at Elsevier
(\url{http://www.elsevier.com/locate/latex}).
The class may be moved or copied to a place, usually,\linebreak
\verb+$TEXMF/tex/latex/elsevier/+, %$%%%%%%%%%%%%%%%%%%%%%%%%%%%%
or a folder which will be read                   
by \LaTeX{} during document compilation.  The \TeX{} file
database needs updation after moving/copying class file.  Usually,
we use commands like \verb+mktexlsr+ or \verb+texhash+ depending
upon the distribution and operating system.

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. 
You are recommended to conform your choice to the journal you 
are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use Bib\TeX\ to generate your bibliography and include DOIs
whenever available.

Here are two sample references: 
\cite{Fortunato2010}
\cite{Fortunato2010,NewmanGirvan2004}
\cite{Fortunato2010,Vehlowetal2013}

\section{Floats}
{Figures} may be included using the command,\linebreak 
\verb+\includegraphics+ in
combination with or without its several options to further control
graphic. \verb+\includegraphics+ is provided by {graphic[s,x].sty}
which is part of any standard \LaTeX{} distribution.
{graphicx.sty} is loaded by default. \LaTeX{} accepts figures in
the postscript format while pdf\LaTeX{} accepts {*.pdf},
{*.mps} (metapost), {*.jpg} and {*.png} formats. 
pdf\LaTeX{} does not accept graphic files in the postscript format.

\begin{figure}
	\centering
		\includegraphics[scale=.75]{figs/Fig1.pdf}
	\caption{The evanescent light - $1S$ quadrupole coupling
	($g_{1,l}$) scaled to the bulk exciton-photon coupling
	($g_{1,2}$). The size parameter $kr_{0}$ is denoted as $x$ and
	the \PMS is placed directly on the cuprous oxide sample ($\delta
	r=0$, See also Table \protect\ref{tbl1}).}
	\label{FIG:1}
\end{figure}

The \verb+table+ environment is handy for marking up tabular
material. If users want to use {multirow.sty},
{array.sty}, etc., to fine control/enhance the tables, they
are welcome to load any package of their choice and
{cas-dc.cls} will work in combination with all loaded
packages.

\begin{table}[width=.9\linewidth,cols=4,pos=h]
\caption{This is a test caption. This is a test caption. This is a test
caption. This is a test caption.}\label{tbl1}
\begin{tabular*}{\tblwidth}{@{} LLLL@{} }
\toprule
Col 1 & Col 2 & Col 3 & Col4\\
\midrule
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
\bottomrule
\end{tabular*}
\end{table}

\section[Theorem and ...]{Theorem and theorem like environments}

{cas-dc.cls} provides a few shortcuts to format theorems and
theorem-like environments with ease. In all commands the options that
are used with the \verb+\newtheorem+ command will work exactly in the same
manner. {cas-dc.cls} provides three commands to format theorem or
theorem-like environments:

\begin{verbatim}
 \newtheorem{theorem}{Theorem}
 \newtheorem{lemma}[theorem]{Lemma}
 \newdefinition{rmk}{Remark}
 \newproof{pf}{Proof}
 \newproof{pot}{Proof of Theorem \ref{thm2}}
\end{verbatim}

The \verb+\newtheorem+ command formats a
theorem in \LaTeX's default style with italicized font, bold font
for theorem heading and theorem number at the right hand side of the
theorem heading.  It also optionally accepts an argument which
will be printed as an extra heading in parentheses.

\begin{verbatim}
  \begin{theorem} 
   For system (8), consensus can be achieved with 
   $\|T_{\omega z}$ ...
     \begin{eqnarray}\label{10}
     ....
     \end{eqnarray}
  \end{theorem}
\end{verbatim}

\newtheorem{theorem}{Theorem}

\begin{theorem}
For system (8), consensus can be achieved with 
$\|T_{\omega z}$ ...
\begin{eqnarray}\label{10}
....
\end{eqnarray}
\end{theorem}

The \verb+\newdefinition+ command is the same in
all respects as its \verb+\newtheorem+ counterpart except that
the font shape is roman instead of italic.  Both
\verb+\newdefinition+ and \verb+\newtheorem+ commands
automatically define counters for the environments defined.

The \verb+\newproof+ command defines proof environments with
upright font shape.  No counters are defined.

\section[Enumerated ...]{Enumerated and Itemized Lists}
{cas-dc.cls} provides an extended list processing macros
which makes the usage a bit more user friendly than the default
\LaTeX{} list macros.   With an optional argument to the
\verb+\begin{enumerate}+ command, you can change the list counter
type and its attributes.

\begin{verbatim}
 \begin{enumerate}[1.]
 \item The enumerate environment starts with an optional
   argument `1.', so that the item counter will be suffixed
   by a period.
 \item You can use `a)' for alphabetical counter and '(i)' 
  for roman counter.
  \begin{enumerate}[a)]
    \item Another level of list with alphabetical counter.
    \item One more item before we start another.
    \item One more item before we start another.
    \item One more item before we start another.
    \item One more item before we start another.
\end{verbatim}

Further, the enhanced list environment allows one to prefix a
string like `step' to all the item numbers.

\begin{verbatim}
 \begin{enumerate}[Step 1.]
  \item This is the first step of the example list.
  \item Obviously this is the second step.
  \item The final step to wind up this example.
 \end{enumerate}
\end{verbatim}

\section{Cross-references}
In electronic publications, articles may be internally
hyperlinked. Hyperlinks are generated from proper
cross-references in the article.  For example, the words
\textcolor{black!80}{Fig.~1} will never be more than simple text,
whereas the proper cross-reference \verb+\ref{tiger}+ may be
turned into a hyperlink to the figure itself:
\textcolor{blue}{Fig.~1}.  In the same way,
the words \textcolor{blue}{Ref.~[1]} will fail to turn into a
hyperlink; the proper cross-reference is \verb+\cite{Knuth96}+.
Cross-referencing is possible in \LaTeX{} for sections,
subsections, formulae, figures, tables, and literature
references.

\section{Bibliography}

Two bibliographic style files (\verb+*.bst+) are provided ---
{model1-num-names.bst} and {model2-names.bst} --- the first one can be
used for the numbered scheme. This can also be used for the numbered
with new options of {natbib.sty}. The second one is for the author year
scheme. When  you use model2-names.bst, the citation commands will be
like \verb+\citep+,  \verb+\citet+, \verb+\citealt+ etc. However when
you use model1-num-names.bst, you may use only \verb+\cite+ command.

\verb+thebibliography+ environment.  Each reference is a\linebreak
\verb+\bibitem+ and each \verb+\bibitem+ is identified by a label,
by which it can be cited in the text:

\noindent In connection with cross-referencing and
possible future hyperlinking it is not a good idea to collect
more that one literature item in one \verb+\bibitem+.  The
so-called Harvard or author-year style of referencing is enabled
by the \LaTeX{} package {natbib}. With this package the
literature can be cited as follows:

\begin{enumerate}[\textbullet]
\item Parenthetical: \verb+\citep{WB96}+ produces (Wettig \& Brown, 1996).
\item Textual: \verb+\citet{ESG96}+ produces Elson et al. (1996).
\item An affix and part of a reference:\break
\verb+\citep[e.g.][Ch. 2]{Gea97}+ produces (e.g. Governato et
al., 1997, Ch. 2).
\end{enumerate}

In the numbered scheme of citation, \verb+\cite{<label>}+ is used,
since \verb+\citep+ or \verb+\citet+ has no relevance in the numbered
scheme.  {natbib} package is loaded by {cas-dc} with
\verb+numbers+ as default option.  You can change this to author-year
or harvard scheme by adding option \verb+authoryear+ in the class
loading command.  If you want to use more options of the {natbib}
package, you can do so with the \verb+\biboptions+ command.  For
details of various options of the {natbib} package, please take a
look at the {natbib} documentation, which is part of any standard
\LaTeX{} installation.

\appendix
\section{My Appendix}
Appendix sections are coded under \verb+\appendix+.

\verb+\printcredits+ command is used after appendix sections to list 
author credit taxonomy contribution roles tagged using \verb+\credit+ 
in frontmatter.

\printcredits

%\vskip3pt

\bio{}
Author biography without author photo.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
\endbio

\bio{figs/pic1}
Author biography with author photo.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
\endbio

\bio{figs/pic1}
Author biography with author photo.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
\endbio

### 2023-08-28 analysis.qmd
Bumpchart reflections

@@
                           message = FALSE)
+    library(cowplot)
     library(here)
     library(patchwork)
+    library(dplyr)
+    library(ggbump)
+    library(ggplot2)
+    library(ggrepel)
+    library(pals)
 ```
@@
 # Analysis
-```{r}
-    source("Functions/TwoModeCA.R")
+```{r Reflections Bump Charts}
+   source(here("Functions", "reflections.R"))
+   source(here("Functions", "ref.long.dat.R"))
+   evens <- function(x) subset(x, x %% 2 == 0)
+   odds <- function(x) subset(x, x %% 2 != 0)
+   a <- reflections(A, iter = 26)
+   p.e.r <- a$p.r[, evens(1:20)]
+   p.o.r <- a$p.r[, odds(1:20)]
+   p.o.r <- p.o.r[, -1]
+   b <- ref.long.dat(a = a$p.r, b = a$g.r, max.iter = 26)
+   plot.dat <- b$person$even
+   p <- ggplot(data = filter(plot.dat, n.iter < 22),
+               mapping = aes(x = n.iter, y = ref, color = person))
+   # geoms
+   p <- p + geom_bump(linewidth = 1, smooth = 8, alpha = 0.4)
+   p <- p + geom_point(size = 4, alpha = 0.8)
+   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
+                                    aes(x = n.iter, label = person), 
+                                    size = 4, nudge_x = -3)
+   p <-    p <- p + geom_text(data = filter(plot.dat, n.iter == 22), 
+                              aes(x = n.iter, label = person), 
+                              size = 4, nudge_x = 0.5)
+
+   # fixings
+   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
+   p <- p + scale_x_continuous(limits = c(-4, 30), breaks = evens(2:22))
+   p <- p + labs(x = "Reflection")
+   p <- p + scale_y_reverse()
+   p <- p + theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.y = element_blank()
+        )
+   p1 <- p + scale_color_manual(values=as.vector(cols25(18)))
+
+   plot.dat <- b$group$even
+   p <- ggplot(data = filter(plot.dat, n.iter < 22),
+               mapping = aes(x = n.iter, y = ref, color = group))
+   #geoms
+   p <- p + geom_bump(linewidth = 1, smooth = 8, alpha = 0.4)
+   p <- p + geom_point(size = 4, alpha = 1)
+   p <-    p <- p + geom_text(data = filter(plot.dat, n.iter == 24), 
+                              aes(x = n.iter, label = group), 
+                              size = 4, nudge_x = -2)
+   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
+                                    aes(x = n.iter, label = group), 
+                                    size = 4, nudge_x = -1)
+   #fixings
+   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
+   p <- p + scale_x_continuous(limits = c(-4, 30), breaks = evens(2:20))
+   p <- p + labs(x = "Reflection")
+   p <- p + scale_y_reverse()
+   p <- p + theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.y = element_blank()
+        )
+   p2 <- p + scale_color_manual(values=as.vector(cols25(15)))
+   png(file = here("Plots", "pg-reflections.png"), width=750, height=950)
+   p1 / p2
+   dev.off()
+```
+
+```{r Main Plot}
+   source(here("Functions", "TwoModeCA.R"))
     tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 3, d.r = 12, d.c = 12)
@@
 ```
-
-```{r}
-    source("Functions/TwoModeCA.R")
-    res <- TwoModeCA(A)
-```

### 2023-08-28 Functions/TwoModeCA.R

@@ -167,6 +167,10 @@ TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
                 sim.plot.r = sim.plot.r,
                 sim.plot.c = sim.plot.c,
                 km.r = km.r,
-                km.c = km.c)
+                km.c = km.c,
+                eigvec.r = eigvec.r,
+                eigvec.c = eigvec.c,
+                eigval.r = CA.r$d,
+                eigval.c = CA.c$d)
                )
         }

### 2023-08-28 main.tex

@@
 On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{groups their members belong to}.
+While these are distinct metrics, in practice the order of nodes in each mode ends up being identical when the ranks ``freeze'' after some number of iterations (proportional to the size of the network).
 
@@
 
+Figure~\ref{fig:pg-reflections} shows the trajectory of reflective centralities for persons and groups in the Southern Women data \citep{davis1941}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.png}
+    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
+    \label{fig:pg-reflections}
+\end{figure}
 \section{Correspondence Analysis and Dual Centralities}

### 2023-08-30 analysis.qmd
New matrix reord plot plus lit

--- 2023-08-28 analysis.qmd
@@
                        message = FALSE)
+   library(cowplot)
    library(here)
    library(patchwork)
+   library(dplyr)
+   library(ggbump)
+   library(ggplot2)
+   library(ggrepel)
+   library(igraph)
+   library(pals)
+   source(here("Functions", "TwoModeCA.R"))
+   source(here("Functions", "reflections.R"))
+   source(here("Functions", "ref.long.dat.R"))
 ```
@@
 ```{r}
-    source("Functions/TwoModeCA.R")
+   source(here("Functions", "TwoModeCA.R"))
+   a <- TwoModeCA(A)
+   b <- CA(A, graph = FALSE)
+   c <- round(a$CA.res$u[, 2], 3)
+   d <- round(b$row$coord[, 1], 3)
+   e <- data.frame(c, d)
+   ggscatter(e, x = "c", y = "d", label = rownames(A), repel = TRUE)
+```
+
+```{r Reflections Bump Charts}
+   evens <- function(x) subset(x, x %% 2 == 0)
+   odds <- function(x) subset(x, x %% 2 != 0)
+   a <- reflections(A, iter = 26)
+   p.e.r <- a$p.r[, evens(1:20)]
+   p.o.r <- a$p.r[, odds(1:20)]
+   p.o.r <- p.o.r[, -1]
+   b <- ref.long.dat(a = a$p.r, b = a$g.r, max.iter = 26)
+   plot.dat <- b$person$even
+   p <- ggplot(data = filter(plot.dat, n.iter < 22),
+               mapping = aes(x = n.iter, y = ref, color = person))
+   # geoms
+   p <- p + geom_bump(linewidth = 1, smooth = 8, alpha = 0.4)
+   p <- p + geom_point(size = 4, alpha = 0.8)
+   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
+                                    aes(x = n.iter, label = person), 
+                                    size = 4, nudge_x = -3)
+   p <-    p <- p + geom_text(data = filter(plot.dat, n.iter == 22), 
+                              aes(x = n.iter, label = person), 
+                              size = 4, nudge_x = 0.5)
+
+   # fixings
+   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
+   p <- p + scale_x_continuous(limits = c(-4, 30), breaks = evens(2:20))
+   p <- p + labs(x = "Reflection")
+   p <- p + scale_y_reverse()
+   p <- p + theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.y = element_blank()
+        )
+   p1 <- p + scale_color_manual(values=as.vector(cols25(18)))
+
+   plot.dat <- b$group$even
+   p <- ggplot(data = filter(plot.dat, n.iter < 22),
+               mapping = aes(x = n.iter, y = ref, color = group))
+   #geoms
+   p <- p + geom_bump(linewidth = 1, smooth = 8, alpha = 0.4)
+   p <- p + geom_point(size = 4, alpha = 1)
+   p <-    p <- p + geom_text(data = filter(plot.dat, n.iter == 24), 
+                              aes(x = n.iter, label = group), 
+                              size = 4, nudge_x = -2)
+   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
+                                    aes(x = n.iter, label = group), 
+                                    size = 4, nudge_x = -1)
+   #fixings
+   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
+   p <- p + scale_x_continuous(limits = c(-4, 30), breaks = evens(2:20))
+   p <- p + labs(x = "Reflection")
+   p <- p + scale_y_reverse()
+   p <- p + theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.y = element_blank()
+        )
+   p2 <- p + scale_color_manual(values=as.vector(cols25(15)))
+   png(file = here("Plots", "pg-reflections.jpg"), width=750, height=950)
+   p1 / p2
+   dev.off()
+```
+
+```{r Comparative scatters}
+   library(Matrix)
+   library(FactoMineR)
+   AAt <- A %*% t(A)
+   AtA <- t(A) %*% A
+   bon.res.r <- svd(AAt)$u[, 1]
+   bon.res.c <- svd(AtA)$u[, 1]
+   names(bon.res.r) <- rownames(A)
+   names(bon.res.c) <- colnames(A)
+   van.res.r <- TwoModeCA(A)$eigvec.r[, 1]
+   van.res.c <- TwoModeCA(A)$eigvec.c[, 1]
+   ref.res.r <- scale(reflections(A, iter = 30)$p.r[, 26])
+   ref.res.c <- scale(reflections(A, iter = 30)$p.c[, 26])
+   ca.res.r <- CA(A, graph = FALSE)$row$coord[, 1]
+   ca.res.c <- CA(A, graph = FALSE)$col$coord[, 1]
+   dat.r <- data.frame(cbind(a = bon.res.r, b = van.res.r, 
+                  c = as.vector(ref.res.r), d = ca.res.r))
+
+```
+
+```{r Main Plot}
     tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 3, d.r = 12, d.c = 12)

### 2023-08-30 Functions/TwoModeCA.R

@@ -72,10 +72,17 @@ TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
         A <- x[ , colSums(x) != 0, drop = FALSE]
         A <- Matrix(A, sparse = TRUE)
         At <- t(A)
+        
         iDr <- matrix(0, nrow(A), nrow(A))
         iDc <- matrix(0, ncol(A), ncol(A))
         diag(iDr) <- 1/rowSums(A)
         diag(iDc) <- 1/colSums(A)
+        
+        iDs.r <- matrix(0, nrow(A), nrow(A))
+        iDs.c <- matrix(0, ncol(A), ncol(A))
+        diag(iDs.r) <- 1/sqrt(rowSums(A))
+        diag(iDs.c) <- 1/sqrt(colSums(A))
+        
         r <- nrow(A)
         c <- ncol(A)
         rn <- rownames(A)
@@ -100,6 +107,13 @@ TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
         colnames(S.r) <- rn
         rownames(S.c) <- cn
         colnames(S.c) <- cn
+        
+        CA.res <- iDs.r %*% A %*% iDs.c
+        CA.res <- svd(CA.res) #row and column CA
+        
+        bon.res <- svd(A) #Bonacich dual eigenvector centrality
+        bon.eigvec.r <- bon.res$u
+        bon.eigvec.c <- bon.res$v
         
         # K-means clustering of similarity matrix
         S.r.dat <- data.frame(svd(S.r[, 1:d.r])$u)
@@ -131,12 +145,19 @@ TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
         eigvec.rc <- cbind(eigvec.rc, km.dat.rc)
         corr.plot <- corr.scatter(eigvec.rc)
         
-        # Affiliation matrix plot
+        # Affiliation matrix plot (CA ordering)
         A.ord <- as.matrix(A[order(eigvec.r[, 1]), order(eigvec.c[, 1])])
         A.plot <- ggcorrplot(t(A.ord)) 
         A.plot <- A.plot + scale_x_discrete(position = "top")
-        A.plot <- A.plot + theme(legend.position = "none",
-                                 axis.text.x = element_text(hjust = -0.2)) 
+        A.plot1 <- A.plot + theme(legend.position = "none",
+                                 axis.text.x = element_text(hjust = -0.2))
+        
[…]

### 2023-08-30 main.tex

--- 2023-08-28 main.tex
@@
 \usepackage{booktabs}
+\usepackage{longtable}
 
@@
 % Main title of the paper
-\title [mode = title]{The Correspondence Analysis of Two-Mode Networks: Applications to Centrality and Positional Analysis}                      
+\title [mode = title]{The Correspondence Analysis of Two-Mode Networks: Centrality and Positional Interpretations}                      
 % Title footnote mark
@@
 \newpage
-
 \section{Introduction}
 While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
-Initially dismissed by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic visualizations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}.
+Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') spatial visualizations of structural connectivity patterns across the two-modes \citep{bonacich1991simultaneous, breiger2000tool, faust2005using}.
+This paper contributes to the stream of previous work applying CA to analyze two-mode data.
+Nevertheless, it departs from the aforementioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been centered or treated in detail before.
 
-This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Still, it departs from the aforementioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been centered or treated in detail before.
+Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
+With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent linear orderings.
+This applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.
 
-Particularly, I show that CA can be thought of as computing a kind of \textit{centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
-I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent linear orderings.
-
-I also show that CA can be interpreted as providing \textit{positional} information on both modes.
-First, I show that there is a clear link between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity" in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
-Second, drawing on recent work by \citet{van2021correspondence}, I show that the first two dimensions provide information on the degree-weighted similarity of nodes in the same mode with respect to their connections to nodes in the other mode.
-Thus, CA can be used as an approach to indirect ``clique'' identification in two-mode networks that connect to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method--- e.g., the discovery of maximally similar subsets of entities.
+I also show that CA can be interpreted as providing \textit{positional} information on people and groups.
+First, I show that there is a clear link between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
+As such, CA recovers clusters of entities (e.g., people) linked by their similar patterns of connectivity to other similar entities in the other mode (e.g., groups).
+Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode.
+Thus, CA can be used as an approach to indirect ``clique'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method---e.g., the discovery of sociometrically similar subsets of entities.
 
@@
 I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I close by providing some pointers as to how CA can be used as a method of structural analysis (e.g., centrality and positional analysis), independently of the usual emphasis on visualization.
+In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint visualization.
 
 \section{Reflective Centralities in Two-Mode Networks}
-In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach to iterated centralities.
+In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach to iterated centralities in two-mode networks.
 Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as ``economic complexity.'' Subsequent shows that there is no logical connection between the formal method and this particular application.
-As such, I introduce it here using the more intuitive---and classical---case of the duality of persons and groups \citep{breiger1974duality}.
+As such, I introduce it here using the more intuitive---and classical from a social science perspective---case of the duality of persons and groups \citep{breiger1974duality}.
 
 If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}.
-A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $P$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.
+A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $G$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.
+
+Given this, the degree-centrality of people is given by:
 
@@
 
-The key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second order centralities" for both people and groups using the centralities of the other mode.
+That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
+As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.
 
-For people, this is:
+For people, these are given by:
 
@@
 Equation \ref{eq:R2_g} says ``groups are more central when the average sum of the number of memberships of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
+Of course, we can keep on going and define third-order reflections.
 
-Of course, we can keep on going and define third-order reflections. For the people these are:
+For the people, these are given by:
 
@@
 Equation \ref{eq:R3_p} says something like ``people are more central when the average of the average number of memberships of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
-Equation \ref{eq:R3_g} says something like ``groups are more central when the average of the average number of members of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
+Equation \ref{eq:R3_g} says, ''groups are more central when the average number of members of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
 
@@
 On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{groups their members belong to}.
+While these are distinct metrics, in practice, the order of nodes in each mode ends up being identical when the ranks ``freeze'' after some number of iterations (proportional to the network size).
 
-More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases):
+More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases).
+
+For people, these are given by:
 
@@
 
+And for groups:
+
 \begin{equation}
@@
 
-\section{Correspondence Analysis and Dual Centralities}
+Equation~\ref{eq:Rq_p} says that the reflective centrality of person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
+Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
+
+\begin{table}
+\caption{Southern Women Data.}
+\label{tab:southern}
+\begin{tabular}[]{lcccccccccccccc}
+\hline
+ & E1 & E2 & E3 & E4 & E5 & E6 & E7 & E8 & E9 & E10 & E11 & E12 & E13 & E14 \\
+ \hline
+EVELYN & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
+LAURA & 1 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+THERESA & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
+BRENDA & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+CHARLOTTE & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+FRANCES & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ELEANOR & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+RUTH & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
+VERNE & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 1 & 0 & 0 \\
+MYRA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 \\
+KATHERINE & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
+SYLVIA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
+NORA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 \\
+HELEN & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 \\
+OLIVIA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
+FLORA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
+PEARL & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
+DOROTHY & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
+\hline
+\end{tabular}
+\textsuperscript{Note: Rows ordered according to the generalized
+blockmodeling solution of Doreian et al. (2004, Table 4).} \\
+\end{table}
+
+\subsection{Empirical Example}
+Figure~\ref{fig:pg-reflections} shows the trajectory of the Hidalgo-Hausmann reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}.
+The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
+
+As the bottom panel of the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to.
+However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
+Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al.
+(2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
+
+The bottom panel of Figure~\ref{fig:pg-reflections} shows the corresponding bump chart for the events.
+Just like for these actors, the equilibrium reflective centralities recover the ordering of the columns according to Doreian et al.
+generalized blockmodel with events 1-6 separated from events 10-15 by events 7-9 (see also \citet{kovacs2010generalized}).
+Figure~\ref{fig:pg-reflections} shows that $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
+Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering the average of the average number of memberships of its members (and so forth).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
+    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
+    \label{fig:pg-reflections}
+\end{figure}
+\section{Correspondence Analysis and Dual Centrality}
+The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
+In that paper, Bonacich also considered CA in passing but dismissed it as providing a measure of centrality.
+Instead, Bonacich noted that ``[r]ather than centrality, what they seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
+
+As we will see, Bonacich was only half right.
+Indeed, the CA scores capture a version of ``clique'' membership, but they---particularly the first dimension---also capture a version of centrality in the most general sense of ranking nodes based on a meaningful criterion \citep{borgatti2006graph}.
+In fact, they capture the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
+
+To see this, recall that the solutions to the following eigensystem give the Bonacich $(C^B)$ two-mode centralities:
+
+\begin{equation}
+    AC^B_p = \lambda C^B_g 
+    \label{eq:bon1}
+\end{equation}
+
+\begin{equation}
+    A^TC^B_g = \lambda C^B_p 
+    \label{eq:bon2}
+\end{equation}
+
+Where $A$ is the original affiliation matrix.
+The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors of the affiliation matrix corresponding to the largest eigenvalue $\lambda$.
+Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of individual person and group centralities.
+
+For people, these are given by:
+
+\begin{equation}
+    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
+    \label{eq:faust1}
+\end{equation}
+
+And for groups:
+
+\begin{equation}
+    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
+    \label{eq:faust2}
+\end{equation}
+
+These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
+These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
+Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
+The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.
+
+As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's reflective centralities can also be obtained as the solution to an eigenvalue problem.
+To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
+Matrix-algebraic reduction of the resulting equations would show that the Bonacich two-mode centralities can also be obtained as solutions to the eigensystem:
+
+\begin{equation}
+    \left(AA^T\right)C^B_p = \lambda^2 C^B_p
+    \label{eq:bon3}
+\end{equation}
+
+\begin{equation}
+    \left(A^TA\right)C^B_g = \lambda^2 C^B_g
+    \label{eq:bon4}
+\end{equation}
+
+Now, consider the $p \times p$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
+In the same way, consider the $g \times g$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
+It can be shown \citep{van2021correspondence}, that in the limit, the iterative Hidalgo-Hausmann \citeyearpar{hidalgo2009building} reflective centralities emerge as the solution of the eigensystem:
+
+\begin{equation}
+    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
+    \label{eq:dam1}
+\end{equation}
+
+\begin{equation}
+    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
+    \label{eq:dam2}
+\end{equation}
+
+Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
+Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups.
+The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode, essentially weighting the projection by node degree before performing the SVD.
+
+This can be clearly seen if we express $\left(Dg^{-1}A^TDp^{-1}A\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
+
+\begin{equation}
+    a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
+    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
+    \label{eq:mealy1}
+\end{equation}
+
+In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share a membership.
+Summed across groups $g$, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
+As noted, the Bonacich dual centralities are obtained from an SVD of this matrix of shared memberships (for people) and shared people (for groups).
+The reflective centralities, on the other hand, are obtained from an SVD of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
+
+The same reasoning applies to groups in~\ref{eq:dam2}, whose entries are given by:
+
+\begin{equation}
+    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
+    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
+    \label{eq:mealy2}
+\end{equation}
+
+Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used before performing the SVD of a two-mode matrix to obtain the corresponding row and column scores in CA.\footnote{More accurately, cells are weighted by inverse of the square root of the product of the row and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, as has recently been shown \citep{van2021correspondence}, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be substantively equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a CA of the original affiliation matrix.\footnote{As \citet[126]{faust2005using}, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of the SVD solution.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
+
+Degree-pre-weighting does not radically alter the nature of the data.
+Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size.
+This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more when the group in question is small than when it is large.
+On the group side, shared members who do not have many affiliations count more than those with many affiliations.
+Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
+In the Bonacich case, SVD is a one-mode projection of the two-mode network; in the reflective case, SVD of the degree-weighted one-mode projection of the same network.
+Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
+Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the centrality of people is a function of the (average) centrality of the groups they belong to.
+The centrality of groups, in turn, is a function of their members' (average) centrality.
+The person and group score ranks given by the first CA dimension are thus better referred to as activity and group-sized adjusted versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.\footnote{Adjusted in the sense described earlier of giving more weight to smaller groups and people with fewer memberships.}
+
+To see the links between CA, reflective centralities, and clustering in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be seen as giving the similarity between pairs of people weighted by the size of the groups they belong to.
+The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}.
 

### 2023-08-30 analysis.qmd
Updated plots

@@
 # Analysis
-```{r}
-   source(here("Functions", "TwoModeCA.R"))
-   a <- TwoModeCA(A)
-   b <- CA(A, graph = FALSE)
-   c <- round(a$CA.res$u[, 2], 3)
-   d <- round(b$row$coord[, 1], 3)
-   e <- data.frame(c, d)
-   ggscatter(e, x = "c", y = "d", label = rownames(A), repel = TRUE)
-```
-
 ```{r Reflections Bump Charts}
@@
 
-```{r Comparative scatters}
-   library(Matrix)
-   library(FactoMineR)
-   AAt <- A %*% t(A)
-   AtA <- t(A) %*% A
-   bon.res.r <- svd(AAt)$u[, 1]
-   bon.res.c <- svd(AtA)$u[, 1]
-   names(bon.res.r) <- rownames(A)
-   names(bon.res.c) <- colnames(A)
-   van.res.r <- TwoModeCA(A)$eigvec.r[, 1]
-   van.res.c <- TwoModeCA(A)$eigvec.c[, 1]
-   ref.res.r <- scale(reflections(A, iter = 30)$p.r[, 26])
-   ref.res.c <- scale(reflections(A, iter = 30)$p.c[, 26])
-   ca.res.r <- CA(A, graph = FALSE)$row$coord[, 1]
-   ca.res.c <- CA(A, graph = FALSE)$col$coord[, 1]
-   dat.r <- data.frame(cbind(a = bon.res.r, b = van.res.r, 
-                  c = as.vector(ref.res.r), d = ca.res.r))
-
+```{r CA versus Bonacich Re-ordered Affiliation Matrices}
+   source(here("Functions", "TwoModeCA.R"))
+   a <- TwoModeCA(A)
+   b <- a$A.plot.ca
+   c <- a$A.plot.bon
+   png(file = here("Plots", "ca-v-bon-reord.png"), width=1000, height=700)
+      (b | c) + plot_annotation(tag_levels = "a")
+   dev.off() 
 ```
@@
 ```{r Main Plot}
-    tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 3, d.r = 12, d.c = 12)
-    a <- tmCA.res$A.plot + theme(axis.text.x = element_text(angle = 0, 
-                                                            size = 9,
-                                                            vjust = -3))
-    b <- tmCA.res$corr.plot
-    c <- tmCA.res$sim.plot.r
-    d <- tmCA.res$sim.plot.c + theme(axis.text.x = element_text(angle = 0))
-    e <- tmCA.res$eigvec.plot.r
-    f <- tmCA.res$eigvec.plot.c
-    g <- tmCA.res$eigval.plot.r
-    h <- tmCA.res$eigval.plot.c
-    e <- e + inset_element(g, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
-    f <- f + inset_element(h, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
-    png(file = here("Plots", "main.png"), width=950, height=1200)
-        (a | e | f) / (c | d) / b
+   source(here("Functions", "TwoModeCA.R"))
+    tmCA.res <- TwoModeCA(A, k.r = 5, k.c = 5, d.r = 9, d.c = 9)
+    a <- tmCA.res$eigvec.plot.r + theme(axis.text = element_blank())
+    b <- tmCA.res$eigvec.plot.c 
+    c <- tmCA.res$eigval.plot.r
+    d <- tmCA.res$eigval.plot.c
+    a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    e <- tmCA.res$sim.plot.r + theme(axis.text.x = element_blank())
+    f <- tmCA.res$sim.plot.c + theme(axis.text.x = element_blank())
+    g <- tmCA.res$corr.plot
+
+    png(file = here("Plots", "eigvec.png"), width=1000, height=600)
+        (a | b) + plot_annotation(tag_levels = "a")
+    dev.off()
+
+    png(file = here("Plots", "sim.png"), width=950, height=700)
+        (e | f) / g + plot_annotation(tag_levels = "a")
     dev.off() 

### 2023-08-30 main.tex

@@
 \end{figure}
-\section{Correspondence Analysis and Dual Centrality}
+\section{Correspondence Analysis, Dual Centrality, and Clique Identification}
 The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
@@
 
-To see the links between CA, reflective centralities, and clustering in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be seen as giving the similarity between pairs of people weighted by the size of the groups they belong to.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Plots/ca-v-bon-reord.png}
+    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to first CA dimension, (b) row and columns re-ordered according to Eigenvector centrality.}
+    \label{fig:ca-v-bon}
+\end{figure}
+
+Figure~\ref{fig:ca-v-bon} illustrates the differences between $C^R$ and $C^B$.
+Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension (a) and the Bonacich centrality (b).
+In the plot, a square is colored red if it has a one in the affiliation matrix and is colored white if it has a zero.
+As we can see, the two matrices have distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
+Accordingly, the $C^R$ centrality rankings reveal a clique-like partition between two groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
+The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side).
+The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward clique partitioning and $C^B$ more focused on finding a core of well-connected persons and groups.
+
+To further appreciate the links between CA, reflective centralities, and clique-like clustering in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be seen as giving the similarity between pairs of people weighted by the size of the groups they belong to.
 The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}.

### 2023-08-31 analysis.qmd
Updated plots

@@
    source(here("Functions", "TwoModeCA.R"))
-    tmCA.res <- TwoModeCA(A, k.r = 5, k.c = 5, d.r = 9, d.c = 9)
+    tmCA.res <- TwoModeCA(A, k.r = 7, k.c = 5, d.r = 9, d.c = 9)
     a <- tmCA.res$eigvec.plot.r + theme(axis.text = element_blank())

### 2023-08-31 Functions/TwoModeCA.R

@@ -46,7 +46,7 @@ TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
                           y = "d2", 
                           point = FALSE,
                           color = "cluster",
-                          palette = "Dark2",
+                          palette = "uchicago",
                           font.label =  10,
                           label = rownames(x), 
                           repel = TRUE)

### 2023-08-31 main.tex
Updates from Overleaf

@@
 Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode.
-Thus, CA can be used as an approach to indirect ``clique'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method---e.g., the discovery of sociometrically similar subsets of entities.
+Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method---e.g., the discovery of sociometrically similar subsets of entities.
 
@@
 \end{figure}
-\section{Correspondence Analysis, Dual Centrality, and Clique Identification}
+\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks}
 The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
@@
 As we will see, Bonacich was only half right.
-Indeed, the CA scores capture a version of ``clique'' membership, but they---particularly the first dimension---also capture a version of centrality in the most general sense of ranking nodes based on a meaningful criterion \citep{borgatti2006graph}.
+Indeed, the CA scores capture a version of ``clique'' (more accurately, community) membership, but they---particularly the first dimension---also capture a version of centrality in the most general sense of ranking nodes based on a meaningful criterion \citep{borgatti2006graph}.
 In fact, they capture the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
@@
 To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
-Matrix-algebraic reduction of the resulting equations would show that the Bonacich two-mode centralities can also be obtained as solutions to the eigensystem:
+Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:
 
@@
 Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size.
-This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more when the group in question is small than when it is large.
-On the group side, shared members who do not have many affiliations count more than those with many affiliations.
+This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large.
+On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
+
 Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
-In the Bonacich case, SVD is a one-mode projection of the two-mode network; in the reflective case, SVD of the degree-weighted one-mode projection of the same network.
+In the Bonacich case, SVD of the one-mode projection of the two-mode network; in the reflective case, SVD of the degree-weighted one-mode projection of the same network.
 Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
@@
 The centrality of groups, in turn, is a function of their members' (average) centrality.
-The person and group score ranks given by the first CA dimension are thus better referred to as activity and group-sized adjusted versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.\footnote{Adjusted in the sense described earlier of giving more weight to smaller groups and people with fewer memberships.}
+The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-sized adjusted} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.\footnote{Adjusted in the sense described earlier of giving more weight to smaller groups and people with fewer memberships.}
 
@@
 Figure~\ref{fig:ca-v-bon} illustrates the differences between $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension (a) and the Bonacich centrality (b).
+Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension (a) and the \citet{bonacich1991simultaneous} centrality (b).
 In the plot, a square is colored red if it has a one in the affiliation matrix and is colored white if it has a zero.
-As we can see, the two matrices have distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-Accordingly, the $C^R$ centrality rankings reveal a clique-like partition between two groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
+As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
+
+Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
 The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side).
-The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward clique partitioning and $C^B$ more focused on finding a core of well-connected persons and groups.
+The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected persons and groups.
 
-To further appreciate the links between CA, reflective centralities, and clique-like clustering in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be seen as giving the similarity between pairs of people weighted by the size of the groups they belong to.
-The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}.
+We can confirm that the first CA axis points toward the underlying partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}, which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
+If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
+It is clear that the Southern Women data belong to the second category.
+In the persons mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth\right]$ on one side, $\left[Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other, and  $\left[Pearl, Dorothy\right]$ in an ambiguous middle.
+Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E8, E9\right]$ in an ambiguous middle position.
+Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Plots/eigvec.png}
+    \caption{Caption}
+    \label{fig:eigvec}
+\end{figure}
+
+To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).
+The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
+As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
+These means that the relative spread of the eigenvalues of the similarity matrix Laplacian gives information concerning the quality of the resulting community partition.
+These are shown as inset plots in Figure~\ref{fig:eigvec} for both people and groups.
+Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{Plots/sim.png}
+    \caption{Caption}
+    \label{fig:sim}
+\end{figure}
+
+To fix the intuition, the top-left panel (a) of Figure~\ref{fig:sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
+In both figures, darker squares indicate more similar node pairs, and closer to the white indicate less similarity; also, the rows of each matrix are sorted according to $C^R$, namely, the scores for the first-non-trivial CA dimension (corresponding to the x-axis of the plots in Figure~\ref{fig:eigvec}) for people and groups.
+As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.
+
+The bottom panel of Figure~\ref{fig:sim} shows the correspondence plot of the first two non-trivial dimensions of the CA of the two-mode network.
+Persons and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of the respective similarity matrices.
+The plot reveals that the usual correspondence plot---usually taken to be a low dimensional representation of the original affiliation matrix---is actually best thought of as a low-dimensional representation of the (other mode) degree-weighted similarity network across people and groups.
+People who have similar values in $S_p$ appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm), as do events that have similar values in $S_g$.
 

### 2023-08-31 analysis.qmd
Updated analysis, functions, and added bonacich similarity and correspondence plots

@@
 
+```{r}
+   source(here("Functions", "TwoModeCA.R"))
+    tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 6, d.r = 6, d.c = 6, b.r = 4, b.c = 4)
+    a <- tmCA.res$bon.sim.plot.r + theme(axis.text.x = element_blank())
+    b <- tmCA.res$bon.sim.plot.c + theme(axis.text.x = element_blank())
+    c <- tmCA.res$bon.corr.plot
+
+    png(file = here("Plots", "bon-sim.png"), width=950, height=700)
+        (a | b) / c + plot_annotation(tag_levels = "a")
+    dev.off() 
+```
+
 ```{r Main Plot}
    source(here("Functions", "TwoModeCA.R"))
-    tmCA.res <- TwoModeCA(A, k.r = 7, k.c = 5, d.r = 9, d.c = 9)
+    tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 6, d.r = 9, d.c = 9)
     a <- tmCA.res$eigvec.plot.r + theme(axis.text = element_blank())

### 2023-08-31 Functions/TwoModeCA.R

@@ -1,4 +1,10 @@
-TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
+TwoModeCA <- function(x, 
+                      k.r = 2, 
+                      k.c = 2, 
+                      b.r = 2,
+                      b.c = 2,
+                      d.r = 2, 
+                      d.c = 2) {
         library(factoextra)
         library(ggcorrplot)
         library(ggpubr)
@@ -60,6 +66,27 @@ TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
                 p <- p + labs(x = "First Axis", y = "Second Axis")
                 return(p)
         }
+        bon.corr.scatter <- function(x) {
+                p <- ggscatter(data.frame(x), 
+                               x = "e1", 
+                               y = "e2", 
+                               point = FALSE,
+                               color = "cluster",
+                               palette = "uchicago",
+                               font.label =  10,
+                               label = rownames(x), 
+                               repel = TRUE)
+                p <- p + theme(legend.position = "none",
+                               axis.line.x = element_blank(),
+                               axis.line.y = element_blank())
+                p <- p + geom_vline(aes(xintercept = 0), 
+                                    color = "gray", linetype = 2)
+                p <- p + geom_hline(aes(yintercept = 0), 
+                                    color = "gray", linetype = 2)
+                p <- p + labs(x = "First Eigenvector", y = "Second Eigenvector")
+                return(p)
+        }
+        
         sim.plot <- function(x) {
                 p <- ggcorrplot(x)
                 p <- p + scale_x_discrete(position = "top")
@@ -91,31 +118,45 @@ TwoModeCA <- function(x, k.r = 2, k.c = 2, d.r = 2, d.c = 2) {
         # Derived matrices and vectors
         CA.r <- svd(iDr %*% A %*% iDc %*% At) #row CA
         CA.c <- svd(iDc %*% At %*% iDr %*% A) #column CA
[…]

### 2023-08-31 main.tex

@@
 While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
-Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') spatial visualizations of structural connectivity patterns across the two-modes \citep{bonacich1991simultaneous, breiger2000tool, faust2005using}.
+Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}.
 This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Nevertheless, it departs from the aforementioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been centered or treated in detail before.
+Nevertheless, it departs from the already-mentioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been as emphasized or treated in detail before, moving beyond the focus on data visualization.
 
 Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
-With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent linear orderings.
+With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
 This applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.
@@
 I also show that CA can be interpreted as providing \textit{positional} information on people and groups.
-First, I show that there is a clear link between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
+First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
 As such, CA recovers clusters of entities (e.g., people) linked by their similar patterns of connectivity to other similar entities in the other mode (e.g., groups).
 Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode.
-Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method---e.g., the discovery of sociometrically similar subsets of entities.
+Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}---e.g., the discovery of sociometrically similar subsets of entities in affiliation networks.
 
@@
 In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
-This leads naturally to an abbreviated way of computing the same scores via singular value decomposition (SVD) of the (inverse degree-weighted) one-mode projection of the affiliation matrix, which is computationally and mathematically equivalent to the CA of the affiliation matrix.
+This leads naturally to an abbreviated way of computing the same scores via singular value decomposition (SVD) of the (inverse degree-weighted) one-mode projection of the affiliation matrix.
+This approach is computationally and mathematically related but \textit{not} equivalent to the CA of the affiliation matrix, which results in the usual Bonacich scores.
 I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint visualization.
+In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
 
 \section{Reflective Centralities in Two-Mode Networks}
-In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach to iterated centralities in two-mode networks.
+In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
 Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as ``economic complexity.'' Subsequent shows that there is no logical connection between the formal method and this particular application.
@@
 The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
-In that paper, Bonacich also considered CA in passing but dismissed it as providing a measure of centrality.
-Instead, Bonacich noted that ``[r]ather than centrality, what they seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
+In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings.
+Instead, Bonacich noted that, in analyzing the Southern Women data shown in Table~\ref{tab:southern}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
 
@@
 Where $A$ is the original affiliation matrix.
-The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors of the affiliation matrix corresponding to the largest eigenvalue $\lambda$.
-Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of individual person and group centralities.
+The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors of the affiliation matrix corresponding to the largest eigenvalue $\lambda$, as given by the usual SVD of the affiliation matrix.
+Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
 
@@
 
-As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's reflective centralities can also be obtained as the solution to an eigenvalue problem.
+As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue problem.
 To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
@@
 
-Now, consider the $p \times p$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
-In the same way, consider the $g \times g$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-It can be shown \citep{van2021correspondence}, that in the limit, the iterative Hidalgo-Hausmann \citeyearpar{hidalgo2009building} reflective centralities emerge as the solution of the eigensystem:
+Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
+In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
+It can be shown \citep{van2021correspondence}, that in the limit, the iterative Hidalgo-Hausmann \citeyearpar{hidalgo2009building} reflective centralities can be obtained as the solution of the eigensystem:
 
@@
 Degree-pre-weighting does not radically alter the nature of the data.
-Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size.
+Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perenial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
 This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large.
@@
 Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
-Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the centrality of people is a function of the (average) centrality of the groups they belong to.
-The centrality of groups, in turn, is a function of their members' (average) centrality.
+Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to.
+The (average) centrality of groups, in turn, is a function of their members' (average) centrality.
 The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-sized adjusted} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.\footnote{Adjusted in the sense described earlier of giving more weight to smaller groups and people with fewer memberships.}

### 2023-09-01 analysis.qmd
New plots based on updated function

@@
 # Analysis
+
+```{r CA versus Bonacich Re-ordered Affiliation Matrices}
+   source(here("Functions", "TwoModeCA.R"))
+   a <- TwoModeCA(A, k.r = 5)
+
+```
+
 ```{r Reflections Bump Charts}
@@
    png(file = here("Plots", "ca-v-bon-reord.png"), width=1000, height=700)
-      (b | c) + plot_annotation(tag_levels = "a")
+      (b | c) + plot_annotation(tag_levels = "a") &
+      theme(plot.tag = element_text(size = 20))
    dev.off() 
@@
 
-```{r}
+```{r Similarity and Correspondence Plot (Bonacich)}
    source(here("Functions", "TwoModeCA.R"))
@@
     png(file = here("Plots", "bon-sim.png"), width=950, height=700)
-        (a | b) / c + plot_annotation(tag_levels = "a")
+      (a | b) / c + plot_annotation(tag_levels = "a") &
+      theme(plot.tag = element_text(size = 20)) 
     dev.off() 
@@
     png(file = here("Plots", "eigvec.png"), width=1000, height=600)
-        (a | b) + plot_annotation(tag_levels = "a")
+        (a | b) + plot_annotation(tag_levels = "a") &
+      theme(plot.tag = element_text(size = 20))
     dev.off()
@@
     png(file = here("Plots", "sim.png"), width=950, height=700)
-        (e | f) / g + plot_annotation(tag_levels = "a")
+        (e | f) / g + plot_annotation(tag_levels = "a") &
+      theme(plot.tag = element_text(size = 20))
     dev.off() 

### 2023-09-01 Functions/TwoModeCA.R

@@ -99,42 +99,55 @@ TwoModeCA <- function(x,
         A <- x[ , colSums(x) != 0, drop = FALSE]
         A <- Matrix(A, sparse = TRUE)
         At <- t(A)
-        
-        iDr <- matrix(0, nrow(A), nrow(A))
-        iDc <- matrix(0, ncol(A), ncol(A))
-        diag(iDr) <- 1/rowSums(A)
-        diag(iDc) <- 1/colSums(A)
-        
-        iDs.r <- matrix(0, nrow(A), nrow(A))
-        iDs.c <- matrix(0, ncol(A), ncol(A))
-        diag(iDs.r) <- 1/sqrt(rowSums(A))
-        diag(iDs.c) <- 1/sqrt(colSums(A))
-        
-        r <- nrow(A)
-        c <- ncol(A)
+        r <- nrow(A) # number of rows
+        c <- ncol(A) # number of columns
         rn <- rownames(A)
         cn <- colnames(A)
+        u.r <- matrix(1, nrow = r, ncol = 1)
+        u.c <- matrix(1, nrow = c, ncol = 1)
+
         
         # Derived matrices and vectors
-        CA.r <- svd(iDr %*% A %*% iDc %*% At) #row CA
-        CA.c <- svd(iDc %*% At %*% iDr %*% A) #column CA
-        CA.res <- svd(iDs.r %*% A %*% iDs.c) #row and column CA
-        bon.res <- svd(A) #Bonacich eigenvector centrality
-        
+        AAt <- A %*% At #row projection
+        AtA <- At %*% A #column projection
+        N <- as.numeric(t(u.r) %*% A %*% u.c)
+        D.r <- diag(as.vector(A %*% u.c), r, r)
+        D.c <- diag(as.vector(At %*% u.r), c, c)
+        iDr <- solve(D.r)
+        iDc <- solve(D.c)
         S.r <- A %*% iDc %*% At #degree-normalized similarity matrix for rows
         S.c <- At %*% iDr %*% A #degree-normalized similarity matrix for columns
         
-        Su.r <- A %*% At #raw similarity matrix for rows
-        Su.c <- At %*% A #raw similarity matrix for columns
+        # Spectral matrices
+        CA.r <- eigen(iDr %*% A %*% iDc %*% At) #row CA
+        CA.c <- eigen(iDc %*% At %*% iDr %*% A) #column CA
+        S.r.eig <- eigen(S.r) #eigendecomposition of row similarity matrix
[…]

### 2023-09-01 Functions/sca.R

@@ -1,0 +1,103 @@
+# K-means clustering of weighted similarity matrices
+S.r.dat <- data.frame(eigen(S.r[, 1:d.r])$vectors)
+S.c.dat <- data.frame(eigen(S.c[, 1:d.c])$vectors)
+rownames(S.r.dat) <- rn
+rownames(S.c.dat) <- cn
+km.r <- hkmeans(S.r.dat, k.r)
+km.c <- hkmeans(S.c.dat, k.c)
+km.dat.r <- data.frame(lab = rn, cluster = factor(km.r$cluster))
+km.dat.c <- data.frame(lab = cn, cluster = factor(km.c$cluster))
+
+# K-means clustering of unweighted similarity matrices
+Su.r.dat <- data.frame(eigen(P.r[, 1:d.r])$vectors)
+Su.c.dat <- data.frame(eigen(P.c[, 1:d.c])$vectors)
+rownames(Su.r.dat) <- rn
+rownames(Su.c.dat) <- cn
+km.r <- hkmeans(Su.r.dat, b.r)
+km.c <- hkmeans(Su.c.dat, b.c)
+bon.km.dat.r <- data.frame(lab = rn, cluster = factor(km.r$cluster))
+bon.km.dat.c <- data.frame(lab = cn, cluster = factor(km.c$cluster))
+
+# Main Eigenvector Plot
+eigvec.dat.r <- data.frame(rank = rank(eigvec.r[, 1]), 
+                           value = eigvec.r[, 1],
+                           lab = rn) %>% 
+     left_join(km.dat.r)
+eigvec.dat.c <- data.frame(rank = rank(eigvec.c[, 1]), 
+                           value = eigvec.c[, 1],
+                           lab = cn)  %>% 
+     left_join(km.dat.c)
+eigvec.plot.r <- eigvec.scatter(eigvec.dat.r)
+eigvec.plot.c <- eigvec.scatter(eigvec.dat.c)
+
+# Correspondence Plots (CA)
+corr.plot.r <- corr.scatter(eigvec.r)
+corr.plot.c <- corr.scatter(eigvec.c)
+eigvec.rc <- rbind(eigvec.r[, 1:2], eigvec.c[, 1:2])
+km.dat.rc <- rbind(km.dat.r, km.dat.c)
+eigvec.rc <- cbind(eigvec.rc, km.dat.rc)
+corr.plot <- corr.scatter(eigvec.rc)
+
[…]

### 2023-09-01 main.tex
Updates from Overleaf

@@
 The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side).
-The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected persons and groups.
+The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
 
@@
 If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
-It is clear that the Southern Women data belong to the second category.
-In the persons mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth\right]$ on one side, $\left[Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other, and  $\left[Pearl, Dorothy\right]$ in an ambiguous middle.
+The Southern Women data clearly belong to the second category.
+In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth\right]$ on one side, $\left[Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other, and  $\left[Pearl, Dorothy\right]$ in an ambiguous middle.
 Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E8, E9\right]$ in an ambiguous middle position.
@@
     \caption{Caption}
-    \label{fig:sim}
+    \label{fig:ca-sim}
 \end{figure}
 
-To fix the intuition, the top-left panel (a) of Figure~\ref{fig:sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
+To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
 In both figures, darker squares indicate more similar node pairs, and closer to the white indicate less similarity; also, the rows of each matrix are sorted according to $C^R$, namely, the scores for the first-non-trivial CA dimension (corresponding to the x-axis of the plots in Figure~\ref{fig:eigvec}) for people and groups.
@@
 Persons and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of the respective similarity matrices.
-The plot reveals that the usual correspondence plot---usually taken to be a low dimensional representation of the original affiliation matrix---is actually best thought of as a low-dimensional representation of the (other mode) degree-weighted similarity network across people and groups.
-People who have similar values in $S_p$ appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm), as do events that have similar values in $S_g$.
+The plot reveals that the usual correspondence plot---usually taken to be a low-dimensional representation of the original affiliation matrix \citep{borgatti1997network}---is best thought of as a low-dimensional representation of the (other mode's) degree-weighted similarity network across people and groups.
+People with similar values in $S_p$ appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm), as do events with similar values in $S_g$.
 
+Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to imply (for instance) that two people attend the same set of events, or that two groups share the same set of members.
+What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar after taking people's activity levels and the sizes of the groups they belong to into account.
+Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
+In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{Plots/bon-sim.png}
+    \caption{Caption}
+    \label{fig:bon-sim}
+\end{figure}
+
+This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
+But as we have seen, this is \textit{not} what CA is designed to do.
+Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes.
+Oddly, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two dimensions of the SVD of this matrix.
+
+Figure~\ref{fig:bon-sim} shows the ``raw'' (unweighted) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score.
+Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix on the top-right panel of Figure~\ref{fig:ca-v-bon}.
+The bottom panel of the figure plots $C^B$ on the x-axis against the second eigenvector of the SVD solution---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}
+
+We can see that the plot of the first two SVD dimensions of $\mathbf{A}$ do a good job of recovering the ``raw'' connectivity structure of the network, partitioning the ``core'' persons and groups (on the lower-right) from the more peripheral ones (on the upper left).
+Moreover, if we are all interested in capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then this plot does a better job showing that than the usual CA correspondence plot in the bottom half of Figure~\ref{fig:ca-sim} For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine\right]$ for $\left[E10, E12, E13, E14\right]$.
+In the same way, core events like $\left[E8, E9, E10\right]$---shown on the upper side of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the lower quadrant.
+The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.
+Conversely, $\left[E1, E2]\right]$, although as poorly attended as other peripheral events, tend to include core members and thus appear closer to those actors in the bottom half of the plot.
 %% Loading bibliography style file

### 2023-09-01 Functions/TwoModeCA.R
Updated bonacich plots

@@ -135,6 +135,10 @@ TwoModeCA <- function(x,
         bon.eigvec.c <- as.matrix(eigen(AtA)$vectors) #column eigenvectors
         bon.eigval.r <- as.matrix(eigen(AAt)$values)  #row eigenvalues
         bon.eigval.c <- as.matrix(eigen(AtA)$values)  #row eigenvalues
+        bon.eigvec.r[, 1] <- bon.eigvec.r[, 1] * -1 #positive values more centrality
+        bon.eigvec.c[, 1] <- bon.eigvec.c[, 1] * -1 #positive values more centrality
+        bon.eigvec.r[, 2] <- bon.eigvec.r[, 2] * -1 #positive values more centrality
+        bon.eigvec.c[, 2] <- bon.eigvec.c[, 2] * -1 #positive values more centrality
         
         # Normalizing row and column vectors
         norm.r <- function(x) {

### 2023-09-07 analysis.qmd
New simrank analysis and plot

@@
    library(dplyr)
+   library(factoextra)
    library(ggbump)
+   library(ggcorrplot)
    library(ggplot2)
+   library(ggpubr)
    library(ggrepel)
@@
    library(pals)
+   library(patchwork)
    source(here("Functions", "TwoModeCA.R"))
+   source(here("Functions", "TwoModeSR.R"))
    source(here("Functions", "reflections.R"))
@@
 # Analysis
-
-```{r CA versus Bonacich Re-ordered Affiliation Matrices}
-   source(here("Functions", "TwoModeCA.R"))
-   a <- TwoModeCA(A, k.r = 5)
-
-```
 
@@
    a <- TwoModeCA(A)
-   b <- a$A.plot.ca
-   c <- a$A.plot.bon
+   b <- a$ca.A
+   c <- a$bon.A
    png(file = here("Plots", "ca-v-bon-reord.png"), width=1000, height=700)
@@
 
-```{r Similarity and Correspondence Plot (Bonacich)}
+```{r Eigenvector Plot}
    source(here("Functions", "TwoModeCA.R"))
-    tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 6, d.r = 6, d.c = 6, b.r = 4, b.c = 4)
-    a <- tmCA.res$bon.sim.plot.r + theme(axis.text.x = element_blank())
-    b <- tmCA.res$bon.sim.plot.c + theme(axis.text.x = element_blank())
-    c <- tmCA.res$bon.corr.plot
-
-    png(file = here("Plots", "bon-sim.png"), width=950, height=700)
-      (a | b) / c + plot_annotation(tag_levels = "a") &
-      theme(plot.tag = element_text(size = 20)) 
-    dev.off() 
-```
-
-```{r Main Plot}
-   source(here("Functions", "TwoModeCA.R"))
-    tmCA.res <- TwoModeCA(A, k.r = 6, k.c = 6, d.r = 9, d.c = 9)
+    tmCA.res <- TwoModeCA(A, k.r = 2, k.c = 2)
     a <- tmCA.res$eigvec.plot.r + theme(axis.text = element_blank())
@@
     b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
-    e <- tmCA.res$sim.plot.r + theme(axis.text.x = element_blank())
-    f <- tmCA.res$sim.plot.c + theme(axis.text.x = element_blank())
-    g <- tmCA.res$corr.plot
 
@@
     dev.off() 
+```
 
-    png(file = here("Plots", "sim.png"), width=950, height=700)
-        (e | f) / g + plot_annotation(tag_levels = "a") &
+```{r Similarity and Correspondence Plot (CA)}
+   source(here("Functions", "TwoModeCA.R"))
+    tmCA.res <- TwoModeCA(A, d.r = 4, d.c = 4, k.r = 3, k.c = 3)
+    a <- tmCA.res$ca.sim.plot.r + theme(axis.text.x = element_blank())
+    b <- tmCA.res$ca.sim.plot.c + theme(axis.text.x = element_blank())
+    c <- tmCA.res$ca.corr.plot
+
+    png(file = here("Plots", "ca-sim.png"), width=950, height=700)
+      (a | b) / c + plot_annotation(tag_levels = "a") &
       theme(plot.tag = element_text(size = 20)) 
@@
 ```
+
+```{r Similarity and Correspondence Plot (Bonacich)}
+   source(here("Functions", "TwoModeCA.R"))
+    tmCA.res <- TwoModeCA(A, d.r = 4, d.c = 4, b.r = 4, b.c = 4)
+    a <- tmCA.res$bon.sim.plot.r + theme(axis.text.x = element_blank())
+    b <- tmCA.res$bon.sim.plot.c + theme(axis.text.x = element_blank())
+    c <- tmCA.res$bon.corr.plot
+
+    png(file = here("Plots", "bon-sim.png"), width=950, height=700)
+      (a | b) / c + 
+       plot_annotation(tag_levels = "a") &
+      theme(plot.tag = element_text(size = 20)) 
+    dev.off() 
+```
+
+```{r SimRank Plot}
+   source(here("Functions", "TwoModeSR.R"))
+   tmSR.res <- TwoModeSR(A, C = 0.8, iter = 10)
+    a <- tmSR.res$eigvec.plot.r + theme(axis.text = element_blank())
+    b <- tmSR.res$eigvec.plot.c 
+    c <- tmSR.res$eigval.plot.r
+    d <- tmSR.res$eigval.plot.c
+    #a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    #b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    e <- tmSR.res$A.plot
+    f <- tmSR.res$corr.plot
+   dat.r <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.r[,1]),
+                       sr = tmSR.res$ov.r,
+                       lab = rownames(A))
+   dat.c <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.c[,1]),
+                       sr = tmSR.res$ov.c,
+                       lab = colnames(A))
+   g <- ggscatter(dat.r, x = "ca", y = "sr", 
+                  label = "lab", repel = TRUE, 
+                  add = "reg.line", 
+                  add.params = list(color = "gray80", linewidth = 0.25))
+   h <- ggscatter(dat.c, x = "ca", y = "sr", 
+                  label = "lab", repel = TRUE, 
+                  add = "reg.line",
+                  add.params = list(color = "gray80", linewidth = 0.25))
+   i <- (a | b | (g / h)) + plot_layout(widths = c(2, 2, 1))
+   j <- (e | f) + plot_layout(widths = c(1, 3))
+
+   png(file = here("Plots", "sr-plot.png"), width=1000, height=1200)
+       (i / j ) +
+       plot_annotation(tag_levels = "a") &
+       theme(plot.tag = element_text(size = 20))
+    dev.off()
+
+```
+
+```{r}
+```
+
+```
+
+```{r}
+
+```

### 2023-09-07 Functions/TwoModeCA.R

@@ -5,11 +5,6 @@ TwoModeCA <- function(x,
                       b.c = 2,
                       d.r = 2, 
                       d.c = 2) {
-        library(factoextra)
-        library(ggcorrplot)
-        library(ggpubr)
-        library(Matrix)
-
         # Plot Functions
         eigval.scatter <- function(x) {
                 p <- ggscatter(x, 
@@ -86,7 +81,6 @@ TwoModeCA <- function(x,
                 p <- p + labs(x = "First Eigenvector", y = "Second Eigenvector")
                 return(p)
         }
-        
         sim.plot <- function(x) {
                 p <- ggcorrplot(x)
                 p <- p + scale_x_discrete(position = "top")
@@ -96,25 +90,23 @@ TwoModeCA <- function(x,
         }
         
         # Initial matrices and vectors
-        A <- x[ , colSums(x) != 0, drop = FALSE]
-        A <- Matrix(A, sparse = TRUE)
-        At <- t(A)
+        A <- x[ , colSums(x) != 0, drop = FALSE] #dropping isolates
+        At <- t(A) #affiliation matrix transpose
         r <- nrow(A) # number of rows
         c <- ncol(A) # number of columns
         rn <- rownames(A)
         cn <- colnames(A)
-        u.r <- matrix(1, nrow = r, ncol = 1)
-        u.c <- matrix(1, nrow = c, ncol = 1)
+        u.r <- matrix(1, nrow = r, ncol = 1) #unit row vector
+        u.c <- matrix(1, nrow = c, ncol = 1) #unit column vector
 
-        
         # Derived matrices and vectors
         AAt <- A %*% At #row projection
         AtA <- At %*% A #column projection
-        N <- as.numeric(t(u.r) %*% A %*% u.c)
-        D.r <- diag(as.vector(A %*% u.c), r, r)
-        D.c <- diag(as.vector(At %*% u.r), c, c)
-        iDr <- solve(D.r)
-        iDc <- solve(D.c)
+        N <- as.numeric(t(u.r) %*% A %*% u.c) #affiliation matrix size
+        D.r <- diag(as.vector(A %*% u.c), r, r) #row diagonal degree matrix
+        D.c <- diag(as.vector(At %*% u.r), c, c) #column diagonal degree matrix
+        iDr <- solve(D.r) #inverse of row diagonal degree matrix
+        iDc <- solve(D.c)  #inverse of column diagonal degree matrix
[…]

### 2023-09-07 Functions/TwoModeSR.R

@@ -1,0 +1,171 @@
+TwoModeSR <- function(A, C = 0.8, iter = 10, k = 4) {
+   nr <- nrow(A)
+   nc <- ncol(A)
+   dr <- rowSums(A)
+   dc <- colSums(A)
+   Sr <- diag(1, nr, nr)
+   Sc <- diag(1, nc, nc)
+   rn <- rownames(A)
+   cn <- colnames(A)
+   rownames(Sr) <- rn
+   colnames(Sr) <- rn
+   rownames(Sc) <- cn
+   colnames(Sc) <- cn
+   m <- 1
+   while(m < iter) {
+      Sr.pre <- Sr
+      Sc.pre <- Sc
+      for(i in 1:nr) {
+         for(j in 1:nr) {
+            if (i != j) {
+               a <- names(which(A[i, ] == 1)) #objects chosen by i
+               b <- names(which(A[j, ] == 1)) #objects chosen by j
+               Scij <- 0
+               for (k in a) {
+                  for (l in b) {
+                     Scij <- Scij + Sc[k, l] #i's similarity to j
+                     }
+                  }
+                  Sr[i, j] <- C/(dr[i] * dr[j]) * Scij
+               }
+            }
+         }
+      for(i in 1:nc) {
+         for(j in 1:nc) {
+            if (i != j) {
+               a <- names(which(A[, i] == 1)) #people who chose object i
+               b <- names(which(A[, j] == 1)) #people who chose object j
+               Srij <- 0
+               for (k in a) {
+                  for (l in b) {
+                     Srij <- Srij + Sr[k, l] #i's similarity to j
+                     }
+                  }
+                  Sc[i, j] <- C/(dc[i] * dc[j]) * Srij
+               }
+            }
+         }
+      m <- m + 1
+   }
+
+   corr.scatter <- function(x) {
+      p <- ggscatter(data.frame(x), 
+                     x = "d1", 
+                     y = "d2", 
+                     point = FALSE,
+                     color = "cluster",
+                     palette = "uchicago",
+                     font.label =  13,
+                     label = "lab", 
+                     repel = TRUE)
+      p <- p + theme(legend.position = "none",
+                     axis.line.x = element_blank(),
+                     axis.line.y = element_blank())
+      p <- p + geom_vline(aes(xintercept = 0), 
+                          color = "gray", linetype = 2)
[…]

### 2023-09-07 Functions/discard/TwoModeCE.R

@@ -1,0 +1,63 @@
+TwoModeCE <- function(W, 
+                      d.1 = 1, d.2 = 2,
+                      k.r = 4, k.c = 4, 
+                      d.r = 8, d.c = 8) {
+     library(igraph)
+     
+     # Initial matrices and vectors
+     A <- W[ , colSums(W) != 0, drop = FALSE] #dropping isolates
+     At <- t(A) #affiliation matrix transpose
+     r <- nrow(A) # number of rows
+     c <- ncol(A) # number of columns
+     rn <- rownames(A)
+     cn <- colnames(A)
+     u.r <- matrix(1, nrow = r, ncol = 1) #unit row vector
+     u.c <- matrix(1, nrow = c, ncol = 1) #unit column vector
+     
+     # Derived matrices and vectors
+     AAt <- A %*% At #row projection
+     AtA <- At %*% A #column projection
+     N <- as.numeric(t(u.r) %*% A %*% u.c) #affiliation matrix size
+     D.r <- diag(as.vector(A %*% u.c), r, r) #row diagonal degree matrix
+     D.c <- diag(as.vector(At %*% u.r), c, c) #column diagonal degree matrix
+     iDr <- solve(D.r) #inverse of row diagonal degree matrix
+     iDc <- solve(D.c)  #inverse of column diagonal degree matrix
+     S.r <- A %*% iDc %*% At #degree-normalized similarity matrix for rows
+     S.c <- At %*% iDr %*% A #degree-normalized similarity matrix for columns
+     
+     Z.r <- matrix(0, r, r)
+     Z.c <- matrix(0, c, c)
+     B <- rbind(cbind(Z.r, A), cbind(At, Z.c))
+     colnames(B) <- rownames(B)
+     B.gr <- graph_from_adjacency_matrix(B)
+     clos <- closeness(B.gr)
+     iCr <- solve(diag(clos[1:r], r, r))
+     a <- r + 1
+     iCc <- solve(diag(clos[a:length(clos)], c, c))
+     CE.r <- eigen(iCr %*% A %*% iCc %*% At)$vectors #row CE
+     CE.c <- eigen(iCc %*% At %*% iCr %*% A)$vectors #column CE
+     S.r <- A %*% iCc %*% At #degree-normalized similarity matrix for rows
+     S.c <- At %*% iCr %*% A #degree-normalized similarity matrix for columns
+     CE.r <- apply(CE.r, 2, as.numeric)
[…]

### 2023-09-07 main.tex

@@
 Summed across groups $g$, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
-As noted, the Bonacich dual centralities are obtained from an SVD of this matrix of shared memberships (for people) and shared people (for groups).
-The reflective centralities, on the other hand, are obtained from an SVD of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
+As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the larger eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
+The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
 
@@
 
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used before performing the SVD of a two-mode matrix to obtain the corresponding row and column scores in CA.\footnote{More accurately, cells are weighted by inverse of the square root of the product of the row and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, as has recently been shown \citep{van2021correspondence}, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be substantively equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a CA of the original affiliation matrix.\footnote{As \citet[126]{faust2005using}, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of the SVD solution.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
+Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by inverse of the square root of the product of the row and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
+9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using}, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
 
@@
 The (average) centrality of groups, in turn, is a function of their members' (average) centrality.
-The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-sized adjusted} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.\footnote{Adjusted in the sense described earlier of giving more weight to smaller groups and people with fewer memberships.}
+The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-sized adjusted} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
 
@@
     \includegraphics[width=0.8\textwidth]{Plots/ca-v-bon-reord.png}
-    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to first CA dimension, (b) row and columns re-ordered according to Eigenvector centrality.}
+    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score.}
     \label{fig:ca-v-bon}
@@
 
-Figure~\ref{fig:ca-v-bon} illustrates the differences between $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension (a) and the \citet{bonacich1991simultaneous} centrality (b).
-In the plot, a square is colored red if it has a one in the affiliation matrix and is colored white if it has a zero.
+Figure~\ref{fig:ca-v-bon} illustrates the key differences between the two versions of dual centrality $C^R$ and $C^B$.
+Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
+In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
 As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-
 Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
@@
 
-We can confirm that the first CA axis points toward the underlying partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}, which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
+We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}, which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
 If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
 The Southern Women data clearly belong to the second category.
-In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth\right]$ on one side, $\left[Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other, and  $\left[Pearl, Dorothy\right]$ in an ambiguous middle.
-Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E8, E9\right]$ in an ambiguous middle position.
+In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other.
+Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
 Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach.
+This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.
 
@@
 As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
-These means that the relative spread of the eigenvalues of the similarity matrix Laplacian gives information concerning the quality of the resulting community partition.
+This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition.
 These are shown as inset plots in Figure~\ref{fig:eigvec} for both people and groups.
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{Plots/sim.png}
+    \includegraphics[width=1.0\textwidth]{Plots/ca-sim.png}
     \caption{Caption}
@@
 To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
-In both figures, darker squares indicate more similar node pairs, and closer to the white indicate less similarity; also, the rows of each matrix are sorted according to $C^R$, namely, the scores for the first-non-trivial CA dimension (corresponding to the x-axis of the plots in Figure~\ref{fig:eigvec}) for people and groups.
+In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity.
+Also, the rows of each matrix are sorted according to $C^R$ for both people and groups.
 As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.
 
-The bottom panel of Figure~\ref{fig:sim} shows the correspondence plot of the first two non-trivial dimensions of the CA of the two-mode network.
-Persons and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of the respective similarity matrices.
-The plot reveals that the usual correspondence plot---usually taken to be a low-dimensional representation of the original affiliation matrix \citep{borgatti1997network}---is best thought of as a low-dimensional representation of the (other mode's) degree-weighted similarity network across people and groups.
+The bottom panel of Figure~\ref{fig:ca-sim} shows the correspondence plot of the first two CA dimensions.
+People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices.
+The plot reveals that the usual correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---is best thought of as a low-dimensional representation of the (other mode's) degree-weighted similarity network across people and groups.
 People with similar values in $S_p$ appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm), as do events with similar values in $S_g$.

### 2023-09-08 analysis.qmd
Updated simrank plot

@@
 
-    png(file = here("Plots", "eigvec.png"), width=1000, height=600)
+    png(file = here("Plots", "ca-eigvec.png"), width=1000, height=600)
         (a | b) + plot_annotation(tag_levels = "a") &
@@
                   add = "reg.line", 
-                  add.params = list(color = "gray80", linewidth = 0.25))
+                  add.params = list(color = "gray80", linewidth = 0.5, linetype = 2))
    h <- ggscatter(dat.c, x = "ca", y = "sr", 
@@
                   add = "reg.line",
-                  add.params = list(color = "gray80", linewidth = 0.25))
-   i <- (a | b | (g / h)) + plot_layout(widths = c(2, 2, 1))
-   j <- (e | f) + plot_layout(widths = c(1, 3))
-
-   png(file = here("Plots", "sr-plot.png"), width=1000, height=1200)
-       (i / j ) +
+                  add.params = list(color = "gray80", linewidth = 0.5, linetype = 2))
+   i <- (a | b | e) 
+   j <- (f | (g / h)) + plot_layout(widths = c(2.5, 1))
+   png(file = here("Plots", "sr-plot.png"), width=1200, height=1200)
+       (i / j) +
        plot_annotation(tag_levels = "a") &
@@
 ```
-
-```{r}
-```
-
-```
-
-```{r}
-
-```

### 2023-09-08 Functions/TwoModeSR.R

@@ -144,12 +144,12 @@ TwoModeSR <- function(A, C = 0.8, iter = 10, k = 4) {
    k.r <- hkmeans(data.frame(e.r$vectors[, 2:2]), k = 2)
    k.c <- hkmeans(data.frame(e.c$vectors[, 2:2]), k = 2)
    
-   eigvec.dat.r <- data.frame(rank = rank(ov.r), 
-                              value = ov.r,
+   eigvec.dat.r <- data.frame(rank = rank(ov.r*-1), 
+                              value = ov.r*-1,
                               lab = rn,
                               cluster = as.factor(k.r$cluster))
-   eigvec.dat.c <- data.frame(rank = rank(ov.c), 
-                              value = ov.c,
+   eigvec.dat.c <- data.frame(rank = rank(ov.c*-1), 
+                              value = ov.c*-1,
                               lab = cn,
                               cluster = as.factor(k.c$cluster))
    eigvec.plot.r <- eigvec.scatter(eigvec.dat.r)

### 2023-09-08 main.tex

@@
 With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
-This applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.
+In social network analysis, this applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.
 
-I also show that CA can be interpreted as providing \textit{positional} information on people and groups.
+Additionally, I argue that CA can also be used to extract latent \textit{positional} information on people and groups.
 First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
-As such, CA recovers clusters of entities (e.g., people) linked by their similar patterns of connectivity to other similar entities in the other mode (e.g., groups).
+As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups).
 Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode.
@@
 In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
-This leads naturally to an abbreviated way of computing the same scores via singular value decomposition (SVD) of the (inverse degree-weighted) one-mode projection of the affiliation matrix.
-This approach is computationally and mathematically related but \textit{not} equivalent to the CA of the affiliation matrix, which results in the usual Bonacich scores.
+This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes.
+This approach is computationally and mathematically related but \textit{not} equivalent to the matrix factorization of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores.
 I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
+I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers instead the core-periphery structure of the two-mode network.
 In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
@@
 In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
-Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as ``economic complexity.'' Subsequent shows that there is no logical connection between the formal method and this particular application.
-As such, I introduce it here using the more intuitive---and classical from a social science perspective---case of the duality of persons and groups \citep{breiger1974duality}.
+Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).
+Subsequent shows that there is no logical connection between the formal method and this particular application.
+As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
 
@@
 That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
+In what follows, I will generally refer to $C^R_p(1)$ as (individual) ``activity'' and $C^R_p(1)$ as (group) ``size.''
+
 As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.
@@
 
-Equation \ref{eq:R1_p} says ``people are more central when the average sum of the memberships of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
-Equation \ref{eq:R2_g} says ``groups are more central when the average sum of the number of memberships of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
+Equation \ref{eq:R2_p} says ``people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
+Equation \ref{eq:R2_g} says ``groups are more central when the average activity of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
 Of course, we can keep on going and define third-order reflections.
@@
 
-Equation \ref{eq:R3_p} says something like ``people are more central when the average of the average number of memberships of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
-Equation \ref{eq:R3_g} says, ''groups are more central when the average number of members of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
+Equation \ref{eq:R3_p} says something like ``people are more central when the average sum of the average activity of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
+Equation \ref{eq:R3_g} says, ''groups are more central when the average sum of the average size of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
 
-Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, their degree centrality).
-On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to.
-In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their degree centrality).
-On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{groups their members belong to}.
-While these are distinct metrics, in practice, the order of nodes in each mode ends up being identical when the ranks ``freeze'' after some number of iterations (proportional to the network size).
+Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes).
+On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to).
+In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity).
+On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size).
+While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).
 
@@
 
-Equation~\ref{eq:Rq_p} says that the reflective centrality of person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
+Equation~\ref{eq:Rq_p} says that the reflective centrality of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
 Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
@@
 \subsection{Empirical Example}
-Figure~\ref{fig:pg-reflections} shows the trajectory of the Hidalgo-Hausmann reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}.
+Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}.
 The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
@@
 Where $A$ is the original affiliation matrix.
-The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors of the affiliation matrix corresponding to the largest eigenvalue $\lambda$, as given by the usual SVD of the affiliation matrix.
+The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, as given by the singular value decomposition (SVD) of the rectangular affiliation matrix.
 Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
@@
 
+Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
 These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
@@
 
-As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue problem.
+As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue (matrix factorization) problem.
 To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
@@
 
+This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.
+
 Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
 In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-It can be shown \citep{van2021correspondence}, that in the limit, the iterative Hidalgo-Hausmann \citeyearpar{hidalgo2009building} reflective centralities can be obtained as the solution of the eigensystem:
+It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as the solution of the eigensystem:
 
@@
 Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups.
-The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode, essentially weighting the projection by node degree before performing the SVD.
+The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before factorizing, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
 
@@
 
-In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share a membership.
-Summed across groups $g$, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
-As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the larger eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
+In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$.
+Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
+As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
 The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
@@
 Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
-In the Bonacich case, SVD of the one-mode projection of the two-mode network; in the reflective case, SVD of the degree-weighted one-mode projection of the same network.
+In the Bonacich case, a factorization of the one-mode projection of the two-mode network; in the reflective case, factorization of the degree-weighted one-mode projection of the same network.
 Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
@@
 The (average) centrality of groups, in turn, is a function of their members' (average) centrality.
-The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-sized adjusted} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
+The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
 
@@
 
-Figure~\ref{fig:ca-v-bon} illustrates the key differences between the two versions of dual centrality $C^R$ and $C^B$.
+\section{Illustrative Analysis of the Southern Women Data}
+\subsection{Row and Column Re-ordered Affiliation Matrices}
+Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
 Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
@@
 
-We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}, which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
+\subsection{Eigenvector Plot}
+We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}(a) and~\ref{fig:eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
 If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
@@
 
-To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).
-The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
+\subsection{Normalized Similarity Plot}
+To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one.
+Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
 As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
@@
 In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity.
-Also, the rows of each matrix are sorted according to $C^R$ for both people and groups.
+Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups.
 As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.
 
-The bottom panel of Figure~\ref{fig:ca-sim} shows the correspondence plot of the first two CA dimensions.
+\subsection{CA Correspondence Plot}
+The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions.
 People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices.
-The plot reveals that the usual correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---is best thought of as a low-dimensional representation of the (other mode's) degree-weighted similarity network across people and groups.
-People with similar values in $S_p$ appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm), as do events with similar values in $S_g$.
+The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups.
+People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm).
+The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart.
+The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.
 
-Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to imply (for instance) that two people attend the same set of events, or that two groups share the same set of members.
-What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar after taking people's activity levels and the sizes of the groups they belong to into account.
-Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
+Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups).
+This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot.
+But this (still common) interpretation is off the mark.
+What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs.
+2.13 and 2.14]{fouss2016algorithms}.
+As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
 In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.
@@
 
+\subsection{One-Mode Projection Matrix and Eigevenctor Plot}
 This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
@@
 Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes.
-Oddly, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two dimensions of the SVD of this matrix.
+Notably, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two eigenvectors of the usual one-mode projections of the matrix (where the first is, of course, the standard Bonacich dual centrality score).
 
-Figure~\ref{fig:bon-sim} shows the ``raw'' (unweighted) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score.
-Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix on the top-right panel of Figure~\ref{fig:ca-v-bon}.
-The bottom panel of the figure plots $C^B$ on the x-axis against the second eigenvector of the SVD solution---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}
+To illustrate, Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by other-mode degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score.
+Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
+Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}  We can see that the plot of the first two eigenvectors does a good job of recovering the raw connectivity structure of the Southern Women affiliation network, partitioning the core persons and groups (on the upper-right) from the more peripheral ones (on the lower-left).
 
-We can see that the plot of the first two SVD dimensions of $\mathbf{A}$ do a good job of recovering the ``raw'' connectivity structure of the network, partitioning the ``core'' persons and groups (on the lower-right) from the more peripheral ones (on the upper left).
-Moreover, if we are all interested in capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then this plot does a better job showing that than the usual CA correspondence plot in the bottom half of Figure~\ref{fig:ca-sim} For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine\right]$ for $\left[E10, E12, E13, E14\right]$.
-In the same way, core events like $\left[E8, E9, E10\right]$---shown on the upper side of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the lower quadrant.
+Moreover, if all we are interested in is capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then Figure~\ref{fig:bon-sim}(c) does a better job of that than the usual CA correspondence plot in Figure~\ref{fig:ca-sim}(c).
+For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine, Sylvia\right]$ for $\left[E10, E12\right]$.
+In the same way, core events like $\left[E8, E9, E10\right]$---shown on the lower half of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the upper half.
 The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.
-Conversely, $\left[E1, E2]\right]$, although as poorly attended as other peripheral events, tend to include core members and thus appear closer to those actors in the bottom half of the plot.
+
+This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's eigenvector centrality.
+For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot.
+Similarly, the main difference between $\left[Ruth\right]$ and $\left[Sylvia\right]$, despite their comparable $C^B$ scores, is that the former's (closer to the upper half of the plot) event profile is composed mainly of core events.
+In contrast, peripheral events dominate the latter's attendance profile (shown in the bottom half of the plot), accordingly, $\left[Ruth\right]$ and $\left[Sylvia\right]$ are assigned to distinct clusters by the k-means algorithm.
+
+\subsection{Correspondence Analysis and Generalized Similarities}
+As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.
+Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}.
+This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence.
+In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.
+
+An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}.
+In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong, and the similarity of groups is a function of the people who belong to them, making the similarity of people and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
+Thus,
+
+\begin{itemize}
+    \item People are \textit{similar} if they belong to \textit{similar} groups.
+    \item Groups are \textit{similar} if they share \textit{similar} members.
+\end{itemize}
+
+Which is consistent with a GSR approach. To do this, \citet[540, eq2 and eq3]{jeh2002simrank} propose that we define the similarity of people as given by:
+
+\begin{equation}
+    S(p, p') = \frac{\alpha}{C^R(1)_pC^R(1)_{p'}}
+    \sum_{i = 1}^{C^R(1)_p} \sum_{j = 1}^{C^R(1)_{p'}} 
+    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
+    \label{eq:simrank1}
+\end{equation}
+
+Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to; $\alpha$ is a free parameter, with the restriction: $0 > \alpha < 1$.
+By construction, $S(p, p) = 1$.
+Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of the groups they belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.
+
+Likewise, for groups, the SimRank similarities are given by:
+
+\begin{equation}
+    S(g, g') = \frac{\alpha}{C^R(1)_gC^R(1)_{g'}}
+    \sum_{i = 1}^{C^R(1)_g} \sum_{j = 1}^{C^R(1)_{g'}} 
+    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
+    \label{eq:simrank2}
+\end{equation}
+
+Where $p(i)_{i \in N(g)}$ is the $i^{th}$ person in the set of members of group $g$, and $p(j)_{j \in N(g')}$ is the $j^{th}$ person in the set of members of group $g'$, and $S(g, g) = 1$.
+Thus, the SimRank similarity between two groups is a function of the sum of the similarities of the people who belong to them, weighted by the reciprocal of the products of their number of members multiplied by $\alpha$.
+
+SimRank generalized similarities for nodes in two-mode networks can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
+Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
+We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}.
+
+\section{Discussion}
 %% Loading bibliography style file

### 2023-09-08 main.tex

@@
 \subsection{Eigenvector Plot}
-We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}(a) and~\ref{fig:eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
+We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
 If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{Plots/eigvec.png}
+    \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}
     \caption{Caption}
-    \label{fig:eigvec}
+    \label{fig:ca-eigvec}
 \end{figure}
@@
 This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition.
-These are shown as inset plots in Figure~\ref{fig:eigvec} for both people and groups.
+These are shown as inset plots in Figure~\ref{fig:ca-eigvec} for both people and groups.
 Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.
@@
 
-Which is consistent with a GSR approach. To do this, \citet[540, eq2 and eq3]{jeh2002simrank} propose that we define the similarity of people as given by:
+Which is consistent with a GRS approach. To do this, \citet[540, eq2 and eq3]{jeh2002simrank} propose that we define the similarity of people as given by:
 
@@
 
-SimRank generalized similarities for nodes in two-mode networks can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
+\begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{Plots/sr-plot.png}
+    \caption{Caption}
+    \label{fig:simrank}
+\end{figure}
+
+In two-mode networks, SimRank scores can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
 Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
 We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}.
+Further Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with the HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
+The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
+Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA.
+
+And indeed, they are not.
+Figure~\ref{fig:simrank}(a) and Figure~\ref{fig:simrank}(b) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the first non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices of both persons and groups.
 

### 2023-09-08 analysis.qmd
Updated SimRank plot and function

@@
    source(here("Functions", "TwoModeSR.R"))
-   tmSR.res <- TwoModeSR(A, C = 0.8, iter = 10)
+   tmSR.res <- TwoModeSR(A)
     a <- tmSR.res$eigvec.plot.r + theme(axis.text = element_blank())
@@
     d <- tmSR.res$eigval.plot.c
-    #a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
-    #b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
     e <- tmSR.res$A.plot
@@
    dat.r <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.r[,1]),
-                       sr = tmSR.res$ov.r,
+                       sr = tmSR.res$ov.r1,
                        lab = rownames(A))
    dat.c <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.c[,1]),
-                       sr = tmSR.res$ov.c,
+                       sr = tmSR.res$ov.c1,
                        lab = colnames(A))
-   g <- ggscatter(dat.r, x = "ca", y = "sr", 
+   cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
+   g <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
                   label = "lab", repel = TRUE, 
                   add = "reg.line", 
-                  add.params = list(color = "gray80", linewidth = 0.5, linetype = 2))
-   h <- ggscatter(dat.c, x = "ca", y = "sr", 
+                  add.params = list(color = "gray80", 
+                                    linewidth = 0.5, linetype = 2)) +
+                  annotate("text", x = -.2, y = .3, 
+                           label = paste("r = ", cor), size = 6) 
+   cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
+   h <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
                   label = "lab", repel = TRUE, 
                   add = "reg.line",
-                  add.params = list(color = "gray80", linewidth = 0.5, linetype = 2))
+                  add.params = list(color = "gray80", 
+                                    linewidth = 0.5, linetype = 2)) +
+                  annotate("text", x = -.2, y = .3, 
+                           label = paste("r = ", cor), size = 6) 
    i <- (a | b | e) 
-   j <- (f | (g / h)) + plot_layout(widths = c(2.5, 1))
+   j <- (f | (g / h)) + plot_layout(widths = c(3, 1))
    png(file = here("Plots", "sr-plot.png"), width=1200, height=1200)

### 2023-09-08 Functions/SimRank.R

@@ -1,0 +1,51 @@
+SimRank <- function(A, C = 0.8, iter = 10) {
+     nr <- nrow(A)
+     nc <- ncol(A)
+     dr <- rowSums(A)
+     dc <- colSums(A)
+     Sr <- diag(1, nr, nr)
+     Sc <- diag(1, nc, nc)
+     rn <- rownames(A)
+     cn <- colnames(A)
+     rownames(Sr) <- rn
+     colnames(Sr) <- rn
+     rownames(Sc) <- cn
+     colnames(Sc) <- cn
+     m <- 1
+     while(m < iter) {
+          Sr.pre <- Sr
+          Sc.pre <- Sc
+          for(i in 1:nr) {
+               for(j in 1:nr) {
+                    if (i != j) {
+                         a <- names(which(A[i, ] == 1)) #objects chosen by i
+                         b <- names(which(A[j, ] == 1)) #objects chosen by j
+                         Scij <- 0
+                         for (k in a) {
+                              for (l in b) {
+                                   Scij <- Scij + Sc[k, l] #i's similarity to j
+                              }
+                         }
+                         Sr[i, j] <- C/(dr[i] * dr[j]) * Scij
+                    }
+               }
+          }
+          for(i in 1:nc) {
+               for(j in 1:nc) {
+                    if (i != j) {
+                         a <- names(which(A[, i] == 1)) #people who chose object i
+                         b <- names(which(A[, j] == 1)) #people who chose object j
+                         Srij <- 0
+                         for (k in a) {
+                              for (l in b) {
+                                   Srij <- Srij + Sr[k, l] #i's similarity to j
+                              }
+                         }
+                         Sc[i, j] <- C/(dc[i] * dc[j]) * Srij
+                    }
+               }
+          }
+          m <- m + 1
+     }
+     return(list(Sr = Sr, Sc = Sc))
+}

### 2023-09-08 Functions/TwoModeSR.R

@@ -1,171 +1,90 @@
-TwoModeSR <- function(A, C = 0.8, iter = 10, k = 4) {
+TwoModeSR <- function(A, k = 4, n.eig = 4) {
+   #loading functions
+   source(here("Functions", "SimRank.R"))
+   source(here("Functions", "eigval.scatter.R"))
+   source(here("Functions", "eigvec.scatter.R"))
+   source(here("Functions", "corr.scatter.R"))
+   
+   # SimRank of matrix A
    nr <- nrow(A)
    nc <- ncol(A)
-   dr <- rowSums(A)
-   dc <- colSums(A)
-   Sr <- diag(1, nr, nr)
-   Sc <- diag(1, nc, nc)
    rn <- rownames(A)
    cn <- colnames(A)
-   rownames(Sr) <- rn
-   colnames(Sr) <- rn
-   rownames(Sc) <- cn
-   colnames(Sc) <- cn
-   m <- 1
-   while(m < iter) {
-      Sr.pre <- Sr
-      Sc.pre <- Sc
-      for(i in 1:nr) {
-         for(j in 1:nr) {
-            if (i != j) {
-               a <- names(which(A[i, ] == 1)) #objects chosen by i
-               b <- names(which(A[j, ] == 1)) #objects chosen by j
-               Scij <- 0
-               for (k in a) {
-                  for (l in b) {
-                     Scij <- Scij + Sc[k, l] #i's similarity to j
-                     }
-                  }
-                  Sr[i, j] <- C/(dr[i] * dr[j]) * Scij
-               }
-            }
-         }
-      for(i in 1:nc) {
-         for(j in 1:nc) {
-            if (i != j) {
-               a <- names(which(A[, i] == 1)) #people who chose object i
-               b <- names(which(A[, j] == 1)) #people who chose object j
-               Srij <- 0
-               for (k in a) {
-                  for (l in b) {
-                     Srij <- Srij + Sr[k, l] #i's similarity to j
-                     }
-                  }
-                  Sc[i, j] <- C/(dc[i] * dc[j]) * Srij
-               }
-            }
-         }
-      m <- m + 1
-   }
-
-   corr.scatter <- function(x) {
-      p <- ggscatter(data.frame(x), 
-                     x = "d1", 
-                     y = "d2", 
-                     point = FALSE,
-                     color = "cluster",
-                     palette = "uchicago",
-                     font.label =  13,
[…]

### 2023-09-08 Functions/eigvec.scatter.R

@@ -1,0 +1,22 @@
+eigvec.scatter <- function(x, s = 10) {
+     p <- ggscatter(x, 
+                    y = "rank", 
+                    x = "value", 
+                    point = FALSE,
+                    font.label =  s,
+                    label = "lab", 
+                    color = "cluster",
+                    palette = "Dark2",
+                    repel = TRUE
+                    #ellipse = TRUE,
+                    #ellipse.type = "convex",
+                    #ellipse.alpha = 0.15,
+                    #ellipse.border.remove = TRUE
+     )
+     p <- p + theme(axis.text.y = element_blank(),
+                    axis.line.x = element_blank(),
+                    axis.line.y = element_blank(),
+                    legend.position = "none")
+     p <- p + labs(y = "Rank", x = "First Axis")
+     return(p)
+     }

### 2023-09-08 main.tex

@@
     \centering
-    \includegraphics[width=1.0\textwidth]{Plots/sr-plot.png}
+    \includegraphics[width=0.8\textwidth]{Plots/sr-plot.png}
     \caption{Caption}
@@
 Figure~\ref{fig:simrank}(a) and Figure~\ref{fig:simrank}(b) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the first non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices of both persons and groups.
+As we can see, the partitioning of the node sets on the first informative eigenvector is substantively the same, separating the same two blocks of people and events.
+Figure~\ref{fig:simrank}(c) shows the original affiliation matrix, this time with rows and columns re-arranged according to the main eigenvector of the SimRank similarity matrix.
+Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two blocks of persons and groups.
+Figure~\ref{fig:simrank}(d) shows the correspondence plot of the first and second informative eigenvector of the SimRank similarity matrices for persons and groups.
+Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c).
 

### 2023-09-08 analysis.qmd
Latest plots

--- 2023-09-08 analysis.qmd
@@
    library(igraph)
-   library(pals)
    library(patchwork)
@@
 ```{r CA versus Bonacich Re-ordered Affiliation Matrices}
-   source(here("Functions", "TwoModeCA.R"))
    a <- TwoModeCA(A)
@@
 
-```{r Eigenvector Plot}
-   source(here("Functions", "TwoModeCA.R"))
+```{r CA Eigenvector Plot}
     tmCA.res <- TwoModeCA(A, k.r = 2, k.c = 2)
@@
     b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
-
     png(file = here("Plots", "ca-eigvec.png"), width=1000, height=600)
@@
 ```{r Similarity and Correspondence Plot (CA)}
-   source(here("Functions", "TwoModeCA.R"))
-    tmCA.res <- TwoModeCA(A, d.r = 4, d.c = 4, k.r = 3, k.c = 3)
+    tmCA.res <- TwoModeCA(A, d.r = 2, d.c = 2, k.r = 5, k.c = 5)
     a <- tmCA.res$ca.sim.plot.r + theme(axis.text.x = element_blank())
@@
 ```{r Similarity and Correspondence Plot (Bonacich)}
-   source(here("Functions", "TwoModeCA.R"))
-    tmCA.res <- TwoModeCA(A, d.r = 4, d.c = 4, b.r = 4, b.c = 4)
+    tmCA.res <- TwoModeCA(A, d.r = 2, d.c = 2, b.r = 4, b.c = 4)
     a <- tmCA.res$bon.sim.plot.r + theme(axis.text.x = element_blank())
@@
     c <- tmCA.res$bon.corr.plot
-
     png(file = here("Plots", "bon-sim.png"), width=950, height=700)
@@
 ```{r SimRank Plot}
-   source(here("Functions", "TwoModeSR.R"))
-   tmSR.res <- TwoModeSR(A, C = 0.8, iter = 10)
+   tmSR.res <- TwoModeSR(A)
     a <- tmSR.res$eigvec.plot.r + theme(axis.text = element_blank())
@@
    dat.r <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.r[,1]),
-                       sr = tmSR.res$ov.r,
+                       sr = tmSR.res$ov.r1,
                        lab = rownames(A))
    dat.c <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.c[,1]),
-                       sr = tmSR.res$ov.c,
+                       sr = tmSR.res$ov.c1,
                        lab = colnames(A))
-   g <- ggscatter(dat.r, x = "ca", y = "sr", 
+   cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
+   g <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
                   label = "lab", repel = TRUE, 
                   add = "reg.line", 
-                  add.params = list(color = "gray80", linewidth = 0.5, linetype = 2))
-   h <- ggscatter(dat.c, x = "ca", y = "sr", 
+                  add.params = list(color = "gray80", 
+                                    linewidth = 0.5, linetype = 2)) +
+                  annotate("text", x = -.2, y = .3, 
+                           label = paste("r = ", cor), size = 6) 
+   cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
+   h <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
                   label = "lab", repel = TRUE, 
                   add = "reg.line",
-                  add.params = list(color = "gray80", linewidth = 0.5, linetype = 2))
-   i <- (a | b | e) 
-   j <- (f | (g / h)) + plot_layout(widths = c(2.5, 1))
+                  add.params = list(color = "gray80", 
+                                    linewidth = 0.5, linetype = 2)) +
+                  annotate("text", x = -.2, y = .3, 
+                           label = paste("r = ", cor), size = 6) 
+   i <- (e | a | b) 
+   j <- (f | (g / h)) + plot_layout(widths = c(3, 1))
    png(file = here("Plots", "sr-plot.png"), width=1200, height=1200)

### 2023-09-08 Functions/TwoModeCA.R

@@ -5,97 +5,20 @@ TwoModeCA <- function(x,
                       b.c = 2,
                       d.r = 2, 
                       d.c = 2) {
-        # Plot Functions
-        eigval.scatter <- function(x) {
-                p <- ggscatter(x, 
-                               x = "k", 
-                               y = "value",
-                               size = 2,
-                               color = "red")
-                p <- p + labs(y = "", x = "k") 
-                p <- p + theme(axis.text.x = element_blank(),
-                               axis.text.y = element_text(size = 8),
-                               axis.line.y = element_blank(),
-                               axis.line.x = element_blank())
-                return(p)
-                }
-        eigvec.scatter <- function(x) {
-                p <- ggscatter(x, 
-                          y = "rank", 
-                          x = "value", 
-                          point = FALSE,
-                          font.label =  12,
-                          label = "lab", 
-                          color = "cluster",
-                          palette = "Dark2",
-                          repel = TRUE
-                          #ellipse = TRUE,
-                          #ellipse.type = "convex",
-                          #ellipse.alpha = 0.15,
-                          #ellipse.border.remove = TRUE
-                          )
-                p <- p + theme(axis.text.y = element_blank(),
-                               axis.line.x = element_blank(),
-                               axis.line.y = element_blank(),
-                               legend.position = "none")
-                p <- p + labs(y = "Rank", x = "First Axis")
-                return(p)
-                }
-        corr.scatter <- function(x) {
-                p <- ggscatter(data.frame(x), 
-                          x = "d1", 
-                          y = "d2", 
-                          point = FALSE,
-                          color = "cluster",
-                          palette = "uchicago",
-                          font.label =  10,
-                          label = rownames(x), 
-                          repel = TRUE)
-                p <- p + theme(legend.position = "none",
-                               axis.line.x = element_blank(),
[…]

### 2023-09-08 main.tex
Updates from Overleaf

--- 2023-09-08 main.tex
@@
 
-\subsection{Correspondence Analysis and Generalized Similarities}
+\subsection{Correspondence Analysis and Generalized Relational Similarity}
 As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.
 Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}.
-This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence.
+This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence \citep{everett1994regular}.
 In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.
@@
 An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}.
-In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong, and the similarity of groups is a function of the people who belong to them, making the similarity of people and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
+In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong, and the similarity of groups is a function of the people who belong to them, making the similarity of persons and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
 Thus,
@@
 
-Which is consistent with a GRS approach. To do this, \citet[540, eq2 and eq3]{jeh2002simrank} propose that we define the similarity of people as given by:
+Which is consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
+To accomplish this, \citet[540, eq.
+2 and eq.
+3]{jeh2002simrank} propose that we define the similarity of each pair of people $S(p, p')$  as given by:
 
@@
 
-Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to; $\alpha$ is a free parameter, with the restriction: $0 > \alpha < 1$.
-By construction, $S(p, p) = 1$.
-Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of the groups they belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.
+Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to, and $\alpha$ is a free parameter obeying the restriction: $0 > \alpha < 1$.
+By construction, $S(p, p) = 1$, for all $p$.
+Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of each unordered pair of groups they both belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.
 
@@
 Where $p(i)_{i \in N(g)}$ is the $i^{th}$ person in the set of members of group $g$, and $p(j)_{j \in N(g')}$ is the $j^{th}$ person in the set of members of group $g'$, and $S(g, g) = 1$.
-Thus, the SimRank similarity between two groups is a function of the sum of the similarities of the people who belong to them, weighted by the reciprocal of the products of their number of members multiplied by $\alpha$.
+Thus, the SimRank similarity between two groups is a function of the sum of the similarities of each pair of people who belong to both groups, weighted by the reciprocal of the products of the number of members of each group multiplied by $\alpha$.
 
@@
 
-In two-mode networks, SimRank scores can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
+In two-mode networks, SimRank scores for each pair of nodes across the two sets can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
 Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
-We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}.
-Further Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with the HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
+We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}, which is confirmed here for the Southern Women data.
+
+Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with the HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
 The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
-Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA.
+Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections \citep{mealy2019interpreting, van2021correspondence}.
 
 And indeed, they are not.
-Figure~\ref{fig:simrank}(a) and Figure~\ref{fig:simrank}(b) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the first non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices of both persons and groups.
+Figure~\ref{fig:simrank}(a) shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices for the row and column nodes.
+Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two blocks of persons and groups, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
+Figure~\ref{fig:simrank}(b) and Figure~\ref{fig:simrank}(d) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
+Looking at the plots from left to right, we can see that the partitioning of the node sets on the first informative eigenvector is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
+
+Figure~\ref{fig:simrank}(f) shows the correspondence plot of the first and second eigenvectors of the SimRank similarity matrices for both persons and groups.
+Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c), suggesting that the first-CA dimensions of the two-mode network recover clusters of ``similar'' entities in the two node sets, where internode similarity is a generalized relational similarity as defined earlier.
+Thus, [$Flora, Olivia$] appear close because they choose similar groups, as do [$Sylvia, Katherine, Helen$].
+In the same way, groups [$E9, E11$] are close in the plot because they are chosen by similar people, as are events [$E10, E12, E13, E14$].
+Figures~\ref{fig:simrank}(g) and (h) show a regression plot of the first CA eigenvector for persons (top) and groups (bottom) on the x-axis and the first eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
+As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that these first eigenvectors capture the same (GSR) information across the two approaches.
 

### 2024-03-29 analysis.qmd
ncsu presentation

@@
    library(igraph)
+   library(pals)
    library(patchwork)
@@
           ncol = 14, byrow = TRUE)
-     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
+     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRNA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
      e <- paste("E", c(1:14), sep = "")
@@
    p1 <- p + scale_color_manual(values=as.vector(cols25(18)))
+   png(file = here("Plots", "p-reflections.png"), width=750, height=500)
+   p1
+   dev.off()
 
@@
    p2 <- p + scale_color_manual(values=as.vector(cols25(15)))
-   png(file = here("Plots", "pg-reflections.jpg"), width=750, height=950)
+   png(file = here("Plots", "g-reflections.png"), width=750, height=500)
+   p2
+   dev.off()
+   png(file = here("Plots", "pg-reflections.png"), width=750, height=950)
    p1 / p2
@@
    c <- a$bon.A
+   png(file = here("Plots", "ca-reord.png"), width=500, height=500)
+   b
+   dev.off() 
+   png(file = here("Plots", "bon-reord.png"), width=500, height=500)
+   c
+   dev.off() 
    png(file = here("Plots", "ca-v-bon-reord.png"), width=1000, height=700)
@@
     b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+
+   png(file = here("Plots", "ca-eigvec-p.png"), width=500, height=500)
+   a
+   dev.off() 
+   png(file = here("Plots", "ca-eigvec-g.png"), width=500, height=500)
+   b
+   dev.off() 
     png(file = here("Plots", "ca-eigvec.png"), width=1000, height=600)
@@
     png(file = here("Plots", "ca-sim.png"), width=950, height=700)
+      (a | b) / c + plot_annotation(tag_levels = "a") &
+      theme(plot.tag = element_text(size = 20)) 
+    dev.off() 
+    png(file = here("Plots", "ca-sim-slide.png"), width=1050, height=700)
       (a | b) / c + plot_annotation(tag_levels = "a") &

### 2024-03-29 ncsu-presentation.qmd
ncsu presentation

---
title: "Coasting on Duality: What We Get When We Do Two-Mode CA"
author: "Omar Lizardo (UCLA)"
format:
  revealjs:
    incremental: false 
    theme: sky 
    smaller: true
---

## CA and Two Mode Networks

- While not common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data.

  - But criticized for being limited and inapplicable for most "sna" tasks*

- CA Proponents point to some important advantages of CA for the analysis of two-mode networks:
  - CA can provide synoptic graphical representations of structural connectivity patterns in two-mode data**
  - (M)CA can be used to "visually explore" two-mode networks***

::: aside
*Borgatti & Everett 1997

**Breiger 2000; Roberts 2000; Faust 2005

***D’Esposito et al. 2014
:::

## CA and Two Mode Networks

- While not common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data.

  - But criticized for being limited and inapplicable for most "sna" tasks*

- This presentation builds on and expands on previous  research on using CA for two-mode data analysis

  - But discusses aspects of CA that have not been emphasized in previous research:

1. CA as computing a type of "(Reflective) Dual Centrality" on the entities in the two-modes**
    - Rediscovered recently under the heading of the Economic Complexity Index (ECI; Hidalgo and Hausman 2009)

::: aside
*Borgatti & Everett 1997

**Distinct from the traditional Bonacich dual centrality
:::

::: {.notes}
I show that CA can be thought of as computing a kind of dual centrality score on both modes---one related to but not mathematically equivalent to the Bonacich two-mode centrality----an approach that has been recently re-invented in some corners of network science under the guise of the "economic complexity index''

I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of ordination---e.g., the discovery of latent one-dimensional orderings among a set of entities.
:::

## CA and Two Mode Networks

- While not common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data.

  - But criticized for being limited and inapplicable for most "sna" tasks*

- This presentation builds on and expands on previous  research on using CA for two-mode data analysis.

  - But discusses aspects of CA that have not been emphasized in previous research:

2. CA as extracting "positional" information the entities in the two modes
    - With positions defined by "Generalized Relational Similarity"**
    - Similar entities have similar patterns of connectivity with entities that are themselves similar

::: aside
*Borgatti & Everett 1997

*Kovacs 2010; Lizardo 2024
:::

::: {.notes}
There is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on "generalized relational similarity'' in two-mode networks. CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups).

Second, I show that the first two CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode. Thus, CA can be used as an approach to indirect "community" identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a clustering method or the discovery of sociometrically similar subsets of entities in affiliation networks. 
:::

## CA and Reflective Centrality

- Hidalgo and Hausmann (2009) introduced a novel way to rank nodes in a two-mode network.
  - Their approach, called "reflective," was originally applied to country-by-product networks*

. . .

  - The method has since been shown to be applicable to other two-mode networks (like cultural networks)**
  - I introduce here using the more familiar sociological example of the **duality of persons and groups*****

::: aside
*Later referred to as the Economic Complexity Index (ECI)

**Lizardo 2018

***Breiger 1974
:::

## Southern Women Data {.scrollable}
```{r Southern women data}
    A <- matrix(
         c(
          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0,

          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,

          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
          ncol = 14, byrow = TRUE)
     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
     e <- paste("E", c(1:14), sep = "")
     rownames(A) <- w
     colnames(A) <- e
     require(kableExtra)
     A <- A[sort(rownames(A)), ]
     kbl(A)
```

## CA and Reflective Centrality
- If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old **degree centrality** (Faust 1997)

. . .

- Notation:
  - A two-mode network composed of a set of persons $P$
  - Their affiliation relations to a set of groups $G$
  - Represented by an affiliation matrix $\mathbf{A}$

. . .

  - $\mathbf{A}$ is of dimensions $|P| \times |G|$ with people along the rows and groups across the columns
    - $|P|$ is the cardinality of the persons set  
    - $|G|$ is the cardinality of the group set

. . .

  - $\mathbf{A}$ has cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.

## CA and Reflective Centrality
- If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old **degree centrality** (Faust 1997)

- Given this, the degree-centrality of people is given by:

$$
C^R_p(1) = \sum_g a_{pg}
$$ {#eq-dp}

. . .

  - And for groups:

$$
   C^R_g(1) = \sum_p a_{pg}
$$ {#eq-dg}

## CA and Reflective Centrality

- The key Hidalgo and Hausman's "reflective" approach is as follows:

   - Once we have the first-order quantities $C^R_p(1)$ and $C^R_g(1)$, we can compute "second-order centralities" $C^R(2)$ for both people and groups:

   - To do that we use the (averaged) first-order centralities of the entities in the other mode they are connected to

. . .

For people, these are given by:

$$
C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
$$ {#eq-r2p}

And for groups:

$$   
C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
$$ {#eq-r2g}

::: {.notes}
Equation @eq-r2p says "people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number). Equation @eq-r2g says "groups are more central when the average activity of their members is large" (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
:::

## CA and Reflective Centrality

- The key Hidalgo and Hausman's "reflective" approach is as follows:

   - Once we have the first-order quantities $C^R_p(1)$ and $C^R_g(1)$, we can compute "second-order centralities" $C^R(2)$ for both people and groups:

  - Of course, we can keep on going and define third-order $C^R(3)$ reflections!

. . .

For people, these are given by:

$$
   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
$$ {#eq-r3p}

And for groups:

$$   
   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
$$ {#eq-r3g}

::: {.notes}
Equation @eq-r3p says something like "people are more central when the average sum of the average activity of the members of the groups they belong to is large" (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number). Equation @eq-r3g says, "groups are more central when the average sum of the average size of the groups their members belong to is large." (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
:::

## CA and Reflective Centrality
- More generally, HS show that we can define a series of reflective quantities for people and groups*

- For people, these are given by:

$$   
C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
$$ {#eq-rip}

- And for groups:

$$   
C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
$$ {#eq-rig}

::: aside
*Whose verbal and substantive interpretation seems to become more elusive as the number of iterations increases
:::

## CA and Reflective Centrality

- For people:
  - Even-numbered reflections measure centrality based on group size 
  - Odd-numbered reflections measure centrality based on the average size of the groups a person belongs to*

. . .

- For groups:
  - Even-numbered reflections measure centrality based on member activity 
  - Odd-numbered reflections measure centrality based on the average group size of a group's members

::: aside
*Although these are nominally distinct metrics, in practice, the rank order of the nodes in each mode for even and odd reflections ends up being identical after a small number of iterations
:::

::: {.notes}
Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the groups they belong to (in this case, the group sizes). On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the members of the groups they belong to (in this case, the average size of the groups they belong to). In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the people who belong to them (in this case, their activity). On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the other groups their members belong to (in this case, their average group size). While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size). 
:::

## CA and Reflective Centrality
![Person Reflection Trajectories](Plots/p-reflections.png)

## CA and Reflective Centrality
![Group Reflection Trajectories](Plots/g-reflections.png)

## CA and Reflective Centrality

- What's the point of going through all these reflective centralities?

- Particularly, since Bonacich already defined dual centralities for two-mode networks based on a similar idea!*

- In that paper, Bonacich also considered CA in passing 
  - However, He dismissed CA as a method to derive centrality rankings:
    - "[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events"**

::: aside
*Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected; Bonacich 1991

** p. 164
:::

## CA and Reflective Centrality

- What's the point of going through all these reflective centralities?

- Particularly, since Bonacich (1991) already defined dual centralities for two-mode networks based on a similar idea!*

- Bonacich was only half right!
  - CA scores do capture "positional" information but not one based purely on cohesion (cliques)
  - But they (particularly the first dimension) also capture a version of centrality** 
  - In fact, they capture the limit $(q \rightarrow \infty)$ of HH's reflective centralities!

::: aside
*Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected; Bonacich 1991
:::

::: aside
**In the most general sense of ranking nodes based on a meaningful criterion; Borgatti 2005
:::

## CA and Bonacich Centrality
- Let's look at the connection between Bonacich, HH Reflections, and CA

- The solutions to the following eigensystem give the Bonacich $(C^B)$ two-mode centralities:

$$
AC^B_p = \lambda C^B_g 
$$ {#eq-bon1}

$$
A^TC^B_g = \lambda C^B_p 
$$ {#eq-bon2}

## CA and Bonacich Centrality

- It is possible to express @eq-bon1 and @eq-bon2 in terms of person and group-specific centralities:*

$$
    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
$$ {#eq-fau1}

$$
    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
$$ {#eq-fau2}

. . .

- The Bonacich Centralities look a lot like the "reflective" ones, minus the averaging by same-mode degree!

::: aside
Faust 1997, p. 170
:::

## CA and Bonacich Centrality

We can solve for $C^B_g$ in @eq-bon2 and $C^B_p$ in @eq-bon1 and substitute the respective solutions into @eq-bon1 and @eq-bon2

- After some algebraic manipulation we get:

$$
\left(AA^T\right)C^B_p = \lambda^2 C^B_p
$$ {#eq-bon3}

$$
\left(A^TA\right)C^B_g = \lambda^2 C^B_g
$$ {#eq-bon4}

- The first eigenvector of the respective one-mode projections also gives us the Bonacich dual centralities!

## CA and Bonacich Centrality
- But what's the connection to reflections and CA after all of this?

- A couple of things first:

  - $Dp$ is a $|P| \times |P|$ matrix containing each person's $C_p^R(1)$ (activity) along the diagonals and zeroes in every other cell
  - $Dg$ is a $|G| \times |G|$ matrix $Dg$ containing each group's $C_g^R(1)$ (size) along the diagonals and zeroes in every other cell

. . .

- In the limit ($q \rightarrow \infty$), the iterative HH reflective centralities are the solution of the eigensystem:*

$$
    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
$$ {#eq-dam1}

$$
    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
$$ {#eq-dam2}

::: aside
*van Dam et al. 2021
:::

::: {.notes}
Note the formal similarity and differences between these and @eq-bon3 and @eq-bon4

Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix

But @eq-dam1 and @eq-dam2 *pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before factorizing*!
:::

## CA and Bonacich Centrality
- The difference between @eq-dam1 and @eq-bon3 can be more clearly seen if we express it in terms of each cell entry*

- For people:

$$
a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
$$ {#eq-mealy1}

- In equation @eq-mealy1, the fraction's numerator is equal to one only when person $p$ and $p'$ are both members of $g$.

- Summed across groups, the number of shared groups between $p$ and $p'$

- But in @eq-mealy1 $\sum_g a_{pg}a_{p'g}$ is *weighted* by group size and each person's number of memberships

::: aside
Mealy 2019, eq. 4
:::

::: {.notes}
In equation @eq-mealy1, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$. Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common.

As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).

The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships. 
:::

## CA and Bonacich Centrality
- The difference between @eq-dam1 and @eq-bon3 can be more clearly seen if we express it in terms of each cell entry*

- For groups:

$$
    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
$$ {#eq-mealy2}

- In equation @eq-mealy2, the fraction's numerator is equal to one only when group $g$ and $g'$ both have $p$ as a member

- Summed across persons, the number of shared members between $g$ and $g'$

- But in @eq-mealy2 $\sum_p a_{pg}a_{pg'}$ is *weighted* by group size and each person's number of memberships

::: aside
Mealy 2019, eq. 4
:::

## CA and Bonacich Centrality

- And here comes out first result/conclusion:

  - Turns out that the $C^R_p$ and $C^R_p$ centralities are the same (to a scaling factor) as the row and column scores obtained from the first dimension of a **simple CA*** of the two-mode affiliation matrix!
     - It is well known among CA users that CA works with the row and *column sum degree weighted* (for each cell) version of the original data matrix
    - Which in the two-mode network case equals degree-weighting!
  - This means that if the reflective centralities are "centralities" then CA produces dual centrality scores on the first dimension
  - These are *different* from the traditional Bonacich centrality scores**

:::aside
*No an MCA as in D'Esposito et al. (2014)
:::

:::aside
**Because of the double-degree pre-weighting
:::

::: {.notes}
Degree-pre-weighting does not radically alter the nature of the data. Indeed, it corresponds to moving from sums to averages, thus "adjusting" for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis. This can be seen by the fact that Equations @eq-mealy1 and @eq-mealy2 show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large. On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.

Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations 
@eq-bon3 and @eq-bon4 "centrality scores" but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' but formally identical. In the Bonacich case, a factorization of the one-mode projection of the two-mode network; in the reflective case, factorization of the degree-weighted one-mode projection of the same network.

Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks. Particularly, the CA centrality retains the duality property, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to.

The (average) centrality of groups, in turn, is a function of their members' (average) centrality. The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich dual centralities for two-mode networks.

:::

## CA and Bonacich Centrality
:::: {.columns}
::: {.column width="50%"}
CA Re-ordered Affiliation Matrix
![](Plots/ca-reord.png)
:::

::: {.column width="50%"}
Bonacich Re-ordered Affiliation Matrix
![](Plots/bon-reord.png)
:::
::::

::: {.notes}
As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups. 
:::

## CA Reveals Dual Communities
- As we know from Breiger (1974) the usual one mode projections are obtained via:

$$
S_p = AA^T
$$

$$
S_g = A^TA
$$

- Which can be thought of as **similarity matrices** based on the shared neighbors in the two mode network

- But we can also obtain *degree weighted* versions of these similarity matrices

$$
S^*_p = ADg^{-1}A^T
$$

$$
S^*_g = A^TDp^{-1}A
$$

## CA Reveals Dual Communities
- As we know from Breiger (1974) the usual one mode projections are obtained via:

$$
S_p = AA^T
$$

$$
S_g = A^TA
$$

- Which can be thought of as **similarity matrices** based on the shared neighbors in the two mode network

- The eigenvalues/first eigenvectors of $S^*_p$ and $S^*_g$ reveal dual communities

$$
S^*_p = ADg^{-1}A^T
$$

$$
S^*_g = A^TDp^{-1}A
$$

## CA Reveals Dual Communities
- As we know from Breiger (1974) the usual one mode projections are obtained via:

$$
S_p = AA^T
$$

$$
S_g = A^TA
$$

- Which can be thought of as **similarity matrices** based on the shared neighbors in the two mode network

- The eigenvalues/first eigenvectors of $S^*_p$ and $S^*_g$ reveal dual communities

- The eigenvectors of $S^*_p$ and $S^*_g$ are equivalent (to a scaling factor) to those obtained from the simple CA of the affiliation matrix!

## CA Reveals Dual Communities
:::: {.columns}
::: {.column width="50%"}
![Second Eigenvector of Degree-Weighted Similarity Matrix Against Rank (Persons)](Plots/ca-eigvec-p.png)
:::

::: {.column width="50%"}
![Second Eigenvector of Degree-Weighted Similarity Matrix Against Rank (Groups)](Plots/ca-eigvec-g.png)
:::
::::

::: {.notes}
We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at the Figures which show the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.

If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution. The Southern Women data clearly belong to the second category. In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other. Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.

Note that this is roughly the same partitioning obtained by Doreian (1979) using a different method (Q analysis) and reproduced by Kovacs and Lizardo (using a generalized relational similarity approach). This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to similar groups, where group similarity is defined dually as having members in common who are themselves similar.

:::

## CA Reveals Dual Communities
:::: {.columns}
::: {.column width="43%"}
![Second Eigenvector of Degree-Weighted Similarity Matrix Against Rank (Persons)](Plots/ca-eigvec-p.png)
:::

::: {.column width="57%"}
![From Doreian (1979, p. 237)](Plots/doreian-sw.png)
:::
::::

## CA Reveals Dual Communities
![](Plots/ca-sim-slide.png)
::: {.notes}

To fix the intuition, the top-left panel (a) of the figure shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups). In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity. Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups. As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.

The bottom panel of the figure shows the usual correspondence plot of the first two CA dimensions. People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices. The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the original affiliation matrix---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups.

People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm). The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart. The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.

Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with "similar" connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups). This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot. But this (still common) interpretation is off the mark.

What the CA correspondence plot distance captures is, instead, people and groups that are surprisingly similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account. As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups. In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.

:::

## Bonacich Reveals Core/Periphery
![](Plots/bon-sim-slide.png)

::: {.notes}
To illustrate, The Figure shows the ``raw'' (unweighted by other-mode degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score. Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix. The Figure plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix---a relatively unusual but not substantively unmotivated practice.

Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.

We can see that the plot of the first two eigenvectors does a good job of recovering the raw connectivity structure of the Southern Women affiliation network, partitioning the core persons and groups (on the upper-right) from the more peripheral ones (on the lower-left).

Moreover, if all we are interested in is capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then the Bonacich plot does a better job of that than the usual CA correspondence plot. For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine, Sylvia\right]$ for $\left[E10, E12\right]$. In the same way, core events like $\left[E8, E9, E10\right]$---shown on the lower half of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the upper half. The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.

This ``preferential attachment'' of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's eigenvector centrality. For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot. Similarly, the main difference between $\left[Ruth\right]$ and $\left[Sylvia\right]$, despite their comparable $C^B$ scores, is that the former's (closer to the upper half of the plot) event profile is composed mainly of core events. In contrast, peripheral events dominate the latter's attendance profile (shown in the bottom half of the plot), accordingly, $\left[Ruth\right]$ and $\left[Sylvia\right]$ are assigned to distinct clusters by the k-means algorithm. 
:::

## CA and Generalized Similarity
- The contrast between CA scores and Bonacich Dual Centralities is that CA picks up patterns of attendance between similar people at similar events

- This is precisely the definition of **generalized relational similarity** (GSR)

- It would be useful to compare CA with one such method for obtaining GSRs in two-mode networks: *SimRank* (Jeh & Widom 2002)

. . .

- According to Jeh & Widom:
  - People are *similar* if they belong to *similar* groups*
  - Groups are *similar* if they share *similar* members*

:::aside
Kovacs (2010); Lizardo (2024)
:::

## CA and Generalized Similarity

- In *SimRank* the similarity between two people is given by:

$$
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
$$ {#eq-simrank1}

. . .

- And for groups:

$$
S(g, g') = \frac{\alpha}{C_g^R(1)C_{g'}^R(1)}
    \sum_{i = 1}^{C_g^R(1)} \sum_{j = 1}^{C_{g'}^R(1)} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
$$ {#eq-simrank2}

. . .

- $S(p, p) = 1$ and $S(g, g) = 1$ 
- $\alpha$ is a free parameter, with the restriction: $0 > \alpha < 1$

::: {.notes}
In two-mode networks, SimRank scores can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation @eq-simrank1 using baseline values. Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset. We then plug those values into equation @eq-simrank2, then loop back to equation @eq-simrank1 with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations
:::

## CA and Generalized Similarity
:::: {.columns}
::: {.column width="30%"}
![SimRank Re-ordered Affiliation Matrix](Plots/simrank-plot-reord.png)
:::

::: {.column width="65%"}
![Simrank Similarity Matrix Eigenvalue Plot](Plots/simrank-plot-eigen.png)
:::
::::

## CA and Generalized Similarity
:::: {.columns}
::: {.column width="65%"}
![Simrank Correspondence Plot](Plots/simrank-plot-corr.png)
:::

::: {.column width="24%"}
![Correlation between CA and SimRank Eigenvectors](Plots/simrank-plot-scatter.png)
:::
::::

## Conclusions

- What have we learned after all of this? A few lessons:
  - You can use CA so do more than "visually explore" two-mode network data!
  - Instead, you an use CA to do "real" sna
    - Compute (a version of) centrality on each node
    - Discover Dual Community Partitions
    - Assign nodes to (generalized) similarity classes
  - Using CA for this kind of "spectral clustering" and "spectral ranking" is not common but feasible
  - Another approach (similar to blockmodeling):
    - Use CA for spectral clustering
    - Then use **specific CA** (on a sub-table) to split clusters into further clusters

### 2024-03-29 Functions/blah.R

@@ -1,0 +1,23 @@
+r <- c("Croatia", "Denmark", "Finland", "France", "Netherlands", "Serbia", "Spain", "Switzerland", "United Kingdom")
+c <- c("Arts", "Institutional", "Values and trad.", "Geographical ID", "Group chars.", "Habits and lifestyles", "Human activity", "History and heritage", "Cultivation", "Good manners")
+
+dat <- matrix(c(0.10,0.42,0.02,0.02,0.10,0.03,0.07,0.03,0.03,0.16, 0.16,0.22,0.10,0.06,0.07,0.06,0.18,0.06,0.04,0.05, 0.21,0.19,0.07,0.05,0.08,0.06,0.19,0.05,0.05,0.04, 0.13,0.25,0.03,0.04,0.02,0.03,0.02,0.05,0.42,0.01, 0.13,0.15,0.14,0.12,0.12,0.16,0.08,0.03,0.02,0.05, 0.10,0.27,0.06,0.03,0.08,0.04,0.05,0.03,0.03,0.31, 0.10,0.12,0.08,0.05,0.06,0.07,0.08,0.04,0.38,0.02, 0.15,0.23,0.07,0.06,0.06,0.05,0.11,0.06,0.16,0.05,0.17,0.07,0.23,0.10,0.18,0.11,0.05,0.06,0.02,0.02), nrow = 9, byrow = TRUE)
+rownames(dat) <- r
+colnames(dat) <- c
+
+library(FactoMineR)
+#res.CA <- CA(dat)
+
+dat.f <- data.frame(country = r, dat) %>% 
+     pivot_longer(
+          cols = 2:11,
+          names_to = "topics",
+          values_to = "prop"
+     ) 
+
+library(ggplot2)
+p <- ggplot(aes(y = country, x = prop, fill = country, group = country), data = dat.f)
+p <- p + geom_bar(stat = "identity")
+p <- p + facet_wrap(~ topics, ncol = 5)
+p <- p + theme_minimal() + theme(legend.position = "none") + labs(y = "", x = "")
+p

### 2024-03-31 ncsu-presentation.qmd
Updated presentation

@@
 ![](Plots/ca-sim-slide.png)
+
 ::: {.notes}
-
 To fix the intuition, the top-left panel (a) of the figure shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
@@
     - Then use **specific CA** (on a sub-table) to split clusters into further clusters
+
+## Thanks!
+
+[https://github.com/olizardo/correspondence-analysis-two-mode-networks](https://github.com/olizardo/correspondence-analysis-two-mode-networks)
+
+olizardo@soc.ucla.edu

### 2024-05-22 analysis.qmd
duality at 50

@@
    library(factoextra)
+   library(FactoMineR)
    library(ggbump)
@@
    library(igraph)
+   library(kableExtra)
    library(pals)
@@
    source(here("Functions", "TwoModeSR.R"))
+   source(here("Functions", "TwoModeGSR.R"))
    source(here("Functions", "reflections.R"))
@@
 
+```{r HTML data table}
+   A <- A[sort(rownames(A)), ]
+   kbl(A, align = "c", format = "html") %>% 
+      kable_minimal(full_width = F, font_size = 26) %>% 
+      row_spec(1:18, extra_css = "border-bottom: 0px;line-height: 20px;") %>% 
+      save_kable(here("southern-women.html"))
+```
+
 # Analysis
+```{r CA of affiliation matrix}
+   ca.res <- CA(A, graph = FALSE)
+   plot.dat <- data.frame(rbind(ca.res$row$coord, ca.res$col$coord)[, 1:2])
+   plot.dat <- cbind(plot.dat, mode = c(rep(1, 18), rep(2, 14)), names = rownames(plot.dat))
+   p <- ggplot(plot.dat, aes(x = Dim.2, y = Dim.1*-1, 
+                             color = factor(mode), label = names))
+   p <- p + geom_vline(xintercept = 0, color = "gray")
+   p <- p + geom_hline(yintercept = 0, color = "gray")
+   p <- p + geom_text_repel(size = 5, max.overlaps = 50)
+   p <- p + theme_minimal() 
+   p <- p + theme(legend.position = "none",
+                  axis.text = element_text(size = 20),
+                  axis.title = element_text(size = 24)
+                  )
+   p <- p + labs(x = "Second Dimension", y = "First Dimension")
+   p <- p + scale_color_manual(values = c('red', 'blue'))
+   png(file = here("Plots", "ca-southern-women.png"), width=600, height=625)
+   p
+   dev.off()
+```
 
-```{r Reflections Bump Charts}
+```{r Reflections Bump Chart (Persons)}
    evens <- function(x) subset(x, x %% 2 == 0)
@@
    plot.dat <- b$person$even
-   p <- ggplot(data = filter(plot.dat, n.iter < 22),
+   p <- ggplot(data = filter(plot.dat, n.iter < 18),
                mapping = aes(x = n.iter, y = ref, color = person))
    # geoms
-   p <- p + geom_bump(linewidth = 1, smooth = 8, alpha = 0.4)
-   p <- p + geom_point(size = 4, alpha = 0.8)
+   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
+   p <- p + geom_point(size = 5)
    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
                                     aes(x = n.iter, label = person), 
-                                    size = 4, nudge_x = -3)
-   p <-    p <- p + geom_text(data = filter(plot.dat, n.iter == 22), 
+                                    size = 5, nudge_x = -3)
+   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 18), 
                               aes(x = n.iter, label = person), 
-                              size = 4, nudge_x = 0.5)
+                              size = 5, nudge_x = 0.01)
 
@@
    p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
-   p <- p + scale_x_continuous(limits = c(-4, 30), breaks = evens(2:20))
+   p <- p + scale_x_continuous(limits = c(-2, 20), breaks = evens(2:16))
    p <- p + labs(x = "Reflection")
@@
         axis.text.y = element_blank(),
-        axis.title.y = element_blank()
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 16)
         )
-   p1 <- p + scale_color_manual(values=as.vector(cols25(18)))
-   png(file = here("Plots", "p-reflections.png"), width=750, height=500)
+   set.seed(145)
+   c <- sample(cols25(25), 18, replace = FALSE)
+   p1 <- p + scale_color_manual(values=as.vector(c))
+   png(file = here("Plots", "p-reflections.png"), width=600, height=625)
    p1
    dev.off()
+```
 
-   plot.dat <- b$group$even
-   p <- ggplot(data = filter(plot.dat, n.iter < 22),
+```{r Reflections Bump Chart (Groups)}
+   plot.dat <- b$group$even %>% 
+      mutate(ref = 15 - ref)
+   p <- ggplot(data = filter(plot.dat, n.iter < 14),
                mapping = aes(x = n.iter, y = ref, color = group))
    #geoms
-   p <- p + geom_bump(linewidth = 1, smooth = 8, alpha = 0.4)
-   p <- p + geom_point(size = 4, alpha = 1)
-   p <-    p <- p + geom_text(data = filter(plot.dat, n.iter == 24), 
+   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
+   p <- p + geom_point(size = 5)
+   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
                                     aes(x = n.iter, label = group), 
-                              size = 4, nudge_x = -2)
-   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
+                                    size = 6, nudge_x = -1)
+   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 14), 
                               aes(x = n.iter, label = group), 
-                                    size = 4, nudge_x = -1)
+                              size = 6, nudge_x = -0.01)
    #fixings
    p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
-   p <- p + scale_x_continuous(limits = c(-4, 30), breaks = evens(2:20))
+   p <- p + scale_x_continuous(limits = c(1, 14), breaks = evens(2:12))
    p <- p + labs(x = "Reflection")
@@
         axis.text.y = element_blank(),
-        axis.title.y = element_blank()
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 16)
         )
-   p2 <- p + scale_color_manual(values=as.vector(cols25(15)))
-   png(file = here("Plots", "g-reflections.png"), width=750, height=500)
+   set.seed(145)
+   c <- sample(cols25(25), 18, replace = FALSE)
+   p2 <- p + scale_color_manual(values=as.vector(c))
+   png(file = here("Plots", "g-reflections.png"), width=600, height=625)
    p2
    dev.off()
-   png(file = here("Plots", "pg-reflections.png"), width=750, height=950)
-   p1 / p2
+```
+
+```{r CA versus Reflections Scatter}
+   tm <- TwoModeCA(A)
+   mr <- reflections(A, iter = 26)
+   p.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.r[, 1], 
+                                  r = mr$p.s[, 22]))
+   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
+   p <- p + geom_point(size = 3, color = "blue") 
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "18th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
+   p
    dev.off()
+
+   p.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1], 
+                                  r = mr$g.s[, 16]))
+   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
+   p <- p + geom_point(size = 3, color = "red") 
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), 
+                            color = "blue", size = 6)
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "14th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
+   p
+   dev.off()
+```
+
+```{r CA via SVD of two mode network}
+   Rp <- diag(sqrt(rowSums(A)))
+   Rc <- diag(sqrt(colSums(A)))
+
+   svd.res <- svd(Rp%*%A%*%Rc)
+
+   t <- sum(rowSums(A))
+   r.rp <- t/rowSums(A)
+   r.cp <- t/colSums(A)
+   s.rp.cp <- sqrt(c(r.rp, r.cp))
+   tv <- t(svd.res$v)
+   x <- c(svd.res$u[, 3]*r.rp, tv[3, ]*r.cp) 
+   y <- c(svd.res$u[, 2]*r.rp, tv[2, ]*r.cp)
+
+   d <- data.frame(x = x, y = y, l = c(rownames(A), colnames(A)))
+   ggscatter(d, x = "x", y = "y", label = "l", repel = TRUE)
 ```
@@
 
-    png(file = here("Plots", "ca-sim.png"), width=950, height=700)
-      (a | b) / c + plot_annotation(tag_levels = "a") &
-      theme(plot.tag = element_text(size = 20)) 
+    png(file = here("Plots", "ca-sim-p.png"), width=600, height=625)
+    a + theme(axis.text.y = element_text(size = 18))
     dev.off()
-    png(file = here("Plots", "ca-sim-slide.png"), width=1050, height=700)
-      (a | b) / c + plot_annotation(tag_levels = "a") &
-      theme(plot.tag = element_text(size = 20)) 
+
+    png(file = here("Plots", "ca-sim-g.png"), width=600, height=625)
+    b + theme(axis.text.y = element_text(size = 22))
+    dev.off()
+
+    png(file = here("Plots", "ca-corr-plot.png"), width=600, height=625)
+    c
     dev.off()
@@
     c <- tmCA.res$bon.corr.plot
-    png(file = here("Plots", "bon-sim.png"), width=950, height=700)
-      (a | b) / c + 
-       plot_annotation(tag_levels = "a") &
-      theme(plot.tag = element_text(size = 20)) 
+    png(file = here("Plots", "bon-sim-p.png"), width=600, height=625)
+      a + theme(axis.text.y = element_text(size = 16))
+    dev.off()
+
+    png(file = here("Plots", "bon-sim-g.png"), width=600, height=625)
+      b + theme(axis.text.y = element_text(size = 16))
+    dev.off()
+
+    png(file = here("Plots", "bon-corr-plot.png"), width=600, height=625)
+      c
     dev.off() 
@@
 ```{r SimRank Plot}
+   source(here("Functions", "TwoModeSR.R"))
    tmSR.res <- TwoModeSR(A)
-    a <- tmSR.res$eigvec.plot.r + theme(axis.text = element_blank())
+    a <- tmSR.res$eigvec.plot.r 
     b <- tmSR.res$eigvec.plot.c 
@@
     e <- tmSR.res$A.plot
-    f <- tmSR.res$corr.plot
-   dat.r <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.r[,1]),
-                       sr = tmSR.res$ov.r1,
+
+   dat.r <- data.frame(ca = as.numeric(tm$ca.eigvec.r[,1]),
+                       sr = tmSR.res$ev.r1,
                        lab = rownames(A))
-   dat.c <- data.frame(ca = as.numeric(tmCA.res$ca.eigvec.c[,1]),
-                       sr = tmSR.res$ov.c1,
+   dat.c <- data.frame(ca = as.numeric(tm$ca.eigvec.c[,1])*-1,
+                       sr = tmSR.res$ev.c1,
                        lab = colnames(A))
@@
                            label = paste("r = ", cor), size = 6)
-   i <- (e | a | b) 
-   j <- (f | (g / h)) + plot_layout(widths = c(3, 1))
-   png(file = here("Plots", "sr-plot.png"), width=1200, height=1200)
-       (i / j) +
-       plot_annotation(tag_levels = "a") &
-       theme(plot.tag = element_text(size = 20))
+
+   png(file = here("Plots", "sr-plot-eigen-p.png"), width=600, height=625)
+   a
    dev.off()
+   png(file = here("Plots", "sr-plot-eigen-g.png"), width=600, height=625)
+   b
+   dev.off()
+   png(file = here("Plots", "sr-plot-reord.png"), width=600, height=625)
+   e
+   dev.off()
+   png(file = here("Plots", "sr-corr-scatter.png"), width=350, height=625)
+   g / h
+   dev.off()
+```
 
+```{r GRS Plot}
+   source(here("Functions", "TwoModeGRS.R"))
+   tmSR.res <- TwoModeGRS(A)
+    a <- tmSR.res$eigvec.plot.r 
+    b <- tmSR.res$eigvec.plot.c 
+    c <- tmSR.res$eigval.plot.r
+    d <- tmSR.res$eigval.plot.c
+    a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+    e <- tmSR.res$A.plot
+
+   dat.r <- data.frame(ca = as.numeric(tm$ca.eigvec.r[,1])*-1,
+                       sr = tmSR.res$ev.r1,
+                       lab = rownames(A))
+   dat.c <- data.frame(ca = as.numeric(tm$ca.eigvec.c[,1])*-1,
+                       sr = tmSR.res$ev.c1,
+                       lab = colnames(A))
+   cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
+   g <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
+                  label = "lab", repel = TRUE, 
+                  add = "reg.line", 
+                  add.params = list(color = "gray80", 
+                                    linewidth = 0.5, linetype = 2)) +
+                  annotate("text", x = -.2, y = .3, 
+                           label = paste("r = ", cor), size = 6) 
+   cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
+   h <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
+                  label = "lab", repel = TRUE, 
+                  add = "reg.line",
+                  add.params = list(color = "gray80", 
+                                    linewidth = 0.5, linetype = 2)) +
+                  annotate("text", x = -.2, y = .3, 
+                           label = paste("r = ", cor), size = 6)
+
+   png(file = here("Plots", "grs-plot-eigen-p.png"), width=600, height=625)
+   a
+   dev.off()
+   png(file = here("Plots", "grs-plot-eigen-g.png"), width=600, height=625)
+   b
+   dev.off()
+   png(file = here("Plots", "grs-plot-reord.png"), width=600, height=625)
+   e
+   dev.off()
+   png(file = here("Plots", "grs-corr-scatter.png"), width=350, height=625)
+   g / h
+   dev.off()
 ```

### 2024-05-22 duality-presentation.qmd
duality at 50

---
title: "Coasting on Duality"
subtitle: "What We Get When We Do Two-Mode Network CA"
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
slide-number: true
slide-level: 2
progress: true
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
    footer: "https://github.com/olizardo/correspondence-analysis-two-mode-networks"
---

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership

::: aside
*Breiger 2000; Roberts 2000; Faust 2005; D’Esposito et al. 2014

**Bonacich 1991; Borgatti & Everett 1997
:::

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership
- CA Proponents point to some important advantages of CA for the analysis of two-mode networks:
  - CA can provide a "joint display" of the structural connectivity patterns in two-mode data
  - (M)CA can be used to "visually explore" two-mode networks

::: aside
*Breiger 2000; Roberts 2000; Faust 2005; D’Esposito et al. 2014

**Bonacich 1991; Borgatti & Everett 1997
:::

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership
- This presentation builds on and expands on previous  research on using CA for two-mode data analysis but discusses aspects of CA that have not been emphasized in previous research:
  - CA as computing a type of "(Reflective) Dual Centrality" on the entities in the two-modes*

:::aside
*Rediscovered recently under the heading of the Economic Complexity Index (ECI; Hidalgo and Hausman 2009)
:::

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership
- This presentation builds on and expands on previous  research on using CA for two-mode data analysis but discusses aspects of CA that have not been emphasized in previous research:
  - CA as extracting "positional" information the entities in the two modes
    - With positions defined by "Generalized Relational Similarity"
    - Similarly central entities have similar patterns of connectivity with entities that are themselves similar**

:::aside
**Jeh & Widom 2002; Kovacs 2010; Lizardo 2024
:::

## Southern Women Data
{{< include southern-women.html >}}

## What do we get when we do CA?
:::: {.columns}
::: {.column width="50%"}
- Taken singly, the first dimension of CA suggests an **ordering**
  - Is this connected to centrality?
  - If so, what is the relation (if any) to existing two-mode centralities?*
- Take jointly, the first two dimensions of CA suggest a **grouping**
  - Connected to group cohesion?**
  - Affinity of actors for particular events?***
:::

::: aside
 *Bonacich 1991

 **Doreian 1979

 ***Faust 2005
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data Obtained from Simple CA of the Affiliation Matrix](Plots/ca-southern-women.png)
:::
::::

## What do we get when we do CA?

- To (try to) shed light on these questions I examine the cases of three relatively recent "accidental reinventions" of CA in the "network science" and "data science" (computer and information science) literatures:
  - The "Economic Complexity Index" (ECI)*
    - Relevant to the question of centrality
  - *SimRank* and Generalized Relational Similarity**
    - Relevant to the question of grouping
- The ECI case has been examined in detail in recent work***
- The second and third are new to this presentation
  - All shed light on what exactly we get when we use CA to analyze two mode data

::: aside
*Hidalgo & Hausmann 2009

**Jeh & Widom 2002; Kovacs (2010)

***van Dam et al. 2021; Mealy 2019
:::

## CA and Reflective Centrality

- Hidalgo & Hausmann (2009) introduced a "novel" way to rank nodes in two-mode networks
  - Their approach, called "reflective," was originally applied to country-by-product networks*

. . .

  - The method has since been shown to be applicable to other two-mode networks (like cultural networks)**
  - I introduce here using the more familiar sociological example of the **duality of persons and groups*****

::: aside
*Referred to as the Economic Complexity Index (ECI)

**Lizardo 2018

***Breiger 1974
:::

## CA and Reflective Centrality

- Notation:
  - A two-mode network composed of a set of persons $P$
  - Their affiliation relations to a set of groups $G$
  - Represented by an affiliation matrix $\mathbf{A}$

. . .

  - $\mathbf{A}$ is of dimensions $|P| \times |G|$ with people along the rows and groups across the columns
    - $|P|$ is the cardinality of the persons set  
    - $|G|$ is the cardinality of the group set

. . .

  - $\mathbf{A}$ has cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.

## CA and Reflective Centrality
- If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old **degree centrality***

- Given this, the degree-centrality of people is given by:

$$
C^R_p(1) = \sum_g a_{pg}
$$ {#eq-dp}

- And for groups:

$$
   C^R_g(1) = \sum_p a_{pg}
$$ {#eq-dg}

:::aside
*Faust 1997, eqs. 4 and 5
:::

## CA and Reflective Centrality

- Once we have the first-order quantities $C^R_p(1)$ and $C^R_g(1)$, we can compute "second-order centralities" $C^R(2)$ for both people and groups

- For people, these are given by:*

$$
C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
$$ {#eq-r2p}

- And for groups:**

$$   
C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
$$ {#eq-r2g}

:::aside
*People are central when the average sum of the size of the groups they belong to is large

**Groups are central when the average activity of the people who belong to them is large
:::

::: {.notes}
Equation @eq-r2p says "people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number). Equation @eq-r2g says "groups are more central when the average activity of their members is large" (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
:::

## CA and Reflective Centrality

- Of course, once we have the second-order quantities $C^R_p(2)$ and $C^R_g(2)$, we can compute "third-order centralities" $C^R(3)$ for both people and groups

- For people, these are given by:

$$
   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
$$ {#eq-r3p}

- And for groups:

$$   
   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
$$ {#eq-r3g}

:::aside
*People are central when the average sum of the average activity of the members of the groups they belong to is large

**Groups are central when the average sum of the average size of the groups their members belong to is large
:::

::: {.notes}
Equation @eq-r3p says something like "people are more central when the average sum of the average activity of the members of the groups they belong to is large" (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number). Equation @eq-r3g says, "groups are more central when the average sum of the average size of the groups their members belong to is large." (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
:::

## CA and Reflective Centrality
- More generally, Hidalgo & Hausmann show that we can define a **series** of reflective quantities* $C^R_p(4), C^R_p(5)...C^R_p(q)$ and $C^R_g(4), C^R_g(5)...C^R_g(q)$

- For people, these are given by:

$$   
C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
$$ {#eq-rip}

- And for groups:

$$   
C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
$$ {#eq-rig}

::: aside
*Whose verbal and substantive interpretation becomes more elusive as the number of iterations increases
:::

## CA and Reflective Centrality

- For people:
  - Even-numbered reflections measure centrality based on the **average size** of the groups a person belongs to* 
  - Odd-numbered reflections measure centrality based the **average number of memberships** of each member of the groups they belong to*

. . .

- For groups:
  - Even-numbered reflections measure centrality based on the **average number of memberships** of the members of the group*
  - Odd-numbered reflections measure centrality based on the **average size** of the group memberships of each member*

::: aside
*and the average of the average of the average of the average...
:::

## Reflection Trajectories for SWD
:::: {.columns}
::: {.column width="50%"}
![Person Reflection Trajectories](Plots/p-reflections.png)
:::

::: {.column width="50%"}
![Group Reflection Trajectories](Plots/g-reflections.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Person Reflection Trajectories](Plots/p-reflections.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Person Reflection Trajectories](Plots/g-reflections.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Scatter of standardized person reflection centralities and person scores on first CA dimension (r = 1.0)](Plots/p-ca-ref-corr.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Scatter of standardized group reflection centralities and group scores on first CA dimension (r = 1.0)](Plots/g-ca-ref-corr.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Method of Reflections versus CA

- So the method of reflections is an accidental re-invention of CA!*

- But *why* is it a reinvention of CA?
  - Short answer: 
    - The final reflection scores are the (second largest) eigenvalues of a matrix, just like CA!

- What's the (slightly) longer answer? 
  - A couple of things first:
    - $Dp$ is a $|P| \times |P|$ square degree matrix containing each person's $C_p^R(1)$ (activity) along the diagonals and zeroes in every other cell
    - $Dg$ is a $|G| \times |G|$ square degree matrix containing each group's $C_g^R(1)$ (size) along the diagonals and zeroes in every other cell

:::aside
*On the first dimension: Mealy 2019; van Dam et al. 2021
:::

## Method of Reflections versus CA
- So the method of reflections is an accidental re-invention of CA!*

- But *why* is it a reinvention of CA?

  - Short answer: 
    - The final reflection scores are the (second largest) eigenvalues of a matrix, just like CA!

- What's the (slightly) longer answer?

  - In the limit ($q \rightarrow \infty$), the iterative Hidalgo-Hausmann reflective centralities are the solution of the eigensystem:*

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
$$ {#eq-dam1}

$$
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
$$ {#eq-dam2}

::: aside
*van Dam et al. 2021
:::

## CA versus Bonacich
- Do the matrices whose eigenvalues yield the equilibrium reflective centralities look familiar to you?

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
$$

$$
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
$$

- They should, because the solutions to the following eigensystem give the Bonacich $(C^B)$ dual centralities for two-mode networks:*

$$
\left(AA^T\right)C^B_p = \lambda C^B_p
$$ {#eq-bon1}

$$
\left(A^TA\right)C^B_g = \lambda C^B_g
$$ {#eq-bon2}

:::aside
*Bonacich 1991
:::

## CA versus Bonacich
- So both Reflective and Bonacich centralities are eigenvectors of the person and group **projections** of the two-mode network

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
\\
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
\\
\left(AA^T\right)C^B_p = \lambda C^B_p
\\
\left(A^TA\right)C^B_g = \lambda C^B_g
$$

- But for reflective centralities the affiliation matrix and its transpose are pre-multiplied by the **inverse of the degree matrix of each mode***

:::aside
* van Dam et al. 2021
:::

## CA versus Bonacich
- So both Reflective and Bonacich centralities are eigenvectors of the person and group **projections** of the two-mode network

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
\\
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
\\
\left(AA^T\right)C^B_p = \lambda C^B_p
\\
\left(A^TA\right)C^B_g = \lambda C^B_g
$$

- Similar to how each entry in the original affiliation matrix is weighted in CA performing the SVD!*

$$
Dp^{-1/2}ADg^{-1/2} = X_p \lambda X_g^T
$$

::: aside
**Faust 2005, eq. 7.8 
:::

## CA versus Bonacich
- We can gain some insight into the equivalence between CA and Reflections by expressing the degree-weighted projections in terms of each cell entry:*

$$
a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
$$ {#eq-mealy1}

  - In @eq-mealy1, the fraction's numerator is equal to one only when person $p$ and $p'$ are both members of $g$.

    - Summed across groups, this is the number of **shared groups** between $p$ and $p'$: 
      - The cell entries in the Bonacich matrix (raw projections)

    - But in @eq-mealy1 the numerator is *weighted* by the product of group size and each person's number of memberships (in contrast to Bonacich)

::: aside
*Mealy 2019, eq. 4
:::

## CA versus Bonacich
- We can gain some insight into the equivalence between CA and Reflections by expressing the degree-weighted projections in terms of each cell entry:*

$$
    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
$$ {#eq-mealy2}

  - In @eq-mealy2, the fraction's numerator is equal to one only when group $g$ and $g'$ both have $p$ as a member

    - Summed across persons, the number of **shared members** between $g$ and $g'$:
      - The cell entries in the Bonacich matrix (raw projections)

    - But in @eq-mealy2 the numerator is *weighted* by the product of group size and each person's number of memberships (in contrast to Bonacich)

::: aside
*Mealy 2019, eq. 4
:::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
![Bonacich Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- The CA re-ordered affiliation matrix is **block-diagonal** 
- Reveals (a well-known and storied*) grouping of persons and groups into two clusters 
![](Plots/doreian-sw.png)
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central events contain the most central people and central people attend the most central events:
  - $E1 = \{Laura, Brenda, Evelyn\}$
- In CA ordering central events are *not* the ones with most members! 
  - $\{E7, E8, E9\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central events contain the most central people and central people attend the most central events:
  - $E1 = \{Laura, Brenda, Evelyn\}$
- But they are the most central in the Bonacich ordering!
  - $\{E7, E8, E9\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Least central events contain many of the least central people:
  - $E11 = \{Helen, Flora, Olivia, \\ Nora\}$
- In CA ordering less central events are *not* the ones with the least members! 
  - $\{E1, E13, E14\}$
  - $E1$ is the most central!
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central people attend the most central events and central events are attended by central people:
  - $Frances = \{E3, E5, E4, E6, E8\}$
- Not necessarily those with the most memberships
  - $\{Theresa, Evelyn, Brenda\}$

:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- The least central person $\{Nora\}$ is not the one with the least memberships
  - Instead $\{Nora\}$ shares many memberships with less central people (those who attend the less central events)
  - $\{Flora,  Olivia, Sylvia, Kat.\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central people attend the most central events and central events are attended by central people:
  - $E1 = \{Laura, Brenda, Evelyn\}$
- In CA ordering central events are *not* the ones with the most members! 
  - e.g., $\{E7, E8, E9\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![Bonacich Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::
::: {.column width="50%"}
- The Bonacich re-ordered affiliation matrix is **upper triangular**
- Arranges persons and groups from top to bottom and left to right by activity and size respectively
  - More central people and groups at the upper right
  - Less central people and groups at the lower left
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![Bonacich Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::
::: {.column width="50%"}
- Bonacich centrality is dual but it is "simple" dual
  - Central people attend well-attended events
  - Central events are attended by people who attend many events
- Centrality in one mode is a weighted *sum* of the other mode's centrality*

$$
    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
$$ {#eq-fau1}

$$
    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
$$ {#eq-fau2}
:::
::::

:::aside
*Faust 1997, p. 170
:::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::
::: {.column width="50%"}
- CA centrality is dual but it is "generalized similarity" dual*
  - Central people attend central (not necessarily the most well-attended) events
  - Central events are attended by central (not necessarily the most active) people
- Some people calculate Bonacich centralities when they have this more "generalized similarity" idea of centrality in mind!
  - They should use CA instead
:::
::::
:::aside
*Kovacs 2010; Lizardo 2024
:::

## CA and Similarity
- Another way to gain insight into what CA is doing is by looking at the similarity of people and groups in the data

- As we all know,* the usual one mode projections are obtained via:

$$
S_p = AA^T
$$ {#eq-brei1}

$$
S_g = A^TA
$$ {#eq-brei1}

- Which can be thought of as **similarity matrices** based on the shared neighbors in the two mode network

::: aside
Breiger 1974
:::

## CA and Similarity
- Another way to gain insight into what CA is doing is by looking at the similarity of people and groups in the data

- But we can also obtain *other-mode degree weighted* versions of these similarity matrices*

$$
S^*_p = ADg^{-1}A^T
$$ {#eq-dam5}

$$
S^*_g = A^TDp^{-1}A
$$ {#eq-dam6}

::: aside
van Dam et al. 2021
:::

## CA and Similarity
:::: {.columns}
::: {.column width="50%"}
![Degree-weighted similarity matrix (persons)](Plots/ca-sim-p.png)
:::
::: {.column width="50%"}
![First eigenvalue of person similarity matrix against rank](Plots/ca-eigvec-p.png)
:::
::::

## CA and Similarity
:::: {.columns}
::: {.column width="50%"}
![Degree-weighted similarity matrix (groups)](Plots/ca-sim-g.png)
:::
::: {.column width="50%"}
![First eigenvalue of group similarity matrix against rank](Plots/ca-eigvec-g.png)
:::
::::

## CA and Similarity
:::: {.columns}
::: {.column width="30%"}
![](Plots/ca-sim-p.png)
![](Plots/ca-sim-g.png)

:::
::: {.column width="55%"}
![CA Correspondence Plot](Plots/ca-corr-plot.png)
:::
::::

## Bonacich and Similarity
:::: {.columns}
::: {.column width="30%"}
![](Plots/bon-sim-p.png)
![](Plots/bon-sim-g.png)

:::
::: {.column width="55%"}
![Bonacich Correspondence Plot](Plots/bon-corr-plot.png)
:::
::::

## CA and Generalized Relational Similarity
- Bonacich orders people according to attendance at central (most-well attended) events and events according to attendance by central (most active) people
- CA orders people according to attendance at **similarly central events** and events according to attendance by **similarly central people** 
  - This is precisely the definition of **generalized relational similarity** (GSR)*

. . .

- In that case proximity of nodes in the first CA dimension may reveal a type of GSR in two-mode networks
  - Let us see how that works in the case of two candidate two-mode GSRs
    - *SimRank* *
    - Similarity Weighted Correlation Distance**

:::aside
*Jeh & Widom 2002

**Kovacs 2010; Lizardo 2024
:::

## CA and SimRank

- To explore the link between CA and GSR we can compare CA with one independently developed approach for obtaining GSRs in two-mode networks:* 
  - *SimRank*

- According to *SimRank*:
  - People are *similar* if they belong to *similar* groups
  - Groups are *similar* if they share *similar* members

- In *SimRank* the similarity between two people is given by:

$$
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
$$ {#eq-simrank1}

:::aside
*Jeh & Widom 2002
:::

## CA and SimRank

- To explore the link between CA and GSR we can compare CA with one independently developed approach for obtaining GSRs in two-mode networks:* 
  - *SimRank*

- According to *SimRank*:
  - People are *similar* if they belong to *similar* groups
  - Groups are *similar* if they share *similar* members

- In *SimRank* the similarity between two groups is given by:

$$
S(g, g') = \frac{\alpha}{C_g^R(1)C_{g'}^R(1)}
    \sum_{i = 1}^{C_g^R(1)} \sum_{j = 1}^{C_{g'}^R(1)} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
$$ {#eq-simrank2}

:::aside
*Jeh & Widom 2002
:::

## CA and SimRank

- *SimRank*  for people:

$$
\small
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
$$

- *SimRank*  for groups:

$$
\small
S(g, g') = \frac{\alpha}{C_g^R(1)C_{g'}^R(1)}
    \sum_{i = 1}^{C_g^R(1)} \sum_{j = 1}^{C_{g'}^R(1)} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
$$

With:*

- $S(p, p) = 1$ and $S(g, g) = 1$ and $0 > \alpha < 1$

:::aside
*Jeh & Widom 2002
:::

::: {.notes}
In two-mode networks, SimRank scores can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation @eq-simrank1 using baseline values. Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset. We then plug those values into equation @eq-simrank2, then loop back to equation @eq-simrank1 with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations
:::

## CA and SimRank
:::: {.columns}
::: {.column width="50%"}
![SimRank Re-ordered Affiliation Matrix](Plots/sr-plot-reord.png)
:::

::: {.column width="50%"}
![Simrank Similarity Matrix Eigenvalue Plot (Persons)](Plots/sr-plot-eigen-p.png)
:::
::::

## CA and SimRank
:::: {.columns}
::: {.column width="50%"}
![SimRank Re-ordered Affiliation Matrix](Plots/sr-plot-reord.png)
:::

::: {.column width="50%"}
![Simrank Similarity Matrix Eigenvalue Plot (Groups)](Plots/sr-plot-eigen-g.png)
:::
::::

## CA and SimRank
:::: {.columns}
::: {.column width="50%"}
![SimRank Re-ordered Affiliation Matrix](Plots/sr-plot-reord.png)
:::

::: {.column width="30%"}
![Correlation between CA and SimRank Similarity Matrix Eigenvectors](Plots/sr-corr-scatter.png)
:::
::::

## CA and Generalized Relational Similarity

- I recently proposed a measure of generalized relational similarity (GSR) metric for nodes in two-mode networks by "tweaking" a previous approach based on the modified correlation distance*

- For pairs of people the GSR is given by:

$$
\begin{equation}
\scriptsize
    S(p, p') = 
    \frac{
    (A_{p\bullet} - \bar{A}_{p\bullet})
    \mathbf{S(G)}
    (A_{p'\bullet} - \bar{A}_{p'\bullet})^T
    }
    {
    \sqrt{
    (A_{p\bullet} - \bar{A}_{p\bullet})
    \mathbf{S(G)}
    (A_{p\bullet} - \bar{A}_{p\bullet})^T
    }
    \sqrt{
    (A_{p'\bullet} - \bar{A}_{p'\bullet})
    \mathbf{S(G)}
    (A_{p'\bullet} - \bar{A}_{p'\bullet})^T
        }
    }
\end{equation}
$${#eq-kovacs1}

- And for groups:

$$
\begin{equation}
\scriptsize 
    S(g, g') = 
    \frac{
    (A_{g\bullet} - \bar{A}_{g\bullet})
    \mathbf{S(P)}
    (A_{g'\bullet} - \bar{A}_{g'\bullet})^T
    }
    {
    \sqrt{
    (A_{g\bullet} - \bar{A}_{g\bullet})
    \mathbf{S(P)}
    (A_{g\bullet} - \bar{A}_{g\bullet})^T
    }
    \sqrt{
    (A_{g'\bullet} - \bar{A}_{g'\bullet})
    \mathbf{S(P)}
    (A_{g'\bullet} - \bar{A}_{g'\bullet})^T
        }
    }
\end{equation}
$${#eq-kovacs2}

:::aside
*Lizardo 2024; Kovacs 2010
:::

## CA and Generalized Relational Similarity

- With the base similarities between people given by:

$$
\mathbf{S(P)}_{pp'} = \frac{AA_{pp'}}{\sqrt{AA_{pp}AA_{p'p'}}}
$${#eq-lizardo1}

- And the base similarities for groups:

$$
\mathbf{S(G)}_{gg'} = \frac{AA_{gg'}}{\sqrt{AA_{gg}AA_{g'g'}}}
$${#eq-lizardo2}

- Namely, the cosine distances between people and groups based on the corresponding cell entries of the one-mode projections*

:::aside
*Lizardo 2024
:::

## CA and Generalized Relational Similarity
:::: {.columns}
::: {.column width="50%"}
![GRS Re-ordered Affiliation Matrix](Plots/grs-plot-reord.png)
:::

::: {.column width="50%"}
![GRS Similarity Matrix Eigenvalue Plot (Persons)](Plots/grs-plot-eigen-p.png)
:::
::::

## CA and Generalized Relational Similarity
:::: {.columns}
::: {.column width="50%"}
![GRS Re-ordered Affiliation Matrix](Plots/grs-plot-reord.png)
:::

::: {.column width="50%"}
![GRS Similarity Matrix Eigenvalue Plot (Groups)](Plots/grs-plot-eigen-g.png)
:::
::::

## CA and Generalized Relational Similarity
:::: {.columns}
::: {.column width="50%"}
![GRS Re-ordered Affiliation Matrix](Plots/grs-plot-reord.png)
:::

::: {.column width="30%"}
![Correlation between CA and GRS Matrix Eigenvectors](Plots/grs-corr-scatter.png)
:::
::::

## Conclusions
- To paraphrase John Martin paraphrasing Levi-Strauss:
  - There is only *one analysis* in two-mode network analysis and it is CA
  - Rediscovered many times by researchers pursuing seemingly incompatible taks

. . .

- Contra skeptics we can do two-mode "sna" when we do CA
  - Rank nodes against a meaningful centrality metric
  - Detect equivalences (generalized similarities) between node clusters
- These are some of the standard "positional" tasks that take up much sna

## Conclusions
- To paraphrase John Martin paraphrasing Levi-Strauss:
  - There is only *one analysis* in two-mode network analysis and it is CA
  - Rediscovered many times by researchers pursuing seemingly incompatible taks

. . .

- We can also go do more than visualization and joint display
- The first CA dimension particularly represents a seldom noted **duality of ranking and equivalence**
  - Ranking via similarity within clusters
  - Bonacich by contrast is a "pure" centrality ranking

## Conclusions
- To paraphrase John Martin paraphrasing Levi-Strauss:
  - There is only *one analysis* in two-mode network analysis and it is CA
  - Rediscovered many times by researchers pursuing seemingly incompatible taks

. . .

- Using CA for this kind of "spectral clustering" and "spectral ranking" is not common but feasible
  - Another approach (similar to blockmodeling):
    - Use CA for spectral clustering
    - Then use **specific CA** (on sub-tables defined by the eigengap heuristic) to split clusters into further clusters

## Thanks!

[https://github.com/olizardo/correspondence-analysis-two-mode-networks](https://github.com/olizardo/correspondence-analysis-two-mode-networks)

olizardo@soc.ucla.edu

### 2024-05-22 reflections-as-CA.qmd
duality at 50

---
title: "Reflections as Correspondence Analysis"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
---

# Setup
```{r setup}
     knitr::opts_chunk$set(include = TRUE, 
                       echo = FALSE, 
                       warning = FALSE, 
                       message = FALSE)
     library(factoextra)
     library(here)
     library(kableExtra)
```

```{r custom functions}
     source(here("Functions", "tmCA.R"))
     source(here("Functions", "reflections.R"))
     source(here("Functions", "sw.dat.R"))
     source(here("Functions", "cult.dat.R"))
     source(here("Functions", "ref.long.dat.R"))
     source(here("Functions", "bump.plot.R"))
```

# Data
```{r Southern Women HTML data table}
   kbl(sw.dat()[sort(rownames(sw.dat())), ], align = "c", format = "html") %>% 
      kable_minimal(full_width = F, font_size = 26) %>% 
      row_spec(1:18, extra_css = "border-bottom: 0px;line-height: 20px;") %>% 
      save_kable(here("southern-women.html"))
```

# Analysis
## Standard Correspondence Analysis of Two Mode Data and Correspondence Plot
```{r CA of affiliation matrix for Southern Women}
   ca.sw <- CA(sw.dat(), graph = FALSE)
   plot.dat <- data.frame(rbind(ca.sw$row$coord, ca.sw$col$coord)[, 1:2])
   plot.dat <- cbind(plot.dat, 
                     mode = c(rep(1, nrow(sw.dat())), 
                                        rep(2, ncol(sw.dat()))), 
                     names = rownames(plot.dat))
   p <- ggplot(plot.dat, aes(x = Dim.2, y = Dim.1*-1, 
                             color = factor(mode), label = names))
   p <- p + geom_vline(xintercept = 0, color = "gray")
   p <- p + geom_hline(yintercept = 0, color = "gray")
   p <- p + geom_text_repel(size = 5, max.overlaps = 50)
   p <- p + theme_minimal() 
   p <- p + theme(legend.position = "none",
                  axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16)
                  )
   p <- p + labs(x = "Second Dimension", y = "First Dimension")
   p <- p + scale_color_manual(values = c('red', 'blue'))
   png(file = here("Plots", "ca-sw.png"), width=600, height=625)
   p
   dev.off()
```

```{r CA of affiliation matrix for cultural taste}
   ca.cult <- CA(cult.dat(), graph = FALSE)
   plot.dat <- data.frame(rbind(ca.cult$row$coord, ca.cult$col$coord)[, 1:2])
   n <- c(rownames(ca.cult$row$coord), rownames(ca.cult$col$coord))
   plot.dat <- cbind(plot.dat, 
                     mode = c(rep(1, nrow(cult.dat())), rep(2, ncol(cult.dat()))), 
                     names = n)
   a <- nrow(C) + 1
   p <- ggplot(plot.dat, aes(x = Dim.2, y = Dim.1*-1, 
                             color = factor(mode), label = names))
   p <- p + geom_vline(xintercept = 0, color = "gray")
   p <- p + geom_hline(yintercept = 0, color = "gray")
   p <- p + geom_text(data = plot.dat[1:nrow(cult.dat()), ], size = 1)
   p <- p + geom_text_repel(data = plot.dat[a:nrow(plot.dat), ], size = 5)
   p <- p + theme_minimal() 
   p <- p + theme(legend.position = "none",
                  axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16)
                  )
   p <- p + labs(x = "Second Dimension", y = "First Dimension")
   p <- p + scale_color_manual(values = c('darkgray', 'blue'))
   png(file = here("Plots", "ca-cult.png"), width=600, height=625)
   p
   dev.off()
```

```{r Reflections Analysis}
   ref.sw <- reflections(, iter = 26)
   ref.cult <- reflections(cult.dat(), iter = 50)
   p.ref.sw <- ref.sw$p.r[, seq(2, 26, by = 2)]
   g.ref.sw <- ref.sw$g.r[, seq(2, 26, by = 2)]
   p.ref.cult <- ref.cult$p.r[, seq(2, 50, by = 2)]
   g.ref.cult <- ref.cult$g.r[, seq(2, 50, by = 2)]
   plot.dat.sw <- ref.long.dat(a = p.ref.sw, b = g.ref.sw, max.iter = 26)
   plot.dat.p.sw <- plot.dat.sw$person
   plot.dat.g.sw <- plot.dat.sw$group
   plot.dat.cult <- ref.long.dat(a = p.ref.cult, b = g.ref.cult, max.iter = 50)
   plot.dat.p.cult <- plot.dat.cult$person
   plot.dat.g.cult <- plot.dat.cult$group
```

```{r Reflections Bump Charts}
   png(file = here("Plots", "p-reflections-sw.png"), width=600, height=625)
   p <- bump.plot(plot.dat.p.sw, c = 20, rev  = 1, n.r = 0.01)
   p
   dev.off()
   png(file = here("Plots", "g-reflections-sw.png"), width=600, height=625)
   p <- bump.plot(plot.dat.g.sw, c = 16, rev  = -1, n.r = -1)
   p
   dev.off()
   png(file = here("Plots", "g-reflections-cult.png"), width=625, height=600)
   p <- bump.plot(plot.dat.g.cult, c = 46, rev  = -1, n.r = 1)
   p
   dev.off()
```

```{r}
   source(here("Functions", "eigvec.plot.R"))
   source(here("Functions", "tmCA.R"))
   #tm <- tmCA(cult.dat())
   #saveRDS(tm, file = here("tm-cult.rds"))
   tm.cult <- readRDS(file = here("tm-cult.rds"))
   tm.sw <- tmCA(sw.dat())
   #p <- eigvec.plot(tm.cult$CA.vec.c[, 2]*-1, s = 14)
```

```{r CA versus Reflections Scatter Plots}
   source(here("Functions", "scatter.line.R"))
   png(file = here("Plots", "p-ca-ref-corr-sw.png"), width=625, height=600)
   p <- scatter.line(c = tm.sw$eigvec.r[, 2], r = ref.sw$p.s[, 26], 
                     p.col = "red", t.col = "blue", ref.num = "26th")
   p
   dev.off()
   png(file = here("Plots", "g-ca-ref-corr-sw.png"), width=625, height=600)
   p <- scatter.line(c = tm.sw$eigvec.c[, 2], r = ref.sw$g.s[, 20], 
                     p.col = "red", t.col = "blue", ref.num = "20th", t.size = 4)
   p
   dev.off()
   png(file = here("Plots", "p-ca-ref-corr-cult.png"), width=625, height=600)
   p <- scatter.line(c = tm.cult$eigvec.r[, 2], r = ref.cult$p.s[, 46], 
                     p.col = "red", t.col = "blue", ref.num = "44th", t.size = 0)
   p
   dev.off()
   png(file = here("Plots", "g-ca-ref-corr-cult.png"), width=625, height=600)
   p <- scatter.line(c = tm.cult$eigvec.c[, 2]*-1, r = ref.cult$g.s[, 46], 
                     p.col = "red", t.col = "blue", ref.num = "44th")
   p
   dev.off()
```

   p.plot.dat <- data.frame(cbind(c = tm.sw$CA.vec.r[, 2], r = ref.res$p.s[, 22]))

   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()
```

   p.plot.dat <- data.frame(cbind(c = tm.sw$ca.eigvec.c[, 2], 
                                  r = mr$g.s[, 16]))
   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
   p <- p + geom_point(size = 3, color = "red") 
   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), 
                            color = "blue", size = 6)
   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "14th Reflection (Standardized)")
   p <- p + theme(axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16))
   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()
```

### 2024-05-22 Functions/SimRank.R

@@ -1,51 +1,0 @@
-SimRank <- function(A, C = 0.8, iter = 10) {
-     nr <- nrow(A)
-     nc <- ncol(A)
-     dr <- rowSums(A)
-     dc <- colSums(A)
-     Sr <- diag(1, nr, nr)
-     Sc <- diag(1, nc, nc)
-     rn <- rownames(A)
-     cn <- colnames(A)
-     rownames(Sr) <- rn
-     colnames(Sr) <- rn
-     rownames(Sc) <- cn
-     colnames(Sc) <- cn
-     m <- 1
-     while(m < iter) {
-          Sr.pre <- Sr
-          Sc.pre <- Sc
-          for(i in 1:nr) {
-               for(j in 1:nr) {
-                    if (i != j) {
-                         a <- names(which(A[i, ] == 1)) #objects chosen by i
-                         b <- names(which(A[j, ] == 1)) #objects chosen by j
-                         Scij <- 0
-                         for (k in a) {
-                              for (l in b) {
-                                   Scij <- Scij + Sc[k, l] #i's similarity to j
-                              }
-                         }
-                         Sr[i, j] <- C/(dr[i] * dr[j]) * Scij
-                    }
-               }
-          }
-          for(i in 1:nc) {
-               for(j in 1:nc) {
-                    if (i != j) {
-                         a <- names(which(A[, i] == 1)) #people who chose object i
-                         b <- names(which(A[, j] == 1)) #people who chose object j
-                         Srij <- 0
-                         for (k in a) {
-                              for (l in b) {
-                                   Srij <- Srij + Sr[k, l] #i's similarity to j
-                              }
-                         }
-                         Sc[i, j] <- C/(dc[i] * dc[j]) * Srij
-                    }
-               }
-          }
-          m <- m + 1
-     }
-     return(list(Sr = Sr, Sc = Sc))
-}

### 2024-05-22 Functions/SimRank1.R

@@ -1,0 +1,55 @@
+SimRank1 <- function(A, alpha = 0.8, iter = 10, mode = "in") {
+     V <- nrow(A)
+     if (mode == "in") {
+          d <- colSums(A)
+          }
+     else if (mode == "out") {
+          d <- rowSums(A)
+          }
+     S <- matrix(0, V, V)
+     n <- rownames(A)
+     rownames(S) <- n
+     colnames(S) <- n
+     m <- 1
+     while(m < iter) {
+          for(i in 1:V) {
+               for(j in 1:V) {
+                    # similarities based on successors
+                    if (d[i] == 0 | d[j] == 0) {
+                         S[i, j] <- 0
+                         }
+                    if (i == j & (d[i] != 0 | d[j] != 0)) {
+                         S[i, j] <- 1
+                         }
+                    if (i != j & d[i] > 0 & d[j] > 0 & mode == "in") {
+                         a <- which(A[, i] == 1) # predecessors of i
+                         b <- which(A[, j] == 1) # predecessors of j
+                         K <- 0
+                         for (k in a) {
+                              for (l in b) { #avg. similarity between ij in-neighbors
+                                   K <- K + (A[k, i]/d[i] * S[k, l] * A[l, j]/d[j])
+                              }
+                         }
+                    S[i, j] <- alpha * K 
+                    }
+                    else if (i != j & d[i] > 0 & d[j] > 0 & mode == "out") {
+                         a <- which(A[i, ] == 1) # successors of i
+                         b <- which(A[j, ] == 1) # successors of j
+                         K <- 0
+                         for (k in a) {
+                              for (l in b) { #avg. similarity between ij out-neighbors
+                                   K <- K + (A[i, k]/d[i] * S[k, l] * A[j, l]/d[j])
+                              }
+                         }
+                    S[i, j] <- alpha * K 
+                    }
+               }
+          }
+     m <- m + 1
+     }
+     return(list(S = round(S, 3), m = m, mode = mode))
+}
+
+
+
+

### 2024-05-22 Functions/SimRank2.R

@@ -1,0 +1,49 @@
+SimRank2 <- function(A, alpha = 0.8, iter = 10) {
+     r <- nrow(A)
+     c <- ncol(A)
+     dr <- rowSums(A)
+     dc <- colSums(A)
+     Sr <- diag(1, r, r)
+     Sc <- diag(1, c, c)
+     rn <- rownames(A)
+     cn <- colnames(A)
+     rownames(Sr) <- rn
+     colnames(Sr) <- rn
+     rownames(Sc) <- cn
+     colnames(Sc) <- cn
+     m <- 1
+     while(m < iter) {
+          for(i in 1:r) {
+               for(j in 1:r) {
+                    if (i != j) {
+                         a <- which(A[i, ] == 1) #objects chosen by i
+                         b <- which(A[j, ] == 1) #objects chosen by j
+                         Kij <- 0
+                         for (k in a) {
+                              for (l in b) { #i's similarity to j 
+                                   Kij <- Kij + Sc[k, l] 
+                              }
+                         }
+                         Sr[i, j] <- alpha/(dr[i] * dr[j]) * Kij
+                    }
+               }
+          }
+          for(i in 1:c) {
+               for(j in 1:c) {
+                    if (i != j) {
+                         a <- which(A[, i] == 1) #people who chose object i
+                         b <- which(A[, j] == 1) #people who chose object j
+                         Klm <- 0
+                         for (k in a) {
+                              for (l in b) {
+                                   Klm <- Klm + Sr[k, l] #i's similarity to j
+                              }
+                         }
+                         Sc[i, j] <- alpha/(dc[i] * dc[j]) * Klm
+                    }
+               }
+          }
+          m <- m + 1
+     }
+     return(list(Sr = round(Sr, 3), Sc = round(Sc, 3)))
+}

### 2024-05-22 Functions/TwoModeCA.R

@@ -36,16 +36,22 @@ TwoModeCA <- function(x,
         # Spectral matrices
         CA.r <- eigen(iDr %*% A %*% iDc %*% At) #row CA
         CA.c <- eigen(iDc %*% At %*% iDr %*% A) #column CA
-        S.r.eig <- eigen(S.r) #eigendecomposition of row similarity matrix
-        S.c.eig <- eigen(S.c) #eigendecomposition of column similarity matrix
-        eigvec.r <- CA.r$vectors[, 2:r] #row coordinates
-        eigvec.c <- CA.c$vectors[, 2:c] #column coordinates
+        S.r.eig <- eigen(S.r) #eigendecomposition of row similarity matrix 
+        S.c.eig <- eigen(S.c) #eigendecomposition of column similarity matrix 
+        L.r.eig <- eigen(D.r - S.r) #eigendecomposition of row similarity matrix Laplacian
+        L.c.eig <- eigen(D.c - S.c) #eigendecomposition of column similarity matrix Laplacian
+        eigvec.r <- Re(CA.r$vectors[, 2:r]) #row coordinates
+        eigvec.c <- Re(CA.c$vectors[, 2:c]) #column coordinates
         eigval.r <- CA.r$values[2:r] #row eigenvalues
         eigval.c <- CA.c$values[2:c] #column eigenvalues
-        eigvec.S.r <- S.r.eig$vectors #row similarity eigenvectors
-        eigvec.S.c <- S.c.eig$vectors #column similarity eigenvectors
+        eigvec.S.r <- Re(S.r.eig$vectors) #row similarity eigenvectors
+        eigvec.S.c <- Re(S.c.eig$vectors) #column similarity eigenvectors
         eigval.S.r <- S.r.eig$values #row similarity eigenvalues 
         eigval.S.c <- S.c.eig$values #column similarity eigenvalues
+        eigvec.L.r <- Re(L.r.eig$vectors) #row Laplacian eigenvectors
+        eigvec.L.c <- Re(L.c.eig$vectors) #column Laplacian eigenvectors
+        eigval.L.r <- Re(L.r.eig$values) #row Laplacian eigenvectors
+        eigval.L.c <- Re(L.c.eig$values) #column Laplacian eigenvectors
         bon.eigvec.r <- as.matrix(eigen(AAt)$vectors) #row eigenvectors
         bon.eigvec.c <- as.matrix(eigen(AtA)$vectors) #column eigenvectors
[…]

### 2024-05-22 Functions/TwoModeSR.R

@@ -1,90 +1,0 @@
-TwoModeSR <- function(A, k = 4, n.eig = 4) {
-   #loading functions
-   source(here("Functions", "SimRank.R"))
-   source(here("Functions", "eigval.scatter.R"))
-   source(here("Functions", "eigvec.scatter.R"))
-   source(here("Functions", "corr.scatter.R"))
-   
-   # SimRank of matrix A
-   nr <- nrow(A)
-   nc <- ncol(A)
-   rn <- rownames(A)
-   cn <- colnames(A)
-   sr.res <- SimRank(A)
-   res.eigen.r <- eigen(sr.res$Sr) #eigenvectors/values of row similarity matrix
-   res.eigen.c <- eigen(sr.res$Sc) #eigenvectors/values of column similarity matrix
-   e.r <- res.eigen.r$vectors #row eigenvectors
-   e.c <- res.eigen.c$vectors #column eigenvectors
-   v.r <- res.eigen.r$values[2:nr] #row eigenvalues minus first
-   v.c <- res.eigen.r$values[2:nc] #column eigenvalues minus first
-   e.r <- e.r[, 2:nr] #dropping first row eigenvector
-   e.c <- e.c[, 2:nc] #dropping first column eigenvector
-   e.r <- e.r * -1 # reversing row eigevectors signs
-   e.c <- e.c * -1 # reversing column eigevectors signs
-   ov.r1 <- e.r[, 1] #main eigenvector of row similarity matrix
-   ov.r2 <- e.r[, 2] #second main eigenvector of row similarity matrix
-   ov.c1 <- e.c[, 1] #main eigenvector of column similarity matrix
-   ov.c2 <- e.c[, 2] #second main eigenvector of column similarity matrix
-   
-   # Affiliation matrix plot (SimRank ordering)
-   A.ord <- as.matrix(A[order(ov.r1), order(ov.c1)])
-   A.plot <- ggcorrplot(t(A.ord)) 
-   A.plot <- A.plot + scale_x_discrete(position = "top")
-   A.plot <- A.plot + theme(legend.position = "none",
-                             axis.text.x = element_text(hjust = -0.2))
-   
-   # Correspondence plot of SimRank similarity matrix
-   k.r <- hkmeans(data.frame(e.r[, 1:n.eig]), k = k)
-   k.c <- hkmeans(data.frame(e.c[, 1:n.eig]), k = k)
-   d.r <- data.frame(lab = rownames(A), 
[…]

### 2024-05-22 Functions/discard/TwoModeGRS.R

@@ -1,0 +1,64 @@
+TwoModeGRS <- function(A, k = 4, n.eig = 4) {
+   #loading functions
+   source(here("Functions", "tm.corr.dist.R"))
+   source(here("Functions", "eigval.scatter.R"))
+   source(here("Functions", "eigvec.scatter.R"))
+   source(here("Functions", "corr.scatter.R"))
+   
+   # GRS of matrix A
+   nr <- nrow(A)
+   nc <- ncol(A)
+   rn <- rownames(A)
+   cn <- colnames(A)
+   sr.res <- tm.corr.dist(A)
+   res.eigen.r <- eigen(sr.res$row.sims) #eigenvectors/values of row similarity matrix
+   res.eigen.c <- eigen(sr.res$col.sims) #eigenvectors/values of column similarity matrix
+   e.r <- res.eigen.r$vector #row eigenvectors
+   e.c <- res.eigen.c$vectors #column eigenvectors 
+   v.r <- res.eigen.r$values #row eigenvalues 
+   v.c <- res.eigen.c$values #column eigenvalues
+   ev.r1 <- e.r[, 1] #main eigenvector of row similarity matrix
+   ev.r2 <- e.r[, 2] #second main eigenvector of row similarity matrix
+   ev.c1 <- e.c[, 1] #main eigenvector of column similarity matrix
+   ev.c2 <- e.c[, 2] #second main eigenvector of column similarity matrix
+   
+   # Affiliation matrix plot (GRS ordering)
+   A.ord <- as.matrix(A[order(ev.r1), order(ev.c1)])
+   A.plot <- ggcorrplot(t(A.ord)) 
+   A.plot <- A.plot + scale_x_discrete(position = "top")
+   A.plot <- A.plot + theme(legend.position = "none",
+                            axis.text.x = element_text(hjust = -0.2))
+   
+   # Eigenvalue plot of GRS similarity matrix
+   eig.dat.r <- data.frame(k = 1:nr, value = v.r) %>% 
+      mutate(value = value / max(value))
+   eig.dat.c <- data.frame(k = 1:nc, value = v.c) %>% 
+      mutate(value = value / max(value))
+   eigval.plot.r <- eigval.scatter(eig.dat.r)
+   eigval.plot.c <- eigval.scatter(eig.dat.c)
+   
+   # Main eigenvector plot of GRS similarity matrix
+   k.r <- hkmeans(data.frame(ev.r1), k = 2)
[…]

### 2024-05-22 Functions/discard/TwoModeGSR.R

@@ -1,0 +1,64 @@
+TwoModeSR <- function(A, k = 4, n.eig = 4) {
+   #loading functions
+   source(here("Functions", "tm.corr.dist.R"))
+   source(here("Functions", "eigval.scatter.R"))
+   source(here("Functions", "eigvec.scatter.R"))
+   source(here("Functions", "corr.scatter.R"))
+   
+   # SimRank of matrix A
+   nr <- nrow(A)
+   nc <- ncol(A)
+   rn <- rownames(A)
+   cn <- colnames(A)
+   sr.res <- tm.corr.dist(A)
+   res.eigen.r <- eigen(sr.res$Sr) #eigenvectors/values of row similarity matrix
+   res.eigen.c <- eigen(sr.res$Sc) #eigenvectors/values of column similarity matrix
+   e.r <- res.eigen.r$vectors[, 2:nr] #row eigenvectors minus first
+   e.c <- res.eigen.c$vectors[, 2:nc] #column eigenvectors minus first
+   v.r <- res.eigen.r$values[2:nr] #row eigenvalues minus first
+   v.c <- res.eigen.r$values[2:nc] #column eigenvalues minus first
+   ev.r1 <- e.r[, 1] #main eigenvector of row similarity matrix
+   ev.r2 <- e.r[, 2] #second main eigenvector of row similarity matrix
+   ev.c1 <- e.c[, 1] #main eigenvector of column similarity matrix
+   ev.c2 <- e.c[, 2] #second main eigenvector of column similarity matrix
+   
+   # Affiliation matrix plot (SimRank ordering)
+   A.ord <- as.matrix(A[order(ev.r1), order(ev.c1)])
+   A.plot <- ggcorrplot(t(A.ord)) 
+   A.plot <- A.plot + scale_x_discrete(position = "top")
+   A.plot <- A.plot + theme(legend.position = "none",
+                            axis.text.x = element_text(hjust = -0.2))
+   
+   # Eigenvalue plot of SimRank similarity matrix
+   eig.dat.r <- data.frame(k = 2:nr, value = v.r) %>% 
+      mutate(value = value / max(value))
+   eig.dat.c <- data.frame(k = 2:nc, value = v.c) %>% 
+      mutate(value = value / max(value))
+   eigval.plot.r <- eigval.scatter(eig.dat.r)
+   eigval.plot.c <- eigval.scatter(eig.dat.c)
+   
+   # Main eigenvector plot of SimRank similarity matrix
[…]

### 2024-05-22 Functions/discard/tm.corr.dist.R

@@ -1,0 +1,85 @@
+tm.corr.dist <- function(x) {
+     library(expm)
+     r <- nrow(x)
+     c <- ncol(x)
+     r.c <- diag(r)
+     c.c <- diag(c)
+     r.m <- rowMeans(x)
+     c.m <- colMeans(x)
+     
+     tm.sim <- function(x) {
+          r <- nrow(x)
+          c <- ncol(x)
+          r.r <- x %*% t(x)
+          c.c <- t(x) %*% x
+          r.s <- diag(r)
+          c.s <- diag(c)
+          
+          sim <- function(a, b, c) {
+               res <- a / sqrt(b*c) #cosine similarity 
+               #res <- a (b + c - a) #Jaccard similarity
+               #res <- (2*a) / (b + c) #Dice similarity
+               return(res)
+          }
+          
+          for (i in 1: r) {
+               for (j in 1:r) {
+                    if (i != j) {
+                         r.s[i, j] <- sim(r.r[i, j], r.r[i, i], r.r[j, j])
+                    }
+               }
+          }
+          for (i in 1: c) {
+               for (j in 1:c) {
+                    if (i != j) {
+                         c.s[i, j] <- sim(c.c[i, j], c.c[i, i], c.c[j, j])
+                    }
+               }
+          }
+          rownames(r.s) <- rownames(x)
+          colnames(r.s) <- rownames(x)
+          rownames(c.s) <- colnames(x)
+          colnames(c.s) <- colnames(x)
+          return(list(row.sims = r.s, col.sims = c.s))
+          }
+     
+     sim.res <- tm.sim(x)
+     r.s <- sim.res[[1]]
+     c.s <- sim.res[[2]]
+     
+     for (i in 1: r) {
+          for (j in 1:r) {
+               if (i != j) {
+                    r.x <- x[i, ] - r.m[i]
+                    r.y <- x[j, ] - r.m[j]
+                    r.xy <- r.x %*% c.s * t(r.y)
+                    r.xx <- r.x %*% c.s * t(r.x)
+                    r.yy <- r.y %*% c.s * t(r.y) 
+                    r.num <- sum(r.xy)
+                    r.den <- sqrt(sum(r.xx)) * sqrt(sum(r.yy))
+                    r.c[i, j] <- round(r.num / r.den, 2)
+                    }
+               }
+          }
+     
+     for (i in 1: c) {
+          for (j in 1:c) {
[…]

### 2024-05-22 Functions/eigvec.plot.R

@@ -1,0 +1,26 @@
+eigvec.plot <- function(x, s = 12, lab = "Second") {
+     # Main Eigenvector Plot (CA)
+     km.res <- hkmeans(x, k = 2)
+     eigvec.dat <- data.frame(rank = rank(x), 
+                           value = as.numeric(x),
+                           lab = names(x),
+                           cluster = factor(km.res$cluster))
+     p <- ggscatter(eigvec.dat, 
+                    y = "rank", 
+                    x = "value", 
+                    point = FALSE,
+                    font.label =  s,
+                    label = "lab", 
+                    color = "cluster",
+                    palette = "Dark2",
+                    repel = TRUE
+     )
+     p <- p + theme(axis.text.y = element_blank(),
+                    axis.line.x = element_blank(),
+                    axis.line.y = element_blank(),
+                    axis.title = element_text(size = 12),
+                    axis.text.x = element_text(size = 12),
+                    legend.position = "none")
+     p <- p + labs(y = "Rank", x = paste(lab, "CA Axis", sep = " "))
+ return(p)
+}

### 2024-05-22 Functions/ref.long.dat.R

@@ -1,32 +1,24 @@
 ref.long.dat <- function(a, b, max.iter) {
      library(tidyverse)
-     evens <- function(x) subset(x, x %% 2 == 0)
-     odds <- function(x) subset(x, x %% 2 != 0)
      pr.long <- data.frame(a)
      pr.long <- pr.long %>% 
-          mutate(person = rownames(pr.long)) %>% 
+          mutate(mode = rownames(pr.long)) %>% 
           pivot_longer(
                cols = 1:ncol(a),
                names_to = "iter",
                values_to = "ref"
-          ) %>% 
-          mutate(iter = factor(iter, ordered = TRUE, levels = names(pr.long))) %>%  
-          mutate(n.iter = as.integer(iter))
-     pr.long.e <- filter(pr.long, n.iter %in% c(evens(2:max.iter)))
-     pr.long.o <- filter(pr.long, n.iter %in% c(odds(3:max.iter)))
-     
+          ) %>%
+       mutate(iter = factor(iter, ordered = TRUE, levels = names(pr.long))) %>%
+       mutate(n.iter = as.integer(iter)*2)
      gr.long <- data.frame(b)
      gr.long <- gr.long %>% 
-          mutate(group = rownames(gr.long)) %>% 
+          mutate(mode = rownames(gr.long)) %>% 
           pivot_longer(
                cols = 1:ncol(b),
                names_to = "iter",
                values_to = "ref"
-          ) %>% 
-          mutate(iter = factor(iter, ordered = TRUE, levels = names(gr.long))) %>%  
-          mutate(n.iter = as.integer(iter))
-     gr.long.e <- filter(gr.long, n.iter %in% c(evens(2:max.iter)))
-     gr.long.o <- filter(gr.long, n.iter %in% c(odds(3:max.iter)))
-     return(list(person = list(all = pr.long, even = pr.long.e, odd = pr.long.o), 
-                 group = list(all = gr.long, even = gr.long.e, odd = gr.long.o)))
+          ) %>%
+       mutate(iter = factor(iter, ordered = TRUE, levels = names(gr.long))) %>%
+       mutate(n.iter = as.integer(iter)*2)
+     return(list(person = pr.long, group = gr.long))
 }

### 2024-05-22 Functions/reflections.R

@@ -38,7 +38,7 @@ reflections <- function(x, iter = 20) { #x is a matrix with people as rows and g
          p.c[i, m] <- sum(x[i, ] * g.c[, k]) * (1/p.c[i, 1]) #assign person avg. centrality groups they belong to
       } #end person loop
       for(j in 1:g) {
-         g.c[j, m] <- sum(x[, j] * p.c[, k]) * (1/g.c[j, 1]) #assign genre avg. person centrality of people in group
+         g.c[j, m] <- sum(x[, j] * p.c[, k]) * (1/g.c[j, 1]) #assign group avg. person centrality of people in group
       } #end group loop
       k <- k + 1 #increase counter
    } #end while loop

### 2024-05-22 Functions/scatter.line.R

@@ -1,0 +1,13 @@
+scatter.line <- function(c, r, p.col, t.col, ref.num, t.size = 3) {
+     dat <- data.frame(c = c, r = r)
+     p <- ggplot(data = dat, aes(x = c, y = r))
+     p <- p + geom_point(size = 3, color = p.col) 
+     p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+     p <- p + geom_text(aes(label = rownames(dat)), color = t.col, size = t.size)
+     p <- p + theme_minimal() 
+     p <- p + labs(x = "First CA Dimension", 
+                   y = paste(ref.num, "Reflection (Standardized)", sep = " "))
+     p <- p + theme(axis.text = element_text(size = 16),
+                    axis.title = element_text(size = 16))
+     return(p)
+     }

### 2024-05-22 Functions/scrap.R

@@ -1,0 +1,125 @@
+
+u.r <- matrix(1, nrow = r, ncol = 1) #unit row vector
+u.c <- matrix(1, nrow = c, ncol = 1) #unit column vector
+
+# Derived matrices and vectors
+AAt <- A %*% At #row projection
+AtA <- At %*% A #column projection
+N <- as.numeric(t(u.r) %*% A %*% u.c) #affiliation matrix size
+D.r <- diag(as.vector(A %*% u.c)) #row diagonal degree matrix
+D.c <- diag(as.vector(At %*% u.r)) #column diagonal degree matrix
+
+iDr <- solve(D.r) #inverse of row diagonal degree matrix
+iDc <- solve(D.c)  #inverse of column diagonal degree matrix
+isDr <- solve(sqrt(D.r)) #inverse of square root of row diagonal degree matrix
+isDc <- solve(sqrt(D.c))  #inverse of square root of column diagonal degree matrix
+S.r <- A %*% iDc %*% At #similarity matrix for rows
+S.c <- At %*% iDr %*% A #similarity matrix for columns
+M.r <- iDr %*% S.r #centrality weighted row similarity matrix
+M.c <- iDc %*% S.c #centrality weighted column similarity matrix
+L.r <- D.r - S.r #row similarity matrix Laplacian
+L.c <- D.c - S.c #column similarity matrix Laplacian
+NL.r <- isDr %*% L.r %*% isDr #normalized row similarity matrix Laplacian
+NL.c <- isDc %*% L.c %*% isDc #normalized row similarity matrix Laplacian
+
+# Spectral matrices
+CA.eig.r <- eigen(iDr %*% A %*% iDc %*% At) #eigendecomposition of centrality weighted row similarity matrix
+CA.eig.c <- eigen(iDc %*% At %*% iDr %*% A) #eigendecomposition of centrality weighted column similarity matrix
+L.eig.r <- eigen(L.r) #eigendecomposition of row similarity matrix Laplacian
+L.eig.c <- eigen(L.c) #eigendecomposition of row similarity matrix Laplacian
+NC.eig.r <- eigen(NL.r) #eigendecomposition of normalized row similarity matrix Laplacian
[…]

### 2024-05-22 Functions/sw.dat.R

@@ -1,0 +1,30 @@
+sw.dat <- function() {
+     A <- matrix(
+          c(
+               1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
+               1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
+               0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
+               1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
+               0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
+               0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
+               0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
+               0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0, 
+               
+               0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
+               0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
+               0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
+               0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
+               0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
+               0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
+               0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
+               0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0, 
+               
+               0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
+               0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
+          ncol = 14, byrow = TRUE)
[…]

### 2024-05-22 southern-women.R

@@ -1,0 +1,61 @@
+require(kableExtra)
+require(here)
+require(FactoMineR)
+
+A <- matrix(
+     c(
+          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
+          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
+          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
+          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
+          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
+          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
+          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
+          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0, 
+          
+          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
+          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
+          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
+          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
+          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
+          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
+          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
+          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0, 
+          
+          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
+          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
+     ncol = 14, byrow = TRUE)
[…]

### 2024-05-22 duality-presentation.qmd
updated duality presentation

@@
 ::: {.column width="50%"}
-![Correspondence Plot of Southern Women Data Obtained from Simple CA of the Affiliation Matrix](Plots/ca-southern-women.png)
+![Correspondence Plot of Southern Women Data Obtained from Simple CA of the Affiliation Matrix](Plots/ca-sw.png)
 :::
@@
 ::: {.column width="50%"}
-![Person Reflection Trajectories](Plots/p-reflections.png)
-:::
-
-::: {.column width="50%"}
-![Group Reflection Trajectories](Plots/g-reflections.png)
+![Person Reflection Trajectories](Plots/p-reflections-sw.png)
+:::
+
+::: {.column width="50%"}
+![Group Reflection Trajectories](Plots/g-reflections-sw.png)
 :::
@@
 ::: {.column width="50%"}
-![Person Reflection Trajectories](Plots/p-reflections.png)
-:::
-
-::: {.column width="50%"}
-![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
+![Person Reflection Trajectories](Plots/p-reflections-sw.png)
+:::
+
+::: {.column width="50%"}
+![Correspondence Plot of Southern Women Data](Plots/ca-sw.png)
 :::
@@
 ::: {.column width="50%"}
-![Person Reflection Trajectories](Plots/g-reflections.png)
-:::
-
-::: {.column width="50%"}
-![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
+![Person Reflection Trajectories](Plots/g-reflections-sw.png)
+:::
+
+::: {.column width="50%"}
+![Correspondence Plot of Southern Women Data](Plots/ca-sw.png)
 :::
@@
 ::: {.column width="50%"}
-![Scatter of standardized person reflection centralities and person scores on first CA dimension (r = 1.0)](Plots/p-ca-ref-corr.png)
-:::
-
-::: {.column width="50%"}
-![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
+![Scatter of standardized person reflection centralities and person scores on first CA dimension (r = 1.0)](Plots/p-ca-ref-corr-sw.png)
+:::
+
+::: {.column width="50%"}
+![Correspondence Plot of Southern Women Data](Plots/ca-sw.png)
 :::
@@
 ::: {.column width="50%"}
-![Scatter of standardized group reflection centralities and group scores on first CA dimension (r = 1.0)](Plots/g-ca-ref-corr.png)
-:::
-
-::: {.column width="50%"}
-![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
+![Scatter of standardized group reflection centralities and group scores on first CA dimension (r = 1.0)](Plots/g-ca-ref-corr-sw.png)
+:::
+
+::: {.column width="50%"}
+![Correspondence Plot of Southern Women Data](Plots/ca-sw.png)
 :::

### 2024-07-10 main.tex
ncsu presentation
Updates from Overleaf

@@
 % Main title of the paper
-\title [mode = title]{The Correspondence Analysis of Two-Mode Networks: Centrality and Positional Interpretations}                      
+\title [mode = title]{The Correspondence Analysis of Two-Mode Networks Revisited}                      
 % Title footnote mark
@@
 %       gplus=<gplus id>]
-\author[1]{Omar Lizardo}[orcid=0000-0001-7511-2910]
+\author[1]{Omar Lizardo}[orcid=0000-0002-5405-3007]
 
@@
 \begin{abstract}
-Lorem...
+This paper reconsiders the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting aspects that have not been previously emphasized, going beyond the use of CA as a visualization tool.
+It argues that CA can be used to compute a dual centrality score on both modes, related to but not mathematically equivalent to the \citet{bonacich1991simultaneous} two-mode centrality.
+This ``reflective'' centrality score connects the use of CA in two-mode network analysis to its use in other disciplines as a method of ordination.
+Additionally, I show CA can extract latent positional information on people and groups, with similarities to recent work on ``generalized relational similarity" in two-mode networks.
+Thus, CA can be used for indirect community or subgroup identification in two-mode networks, connecting to its use in some disciplines as a clustering method.
+The paper demonstrates these applications of CA, comparing them with the Bonacich dual centrality scores, and provides guidance on using CA for structural analysis in network analysis.
 \end{abstract}
@@
 \begin{highlights}
-    \item 
-    \item 
-    \item 
+  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
+  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
+  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
 \end{highlights}
@@
 \begin{keywords}
-Generalized similarity \sep Duality \sep Two-mode networks \sep Correspondence Analysis
+Correspondence Analysis \sep Generalized Relational Similarity \sep Duality \sep Two-mode networks \sep Centrality 
 \end{keywords}
@@
 \newpage
-\section{Introduction}
+\section{Introduction} \label{sec:intro}
 While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
@@
 This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Nevertheless, it departs from the already-mentioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been as emphasized or treated in detail before, moving beyond the focus on data visualization.
+Nevertheless, it departs from the already-mentioned previous efforts in highlighting aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before, moving beyond the focus on data visualization.
 
 Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
-With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
-In social network analysis, this applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.
+With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general centrality score connecting the use of CA in two-mode network analysis to way CA features in some disciplines also concerned with two-mode data (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
+In social network analysis, this applies to conceptualizing the venerable duality between people and the groups they join \citep{breiger1974duality}.
 
-Additionally, I argue that CA can also be used to extract latent \textit{positional} information on people and groups.
-First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
+In addition to its capacity to reveal latent ordinal structure in two-mode data, I also argue that CA can also be used to extract latent \textit{positional} information on people and groups.
+Particularly, there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
 As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups).
-Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode.
-Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}---e.g., the discovery of sociometrically similar subsets of entities in affiliation networks.
+Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks connecting to how CA is used in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}, namely, the discovery of sociometrically similar subsets of entities in affiliation networks.
+In this way, CA emerges as a tool central to the usual network-analytic tasks and not just as a visualization or data-summarization tool.
 
-\subsection{Organization of the Paper}
+\subsection{Organization of the Paper} \label{subsec:org}
 The rest of the paper is organized as follows.
@@
 This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes.
-This approach is computationally and mathematically related but \textit{not} equivalent to the matrix factorization of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores.
+This approach is computationally and mathematically related but \textit{not} equivalent to the eigenvalue decomposition of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores.
 I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers instead the core-periphery structure of the two-mode network.
+I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers the core-periphery structure of the two-mode network instead.
 In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
 
-\section{Reflective Centralities in Two-Mode Networks}
+\section{Reflective Centralities in Two-Mode Networks} \label{sec:ref2mode}
 In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
 Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).
-Subsequent shows that there is no logical connection between the formal method and this particular application.
+Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
 As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
@@
 That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
-In what follows, I will generally refer to $C^R_p(1)$ as (individual) ``activity'' and $C^R_p(1)$ as (group) ``size.''
-
 As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.
@@
 
-\subsection{Empirical Example}
+\subsection{Empirical Example} \label{subsec:ex}
 Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}.
-The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
+The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
 
@@
 \end{figure}
-\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks}
+
+\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks} \label{sec:ca}
 The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
@@
 
-To see this, recall that the solutions to the following eigensystem give the Bonacich $(C^B)$ two-mode centralities:
+To see this, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:
 
@@
 Where $A$ is the original affiliation matrix.
-The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, as given by the singular value decomposition (SVD) of the rectangular affiliation matrix.
+Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, obtained from the singular value decomposition (SVD) of the rectangular affiliation matrix $A$.
 Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
@@
 Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
+These equations make clear that in the Bonacich two-mode, the centrality of people is proportional to the centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
 These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
@@
 The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.
+We will return to this crucial point later.
 
-As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue (matrix factorization) problem.
+\subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
+To see the connection between the reflective and Bonacich two-mode centralities more clearly, we can motivate the Bonacich two-mode centrality using the same ``reflective'' approach we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}.
+Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centrality for two-mode networks~\citep{bonacich1991simultaneous}, but it will help clarify the similarities and differences between the HH and the Bonacich approaches.
+Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:
+
+\begin{equation}
+   C^B_p(2) = \sum_g a_{pg}C^R_g(1)
+   \label{eq:BR2_p}
+\end{equation}
+
+\begin{equation}
+   C^B_g(2) = \sum_p a_{pg}C^R_p(1)
+   \label{eq:BR2_g}
+\end{equation}
+
+Equation~\ref{eq:BR2_p} says that people are central when the sum of the number of members of the groups they belong to is large.
+Equation~\ref{eq:BR2_g} says groups are central when the sum of the number of memberships of the people who belong to them is large.
+
+As before, we can keep on going and define a third-order reflection using the formulas:
+
+\begin{equation}
+   C^B_p(3) = \sum_g a_{pg}C^R_g(2)
+   \label{eq:BR3_p}
+\end{equation}
+
+\begin{equation}
+   C^B_g(3) = \sum_p a_{pg}C^R_p(2)
+   \label{eq:BR3_g}
+\end{equation}
+
+Equation~\ref{eq:BR3_p} says that people are central when the sum of the sum of the number of memberships held by the people who belong to the groups they belong to is large.
+Equation~\ref{eq:BR3_g} says that groups are central when the sum of the sum of the number of members of the groups their members belong to is large.
+Once again, we can keep going and define fourth order, fifth order, and higher reflections $C^R_p(4), C^R_p(4) \ldots C^R_p(q)$, where the centralities of nodes in one mode are based on the sums, of the sums, of the sums, of the centralities of nodes in the other mode.
+
+More generally, and in parallel with equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}, the reflective Bonacich centralities for persons and groups are given by:
+
+\begin{equation}
+   C^B_p(q) = \sum_g a_{pg}C^R_g(q-1)
+   \label{eq:BRq_p}
+\end{equation}
+
+\begin{equation}
+   C^B_g(q) = \sum_p a_{pg}C^R_p(q-1)
+   \label{eq:BRq_g}
+\end{equation}
+
+This just says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
+
+Of course, it is evident that these sum of sums would diverge to a bigger and bigger quantity at each step $q$.
+To prevent this and guarantee convergence, we normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:
+
+\begin{equation}
+   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
+   \label{eq:BRq_pn}
+\end{equation}
+
+\begin{equation}
+   C^B_g(q) = \frac{C^B_g(q)}{||C^B_g(q)||_2}
+   \label{eq:BRq_gn}
+\end{equation}
+
+Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}.
+Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums around steady values after a few iterations.
+These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach.
+Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power'' method of \citet{mises1929praktische}).
+
+This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' centralities and the Bonacich two-mode centralities.
+In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.
+
+\subsection{HH Reflective Centrality as an Eigenvector Centrality} \label{subsec:refeigen}
+Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem.
+As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.
+
 To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
@@
 Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups.
-The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before factorizing, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
+The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
 
@@
 
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by inverse of the square root of the product of the row and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using}, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
+Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
+9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
 
 Degree-pre-weighting does not radically alter the nature of the data.
-Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perenial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
+Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
 This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large.
@@
 Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
-In the Bonacich case, a factorization of the one-mode projection of the two-mode network; in the reflective case, factorization of the degree-weighted one-mode projection of the same network.
+In the Bonacich case, an eigendecomposition of the one-mode projection of the two-mode network; in the reflective case, an eigendecomposition of the degree-weighted one-mode projection of the same network.
 Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
@@
 
-\section{Illustrative Analysis of the Southern Women Data}
+\section{Illustrative Analysis of the Southern Women Data} \label{sec:anal}
 \subsection{Row and Column Re-ordered Affiliation Matrices}
@@
 Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
-The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side).
+The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side).
 The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
 
-\subsection{Eigenvector Plot}
+\subsection{Eigenvector Plot} \label{subsec:eigplot}
 We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
@@
     \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}
-    \caption{Caption}
+    \caption{CA Eigenvector plot for persons (a) and groups (c).
+The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis.
+Insets show the eigenvalue orderings on the x-axis ($k$) for persons (b) and groups (d).}
     \label{fig:ca-eigvec}
@@
 
-\subsection{Normalized Similarity Plot}
+\subsection{Normalized Similarity Plot} \label{subsec:normsim}
 To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one.
@@
     \includegraphics[width=1.0\textwidth]{Plots/ca-sim.png}
-    \caption{Caption}
+    \caption{CA similarity plots for persons (a) and groups (b) on the left and right top panel.
+The bottom panel (c) shows the correspondence plot for persons and groups with scores corresponding to the first dimension on the horizontal axis and scores corresponding to the second dimension on the vertical axis.}
     \label{fig:ca-sim}
@@
 
-\subsection{CA Correspondence Plot}
+\subsection{CA Correspondence Plot} \label{subsec:caplot}
 The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions.
@@
     \includegraphics[width=1.0\textwidth]{Plots/bon-sim.png}
-    \caption{Caption}
+    \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel.
+The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis.}
     \label{fig:bon-sim}
@@
 
-This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's eigenvector centrality.
+This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's Bonacich eigenvector centrality.
 For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot.
@@
 
-\subsection{Correspondence Analysis and Generalized Relational Similarity}
+\subsection{Correspondence Analysis and Generalized Relational Similarity} \label{subsec:cagensim}
 As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.
-Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}.
+Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}, where objects' similarities are given by their pattern of connections to objects in the other mode.
 This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence \citep{everett1994regular}.
-In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.
+In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.\footnote{\citet{lizardo2024two} shows the connection between Kovacs's idea of generalized relations similarity and the two-mode network projection.}
 
 An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}.
-In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong, and the similarity of groups is a function of the people who belong to them, making the similarity of persons and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
+In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong to, and the similarity of groups is a function of the people who belong to them, making the similarity of persons and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
 Thus,
@@
     \includegraphics[width=0.8\textwidth]{Plots/sr-plot.png}
-    \caption{Caption}
+    \caption{Simrank versus CA comparison.
+Panel (a) shows the affiliation matrix re-ordered according to the first eigenvector of the SVD of the SimRank similarity matrix.
+Panels (b) and (d) show the corresponding eigenvector and eigenvalue---panels (c) and (e)---plots for persons and groups, respectively, of the SimRank similarity matrix.
+Panel (f) shows the correspondence plot based on the eigenvector decomposition of the SimRank similarity matrix.
+Panels (g) and (h) show the Pearson correlation between the scores corresponding to the first eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis).}
     \label{fig:simrank}
@@
 
-Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with the HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
+Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
 The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
@@
 And indeed, they are not.
-Figure~\ref{fig:simrank}(a) shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices for the row and column nodes.
-Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two blocks of persons and groups, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
+Figure~\ref{fig:simrank}(a) shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices---with $\alpha$ set to $0.8$---for the row and column nodes.
+Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two communities of persons and groups revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
+
 Figure~\ref{fig:simrank}(b) and Figure~\ref{fig:simrank}(d) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
@@
 
-\section{Discussion}
-%% Loading bibliography style file
+\section{Discussion and Concluding Remarks} \label{sec:disc}
+This paper reconsiders the role of CA in the analysis of two-mode network data.
+We began with an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to simple CA of the affiliation matrix.
+Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}.
+I showed that just like the reflective centralities that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums.
+This exercise clarifies the links between CA and the dual two-mode eigenvector centrality in a more coherent and systematic way, an issue that was left open and somewhat unclear in Bonacich's (1991) classic paper.
+Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.
+
+But CA does more than reveal a latent dual ordering of nodes in the two-mode network.
+In addition to doing this, CA can be shown to reveal \textit{groupings} of nodes based on some conception of the similarity of their connections to nodes in the other mode.
+This is evident once we use the scores of the first CA dimension to re-order the rows and columns of the original affiliation matrix.
+When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data.
+This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events.
+
+This is another way in which CA and the Bonacich dual centrality approach systematically differ; one is geared to uncovering blocks of actors with similar connections to events in the other mode (and vice versa), while the other reveals blocks of actors who are the most active and who attend the largest events.
+Both are, of course, legitimate ways of analyzing the structure of a two-mode network.
+Still, they differ in terms of the structural patterns that they are sensitive to, with the CA approach closer to a community partition where nodes that are surprisingly similar end up in the same clusters---where ``surprising'' means similarity based on deviations from a suitable null model based on random mixing given their first-order degrees.
+
+But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to?
+Here we argued that CA recovers a type of generalized relational similarity (GRS).
+That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar.
+Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, we saw that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach.
+Thus, we can clarify that two-mode network CA reveals latent groupings of nodes based on a GRS criterion.
+
+Overall, the preceding has shown that CA can be upgraded from a method designed to generate joint plots and visualization of two-mode data to one that can be seen as more ``central'' to the usual social-network-analytic tasks, like ordering the nodes in the two sets according to some substantively meaningful rank---centrality analysis---or finding sets of similar actors in the network (subgroup or community detection).
+At the very least, it seems like the Bonacich style dual centrality based on the eigenvector decomposition of the raw affiliation matrix should not be the default ``reflective'' centrality approach unless the analyst has the explicit analytic goal of exploring center-periphery partitioning in the network.
+
+A better approach, similar to the one exemplified here, would be to present a comparison of the reflective centralities obtained by CA and Bonacich side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering.
+Of course, suppose the analyst is more interested in such a ``subgroup'' analytic approach.
+In that case, the Bonacich dual centrality approach is less helpful (even if multiple dimensions are considered), and CA should be the first line of attack.
+
+%% Loading bibliography style file]
 %\bibliographystyle{model1-num-names}

### 2024-07-10 duality-presentation.tex
updated!

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usetheme[]{simple}
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
\usepackage{caption}
% Make caption package work with longtable
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\@ifpackageloaded{sidenotes}{}{\usepackage{sidenotes}}
\@ifpackageloaded{marginnote}{}{\usepackage{marginnote}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Coasting on Duality},
  pdfauthor={Omar Lizardo},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Coasting on Duality}
\subtitle{What We Get When We Do Two-Mode Network CA}
\author{Omar Lizardo}
\date{5/22/24}
\institute{\emph{University of California, Los Angeles}}

\begin{document}
\frame{\titlepage}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, interior hidden]}{\end{tcolorbox}}\fi

\begin{frame}{CA and Two Mode Networks}
\protect\hypertarget{ca-and-two-mode-networks}{}
\begin{itemize}
\tightlist
\item
  While still not very common, Correspondence Analysis (CA) has been
  sporadically used in \emph{analyzing} two-mode network data*
\item
  But criticized for being limited and inapplicable for most ``SNA''
  tasks**

  \begin{itemize}
  \tightlist
  \item
    Computing (Centrality-Like) Measures of Node Position
  \item
    Computing Measures of Group Membership
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*Breiger 2000; Roberts 2000; Faust 2005; D'Esposito et al.~2014

**Bonacich 1991; Borgatti \& Everett 1997

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Two Mode Networks}
\protect\hypertarget{ca-and-two-mode-networks-1}{}
\begin{itemize}
\tightlist
\item
  While still not very common, Correspondence Analysis (CA) has been
  sporadically used in \emph{analyzing} two-mode network data*
\item
  But criticized for being limited and inapplicable for most ``SNA''
  tasks**

  \begin{itemize}
  \tightlist
  \item
    Computing (Centrality-Like) Measures of Node Position
  \item
    Computing Measures of Group Membership
  \end{itemize}
\item
  CA Proponents point to some important advantages of CA for the
  analysis of two-mode networks:

  \begin{itemize}
  \tightlist
  \item
    CA can provide a ``joint display'' of the structural connectivity
    patterns in two-mode data
  \item
    (M)CA can be used to ``visually explore'' two-mode networks
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*Breiger 2000; Roberts 2000; Faust 2005; D'Esposito et al.~2014

**Bonacich 1991; Borgatti \& Everett 1997

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Two Mode Networks}
\protect\hypertarget{ca-and-two-mode-networks-2}{}
\begin{itemize}
\tightlist
\item
  While still not very common, Correspondence Analysis (CA) has been
  sporadically used in \emph{analyzing} two-mode network data*
\item
  But criticized for being limited and inapplicable for most ``SNA''
  tasks**

  \begin{itemize}
  \tightlist
  \item
    Computing (Centrality-Like) Measures of Node Position
  \item
    Computing Measures of Group Membership
  \end{itemize}
\item
  This presentation builds on and expands on previous research on using
  CA for two-mode data analysis but discusses aspects of CA that have
  not been emphasized in previous research:

  \begin{itemize}
  \tightlist
  \item
    CA as computing a type of ``(Reflective) Dual Centrality'' on the
    entities in the two-modes*
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*Rediscovered recently under the heading of the Economic Complexity
Index (ECI; Hidalgo and Hausman 2009)

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Two Mode Networks}
\protect\hypertarget{ca-and-two-mode-networks-3}{}
\begin{itemize}
\tightlist
\item
  While still not very common, Correspondence Analysis (CA) has been
  sporadically used in \emph{analyzing} two-mode network data*
\item
  But criticized for being limited and inapplicable for most ``SNA''
  tasks**

  \begin{itemize}
  \tightlist
  \item
    Computing (Centrality-Like) Measures of Node Position
  \item
    Computing Measures of Group Membership
  \end{itemize}
\item
  This presentation builds on and expands on previous research on using
  CA for two-mode data analysis but discusses aspects of CA that have
  not been emphasized in previous research:

  \begin{itemize}
  \tightlist
  \item
    CA as extracting ``positional'' information the entities in the two
    modes

    \begin{itemize}
    \tightlist
    \item
      With positions defined by ``Generalized Relational Similarity''
    \item
      Similarly central entities have similar patterns of connectivity
      with entities that are themselves similar**
    \end{itemize}
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

**Jeh \& Widom 2002; Kovacs 2010; Lizardo 2024

\end{footnotesize}}
\end{frame}

\begin{frame}{Southern Women Data}
\protect\hypertarget{southern-women-data}{}
table output

","

"{]},col:{[}2,"

","

"{]},tr:{[}2,"

","

"{]},td:{[}3,"

","

E1

E2

E3

E4

E5

E6

E7

E8

E9

E10

E11

E12

E13

E14

BRENDA

1

0

1

1

1

1

1

1

0

0

0

0

0

0

CHARLOTTE

0

0

1

1

1

0

1

0

0

0

0

0

0

0

DOROTHY

0

0

0

0

0

0

0

1

1

0

0

0

0

0

ELEANOR

0

0

0

1

1

1

1

1

0

0

0

0

0

0

EVELYN

1

1

1

1

1

1

0

1

1

0

0

0

0

0

FLORA

0

0

0

0

0

0

0

0

1

0

1

0

0

0

FRANCES

0

0

1

1

1

1

0

1

0

0

0

0

0

0

HELEN

0

0

0

0

0

0

1

1

0

1

1

1

0

0

KATHERINE

0

0

0

0

0

0

0

1

1

1

0

1

1

1

LAURA

1

1

1

0

1

1

1

1

0

0

0

0

0

0

MYRNA

0

0

0

0

0

0

0

1

1

1

0

1

0

0

NORA

0

0

0

0

0

0

1

0

1

1

1

1

1

1

OLIVIA

0

0

0

0

0

0

0

0

1

0

1

0

0

0

PEARL

0

0

0

0

0

1

0

1

1

0

0

0

0

0

RUTH

0

0

0

1

1

0

1

1

1

0

0

0

0

0

SYLVIA

0

0

0

0

0

0

1

1

1

1

0

1

1

1

THERESA

0

1

1

1

1

1

1

1

1

0

0

0

0

0

VERNE

0

0

0

0

0

0

1

1

1

0

0

1

0

0
\end{frame}

\begin{frame}{What do we get when we do CA?}
\protect\hypertarget{what-do-we-get-when-we-do-ca}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  Taken singly, the first dimension of CA suggests an \textbf{ordering}

  \begin{itemize}
  \tightlist
  \item
    Is this connected to centrality?
  \item
    If so, what is the relation (if any) to existing two-mode
    centralities?*
  \end{itemize}
\item
  Take jointly, the first two dimensions of CA suggest a
  \textbf{grouping}

  \begin{itemize}
  \tightlist
  \item
    Connected to group cohesion?**
  \item
    Affinity of actors for particular events?***
  \end{itemize}
\end{itemize}
\end{column}

\marginnote{\begin{footnotesize}

*Bonacich 1991

**Doreian 1979

***Faust 2005

\end{footnotesize}}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-sw.png}

}

\caption{Correspondence Plot of Southern Women Data Obtained from Simple
CA of the Affiliation Matrix}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{What do we get when we do CA?}
\protect\hypertarget{what-do-we-get-when-we-do-ca-1}{}
\begin{itemize}
\tightlist
\item
  To (try to) shed light on these questions I examine the cases of three
  relatively recent ``accidental reinventions'' of CA in the ``network
  science'' and ``data science'' (computer and information science)
  literatures:

  \begin{itemize}
  \tightlist
  \item
    The ``Economic Complexity Index'' (ECI)*

    \begin{itemize}
    \tightlist
    \item
      Relevant to the question of centrality
    \end{itemize}
  \item
    \emph{SimRank} and Generalized Relational Similarity**

    \begin{itemize}
    \tightlist
    \item
      Relevant to the question of grouping
    \end{itemize}
  \end{itemize}
\item
  The ECI case has been examined in detail in recent work***
\item
  The second and third are new to this presentation

  \begin{itemize}
  \tightlist
  \item
    All shed light on what exactly we get when we use CA to analyze two
    mode data
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*Hidalgo \& Hausmann 2009

**Jeh \& Widom 2002; Kovacs (2010)

***van Dam et al.~2021; Mealy 2019

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Reflective Centrality}
\protect\hypertarget{ca-and-reflective-centrality}{}
\begin{itemize}
\tightlist
\item
  Hidalgo \& Hausmann (2009) introduced a ``novel'' way to rank nodes in
  two-mode networks

  \begin{itemize}
  \tightlist
  \item
    Their approach, called ``reflective,'' was originally applied to
    country-by-product networks*
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  The method has since been shown to be applicable to other two-mode
  networks (like cultural networks)**
\item
  I introduce here using the more familiar sociological example of the
  \textbf{duality of persons and groups}***
\end{itemize}

\marginnote{\begin{footnotesize}

*Referred to as the Economic Complexity Index (ECI)

**Lizardo 2018

***Breiger 1974

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Reflective Centrality}
\protect\hypertarget{ca-and-reflective-centrality-1}{}
\begin{itemize}
\tightlist
\item
  Notation:

  \begin{itemize}
  \tightlist
  \item
    A two-mode network composed of a set of persons \(P\)
  \item
    Their affiliation relations to a set of groups \(G\)
  \item
    Represented by an affiliation matrix \(\mathbf{A}\)
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  \(\mathbf{A}\) is of dimensions \(|P| \times |G|\) with people along
  the rows and groups across the columns

  \begin{itemize}
  \tightlist
  \item
    \(|P|\) is the cardinality of the persons set\\
  \item
    \(|G|\) is the cardinality of the group set
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  \(\mathbf{A}\) has cell entries \(a_{pg}= 1\) if person \textit{p} is
  affiliated with group \textit{g} and \(a_{pg}= 0\) otherwise.
\end{itemize}
\end{frame}

\begin{frame}{CA and Reflective Centrality}
\protect\hypertarget{ca-and-reflective-centrality-2}{}
\begin{itemize}
\item
  If we are going to compute the centrality of nodes in a two-mode
  network, the most natural place to start is with the good old
  \textbf{degree centrality}*
\item
  Given this, the degree-centrality of people is given by:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-dp}{}{
C^R_p(1) = \sum_g a_{pg}
}\label{eq-dp}\end{equation}

\begin{itemize}
\tightlist
\item
  And for groups:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-dg}{}{
   C^R_g(1) = \sum_p a_{pg}
}\label{eq-dg}\end{equation}

\marginnote{\begin{footnotesize}

*Faust 1997, eqs. 4 and 5

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Reflective Centrality}
\protect\hypertarget{ca-and-reflective-centrality-3}{}
\begin{itemize}
\item
  Once we have the first-order quantities \(C^R_p(1)\) and \(C^R_g(1)\),
  we can compute ``second-order centralities'' \(C^R(2)\) for both
  people and groups
\item
  For people, these are given by:*
\end{itemize}

\begin{equation}\protect\hypertarget{eq-r2p}{}{
C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
}\label{eq-r2p}\end{equation}

\begin{itemize}
\tightlist
\item
  And for groups:**
\end{itemize}

\begin{equation}\protect\hypertarget{eq-r2g}{}{   
C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
}\label{eq-r2g}\end{equation}

\marginnote{\begin{footnotesize}

*People are central when the average sum of the size of the groups they
belong to is large

**Groups are central when the average activity of the people who belong
to them is large

\end{footnotesize}}

\note{Equation Equation~\ref{eq-r2p} says ``people are more central when
the average sum of the size of the groups they belong to is large'\,'
(e.g., whenever \(a_{pg} = 1\) and \(C^R_g(1)\) is a big number).
Equation Equation~\ref{eq-r2g} says''groups are more central when the
average activity of their members is large'' (e.g., whenever
\(a_{pg} = 1\) and \(C^R_p(1)\) is a big number).}
\end{frame}

\begin{frame}{CA and Reflective Centrality}
\protect\hypertarget{ca-and-reflective-centrality-4}{}
\begin{itemize}
\item
  Of course, once we have the second-order quantities \(C^R_p(2)\) and
  \(C^R_g(2)\), we can compute ``third-order centralities'' \(C^R(3)\)
  for both people and groups
\item
  For people, these are given by:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-r3p}{}{
   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
}\label{eq-r3p}\end{equation}

\begin{itemize}
\tightlist
\item
  And for groups:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-r3g}{}{   
   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
}\label{eq-r3g}\end{equation}

\marginnote{\begin{footnotesize}

*People are central when the average sum of the average activity of the
members of the groups they belong to is large

**Groups are central when the average sum of the average size of the
groups their members belong to is large

\end{footnotesize}}

\note{Equation Equation~\ref{eq-r3p} says something like ``people are
more central when the average sum of the average activity of the members
of the groups they belong to is large'' (e.g., \(a_{pg} = 1\) and
\(C^R_g(2)\) is a big number). Equation Equation~\ref{eq-r3g} says,
``groups are more central when the average sum of the average size of
the groups their members belong to is large.'' (e.g., \(a_{pg} = 1\) and
\(C^R_p(2)\) is a big number).}
\end{frame}

\begin{frame}{CA and Reflective Centrality}
\protect\hypertarget{ca-and-reflective-centrality-5}{}
\begin{itemize}
\item
  More generally, Hidalgo \& Hausmann show that we can define a
  \textbf{series} of reflective quantities*
  \(C^R_p(4), C^R_p(5)...C^R_p(q)\) and
  \(C^R_g(4), C^R_g(5)...C^R_g(q)\)
\item
  For people, these are given by:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-rip}{}{   
C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
}\label{eq-rip}\end{equation}

\begin{itemize}
\tightlist
\item
  And for groups:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-rig}{}{   
C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
}\label{eq-rig}\end{equation}

\marginnote{\begin{footnotesize}

*Whose verbal and substantive interpretation becomes more elusive as the
number of iterations increases

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Reflective Centrality}
\protect\hypertarget{ca-and-reflective-centrality-6}{}
\begin{itemize}
\tightlist
\item
  For people:

  \begin{itemize}
  \tightlist
  \item
    Even-numbered reflections measure centrality based on the
    \textbf{average size} of the groups a person belongs to*
  \item
    Odd-numbered reflections measure centrality based the
    \textbf{average number of memberships} of each member of the groups
    they belong to*
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  For groups:

  \begin{itemize}
  \tightlist
  \item
    Even-numbered reflections measure centrality based on the
    \textbf{average number of memberships} of the members of the group*
  \item
    Odd-numbered reflections measure centrality based on the
    \textbf{average size} of the group memberships of each member*
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*and the average of the average of the average of the average\ldots{}

\end{footnotesize}}
\end{frame}

\begin{frame}{Reflection Trajectories for SWD}
\protect\hypertarget{reflection-trajectories-for-swd}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/p-reflections-sw.png}

}

\caption{Person Reflection Trajectories}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/g-reflections-sw.png}

}

\caption{Group Reflection Trajectories}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Final Reflection Ordering versus CA}
\protect\hypertarget{final-reflection-ordering-versus-ca}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/p-reflections-sw.png}

}

\caption{Person Reflection Trajectories}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-sw.png}

}

\caption{Correspondence Plot of Southern Women Data}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Final Reflection Ordering versus CA}
\protect\hypertarget{final-reflection-ordering-versus-ca-1}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/g-reflections-sw.png}

}

\caption{Person Reflection Trajectories}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-sw.png}

}

\caption{Correspondence Plot of Southern Women Data}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Final Reflection Ordering versus CA}
\protect\hypertarget{final-reflection-ordering-versus-ca-2}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/p-ca-ref-corr-sw.png}

}

\caption{Scatter of standardized person reflection centralities and
person scores on first CA dimension (r = 1.0)}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-sw.png}

}

\caption{Correspondence Plot of Southern Women Data}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Final Reflection Ordering versus CA}
\protect\hypertarget{final-reflection-ordering-versus-ca-3}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/g-ca-ref-corr-sw.png}

}

\caption{Scatter of standardized group reflection centralities and group
scores on first CA dimension (r = 1.0)}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-sw.png}

}

\caption{Correspondence Plot of Southern Women Data}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Method of Reflections versus CA}
\protect\hypertarget{method-of-reflections-versus-ca}{}
\begin{itemize}
\item
  So the method of reflections is an accidental re-invention of CA!*
\item
  But \emph{why} is it a reinvention of CA?

  \begin{itemize}
  \tightlist
  \item
    Short answer:

    \begin{itemize}
    \tightlist
    \item
      The final reflection scores are the (second largest) eigenvalues
      of a matrix, just like CA!
    \end{itemize}
  \end{itemize}
\item
  What's the (slightly) longer answer?

  \begin{itemize}
  \tightlist
  \item
    A couple of things first:

    \begin{itemize}
    \tightlist
    \item
      \(Dp\) is a \(|P| \times |P|\) square degree matrix containing
      each person's \(C_p^R(1)\) (activity) along the diagonals and
      zeroes in every other cell
    \item
      \(Dg\) is a \(|G| \times |G|\) square degree matrix containing
      each group's \(C_g^R(1)\) (size) along the diagonals and zeroes in
      every other cell
    \end{itemize}
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*On the first dimension: Mealy 2019; van Dam et al.~2021

\end{footnotesize}}
\end{frame}

\begin{frame}{Method of Reflections versus CA}
\protect\hypertarget{method-of-reflections-versus-ca-1}{}
\begin{itemize}
\item
  So the method of reflections is an accidental re-invention of CA!*
\item
  But \emph{why} is it a reinvention of CA?

  \begin{itemize}
  \tightlist
  \item
    Short answer:

    \begin{itemize}
    \tightlist
    \item
      The final reflection scores are the (second largest) eigenvalues
      of a matrix, just like CA!
    \end{itemize}
  \end{itemize}
\item
  What's the (slightly) longer answer?

  \begin{itemize}
  \tightlist
  \item
    In the limit (\(q \rightarrow \infty\)), the iterative
    Hidalgo-Hausmann reflective centralities are the solution of the
    eigensystem:*
  \end{itemize}
\end{itemize}

\begin{equation}\protect\hypertarget{eq-dam1}{}{
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
}\label{eq-dam1}\end{equation}

\begin{equation}\protect\hypertarget{eq-dam2}{}{
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
}\label{eq-dam2}\end{equation}

\marginnote{\begin{footnotesize}

*van Dam et al.~2021

\end{footnotesize}}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich}{}
\begin{itemize}
\tightlist
\item
  Do the matrices whose eigenvalues yield the equilibrium reflective
  centralities look familiar to you?
\end{itemize}

\[
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
\]

\[
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
\]

\begin{itemize}
\tightlist
\item
  They should, because the solutions to the following eigensystem give
  the Bonacich \((C^B)\) dual centralities for two-mode networks:*
\end{itemize}

\begin{equation}\protect\hypertarget{eq-bon1}{}{
\left(AA^T\right)C^B_p = \lambda C^B_p
}\label{eq-bon1}\end{equation}

\begin{equation}\protect\hypertarget{eq-bon2}{}{
\left(A^TA\right)C^B_g = \lambda C^B_g
}\label{eq-bon2}\end{equation}

\marginnote{\begin{footnotesize}

*Bonacich 1991

\end{footnotesize}}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-1}{}
\begin{itemize}
\tightlist
\item
  So both Reflective and Bonacich centralities are eigenvectors of the
  person and group \textbf{projections} of the two-mode network
\end{itemize}

\[
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
\\
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
\\
\left(AA^T\right)C^B_p = \lambda C^B_p
\\
\left(A^TA\right)C^B_g = \lambda C^B_g
\]

\begin{itemize}
\tightlist
\item
  But for reflective centralities the affiliation matrix and its
  transpose are pre-multiplied by the \textbf{inverse of the degree
  matrix of each mode}*
\end{itemize}

\marginnote{\begin{footnotesize}

\begin{itemize}
\tightlist
\item
  van Dam et al.~2021
\end{itemize}

\end{footnotesize}}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-2}{}
\begin{itemize}
\tightlist
\item
  So both Reflective and Bonacich centralities are eigenvectors of the
  person and group \textbf{projections} of the two-mode network
\end{itemize}

\[
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
\\
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
\\
\left(AA^T\right)C^B_p = \lambda C^B_p
\\
\left(A^TA\right)C^B_g = \lambda C^B_g
\]

\begin{itemize}
\tightlist
\item
  Similar to how each entry in the original affiliation matrix is
  weighted in CA performing the SVD!*
\end{itemize}

\[
Dp^{-1/2}ADg^{-1/2} = X_p \lambda X_g^T
\]

\marginnote{\begin{footnotesize}

**Faust 2005, eq. 7.8

\end{footnotesize}}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-3}{}
\begin{itemize}
\tightlist
\item
  We can gain some insight into the equivalence between CA and
  Reflections by expressing the degree-weighted projections in terms of
  each cell entry:*
\end{itemize}

\begin{equation}\protect\hypertarget{eq-mealy1}{}{
a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
}\label{eq-mealy1}\end{equation}

\begin{itemize}
\item
  In Equation~\ref{eq-mealy1}, the fraction's numerator is equal to one
  only when person \(p\) and \(p'\) are both members of \(g\).

  \begin{itemize}
  \tightlist
  \item
    Summed across groups, this is the number of \textbf{shared groups}
    between \(p\) and \(p'\):

    \begin{itemize}
    \tightlist
    \item
      The cell entries in the Bonacich matrix (raw projections)
    \end{itemize}
  \item
    But in Equation~\ref{eq-mealy1} the numerator is \emph{weighted} by
    the product of group size and each person's number of memberships
    (in contrast to Bonacich)
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*Mealy 2019, eq. 4

\end{footnotesize}}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-4}{}
\begin{itemize}
\tightlist
\item
  We can gain some insight into the equivalence between CA and
  Reflections by expressing the degree-weighted projections in terms of
  each cell entry:*
\end{itemize}

\begin{equation}\protect\hypertarget{eq-mealy2}{}{
    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
}\label{eq-mealy2}\end{equation}

\begin{itemize}
\item
  In Equation~\ref{eq-mealy2}, the fraction's numerator is equal to one
  only when group \(g\) and \(g'\) both have \(p\) as a member

  \begin{itemize}
  \tightlist
  \item
    Summed across persons, the number of \textbf{shared members} between
    \(g\) and \(g'\):

    \begin{itemize}
    \tightlist
    \item
      The cell entries in the Bonacich matrix (raw projections)
    \end{itemize}
  \item
    But in Equation~\ref{eq-mealy2} the numerator is \emph{weighted} by
    the product of group size and each person's number of memberships
    (in contrast to Bonacich)
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*Mealy 2019, eq. 4

\end{footnotesize}}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-5}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/bon-reord.png}

}

\caption{Bonacich Re-ordered Affiliation Matrix}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-6}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  The CA re-ordered affiliation matrix is \textbf{block-diagonal}
\item
  Reveals (a well-known and storied*) grouping of persons and groups
  into two clusters \includegraphics{Plots/doreian-sw.png}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-7}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  First CA dimension is also an \textbf{ordering} so it counts as a
  centrality (a rank over the nodes)
\item
  It is an ordering that uses a logic of grouping in order to rank

  \begin{itemize}
  \tightlist
  \item
    Tasks usually segregated in sna
  \end{itemize}
\item
  Central events contain the most central people and central people
  attend the most central events:

  \begin{itemize}
  \tightlist
  \item
    \(E1 = \{Laura, Brenda, Evelyn\}\)
  \end{itemize}
\item
  In CA ordering central events are \emph{not} the ones with most
  members!

  \begin{itemize}
  \tightlist
  \item
    \(\{E7, E8, E9\}\)
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-8}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/bon-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  First CA dimension is also an \textbf{ordering} so it counts as a
  centrality (a rank over the nodes)
\item
  It is an ordering that uses a logic of grouping in order to rank

  \begin{itemize}
  \tightlist
  \item
    Tasks usually segregated in sna
  \end{itemize}
\item
  Central events contain the most central people and central people
  attend the most central events:

  \begin{itemize}
  \tightlist
  \item
    \(E1 = \{Laura, Brenda, Evelyn\}\)
  \end{itemize}
\item
  But they are the most central in the Bonacich ordering!

  \begin{itemize}
  \tightlist
  \item
    \(\{E7, E8, E9\}\)
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-9}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  First CA dimension is also an \textbf{ordering} so it counts as a
  centrality (a rank over the nodes)
\item
  It is an ordering that uses a logic of grouping in order to rank

  \begin{itemize}
  \tightlist
  \item
    Tasks usually segregated in sna
  \end{itemize}
\item
  Least central events contain many of the least central people:

  \begin{itemize}
  \tightlist
  \item
    \(E11 = \{Helen, Flora, Olivia, \\ Nora\}\)
  \end{itemize}
\item
  In CA ordering less central events are \emph{not} the ones with the
  least members!

  \begin{itemize}
  \tightlist
  \item
    \(\{E1, E13, E14\}\)
  \item
    \(E1\) is the most central!
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-10}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  First CA dimension is also an \textbf{ordering} so it counts as a
  centrality (a rank over the nodes)
\item
  It is an ordering that uses a logic of grouping in order to rank

  \begin{itemize}
  \tightlist
  \item
    Tasks usually segregated in sna
  \end{itemize}
\item
  Central people attend the most central events and central events are
  attended by central people:

  \begin{itemize}
  \tightlist
  \item
    \(Frances = \{E3, E5, E4, E6, E8\}\)
  \end{itemize}
\item
  Not necessarily those with the most memberships

  \begin{itemize}
  \tightlist
  \item
    \(\{Theresa, Evelyn, Brenda\}\)
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-11}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  First CA dimension is also an \textbf{ordering} so it counts as a
  centrality (a rank over the nodes)
\item
  It is an ordering that uses a logic of grouping in order to rank

  \begin{itemize}
  \tightlist
  \item
    Tasks usually segregated in sna
  \end{itemize}
\item
  The least central person \(\{Nora\}\) is not the one with the least
  memberships

  \begin{itemize}
  \tightlist
  \item
    Instead \(\{Nora\}\) shares many memberships with less central
    people (those who attend the less central events)
  \item
    \(\{Flora, Olivia, Sylvia, Kat.\}\)
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-12}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  First CA dimension is also an \textbf{ordering} so it counts as a
  centrality (a rank over the nodes)
\item
  It is an ordering that uses a logic of grouping in order to rank

  \begin{itemize}
  \tightlist
  \item
    Tasks usually segregated in sna
  \end{itemize}
\item
  Central people attend the most central events and central events are
  attended by central people:

  \begin{itemize}
  \tightlist
  \item
    \(E1 = \{Laura, Brenda, Evelyn\}\)
  \end{itemize}
\item
  In CA ordering central events are \emph{not} the ones with the most
  members!

  \begin{itemize}
  \tightlist
  \item
    e.g., \(\{E7, E8, E9\}\)
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-13}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/bon-reord.png}

}

\caption{Bonacich Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  The Bonacich re-ordered affiliation matrix is \textbf{upper
  triangular}
\item
  Arranges persons and groups from top to bottom and left to right by
  activity and size respectively

  \begin{itemize}
  \tightlist
  \item
    More central people and groups at the upper right
  \item
    Less central people and groups at the lower left
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-14}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/bon-reord.png}

}

\caption{Bonacich Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  Bonacich centrality is dual but it is ``simple'' dual

  \begin{itemize}
  \tightlist
  \item
    Central people attend well-attended events
  \item
    Central events are attended by people who attend many events
  \end{itemize}
\item
  Centrality in one mode is a weighted \emph{sum} of the other mode's
  centrality*
\end{itemize}

\begin{equation}\protect\hypertarget{eq-fau1}{}{
    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
}\label{eq-fau1}\end{equation}

\begin{equation}\protect\hypertarget{eq-fau2}{}{
    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
}\label{eq-fau2}\end{equation}
\end{column}
\end{columns}

\marginnote{\begin{footnotesize}

*Faust 1997, p.~170

\end{footnotesize}}
\end{frame}

\begin{frame}{CA versus Bonacich}
\protect\hypertarget{ca-versus-bonacich-15}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-reord.png}

}

\caption{CA Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{itemize}
\tightlist
\item
  CA centrality is dual but it is ``generalized similarity'' dual*

  \begin{itemize}
  \tightlist
  \item
    Central people attend central (not necessarily the most
    well-attended) events
  \item
    Central events are attended by central (not necessarily the most
    active) people
  \end{itemize}
\item
  Some people calculate Bonacich centralities when they have this more
  ``generalized similarity'' idea of centrality in mind!

  \begin{itemize}
  \tightlist
  \item
    They should use CA instead
  \end{itemize}
\end{itemize}
\end{column}
\end{columns}

\marginnote{\begin{footnotesize}

*Kovacs 2010; Lizardo 2024

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Similarity}
\protect\hypertarget{ca-and-similarity}{}
\begin{itemize}
\item
  Another way to gain insight into what CA is doing is by looking at the
  similarity of people and groups in the data
\item
  As we all know,* the usual one mode projections are obtained via:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-brei1}{}{
S_p = AA^T
}\label{eq-brei1}\end{equation}

\begin{equation}\protect\hypertarget{eq-brei1}{}{
S_g = A^TA
}\label{eq-brei1}\end{equation}

\begin{itemize}
\tightlist
\item
  Which can be thought of as \textbf{similarity matrices} based on the
  shared neighbors in the two mode network
\end{itemize}

\marginnote{\begin{footnotesize}

Breiger 1974

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Similarity}
\protect\hypertarget{ca-and-similarity-1}{}
\begin{itemize}
\item
  Another way to gain insight into what CA is doing is by looking at the
  similarity of people and groups in the data
\item
  But we can also obtain \emph{other-mode degree weighted} versions of
  these similarity matrices*
\end{itemize}

\begin{equation}\protect\hypertarget{eq-dam5}{}{
S^*_p = ADg^{-1}A^T
}\label{eq-dam5}\end{equation}

\begin{equation}\protect\hypertarget{eq-dam6}{}{
S^*_g = A^TDp^{-1}A
}\label{eq-dam6}\end{equation}

\marginnote{\begin{footnotesize}

van Dam et al.~2021

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Similarity}
\protect\hypertarget{ca-and-similarity-2}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-sim-p.png}

}

\caption{Degree-weighted similarity matrix (persons)}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-eigvec-p.png}

}

\caption{First eigenvalue of person similarity matrix against rank}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and Similarity}
\protect\hypertarget{ca-and-similarity-3}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-sim-g.png}

}

\caption{Degree-weighted similarity matrix (groups)}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-eigvec-g.png}

}

\caption{First eigenvalue of group similarity matrix against rank}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and Similarity}
\protect\hypertarget{ca-and-similarity-4}{}
\begin{columns}[T]
\begin{column}{0.3\textwidth}
\includegraphics{Plots/ca-sim-p.png}
\includegraphics{Plots/ca-sim-g.png}
\end{column}

\begin{column}{0.55\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/ca-corr-plot.png}

}

\caption{CA Correspondence Plot}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Bonacich and Similarity}
\protect\hypertarget{bonacich-and-similarity}{}
\begin{columns}[T]
\begin{column}{0.3\textwidth}
\includegraphics{Plots/bon-sim-p.png}
\includegraphics{Plots/bon-sim-g.png}
\end{column}

\begin{column}{0.55\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/bon-corr-plot.png}

}

\caption{Bonacich Correspondence Plot}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and Generalized Relational Similarity}
\protect\hypertarget{ca-and-generalized-relational-similarity}{}
\begin{itemize}
\tightlist
\item
  Bonacich orders people according to attendance at central (most-well
  attended) events and events according to attendance by central (most
  active) people
\item
  CA orders people according to attendance at \textbf{similarly central
  events} and events according to attendance by \textbf{similarly
  central people}

  \begin{itemize}
  \tightlist
  \item
    This is precisely the definition of \textbf{generalized relational
    similarity} (GSR)*
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  In that case proximity of nodes in the first CA dimension may reveal a
  type of GSR in two-mode networks

  \begin{itemize}
  \tightlist
  \item
    Let us see how that works in the case of two candidate two-mode GSRs

    \begin{itemize}
    \tightlist
    \item
      \emph{SimRank} *
    \item
      Similarity Weighted Correlation Distance**
    \end{itemize}
  \end{itemize}
\end{itemize}

\marginnote{\begin{footnotesize}

*Jeh \& Widom 2002

**Kovacs 2010; Lizardo 2024

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and SimRank}
\protect\hypertarget{ca-and-simrank}{}
\begin{itemize}
\tightlist
\item
  To explore the link between CA and GSR we can compare CA with one
  independently developed approach for obtaining GSRs in two-mode
  networks:*

  \begin{itemize}
  \tightlist
  \item
    \emph{SimRank}
  \end{itemize}
\item
  According to \emph{SimRank}:

  \begin{itemize}
  \tightlist
  \item
    People are \emph{similar} if they belong to \emph{similar} groups
  \item
    Groups are \emph{similar} if they share \emph{similar} members
  \end{itemize}
\item
  In \emph{SimRank} the similarity between two people is given by:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-simrank1}{}{
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
}\label{eq-simrank1}\end{equation}

\marginnote{\begin{footnotesize}

*Jeh \& Widom 2002

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and SimRank}
\protect\hypertarget{ca-and-simrank-1}{}
\begin{itemize}
\tightlist
\item
  To explore the link between CA and GSR we can compare CA with one
  independently developed approach for obtaining GSRs in two-mode
  networks:*

  \begin{itemize}
  \tightlist
  \item
    \emph{SimRank}
  \end{itemize}
\item
  According to \emph{SimRank}:

  \begin{itemize}
  \tightlist
  \item
    People are \emph{similar} if they belong to \emph{similar} groups
  \item
    Groups are \emph{similar} if they share \emph{similar} members
  \end{itemize}
\item
  In \emph{SimRank} the similarity between two groups is given by:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-simrank2}{}{
S(g, g') = \frac{\alpha}{C_g^R(1)C_{g'}^R(1)}
    \sum_{i = 1}^{C_g^R(1)} \sum_{j = 1}^{C_{g'}^R(1)} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
}\label{eq-simrank2}\end{equation}

\marginnote{\begin{footnotesize}

*Jeh \& Widom 2002

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and SimRank}
\protect\hypertarget{ca-and-simrank-2}{}
\begin{itemize}
\tightlist
\item
  \emph{SimRank} for people:
\end{itemize}

\[
\small
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
\]

\begin{itemize}
\tightlist
\item
  \emph{SimRank} for groups:
\end{itemize}

\[
\small
S(g, g') = \frac{\alpha}{C_g^R(1)C_{g'}^R(1)}
    \sum_{i = 1}^{C_g^R(1)} \sum_{j = 1}^{C_{g'}^R(1)} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
\]

With:*

\begin{itemize}
\tightlist
\item
  \(S(p, p) = 1\) and \(S(g, g) = 1\) and \(0 > \alpha < 1\)
\end{itemize}

\marginnote{\begin{footnotesize}

*Jeh \& Widom 2002

\end{footnotesize}}

\note{In two-mode networks, SimRank scores can be estimated via a simple
algorithm, in which we first estimate \(S(p, p')\) in equation
Equation~\ref{eq-simrank1} using baseline values. Hence, only groups
that two people share contribute to the initial values of \(S(p, p')\)
since only \(S(g, g)>0\) at the outset. We then plug those values into
equation Equation~\ref{eq-simrank2}, then loop back to equation
Equation~\ref{eq-simrank1} with the resulting \(S(g, g')\) values, and
continue iterating until convergence---generally achieved after five
iterations}
\end{frame}

\begin{frame}{CA and SimRank}
\protect\hypertarget{ca-and-simrank-3}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/sr-plot-reord.png}

}

\caption{SimRank Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/sr-plot-eigen-p.png}

}

\caption{Simrank Similarity Matrix Eigenvalue Plot (Persons)}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and SimRank}
\protect\hypertarget{ca-and-simrank-4}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/sr-plot-reord.png}

}

\caption{SimRank Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/sr-plot-eigen-g.png}

}

\caption{Simrank Similarity Matrix Eigenvalue Plot (Groups)}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and SimRank}
\protect\hypertarget{ca-and-simrank-5}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/sr-plot-reord.png}

}

\caption{SimRank Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.3\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/sr-corr-scatter.png}

}

\caption{Correlation between CA and SimRank Similarity Matrix
Eigenvectors}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and Generalized Relational Similarity}
\protect\hypertarget{ca-and-generalized-relational-similarity-1}{}
\begin{itemize}
\item
  I recently proposed a measure of generalized relational similarity
  (GSR) metric for nodes in two-mode networks by ``tweaking'' a previous
  approach based on the modified correlation distance*
\item
  For pairs of people the GSR is given by:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-kovacs1}{}{
\begin{equation}
\scriptsize
    S(p, p') = 
    \frac{
    (A_{p\bullet} - \bar{A}_{p\bullet})
    \mathbf{S(G)}
    (A_{p'\bullet} - \bar{A}_{p'\bullet})^T
    }
    {
    \sqrt{
    (A_{p\bullet} - \bar{A}_{p\bullet})
    \mathbf{S(G)}
    (A_{p\bullet} - \bar{A}_{p\bullet})^T
    }
    \sqrt{
    (A_{p'\bullet} - \bar{A}_{p'\bullet})
    \mathbf{S(G)}
    (A_{p'\bullet} - \bar{A}_{p'\bullet})^T
        }
    }
\end{equation}
}\label{eq-kovacs1}\end{equation}

\begin{itemize}
\tightlist
\item
  And for groups:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-kovacs2}{}{
\begin{equation}
\scriptsize 
    S(g, g') = 
    \frac{
    (A_{g\bullet} - \bar{A}_{g\bullet})
    \mathbf{S(P)}
    (A_{g'\bullet} - \bar{A}_{g'\bullet})^T
    }
    {
    \sqrt{
    (A_{g\bullet} - \bar{A}_{g\bullet})
    \mathbf{S(P)}
    (A_{g\bullet} - \bar{A}_{g\bullet})^T
    }
    \sqrt{
    (A_{g'\bullet} - \bar{A}_{g'\bullet})
    \mathbf{S(P)}
    (A_{g'\bullet} - \bar{A}_{g'\bullet})^T
        }
    }
\end{equation}
}\label{eq-kovacs2}\end{equation}

\marginnote{\begin{footnotesize}

*Lizardo 2024; Kovacs 2010

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Generalized Relational Similarity}
\protect\hypertarget{ca-and-generalized-relational-similarity-2}{}
\begin{itemize}
\tightlist
\item
  With the base similarities between people given by:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-lizardo1}{}{
\mathbf{S(P)}_{pp'} = \frac{AA_{pp'}}{\sqrt{AA_{pp}AA_{p'p'}}}
}\label{eq-lizardo1}\end{equation}

\begin{itemize}
\tightlist
\item
  And the base similarities for groups:
\end{itemize}

\begin{equation}\protect\hypertarget{eq-lizardo2}{}{
\mathbf{S(G)}_{gg'} = \frac{AA_{gg'}}{\sqrt{AA_{gg}AA_{g'g'}}}
}\label{eq-lizardo2}\end{equation}

\begin{itemize}
\tightlist
\item
  Namely, the cosine distances between people and groups based on the
  corresponding cell entries of the one-mode projections*
\end{itemize}

\marginnote{\begin{footnotesize}

*Lizardo 2024

\end{footnotesize}}
\end{frame}

\begin{frame}{CA and Generalized Relational Similarity}
\protect\hypertarget{ca-and-generalized-relational-similarity-3}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/grs-plot-reord.png}

}

\caption{GRS Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/grs-plot-eigen-p.png}

}

\caption{GRS Similarity Matrix Eigenvalue Plot (Persons)}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and Generalized Relational Similarity}
\protect\hypertarget{ca-and-generalized-relational-similarity-4}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/grs-plot-reord.png}

}

\caption{GRS Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/grs-plot-eigen-g.png}

}

\caption{GRS Similarity Matrix Eigenvalue Plot (Groups)}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{CA and Generalized Relational Similarity}
\protect\hypertarget{ca-and-generalized-relational-similarity-5}{}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/grs-plot-reord.png}

}

\caption{GRS Re-ordered Affiliation Matrix}

\end{figure}
\end{column}

\begin{column}{0.3\textwidth}
\begin{figure}

{\centering \includegraphics{Plots/grs-corr-scatter.png}

}

\caption{Correlation between CA and GRS Matrix Eigenvectors}

\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Conclusions}
\protect\hypertarget{conclusions}{}
\begin{itemize}
\tightlist
\item
  To paraphrase John Martin paraphrasing Levi-Strauss:

  \begin{itemize}
  \tightlist
  \item
    There is only \emph{one analysis} in two-mode network analysis and
    it is CA
  \item
    Rediscovered many times by researchers pursuing seemingly
    incompatible taks
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  Contra skeptics we can do two-mode ``sna'' when we do CA

  \begin{itemize}
  \tightlist
  \item
    Rank nodes against a meaningful centrality metric
  \item
    Detect equivalences (generalized similarities) between node clusters
  \end{itemize}
\item
  These are some of the standard ``positional'' tasks that take up much
  sna
\end{itemize}
\end{frame}

\begin{frame}{Conclusions}
\protect\hypertarget{conclusions-1}{}
\begin{itemize}
\tightlist
\item
  To paraphrase John Martin paraphrasing Levi-Strauss:

  \begin{itemize}
  \tightlist
  \item
    There is only \emph{one analysis} in two-mode network analysis and
    it is CA
  \item
    Rediscovered many times by researchers pursuing seemingly
    incompatible taks
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  We can also go do more than visualization and joint display
\item
  The first CA dimension particularly represents a seldom noted
  \textbf{duality of ranking and equivalence}

  \begin{itemize}
  \tightlist
  \item
    Ranking via similarity within clusters
  \item
    Bonacich by contrast is a ``pure'' centrality ranking
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Conclusions}
\protect\hypertarget{conclusions-2}{}
\begin{itemize}
\tightlist
\item
  To paraphrase John Martin paraphrasing Levi-Strauss:

  \begin{itemize}
  \tightlist
  \item
    There is only \emph{one analysis} in two-mode network analysis and
    it is CA
  \item
    Rediscovered many times by researchers pursuing seemingly
    incompatible taks
  \end{itemize}
\end{itemize}

\pause

\begin{itemize}
\tightlist
\item
  Using CA for this kind of ``spectral clustering'' and ``spectral
  ranking'' is not common but feasible

  \begin{itemize}
  \tightlist
  \item
    Another approach (similar to blockmodeling):

    \begin{itemize}
    \tightlist
    \item
      Use CA for spectral clustering
    \item
      Then use \textbf{specific CA} (on sub-tables defined by the
      eigengap heuristic) to split clusters into further clusters
    \end{itemize}
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Thanks!}
\protect\hypertarget{thanks}{}
\url{https://github.com/olizardo/correspondence-analysis-two-mode-networks}

olizardo@soc.ucla.edu
\end{frame}

\end{document}

### 2024-07-10 highlights.tex
do it!

--- 2023-09-08 main.tex
@@
-%% 
-%% Copyright 2019-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'CAS Bundle'.
-%% --------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'CAS Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for cas-dc documentclass for 
-%% double column output.
+\documentclass{article}
+\begin{document}
 
-%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
-\documentclass[a4paper,fleqn]{cas-sc}
-
-%\usepackage[numbers]{natbib}
-%\usepackage[authoryear]{natbib}
-\usepackage[authoryear,longnamesfirst]{natbib}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage{longtable}
-
-%%%Author definitions
-\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
-\tsc{WGM}
-\tsc{QE}
-\tsc{EP}
-\tsc{PMS}
-\tsc{BEC}
-\tsc{DE}
-%%%
-
-% Uncomment and use as if needed
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdefinition{rmk}{Remark}
-%\newproof{pf}{Proof}
-%\newproof{pot}{Proof of Theorem \ref{thm}}
-
-\begin{document}
-\let\WriteBookmarks\relax
-\def\floatpagepagefraction{1}
-\def\textpagefraction{.001}
-
-% Short title
-\shorttitle{Correspondence Analysis of Two-Mode Networks}
-
-% Short author
-%\shortauthors{O. Lizardo}
-
-% Main title of the paper
-\title [mode = title]{The Correspondence Analysis of Two-Mode Networks: Centrality and Positional Interpretations}                      
-% Title footnote mark
-% eg: \tnotemark[1]
-%\tnotemark[1,2]
-
-% Title footnote 1.
-% eg: \tnotetext[1]{Title footnote text}
-% \tnotetext[<tnote number>]{<tnote text>} 
-%\tnotetext[1]{}
-
-%\tnotetext[2]{.}
-
-% First author
-%
-% Options: Use if required
-% eg: \author[1,3]{Author Name}[type=editor,
-%       style=chinese,
-%       auid=000,
-%       bioid=1,
-%       prefix=Sir,
-       %orcid=0000-0002-5405-3007,
-%       facebook=<facebook id>,
-%       twitter=<twitter id>,
-%       linkedin=<linkedin id>,
-%       gplus=<gplus id>]
-\author[1]{Omar Lizardo}[orcid=0000-0001-7511-2910]
-
-% Corresponding author indication
-\cormark[1]
-
-% Footnote of the first author
-\fnmark[1]
-
-% Email id of the first author
-\ead{olizardo@soc.ucla.edu}
-
-% URL of the first author
-\ead[url]{http://olizardo.bol.ucla.edu/}
-
-%  Credit authorship
-\credit{Conceptualization of this study, Methodology, Software}
-
-%Address/affiliation
-\affiliation[1]{organization={University of California, Los Angeles},
-    addressline={264 Haines Hall}, 
-    city={Los Angeles},
-    %citysep={},  Uncomment if no comma needed between city and postcode
-     postcode={90095}, 
-     state={CA},
-     country={USA}}
-
- %Corresponding author text
-\cortext[cor1]{Corresponding author}
-%\cortext[cor2]{Principal corresponding author}
-
-%Footnote text
-\fntext[fn1]{This paper was written while the author was partially supported by National Science Foundation grant HNDS-R \#2214215.}
- % author as well.}
-%\fntext[fn2]{Another author footnote, this is a very long footnote and
- % it should be a really long footnote. But this footnote is not yet
- % sufficiently long enough to make two lines of footnote text.}
-
-% For a title note without a number/mark
-%\nonumnote{This note has no numbers. In this work we demonstrate $a_b$
-%  the formation Y\_1 of a new type of polariton on the interface
- % between a cuprous oxide slab and a polystyrene micro-sphere placed
- % on the slab.
- % }
-% Here goes the abstract
-\begin{abstract}
-Lorem...
-\end{abstract}
-
-% Use if graphical abstract is present
-% \begin{graphicalabstract}
-% \includegraphics{figs/grabs.pdf}
-% \end{graphicalabstract}
-
-%Research highlights
-\begin{highlights}
-    \item 
-    \item 
-    \item 
-\end{highlights}
-
-% Keywords
-% Each keyword is separated by \sep
-\begin{keywords}
-Generalized similarity \sep Duality \sep Two-mode networks \sep Correspondence Analysis
-\end{keywords}
-
-\maketitle
-\newpage
-\section{Introduction}
-While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
-Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}.
-This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Nevertheless, it departs from the already-mentioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been as emphasized or treated in detail before, moving beyond the focus on data visualization.
-
-Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
-With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
-In social network analysis, this applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.
-
-Additionally, I argue that CA can also be used to extract latent \textit{positional} information on people and groups.
-First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
-As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups).
-Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode.
-Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}---e.g., the discovery of sociometrically similar subsets of entities in affiliation networks.
-
-\subsection{Organization of the Paper}
-The rest of the paper is organized as follows.
-In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
-This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes.
-This approach is computationally and mathematically related but \textit{not} equivalent to the matrix factorization of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores.
-I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers instead the core-periphery structure of the two-mode network.
-In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
-
-\section{Reflective Centralities in Two-Mode Networks}
-In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
-Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).
-Subsequent shows that there is no logical connection between the formal method and this particular application.
-As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
-
-If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}.
-A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $G$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.
-
-Given this, the degree-centrality of people is given by:
-
-\begin{equation}
-    C^R_p(1) = \sum_g a_{pg}
-    \label{eq:R1_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(1) = \sum_p a_{pg}
-  \label{eq:R1_g}
-\end{equation}
-
-That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
-In what follows, I will generally refer to $C^R_p(1)$ as (individual) ``activity'' and $C^R_p(1)$ as (group) ``size.''
-
-As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.
-
-For people, these are given by:
-
-\begin{equation}
-    C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
-    \label{eq:R2_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
-    \label{eq:R2_g}
-\end{equation}
-
-Equation \ref{eq:R2_p} says ``people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
-Equation \ref{eq:R2_g} says ``groups are more central when the average activity of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
-Of course, we can keep on going and define third-order reflections.
-
-For the people, these are given by:
-
-\begin{equation}
-   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
-   \label{eq:R3_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
-   \label{eq:R3_g}
-\end{equation}
-
-Equation \ref{eq:R3_p} says something like ``people are more central when the average sum of the average activity of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
-Equation \ref{eq:R3_g} says, ''groups are more central when the average sum of the average size of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
-
-Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes).
-On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to).
-In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity).
-On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size).
-While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).
-
-More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases).
-
-For people, these are given by:
-
-\begin{equation}   
-    C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
-   \label{eq:Rq_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
-   \label{eq:Rq_g}
-\end{equation}
-
-Equation~\ref{eq:Rq_p} says that the reflective centrality of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
-Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
-
-\begin{table}
-\caption{Southern Women Data.}
-\label{tab:southern}
-\begin{tabular}[]{lcccccccccccccc}
-\hline
- & E1 & E2 & E3 & E4 & E5 & E6 & E7 & E8 & E9 & E10 & E11 & E12 & E13 & E14 \\
- \hline
-EVELYN & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-LAURA & 1 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-THERESA & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-BRENDA & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-CHARLOTTE & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-FRANCES & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-ELEANOR & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-RUTH & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-VERNE & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 1 & 0 & 0 \\
-MYRA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 \\
-KATHERINE & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-SYLVIA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-NORA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 \\
-HELEN & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 \\
-OLIVIA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-FLORA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-PEARL & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-DOROTHY & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-\hline
-\end{tabular}
-\textsuperscript{Note: Rows ordered according to the generalized
-blockmodeling solution of Doreian et al. (2004, Table 4).} \\
-\end{table}
-
-\subsection{Empirical Example}
-Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}.
-The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
-
-As the bottom panel of the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to.
-However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
-Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al.
-(2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
-
-The bottom panel of Figure~\ref{fig:pg-reflections} shows the corresponding bump chart for the events.
-Just like for these actors, the equilibrium reflective centralities recover the ordering of the columns according to Doreian et al.
-generalized blockmodel with events 1-6 separated from events 10-15 by events 7-9 (see also \citet{kovacs2010generalized}).
-Figure~\ref{fig:pg-reflections} shows that $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
-Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering the average of the average number of memberships of its members (and so forth).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
-    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
-    \label{fig:pg-reflections}
-\end{figure}
-\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks}
-The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
-In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings.
-Instead, Bonacich noted that, in analyzing the Southern Women data shown in Table~\ref{tab:southern}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
-
-As we will see, Bonacich was only half right.
-Indeed, the CA scores capture a version of ``clique'' (more accurately, community) membership, but they---particularly the first dimension---also capture a version of centrality in the most general sense of ranking nodes based on a meaningful criterion \citep{borgatti2006graph}.
-In fact, they capture the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
-
-To see this, recall that the solutions to the following eigensystem give the Bonacich $(C^B)$ two-mode centralities:
-
-\begin{equation}
-    AC^B_p = \lambda C^B_g 
-    \label{eq:bon1}
-\end{equation}
-
-\begin{equation}
-    A^TC^B_g = \lambda C^B_p 
-    \label{eq:bon2}
-\end{equation}
-
-Where $A$ is the original affiliation matrix.
-The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, as given by the singular value decomposition (SVD) of the rectangular affiliation matrix.
-Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
-
-For people, these are given by:
-
-\begin{equation}
-    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
-    \label{eq:faust1}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
-    \label{eq:faust2}
-\end{equation}
-
-Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
-These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
-Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
-The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.
-
-As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue (matrix factorization) problem.
-To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
-Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:
-
-\begin{equation}
-    \left(AA^T\right)C^B_p = \lambda^2 C^B_p
-    \label{eq:bon3}
-\end{equation}
-
-\begin{equation}
-    \left(A^TA\right)C^B_g = \lambda^2 C^B_g
-    \label{eq:bon4}
-\end{equation}
-
-This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.
-
-Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
-In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as the solution of the eigensystem:
-
-\begin{equation}
-    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
-    \label{eq:dam1}
-\end{equation}
-
-\begin{equation}
-    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
-    \label{eq:dam2}
-\end{equation}
-
-Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups.
-The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before factorizing, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
-
-This can be clearly seen if we express $\left(Dg^{-1}A^TDp^{-1}A\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
-
-\begin{equation}
-    a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
-    \label{eq:mealy1}
-\end{equation}
-
-In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$.
-Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
-As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
-The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
-
-The same reasoning applies to groups in~\ref{eq:dam2}, whose entries are given by:
-
-\begin{equation}
-    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
-    \label{eq:mealy2}
-\end{equation}
-
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by inverse of the square root of the product of the row and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using}, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
-
-Degree-pre-weighting does not radically alter the nature of the data.
-Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perenial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
-This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large.
-On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
-
-Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
-In the Bonacich case, a factorization of the one-mode projection of the two-mode network; in the reflective case, factorization of the degree-weighted one-mode projection of the same network.
-Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
-Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to.
-The (average) centrality of groups, in turn, is a function of their members' (average) centrality.
-The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{Plots/ca-v-bon-reord.png}
-    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score.}
-    \label{fig:ca-v-bon}
-\end{figure}
-
-\section{Illustrative Analysis of the Southern Women Data}
-\subsection{Row and Column Re-ordered Affiliation Matrices}
-Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
-In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
-As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
-The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side).
-The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
-
-\subsection{Eigenvector Plot}
-We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
-If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
-The Southern Women data clearly belong to the second category.
-In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other.
-Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
-Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach.
-This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}
-    \caption{Caption}
-    \label{fig:ca-eigvec}
-\end{figure}
-
-\subsection{Normalized Similarity Plot}
-To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one.
-Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
-As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
-This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition.
-These are shown as inset plots in Figure~\ref{fig:ca-eigvec} for both people and groups.
-Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.0\textwidth]{Plots/ca-sim.png}
-    \caption{Caption}
-    \label{fig:ca-sim}
-\end{figure}
-
-To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
-In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity.
-Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups.
-As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.
-
-\subsection{CA Correspondence Plot}
-The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions.
-People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices.
-The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups.
-People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm).
-The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart.
-The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.
-
-Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups).
-This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot.
-But this (still common) interpretation is off the mark.
-What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs.
-2.13 and 2.14]{fouss2016algorithms}.
-As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
-In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.0\textwidth]{Plots/bon-sim.png}
-    \caption{Caption}
-    \label{fig:bon-sim}
-\end{figure}
-
-\subsection{One-Mode Projection Matrix and Eigevenctor Plot}
-This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
-But as we have seen, this is \textit{not} what CA is designed to do.
-Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes.
-Notably, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two eigenvectors of the usual one-mode projections of the matrix (where the first is, of course, the standard Bonacich dual centrality score).
-
-To illustrate, Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by other-mode degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score.
-Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
-Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}  We can see that the plot of the first two eigenvectors does a good job of recovering the raw connectivity structure of the Southern Women affiliation network, partitioning the core persons and groups (on the upper-right) from the more peripheral ones (on the lower-left).
-
-Moreover, if all we are interested in is capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then Figure~\ref{fig:bon-sim}(c) does a better job of that than the usual CA correspondence plot in Figure~\ref{fig:ca-sim}(c).
-For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine, Sylvia\right]$ for $\left[E10, E12\right]$.
-In the same way, core events like $\left[E8, E9, E10\right]$---shown on the lower half of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the upper half.
-The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.
-
-This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's eigenvector centrality.
-For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot.
-Similarly, the main difference between $\left[Ruth\right]$ and $\left[Sylvia\right]$, despite their comparable $C^B$ scores, is that the former's (closer to the upper half of the plot) event profile is composed mainly of core events.
-In contrast, peripheral events dominate the latter's attendance profile (shown in the bottom half of the plot), accordingly, $\left[Ruth\right]$ and $\left[Sylvia\right]$ are assigned to distinct clusters by the k-means algorithm.
-
-\subsection{Correspondence Analysis and Generalized Similarities}
-As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.
-Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}.
-This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence.
-In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.
-
-An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}.
-In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong, and the similarity of groups is a function of the people who belong to them, making the similarity of people and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
-Thus,
+\section*{Highglights}
 
 \begin{itemize}
-    \item People are \textit{similar} if they belong to \textit{similar} groups.
-    \item Groups are \textit{similar} if they share \textit{similar} members.
+  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
+  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
+  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
 \end{itemize}
 
-Which is consistent with a GRS approach. To do this, \citet[540, eq2 and eq3]{jeh2002simrank} propose that we define the similarity of people as given by:
-
-\begin{equation}
-    S(p, p') = \frac{\alpha}{C^R(1)_pC^R(1)_{p'}}
-    \sum_{i = 1}^{C^R(1)_p} \sum_{j = 1}^{C^R(1)_{p'}} 
-    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
-    \label{eq:simrank1}
-\end{equation}
-
-Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to; $\alpha$ is a free parameter, with the restriction: $0 > \alpha < 1$.
-By construction, $S(p, p) = 1$.
-Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of the groups they belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.
-
-Likewise, for groups, the SimRank similarities are given by:
-
-\begin{equation}
-    S(g, g') = \frac{\alpha}{C^R(1)_gC^R(1)_{g'}}
-    \sum_{i = 1}^{C^R(1)_g} \sum_{j = 1}^{C^R(1)_{g'}} 
-    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
-    \label{eq:simrank2}
-\end{equation}
-
-Where $p(i)_{i \in N(g)}$ is the $i^{th}$ person in the set of members of group $g$, and $p(j)_{j \in N(g')}$ is the $j^{th}$ person in the set of members of group $g'$, and $S(g, g) = 1$.
-Thus, the SimRank similarity between two groups is a function of the sum of the similarities of the people who belong to them, weighted by the reciprocal of the products of their number of members multiplied by $\alpha$.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{Plots/sr-plot.png}
-    \caption{Caption}
-    \label{fig:simrank}
-\end{figure}
-
-In two-mode networks, SimRank scores can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
-Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
-We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}.
-Further Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with the HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
-The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
-Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA.
-
-And indeed, they are not.
-Figure~\ref{fig:simrank}(a) and Figure~\ref{fig:simrank}(b) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the first non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices of both persons and groups.
-As we can see, the partitioning of the node sets on the first informative eigenvector is substantively the same, separating the same two blocks of people and events.
-Figure~\ref{fig:simrank}(c) shows the original affiliation matrix, this time with rows and columns re-arranged according to the main eigenvector of the SimRank similarity matrix.
-Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two blocks of persons and groups.
-Figure~\ref{fig:simrank}(d) shows the correspondence plot of the first and second informative eigenvector of the SimRank similarity matrices for persons and groups.
-Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c).
-
-\section{Discussion}
-%% Loading bibliography style file
-%\bibliographystyle{model1-num-names}
-\bibliographystyle{cas-model2-names}
-% Loading bibliography database
-\bibliography{ca.bib}
 \end{document}
-
-The Elsevier cas-dc class is based on the
-standard article class and supports almost all of the functionality of
-that class. In addition, it features commands and options to format the
-\begin{itemize} \item document style \item baselineskip \item front
-matter \item keywords and MSC codes \item theorems, definitions and
-proofs \item labels of enumerations \item citation style and labeling.
-\end{itemize}
-
-This class depends on the following packages
-for its proper functioning:
-
-\begin{enumerate}
-\itemsep=0pt
-\item {natbib.sty} for citation processing;
-\item {geometry.sty} for margin settings;
-\item {fleqn.clo} for left aligned equations;
-\item {graphicx.sty} for graphics inclusion;
-\item {hyperref.sty} optional packages if hyperlinking is
-  required in the document;
-\end{enumerate}
-
-All the above packages are part of any
-standard \LaTeX{} installation.
-Therefore, the users need not be
-bothered about downloading any extra packages.
-
-\section{Installation}
-
-The package is available at author resources page at Elsevier
-(\url{http://www.elsevier.com/locate/latex}).
-The class may be moved or copied to a place, usually,\linebreak
-\verb+$TEXMF/tex/latex/elsevier/+, %$%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-or a folder which will be read                   
-by \LaTeX{} during document compilation.  The \TeX{} file
-database needs updation after moving/copying class file.  Usually,
-we use commands like \verb+mktexlsr+ or \verb+texhash+ depending
-upon the distribution and operating system.
-
-\section{Front matter}
-
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. 
-You are recommended to conform your choice to the journal you 
-are submitting to.
-
-\section{Bibliography styles}
-
-There are various bibliography styles available. You can select the
-style of your choice in the preamble of this document. These styles are
-Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use Bib\TeX\ to generate your bibliography and include DOIs
-whenever available.
-
-Here are two sample references: 
-\cite{Fortunato2010}
-\cite{Fortunato2010,NewmanGirvan2004}
-\cite{Fortunato2010,Vehlowetal2013}
-
-\section{Floats}
-{Figures} may be included using the command,\linebreak 
-\verb+\includegraphics+ in
-combination with or without its several options to further control
-graphic. \verb+\includegraphics+ is provided by {graphic[s,x].sty}
-which is part of any standard \LaTeX{} distribution.
-{graphicx.sty} is loaded by default. \LaTeX{} accepts figures in
-the postscript format while pdf\LaTeX{} accepts {*.pdf},
-{*.mps} (metapost), {*.jpg} and {*.png} formats. 
-pdf\LaTeX{} does not accept graphic files in the postscript format.
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=.75]{figs/Fig1.pdf}
-	\caption{The evanescent light - $1S$ quadrupole coupling
-	($g_{1,l}$) scaled to the bulk exciton-photon coupling
-	($g_{1,2}$). The size parameter $kr_{0}$ is denoted as $x$ and
-	the \PMS is placed directly on the cuprous oxide sample ($\delta
-	r=0$, See also Table \protect\ref{tbl1}).}
-	\label{FIG:1}
-\end{figure}
-
-The \verb+table+ environment is handy for marking up tabular
-material. If users want to use {multirow.sty},
-{array.sty}, etc., to fine control/enhance the tables, they
-are welcome to load any package of their choice and
-{cas-dc.cls} will work in combination with all loaded
-packages.
-
-\begin{table}[width=.9\linewidth,cols=4,pos=h]
-\caption{This is a test caption. This is a test caption. This is a test
-caption. This is a test caption.}\label{tbl1}
-\begin{tabular*}{\tblwidth}{@{} LLLL@{} }
-\toprule
-Col 1 & Col 2 & Col 3 & Col4\\
-\midrule
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-\bottomrule
-\end{tabular*}
-\end{table}
-
-\section[Theorem and ...]{Theorem and theorem like environments}
-
-{cas-dc.cls} provides a few shortcuts to format theorems and
-theorem-like environments with ease. In all commands the options that
-are used with the \verb+\newtheorem+ command will work exactly in the same
-manner. {cas-dc.cls} provides three commands to format theorem or
-theorem-like environments:
-
-\begin{verbatim}
- \newtheorem{theorem}{Theorem}
- \newtheorem{lemma}[theorem]{Lemma}
- \newdefinition{rmk}{Remark}
- \newproof{pf}{Proof}
- \newproof{pot}{Proof of Theorem \ref{thm2}}
-\end{verbatim}
-
-The \verb+\newtheorem+ command formats a
-theorem in \LaTeX's default style with italicized font, bold font
-for theorem heading and theorem number at the right hand side of the
-theorem heading.  It also optionally accepts an argument which
-will be printed as an extra heading in parentheses.
-
-\begin{verbatim}
-  \begin{theorem} 
-   For system (8), consensus can be achieved with 
-   $\|T_{\omega z}$ ...
-     \begin{eqnarray}\label{10}
-     ....
-     \end{eqnarray}
-  \end{theorem}
-\end{verbatim}
-
-\newtheorem{theorem}{Theorem}
-
-\begin{theorem}
-For system (8), consensus can be achieved with 
-$\|T_{\omega z}$ ...
-\begin{eqnarray}\label{10}
-....
-\end{eqnarray}
-\end{theorem}
-
-The \verb+\newdefinition+ command is the same in
-all respects as its \verb+\newtheorem+ counterpart except that
-the font shape is roman instead of italic.  Both
-\verb+\newdefinition+ and \verb+\newtheorem+ commands
-automatically define counters for the environments defined.
-
-The \verb+\newproof+ command defines proof environments with
-upright font shape.  No counters are defined.
-
-\section[Enumerated ...]{Enumerated and Itemized Lists}
-{cas-dc.cls} provides an extended list processing macros
-which makes the usage a bit more user friendly than the default
-\LaTeX{} list macros.   With an optional argument to the
-\verb+\begin{enumerate}+ command, you can change the list counter
-type and its attributes.
-
-\begin{verbatim}
- \begin{enumerate}[1.]
- \item The enumerate environment starts with an optional
-   argument `1.', so that the item counter will be suffixed
-   by a period.
- \item You can use `a)' for alphabetical counter and '(i)' 
-  for roman counter.
-  \begin{enumerate}[a)]
-    \item Another level of list with alphabetical counter.
-    \item One more item before we start another.
-    \item One more item before we start another.
-    \item One more item before we start another.
-    \item One more item before we start another.
-\end{verbatim}
-
-Further, the enhanced list environment allows one to prefix a
-string like `step' to all the item numbers.
-
-\begin{verbatim}
- \begin{enumerate}[Step 1.]
-  \item This is the first step of the example list.
-  \item Obviously this is the second step.
-  \item The final step to wind up this example.
- \end{enumerate}
-\end{verbatim}
-
-\section{Cross-references}
-In electronic publications, articles may be internally
-hyperlinked. Hyperlinks are generated from proper
-cross-references in the article.  For example, the words
-\textcolor{black!80}{Fig.~1} will never be more than simple text,
-whereas the proper cross-reference \verb+\ref{tiger}+ may be
-turned into a hyperlink to the figure itself:
-\textcolor{blue}{Fig.~1}.  In the same way,
-the words \textcolor{blue}{Ref.~[1]} will fail to turn into a
-hyperlink; the proper cross-reference is \verb+\cite{Knuth96}+.
-Cross-referencing is possible in \LaTeX{} for sections,
-subsections, formulae, figures, tables, and literature
-references.
-
-\section{Bibliography}
-
-Two bibliographic style files (\verb+*.bst+) are provided ---
-{model1-num-names.bst} and {model2-names.bst} --- the first one can be
-used for the numbered scheme. This can also be used for the numbered
-with new options of {natbib.sty}. The second one is for the author year
-scheme. When  you use model2-names.bst, the citation commands will be
-like \verb+\citep+,  \verb+\citet+, \verb+\citealt+ etc. However when
-you use model1-num-names.bst, you may use only \verb+\cite+ command.
-
-\verb+thebibliography+ environment.  Each reference is a\linebreak
-\verb+\bibitem+ and each \verb+\bibitem+ is identified by a label,
-by which it can be cited in the text:
-
-\noindent In connection with cross-referencing and
-possible future hyperlinking it is not a good idea to collect
-more that one literature item in one \verb+\bibitem+.  The
-so-called Harvard or author-year style of referencing is enabled
-by the \LaTeX{} package {natbib}. With this package the
-literature can be cited as follows:
-
-\begin{enumerate}[\textbullet]
-\item Parenthetical: \verb+\citep{WB96}+ produces (Wettig \& Brown, 1996).
-\item Textual: \verb+\citet{ESG96}+ produces Elson et al. (1996).
-\item An affix and part of a reference:\break
-\verb+\citep[e.g.][Ch. 2]{Gea97}+ produces (e.g. Governato et
-al., 1997, Ch. 2).
-\end{enumerate}
-
-In the numbered scheme of citation, \verb+\cite{<label>}+ is used,
-since \verb+\citep+ or \verb+\citet+ has no relevance in the numbered
-scheme.  {natbib} package is loaded by {cas-dc} with
-\verb+numbers+ as default option.  You can change this to author-year
-or harvard scheme by adding option \verb+authoryear+ in the class
-loading command.  If you want to use more options of the {natbib}
-package, you can do so with the \verb+\biboptions+ command.  For
-details of various options of the {natbib} package, please take a
-look at the {natbib} documentation, which is part of any standard
-\LaTeX{} installation.
-
-\appendix
-\section{My Appendix}
-Appendix sections are coded under \verb+\appendix+.
-
-\verb+\printcredits+ command is used after appendix sections to list 
-author credit taxonomy contribution roles tagged using \verb+\credit+ 
-in frontmatter.
-
-\printcredits
-
-%\vskip3pt
-
-\bio{}
-Author biography without author photo.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-\endbio
-
-\bio{figs/pic1}
-Author biography with author photo.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-\endbio
-
-\bio{figs/pic1}
-Author biography with author photo.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-\endbio

### 2024-07-10 main.tex
the whole shebang

@@
 %       gplus=<gplus id>]
-\author[1]{Omar Lizardo}[orcid=0000-0002-5405-3007]
+%\author[1]{Omar Lizardo}[orcid=0000-0002-5405-3007]
 
 % Corresponding author indication
-\cormark[1]
+%\cormark[1]
 
 % Footnote of the first author
-\fnmark[1]
+%\fnmark[1]
 
 % Email id of the first author
-\ead{olizardo@soc.ucla.edu}
+%\ead{olizardo@soc.ucla.edu}
 
 % URL of the first author
-\ead[url]{http://olizardo.bol.ucla.edu/}
+%\ead[url]{http://olizardo.bol.ucla.edu/}
 
 %  Credit authorship
-\credit{Conceptualization of this study, Methodology, Software}
+%\credit{Conceptualization of this study, Methodology, Software}
 
 %Address/affiliation
-\affiliation[1]{organization={University of California, Los Angeles},
-    addressline={264 Haines Hall}, 
-    city={Los Angeles},
+%\affiliation[1]{organization={University of California, Los Angeles},
+   % addressline={264 Haines Hall}, 
+   % city={Los Angeles},
     %citysep={},  Uncomment if no comma needed between city and postcode
-     postcode={90095}, 
-     state={CA},
-     country={USA}}
+    % postcode={90095}, 
+    % state={CA},
+    % country={USA}}
 
  %Corresponding author text
-\cortext[cor1]{Corresponding author}
+%\cortext[cor1]{Corresponding author}
 %\cortext[cor2]{Principal corresponding author}
@@
 %Footnote text
-\fntext[fn1]{This paper was written while the author was partially supported by National Science Foundation grant HNDS-R \#2214215.}
+%\fntext[fn1]{This paper was written while the author was partially supported by National Science Foundation grant HNDS-R \#2214215.}
  % author as well.}

### 2024-07-10 title-page.tex
the whole shebang

--- 2023-09-08 main.tex
@@
 % Short author
-%\shortauthors{O. Lizardo}
+\shortauthors{O. Lizardo}
 
 % Main title of the paper
-\title [mode = title]{The Correspondence Analysis of Two-Mode Networks: Centrality and Positional Interpretations}                      
+\title [mode = title]{The Correspondence Analysis of Two-Mode Networks Revisited}                      
 % Title footnote mark
@@
     city={Los Angeles},
-    %citysep={},  Uncomment if no comma needed between city and postcode
+    citysep={},  Uncomment if no comma needed between city and postcode
      postcode={90095}, 
@@
 %Research highlights
-\begin{highlights}
-    \item 
-    \item 
-    \item 
-\end{highlights}
 
@@
 \maketitle
-\newpage
-\section{Introduction}
-While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
-Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}.
-This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Nevertheless, it departs from the already-mentioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been as emphasized or treated in detail before, moving beyond the focus on data visualization.
 
-Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
-With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
-In social network analysis, this applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.
-
-Additionally, I argue that CA can also be used to extract latent \textit{positional} information on people and groups.
-First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
-As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups).
-Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode.
-Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}---e.g., the discovery of sociometrically similar subsets of entities in affiliation networks.
-
-\subsection{Organization of the Paper}
-The rest of the paper is organized as follows.
-In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
-This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes.
-This approach is computationally and mathematically related but \textit{not} equivalent to the matrix factorization of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores.
-I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers instead the core-periphery structure of the two-mode network.
-In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
-
-\section{Reflective Centralities in Two-Mode Networks}
-In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
-Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).
-Subsequent shows that there is no logical connection between the formal method and this particular application.
-As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
-
-If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}.
-A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $G$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.
-
-Given this, the degree-centrality of people is given by:
-
-\begin{equation}
-    C^R_p(1) = \sum_g a_{pg}
-    \label{eq:R1_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(1) = \sum_p a_{pg}
-  \label{eq:R1_g}
-\end{equation}
-
-That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
-In what follows, I will generally refer to $C^R_p(1)$ as (individual) ``activity'' and $C^R_p(1)$ as (group) ``size.''
-
-As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.
-
-For people, these are given by:
-
-\begin{equation}
-    C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
-    \label{eq:R2_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
-    \label{eq:R2_g}
-\end{equation}
-
-Equation \ref{eq:R2_p} says ``people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
-Equation \ref{eq:R2_g} says ``groups are more central when the average activity of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
-Of course, we can keep on going and define third-order reflections.
-
-For the people, these are given by:
-
-\begin{equation}
-   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
-   \label{eq:R3_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
-   \label{eq:R3_g}
-\end{equation}
-
-Equation \ref{eq:R3_p} says something like ``people are more central when the average sum of the average activity of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
-Equation \ref{eq:R3_g} says, ''groups are more central when the average sum of the average size of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
-
-Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes).
-On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to).
-In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity).
-On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size).
-While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).
-
-More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases).
-
-For people, these are given by:
-
-\begin{equation}   
-    C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
-   \label{eq:Rq_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
-   \label{eq:Rq_g}
-\end{equation}
-
-Equation~\ref{eq:Rq_p} says that the reflective centrality of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
-Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
-
-\begin{table}
-\caption{Southern Women Data.}
-\label{tab:southern}
-\begin{tabular}[]{lcccccccccccccc}
-\hline
- & E1 & E2 & E3 & E4 & E5 & E6 & E7 & E8 & E9 & E10 & E11 & E12 & E13 & E14 \\
- \hline
-EVELYN & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-LAURA & 1 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-THERESA & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-BRENDA & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-CHARLOTTE & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-FRANCES & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-ELEANOR & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-RUTH & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-VERNE & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 1 & 0 & 0 \\
-MYRA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 \\
-KATHERINE & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-SYLVIA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-NORA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 \\
-HELEN & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 \\
-OLIVIA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-FLORA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-PEARL & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-DOROTHY & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-\hline
-\end{tabular}
-\textsuperscript{Note: Rows ordered according to the generalized
-blockmodeling solution of Doreian et al. (2004, Table 4).} \\
-\end{table}
-
-\subsection{Empirical Example}
-Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}.
-The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
-
-As the bottom panel of the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to.
-However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
-Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al.
-(2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
-
-The bottom panel of Figure~\ref{fig:pg-reflections} shows the corresponding bump chart for the events.
-Just like for these actors, the equilibrium reflective centralities recover the ordering of the columns according to Doreian et al.
-generalized blockmodel with events 1-6 separated from events 10-15 by events 7-9 (see also \citet{kovacs2010generalized}).
-Figure~\ref{fig:pg-reflections} shows that $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
-Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering the average of the average number of memberships of its members (and so forth).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
-    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
-    \label{fig:pg-reflections}
-\end{figure}
-\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks}
-The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
-In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings.
-Instead, Bonacich noted that, in analyzing the Southern Women data shown in Table~\ref{tab:southern}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
-
-As we will see, Bonacich was only half right.
-Indeed, the CA scores capture a version of ``clique'' (more accurately, community) membership, but they---particularly the first dimension---also capture a version of centrality in the most general sense of ranking nodes based on a meaningful criterion \citep{borgatti2006graph}.
-In fact, they capture the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
-
-To see this, recall that the solutions to the following eigensystem give the Bonacich $(C^B)$ two-mode centralities:
-
-\begin{equation}
-    AC^B_p = \lambda C^B_g 
-    \label{eq:bon1}
-\end{equation}
-
-\begin{equation}
-    A^TC^B_g = \lambda C^B_p 
-    \label{eq:bon2}
-\end{equation}
-
-Where $A$ is the original affiliation matrix.
-The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, as given by the singular value decomposition (SVD) of the rectangular affiliation matrix.
-Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
-
-For people, these are given by:
-
-\begin{equation}
-    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
-    \label{eq:faust1}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
-    \label{eq:faust2}
-\end{equation}
-
-Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
-These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
-Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
-The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.
-
-As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue (matrix factorization) problem.
-To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
-Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:
-
-\begin{equation}
-    \left(AA^T\right)C^B_p = \lambda^2 C^B_p
-    \label{eq:bon3}
-\end{equation}
-
-\begin{equation}
-    \left(A^TA\right)C^B_g = \lambda^2 C^B_g
-    \label{eq:bon4}
-\end{equation}
-
-This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.
-
-Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
-In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as the solution of the eigensystem:
-
-\begin{equation}
-    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
-    \label{eq:dam1}
-\end{equation}
-
-\begin{equation}
-    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
-    \label{eq:dam2}
-\end{equation}
-
-Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups.
-The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before factorizing, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
-
-This can be clearly seen if we express $\left(Dg^{-1}A^TDp^{-1}A\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
-
-\begin{equation}
-    a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
-    \label{eq:mealy1}
-\end{equation}
-
-In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$.
-Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
-As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
-The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
-
-The same reasoning applies to groups in~\ref{eq:dam2}, whose entries are given by:
-
-\begin{equation}
-    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
-    \label{eq:mealy2}
-\end{equation}
-
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by inverse of the square root of the product of the row and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using}, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
-
-Degree-pre-weighting does not radically alter the nature of the data.
-Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perenial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
-This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large.
-On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
-
-Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
-In the Bonacich case, a factorization of the one-mode projection of the two-mode network; in the reflective case, factorization of the degree-weighted one-mode projection of the same network.
-Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
-Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to.
-The (average) centrality of groups, in turn, is a function of their members' (average) centrality.
-The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{Plots/ca-v-bon-reord.png}
-    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score.}
-    \label{fig:ca-v-bon}
-\end{figure}
-
-\section{Illustrative Analysis of the Southern Women Data}
-\subsection{Row and Column Re-ordered Affiliation Matrices}
-Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
-In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
-As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
-The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side).
-The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
-
-\subsection{Eigenvector Plot}
-We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
-If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
-The Southern Women data clearly belong to the second category.
-In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other.
-Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
-Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach.
-This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}
-    \caption{Caption}
-    \label{fig:ca-eigvec}
-\end{figure}
-
-\subsection{Normalized Similarity Plot}
-To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one.
-Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
-As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
-This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition.
-These are shown as inset plots in Figure~\ref{fig:ca-eigvec} for both people and groups.
-Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.0\textwidth]{Plots/ca-sim.png}
-    \caption{Caption}
-    \label{fig:ca-sim}
-\end{figure}
-
-To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
-In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity.
-Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups.
-As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.
-
-\subsection{CA Correspondence Plot}
-The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions.
-People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices.
-The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups.
-People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm).
-The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart.
-The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.
-
-Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups).
-This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot.
-But this (still common) interpretation is off the mark.
-What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs.
-2.13 and 2.14]{fouss2016algorithms}.
-As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
-In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.0\textwidth]{Plots/bon-sim.png}
-    \caption{Caption}
-    \label{fig:bon-sim}
-\end{figure}
-
-\subsection{One-Mode Projection Matrix and Eigevenctor Plot}
-This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
-But as we have seen, this is \textit{not} what CA is designed to do.
-Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes.
-Notably, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two eigenvectors of the usual one-mode projections of the matrix (where the first is, of course, the standard Bonacich dual centrality score).
-
-To illustrate, Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by other-mode degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score.
-Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
-Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}  We can see that the plot of the first two eigenvectors does a good job of recovering the raw connectivity structure of the Southern Women affiliation network, partitioning the core persons and groups (on the upper-right) from the more peripheral ones (on the lower-left).
-
-Moreover, if all we are interested in is capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then Figure~\ref{fig:bon-sim}(c) does a better job of that than the usual CA correspondence plot in Figure~\ref{fig:ca-sim}(c).
-For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine, Sylvia\right]$ for $\left[E10, E12\right]$.
-In the same way, core events like $\left[E8, E9, E10\right]$---shown on the lower half of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the upper half.
-The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.
-
-This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's eigenvector centrality.
-For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot.
-Similarly, the main difference between $\left[Ruth\right]$ and $\left[Sylvia\right]$, despite their comparable $C^B$ scores, is that the former's (closer to the upper half of the plot) event profile is composed mainly of core events.
-In contrast, peripheral events dominate the latter's attendance profile (shown in the bottom half of the plot), accordingly, $\left[Ruth\right]$ and $\left[Sylvia\right]$ are assigned to distinct clusters by the k-means algorithm.
-
-\subsection{Correspondence Analysis and Generalized Relational Similarity}
-As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.
-Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}.
-This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence \citep{everett1994regular}.
-In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.
-
-An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}.
-In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong, and the similarity of groups is a function of the people who belong to them, making the similarity of persons and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
-Thus,
-
-\begin{itemize}
-    \item People are \textit{similar} if they belong to \textit{similar} groups.
-    \item Groups are \textit{similar} if they share \textit{similar} members.
-\end{itemize}
-
-Which is consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
-To accomplish this, \citet[540, eq.
-2 and eq.
-3]{jeh2002simrank} propose that we define the similarity of each pair of people $S(p, p')$  as given by:
-
-\begin{equation}
-    S(p, p') = \frac{\alpha}{C^R(1)_pC^R(1)_{p'}}
-    \sum_{i = 1}^{C^R(1)_p} \sum_{j = 1}^{C^R(1)_{p'}} 
-    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
-    \label{eq:simrank1}
-\end{equation}
-
-Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to, and $\alpha$ is a free parameter obeying the restriction: $0 > \alpha < 1$.
-By construction, $S(p, p) = 1$, for all $p$.
-Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of each unordered pair of groups they both belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.
-
-Likewise, for groups, the SimRank similarities are given by:
-
-\begin{equation}
-    S(g, g') = \frac{\alpha}{C^R(1)_gC^R(1)_{g'}}
-    \sum_{i = 1}^{C^R(1)_g} \sum_{j = 1}^{C^R(1)_{g'}} 
-    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
-    \label{eq:simrank2}
-\end{equation}
-
-Where $p(i)_{i \in N(g)}$ is the $i^{th}$ person in the set of members of group $g$, and $p(j)_{j \in N(g')}$ is the $j^{th}$ person in the set of members of group $g'$, and $S(g, g) = 1$.
-Thus, the SimRank similarity between two groups is a function of the sum of the similarities of each pair of people who belong to both groups, weighted by the reciprocal of the products of the number of members of each group multiplied by $\alpha$.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{Plots/sr-plot.png}
-    \caption{Caption}
-    \label{fig:simrank}
-\end{figure}
-
-In two-mode networks, SimRank scores for each pair of nodes across the two sets can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
-Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
-We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}, which is confirmed here for the Southern Women data.
-
-Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with the HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
-The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
-Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections \citep{mealy2019interpreting, van2021correspondence}.
-
-And indeed, they are not.
-Figure~\ref{fig:simrank}(a) shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices for the row and column nodes.
-Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two blocks of persons and groups, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
-Figure~\ref{fig:simrank}(b) and Figure~\ref{fig:simrank}(d) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
-Looking at the plots from left to right, we can see that the partitioning of the node sets on the first informative eigenvector is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
-
-Figure~\ref{fig:simrank}(f) shows the correspondence plot of the first and second eigenvectors of the SimRank similarity matrices for both persons and groups.
-Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c), suggesting that the first-CA dimensions of the two-mode network recover clusters of ``similar'' entities in the two node sets, where internode similarity is a generalized relational similarity as defined earlier.
-Thus, [$Flora, Olivia$] appear close because they choose similar groups, as do [$Sylvia, Katherine, Helen$].
-In the same way, groups [$E9, E11$] are close in the plot because they are chosen by similar people, as are events [$E10, E12, E13, E14$].
-Figures~\ref{fig:simrank}(g) and (h) show a regression plot of the first CA eigenvector for persons (top) and groups (bottom) on the x-axis and the first eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
-As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that these first eigenvectors capture the same (GSR) information across the two approaches.
-
-\section{Discussion}
-%% Loading bibliography style file
-%\bibliographystyle{model1-num-names}
-\bibliographystyle{cas-model2-names}
-% Loading bibliography database
-\bibliography{ca.bib}
 \end{document}

### 2024-07-12 analysis.qmd
Move around some things

@@
    library(patchwork)
-   source(here("Functions", "TwoModeCA.R"))
-   source(here("Functions", "TwoModeSR.R"))
-   source(here("Functions", "TwoModeGSR.R"))
-   source(here("Functions", "reflections.R"))
-   source(here("Functions", "ref.long.dat.R"))
 ```
 
-# Data
-```{r Southern women data}
-    A <- matrix(
-         c(
-          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
-          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
-          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
-          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0,
-
-          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
-
-          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
-          ncol = 14, byrow = TRUE)
-     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRNA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-     e <- paste("E", c(1:14), sep = "")
-     rownames(A) <- w
-     colnames(A) <- e
-```
-
+# Southern Women Data Table
 ```{r HTML data table}
-   A <- A[sort(rownames(A)), ]
+   source(here("Functions", "sw.dat.R"))
+   A <- A[sort(rownames(sw.dat())), ]
    kbl(A, align = "c", format = "html") %>% 
@@
 ```{r Reflections Bump Chart (Persons)}
+   source(here("Functions", "reflections.R"))
+   source(here("Functions", "ref.long.dat.R"))
    evens <- function(x) subset(x, x %% 2 == 0)
@@
    p.e.r <- a$p.r[, evens(1:20)]
-   p.o.r <- a$p.r[, odds(1:20)]
-   p.o.r <- p.o.r[, -1]
-   b <- ref.long.dat(a = a$p.r, b = a$g.r, max.iter = 26)
-   plot.dat <- b$person$even
+   g.e.r <- a$g.r[, evens(1:20)]
+   b <- ref.long.dat(a = p.e.r, b = g.e.r, max.iter = 26)
+   plot.dat <- b$person
    p <- ggplot(data = filter(plot.dat, n.iter < 18),
@@
 ```{r Reflections Bump Chart (Groups)}
-   plot.dat <- b$group$even %>% 
-      mutate(ref = 15 - ref)
+   plot.dat <- b$group
    p <- ggplot(data = filter(plot.dat, n.iter < 14),

### 2024-07-12 slides/duality-presentation.qmd
changes

--- 2024-05-22 duality-presentation.qmd
@@
 date: "`r Sys.Date()`"
+format:
+  beamer:
+    incremental: false 
+    theme: simple 
+    smaller: true
     slide-number: true
@@
     progress: true
-format:
-  revealjs:
-    incremental: false 
-    theme: simple 
-    smaller: true
     footer: "https://github.com/olizardo/correspondence-analysis-two-mode-networks"

### 2024-07-12 slides/ncsu-presentation.qmd
changes

--- 2024-03-31 ncsu-presentation.qmd
@@
     incremental: false 
-    theme: sky 
+    theme: simple 
     smaller: true
 ---
-
-## CA and Two Mode Networks
-
-- While not common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data.
-
-  - But criticized for being limited and inapplicable for most "sna" tasks*
-
-- CA Proponents point to some important advantages of CA for the analysis of two-mode networks:
-  - CA can provide synoptic graphical representations of structural connectivity patterns in two-mode data**
-  - (M)CA can be used to "visually explore" two-mode networks***
-
-::: aside
-*Borgatti & Everett 1997
-
-**Breiger 2000; Roberts 2000; Faust 2005
-
-***D’Esposito et al. 2014
-:::
-
-## CA and Two Mode Networks
-
-- While not common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data.
-
-  - But criticized for being limited and inapplicable for most "sna" tasks*
-
-- This presentation builds on and expands on previous  research on using CA for two-mode data analysis
-
-  - But discusses aspects of CA that have not been emphasized in previous research:
-
-1. CA as computing a type of "(Reflective) Dual Centrality" on the entities in the two-modes**
-    - Rediscovered recently under the heading of the Economic Complexity Index (ECI; Hidalgo and Hausman 2009)
 
@@
 
-2. CA as extracting "positional" information the entities in the two modes
-    - With positions defined by "Generalized Relational Similarity"**
-    - Similar entities have similar patterns of connectivity with entities that are themselves similar
-
 ::: aside
@@
 
-*Kovacs 2010; Lizardo 2024
 :::
@@
 ::: aside
-*Later referred to as the Economic Complexity Index (ECI)
+*Referred to as the Economic Complexity Index (ECI)
 
@@
 :::
-
-## Southern Women Data {.scrollable}
-```{r Southern women data}
-    A <- matrix(
-         c(
-          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
-          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
-          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
-          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0,
-
-          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
-
-          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
-          ncol = 14, byrow = TRUE)
-     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-     e <- paste("E", c(1:14), sep = "")
-     rownames(A) <- w
-     colnames(A) <- e
-     require(kableExtra)
-     A <- A[sort(rownames(A)), ]
-     kbl(A)
-```
 
@@
 ## CA and Bonacich Centrality
-- But what's the connection to reflections and CA after all of this?
-
-- A couple of things first:
-
-  - $Dp$ is a $|P| \times |P|$ matrix containing each person's $C_p^R(1)$ (activity) along the diagonals and zeroes in every other cell
-  - $Dg$ is a $|G| \times |G|$ matrix $Dg$ containing each group's $C_g^R(1)$ (size) along the diagonals and zeroes in every other cell
-
-. . .
-
-- In the limit ($q \rightarrow \infty$), the iterative HH reflective centralities are the solution of the eigensystem:*
-
-$$
-    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
-$$ {#eq-dam1}
-
-$$
-    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
-$$ {#eq-dam2}
-
-::: aside
-*van Dam et al. 2021
-:::
 
@@
 ::: aside
-Mealy 2019, eq. 4
+*Mealy 2019, eq. 4
 :::
@@
 :::
-
+ be thought of as **similarity matrices** based on the shared neighbors in the two mode network
 ## CA Reveals Dual Communities
@@
 
-- Which can be thought of as **similarity matrices** based on the shared neighbors in the two mode network
+- Which can
 
@@
 
-. . .
-
 - According to Jeh & Widom:
@@
 ## CA and Generalized Similarity
-
-- In *SimRank* the similarity between two people is given by:
-
-$$
-S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
-    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
-    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
-$$ {#eq-simrank1}
 

### 2024-07-12 Functions/discard/TwoModeCA.R

@@ -1,0 +1,173 @@
+TwoModeCA <- function(x, 
+                      k.r = 2, 
+                      k.c = 2, 
+                      b.r = 2,
+                      b.c = 2,
+                      d.r = 2, 
+                      d.c = 2) {
+        
+        #loading functions
+        source(here("Functions", "eigval.scatter.R"))
+        source(here("Functions", "eigvec.scatter.R"))
+        source(here("Functions", "corr.scatter.R"))
+        source(here("Functions", "simplot.R"))
+        
+        # Initial matrices and vectors
+        A <- x[ , colSums(x) != 0, drop = FALSE] #dropping isolates
+        At <- t(A) #affiliation matrix transpose
+        r <- nrow(A) #number of rows
+        c <- ncol(A) #number of columns
+        rn <- rownames(A) #row name vector
+        cn <- colnames(A) #column name vector
+        u.r <- matrix(1, nrow = r, ncol = 1) #unit row vector
+        u.c <- matrix(1, nrow = c, ncol = 1) #unit column vector
+
+        # Derived matrices and vectors
+        AAt <- A %*% At #row projection
+        AtA <- At %*% A #column projection
+        N <- as.numeric(t(u.r) %*% A %*% u.c) #affiliation matrix size
+        D.r <- diag(as.vector(A %*% u.c), r, r) #row diagonal degree matrix
+        D.c <- diag(as.vector(At %*% u.r), c, c) #column diagonal degree matrix
+        iDr <- solve(D.r) #inverse of row diagonal degree matrix
+        iDc <- solve(D.c)  #inverse of column diagonal degree matrix
+        S.r <- A %*% iDc %*% At #degree-normalized similarity matrix for rows
+        S.c <- At %*% iDr %*% A #degree-normalized similarity matrix for columns
+        
+        # Spectral matrices
+        CA.r <- eigen(iDr %*% A %*% iDc %*% At) #row CA
+        CA.c <- eigen(iDc %*% At %*% iDr %*% A) #column CA
+        S.r.eig <- eigen(S.r) #eigendecomposition of row similarity matrix
+        S.c.eig <- eigen(S.c) #eigendecomposition of column similarity matrix
+        eigvec.r <- CA.r$vectors[, 2:r] #row coordinates
[…]

### 2024-07-12 Functions/ref.long.dat.R

@@ -2,7 +2,7 @@ ref.long.dat <- function(a, b, max.iter) {
      library(tidyverse)
      pr.long <- data.frame(a)
      pr.long <- pr.long %>% 
-          mutate(mode = rownames(pr.long)) %>% 
+          mutate(person = rownames(pr.long)) %>% 
           pivot_longer(
                cols = 1:ncol(a),
                names_to = "iter",
@@ -12,7 +12,7 @@ ref.long.dat <- function(a, b, max.iter) {
        mutate(n.iter = as.integer(iter)*2)
      gr.long <- data.frame(b)
      gr.long <- gr.long %>% 
-          mutate(mode = rownames(gr.long)) %>% 
+          mutate(group = rownames(gr.long)) %>% 
           pivot_longer(
                cols = 1:ncol(b),
                names_to = "iter",

### 2024-07-12 southern-women.R

@@ -1,61 +1,0 @@
-require(kableExtra)
-require(here)
-require(FactoMineR)
-
-A <- matrix(
-     c(
-          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
-          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
-          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
-          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
-          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0, 
-          
-          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
-          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0, 
-          
-          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
-          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
-     ncol = 14, byrow = TRUE)
[…]

### 2024-07-13 analysis.qmd
social networks paper

@@
    source(here("Functions", "sw.dat.R"))
-   A <- A[sort(rownames(sw.dat())), ]
+   A <- sw.dat()
+   A <- A[sort(rownames(A)), ]
    kbl(A, align = "c", format = "html") %>% 
@@
 ```{r CA versus Reflections Scatter}
+   source(here("Functions", "TwoModeCA.R"))
    tm <- TwoModeCA(A)
-   mr <- reflections(A, iter = 26)
-   p.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.r[, 1], 
-                                  r = mr$p.s[, 22]))
+   p.plot.dat <- data.frame(c = tm$ca.eigvec.r[, 1], r = a$p.s[, 26])
    p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
@@
 
-   p.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1], 
-                                  r = mr$g.s[, 16]))
-   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
+   g.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1], 
+                                  r = a$g.s[, 26]))
+   p <- ggplot(data = g.plot.dat, aes(x = c, y = r))
    p <- p + geom_point(size = 3, color = "red") 
    p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
-   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), 
+   p <- p + geom_text_repel(aes(label = rownames(g.plot.dat)), 
                             color = "blue", size = 6)
@@
    dev.off()
-```
-
-```{r CA via SVD of two mode network}
-   Rp <- diag(sqrt(rowSums(A)))
-   Rc <- diag(sqrt(colSums(A)))
-
-   svd.res <- svd(Rp%*%A%*%Rc)
-
-   t <- sum(rowSums(A))
-   r.rp <- t/rowSums(A)
-   r.cp <- t/colSums(A)
-   s.rp.cp <- sqrt(c(r.rp, r.cp))
-   tv <- t(svd.res$v)
-   x <- c(svd.res$u[, 3]*r.rp, tv[3, ]*r.cp) 
-   y <- c(svd.res$u[, 2]*r.rp, tv[2, ]*r.cp)
-
-   d <- data.frame(x = x, y = y, l = c(rownames(A), colnames(A)))
-   ggscatter(d, x = "x", y = "y", label = "l", repel = TRUE)
 ```
@@
    dev.off() 
-   png(file = here("Plots", "ca-v-bon-reord.png"), width=1000, height=700)
-      (b | c) + plot_annotation(tag_levels = "a") &
-      theme(plot.tag = element_text(size = 20))
-   dev.off() 
 ```
@@
 
-   png(file = here("Plots", "ca-eigvec-p.png"), width=500, height=500)
+   png(file = here("Plots", "p-ca-eigvec-p.png"), width=500, height=500)
    a
    dev.off() 
-   png(file = here("Plots", "ca-eigvec-g.png"), width=500, height=500)
+   png(file = here("Plots", "g-ca-eigvec-g.png"), width=500, height=500)
    b
-   dev.off() 
-    png(file = here("Plots", "ca-eigvec.png"), width=1000, height=600)
-        (a | b) + plot_annotation(tag_levels = "a") &
-      theme(plot.tag = element_text(size = 20))
    dev.off() 
@@
 
-    png(file = here("Plots", "ca-sim-p.png"), width=600, height=625)
+    png(file = here("Plots", "p-ca-sim.png"), width=600, height=625)
     a + theme(axis.text.y = element_text(size = 18))
@@
 
-    png(file = here("Plots", "ca-sim-g.png"), width=600, height=625)
+    png(file = here("Plots", "g-ca-sim.png"), width=600, height=625)
     b + theme(axis.text.y = element_text(size = 22))
@@
     c <- tmCA.res$bon.corr.plot
-    png(file = here("Plots", "bon-sim-p.png"), width=600, height=625)
+    png(file = here("Plots", "p-bon-sim.png"), width=600, height=625)
       a + theme(axis.text.y = element_text(size = 16))
@@
 
-    png(file = here("Plots", "bon-sim-g.png"), width=600, height=625)
+    png(file = here("Plots", "g-bon-sim.png"), width=600, height=625)
       b + theme(axis.text.y = element_text(size = 16))
@@
 
-   png(file = here("Plots", "sr-plot-eigen-p.png"), width=600, height=625)
+   png(file = here("Plots", "p-sr-plot-eigen.png"), width=600, height=625)
    a
    dev.off()
-   png(file = here("Plots", "sr-plot-eigen-g.png"), width=600, height=625)
+   png(file = here("Plots", "g-sr-plot-eigen.png"), width=600, height=625)
    b
@@
 ```
-
-```{r GRS Plot}
-   source(here("Functions", "TwoModeGRS.R"))
-   tmSR.res <- TwoModeGRS(A)
-    a <- tmSR.res$eigvec.plot.r 
-    b <- tmSR.res$eigvec.plot.c 
-    c <- tmSR.res$eigval.plot.r
-    d <- tmSR.res$eigval.plot.c
-    a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
-    b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
-    e <- tmSR.res$A.plot
-
-   dat.r <- data.frame(ca = as.numeric(tm$ca.eigvec.r[,1])*-1,
-                       sr = tmSR.res$ev.r1,
-                       lab = rownames(A))
-   dat.c <- data.frame(ca = as.numeric(tm$ca.eigvec.c[,1])*-1,
-                       sr = tmSR.res$ev.c1,
-                       lab = colnames(A))
-   cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
-   g <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
-                  label = "lab", repel = TRUE, 
-                  add = "reg.line", 
-                  add.params = list(color = "gray80", 
-                                    linewidth = 0.5, linetype = 2)) +
-                  annotate("text", x = -.2, y = .3, 
-                           label = paste("r = ", cor), size = 6) 
-   cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
-   h <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
-                  label = "lab", repel = TRUE, 
-                  add = "reg.line",
-                  add.params = list(color = "gray80", 
-                                    linewidth = 0.5, linetype = 2)) +
-                  annotate("text", x = -.2, y = .3, 
-                           label = paste("r = ", cor), size = 6)
-
-   png(file = here("Plots", "grs-plot-eigen-p.png"), width=600, height=625)
-   a
-   dev.off()
-   png(file = here("Plots", "grs-plot-eigen-g.png"), width=600, height=625)
-   b
-   dev.off()
-   png(file = here("Plots", "grs-plot-reord.png"), width=600, height=625)
-   e
-   dev.off()
-   png(file = here("Plots", "grs-corr-scatter.png"), width=350, height=625)
-   g / h
-   dev.off()
-```

### 2024-07-13 Functions/TwoModeCA.R

@@ -1,0 +1,187 @@
+TwoModeCA <- function(x, 
+                      k.r = 2, 
+                      k.c = 2, 
+                      b.r = 2,
+                      b.c = 2,
+                      d.r = 2, 
+                      d.c = 2) {
+        
+        #loading functions
+        source(here("Functions", "eigval.scatter.R"))
+        source(here("Functions", "eigvec.scatter.R"))
+        source(here("Functions", "corr.scatter.R"))
+        source(here("Functions", "simplot.R"))
+        
+        # Initial matrices and vectors
+        A <- x[ , colSums(x) != 0, drop = FALSE] #dropping isolates
+        At <- t(A) #affiliation matrix transpose
+        r <- nrow(A) #number of rows
+        c <- ncol(A) #number of columns
+        rn <- rownames(A) #row name vector
+        cn <- colnames(A) #column name vector
+        u.r <- matrix(1, nrow = r, ncol = 1) #unit row vector
+        u.c <- matrix(1, nrow = c, ncol = 1) #unit column vector
+
+        # Derived matrices and vectors
+        AAt <- A %*% At #row projection
+        AtA <- At %*% A #column projection
+        N <- as.numeric(t(u.r) %*% A %*% u.c) #affiliation matrix size
+        D.r <- diag(as.vector(A %*% u.c), r, r) #row diagonal degree matrix
+        D.c <- diag(as.vector(At %*% u.r), c, c) #column diagonal degree matrix
+        iDr <- solve(D.r) #inverse of row diagonal degree matrix
+        iDc <- solve(D.c)  #inverse of column diagonal degree matrix
+        S.r <- A %*% iDc %*% At #degree-normalized similarity matrix for rows
+        S.c <- At %*% iDr %*% A #degree-normalized similarity matrix for columns
+        
+        # Spectral matrices
+        CA.r <- eigen(iDr %*% A %*% iDc %*% At) #row CA
+        CA.c <- eigen(iDc %*% At %*% iDr %*% A) #column CA
+        S.r.eig <- eigen(S.r) #eigendecomposition of row similarity matrix 
+        S.c.eig <- eigen(S.c) #eigendecomposition of column similarity matrix 
[…]

### 2024-07-13 Functions/TwoModeSR.R

@@ -1,0 +1,64 @@
+TwoModeSR <- function(A, k = 4, n.eig = 4) {
+   #loading functions
+   source(here("Functions", "SimRank.R"))
+   source(here("Functions", "eigval.scatter.R"))
+   source(here("Functions", "eigvec.scatter.R"))
+   source(here("Functions", "corr.scatter.R"))
+   
+   # SimRank of matrix A
+   nr <- nrow(A)
+   nc <- ncol(A)
+   rn <- rownames(A)
+   cn <- colnames(A)
+   sr.res <- SimRank(A)
+   res.eigen.r <- eigen(sr.res$Sr) #eigenvectors/values of row similarity matrix
+   res.eigen.c <- eigen(sr.res$Sc) #eigenvectors/values of column similarity matrix
+   e.r <- res.eigen.r$vectors[, 2:nr] #row eigenvectors minus first
+   e.c <- res.eigen.c$vectors[, 2:nc] #column eigenvectors minus first
+   v.r <- res.eigen.r$values[2:nr] #row eigenvalues minus first
+   v.c <- res.eigen.c$values[2:nc] #column eigenvalues minus first
+   ev.r1 <- e.r[, 1] #main eigenvector of row similarity matrix
+   ev.r2 <- e.r[, 2] #second main eigenvector of row similarity matrix
+   ev.c1 <- e.c[, 1] #main eigenvector of column similarity matrix
+   ev.c2 <- e.c[, 2] #second main eigenvector of column similarity matrix
+   
+   # Affiliation matrix plot (SimRank ordering)
+   A.ord <- as.matrix(A[order(ev.r1), order(ev.c1)])
+   A.plot <- ggcorrplot(t(A.ord)) 
+   A.plot <- A.plot + scale_x_discrete(position = "top")
+   A.plot <- A.plot + theme(legend.position = "none",
+                             axis.text.x = element_text(hjust = -0.2))
+   
+   # Eigenvalue plot of SimRank similarity matrix
+   eig.dat.r <- data.frame(k = 2:nr, value = v.r) %>% 
+      mutate(value = value / max(value))
+   eig.dat.c <- data.frame(k = 2:nc, value = v.c) %>% 
+      mutate(value = value / max(value))
+   eigval.plot.r <- eigval.scatter(eig.dat.r)
+   eigval.plot.c <- eigval.scatter(eig.dat.c)
+   
+   # Main eigenvector plot of SimRank similarity matrix
[…]

### 2024-07-13 Functions/corr.scatter.R

@@ -1,0 +1,20 @@
+corr.scatter <- function(x, s = 12) {
+     p <- ggscatter(data.frame(x), 
+                    x = "d1", 
+                    y = "d2", 
+                    point = FALSE,
+                    color = "cluster",
+                    palette = "uchicago",
+                    font.label =  s,
+                    label = "lab", 
+                    repel = TRUE)
+     p <- p + theme(legend.position = "none",
+                    axis.line.x = element_blank(),
+                    axis.line.y = element_blank())
+     p <- p + geom_vline(aes(xintercept = 0), 
+                         color = "gray", linetype = 2)
+     p <- p + geom_hline(aes(yintercept = 0), 
+                         color = "gray", linetype = 2)
+     p <- p + labs(x = "First Axis", y = "Second Axis")
+     return(p)
+     }

### 2024-07-13 Functions/eigval.scatter.R

@@ -1,0 +1,13 @@
+eigval.scatter <- function(x, s = 2) {
+     p <- ggscatter(x, 
+                    x = "k", 
+                    y = "value",
+                    size = s,
+                    color = "red")
+     p <- p + labs(y = "", x = "k") 
+     p <- p + theme(axis.text.x = element_blank(),
+                    axis.text.y = element_text(size = 8),
+                    axis.line.y = element_blank(),
+                    axis.line.x = element_blank())
+     return(p)
+     }

### 2024-07-13 Functions/simplot.R

@@ -1,0 +1,7 @@
+sim.plot <- function(x) {
+     p <- ggcorrplot(x)
+     p <- p + scale_x_discrete(position = "top")
+     p <- p + theme(legend.position = "none",
+                    axis.text.x = element_text(hjust = -0.2)) 
+     return(p)
+}

### 2024-07-13 Functions/discard/.R
re-arrange

@@ -1,187 +1,0 @@
-TwoModeCA <- function(x, 
-                      k.r = 2, 
-                      k.c = 2, 
-                      b.r = 2,
-                      b.c = 2,
-                      d.r = 2, 
-                      d.c = 2) {
-        
-        #loading functions
-        source(here("Functions", "eigval.scatter.R"))
-        source(here("Functions", "eigvec.scatter.R"))
-        source(here("Functions", "corr.scatter.R"))
-        source(here("Functions", "simplot.R"))
-        
-        # Initial matrices and vectors
-        A <- x[ , colSums(x) != 0, drop = FALSE] #dropping isolates
-        At <- t(A) #affiliation matrix transpose
-        r <- nrow(A) #number of rows
-        c <- ncol(A) #number of columns
-        rn <- rownames(A) #row name vector
-        cn <- colnames(A) #column name vector
-        u.r <- matrix(1, nrow = r, ncol = 1) #unit row vector
-        u.c <- matrix(1, nrow = c, ncol = 1) #unit column vector
-
-        # Derived matrices and vectors
-        AAt <- A %*% At #row projection
-        AtA <- At %*% A #column projection
-        N <- as.numeric(t(u.r) %*% A %*% u.c) #affiliation matrix size
-        D.r <- diag(as.vector(A %*% u.c), r, r) #row diagonal degree matrix
-        D.c <- diag(as.vector(At %*% u.r), c, c) #column diagonal degree matrix
-        iDr <- solve(D.r) #inverse of row diagonal degree matrix
-        iDc <- solve(D.c)  #inverse of column diagonal degree matrix
-        S.r <- A %*% iDc %*% At #degree-normalized similarity matrix for rows
-        S.c <- At %*% iDr %*% A #degree-normalized similarity matrix for columns
-        
-        # Spectral matrices
-        CA.r <- eigen(iDr %*% A %*% iDc %*% At) #row CA
-        CA.c <- eigen(iDc %*% At %*% iDr %*% A) #column CA
-        S.r.eig <- eigen(S.r) #eigendecomposition of row similarity matrix 
-        S.c.eig <- eigen(S.c) #eigendecomposition of column similarity matrix 
[…]

### 2024-07-13 Functions/discard/TwoModeSR.R

@@ -1,64 +1,0 @@
-TwoModeSR <- function(A, k = 4, n.eig = 4) {
-   #loading functions
-   source(here("Functions", "SimRank.R"))
-   source(here("Functions", "eigval.scatter.R"))
-   source(here("Functions", "eigvec.scatter.R"))
-   source(here("Functions", "corr.scatter.R"))
-   
-   # SimRank of matrix A
-   nr <- nrow(A)
-   nc <- ncol(A)
-   rn <- rownames(A)
-   cn <- colnames(A)
-   sr.res <- SimRank(A)
-   res.eigen.r <- eigen(sr.res$Sr) #eigenvectors/values of row similarity matrix
-   res.eigen.c <- eigen(sr.res$Sc) #eigenvectors/values of column similarity matrix
-   e.r <- res.eigen.r$vectors[, 2:nr] #row eigenvectors minus first
-   e.c <- res.eigen.c$vectors[, 2:nc] #column eigenvectors minus first
-   v.r <- res.eigen.r$values[2:nr] #row eigenvalues minus first
-   v.c <- res.eigen.c$values[2:nc] #column eigenvalues minus first
-   ev.r1 <- e.r[, 1] #main eigenvector of row similarity matrix
-   ev.r2 <- e.r[, 2] #second main eigenvector of row similarity matrix
-   ev.c1 <- e.c[, 1] #main eigenvector of column similarity matrix
-   ev.c2 <- e.c[, 2] #second main eigenvector of column similarity matrix
-   
-   # Affiliation matrix plot (SimRank ordering)
-   A.ord <- as.matrix(A[order(ev.r1), order(ev.c1)])
-   A.plot <- ggcorrplot(t(A.ord)) 
-   A.plot <- A.plot + scale_x_discrete(position = "top")
-   A.plot <- A.plot + theme(legend.position = "none",
-                             axis.text.x = element_text(hjust = -0.2))
-   
-   # Eigenvalue plot of SimRank similarity matrix
-   eig.dat.r <- data.frame(k = 2:nr, value = v.r) %>% 
-      mutate(value = value / max(value))
-   eig.dat.c <- data.frame(k = 2:nc, value = v.c) %>% 
-      mutate(value = value / max(value))
-   eigval.plot.r <- eigval.scatter(eig.dat.r)
-   eigval.plot.c <- eigval.scatter(eig.dat.c)
-   
-   # Main eigenvector plot of SimRank similarity matrix
[…]

### 2024-07-13 Functions/discard/blah.R

@@ -1,23 +1,0 @@
-r <- c("Croatia", "Denmark", "Finland", "France", "Netherlands", "Serbia", "Spain", "Switzerland", "United Kingdom")
-c <- c("Arts", "Institutional", "Values and trad.", "Geographical ID", "Group chars.", "Habits and lifestyles", "Human activity", "History and heritage", "Cultivation", "Good manners")
-
-dat <- matrix(c(0.10,0.42,0.02,0.02,0.10,0.03,0.07,0.03,0.03,0.16, 0.16,0.22,0.10,0.06,0.07,0.06,0.18,0.06,0.04,0.05, 0.21,0.19,0.07,0.05,0.08,0.06,0.19,0.05,0.05,0.04, 0.13,0.25,0.03,0.04,0.02,0.03,0.02,0.05,0.42,0.01, 0.13,0.15,0.14,0.12,0.12,0.16,0.08,0.03,0.02,0.05, 0.10,0.27,0.06,0.03,0.08,0.04,0.05,0.03,0.03,0.31, 0.10,0.12,0.08,0.05,0.06,0.07,0.08,0.04,0.38,0.02, 0.15,0.23,0.07,0.06,0.06,0.05,0.11,0.06,0.16,0.05,0.17,0.07,0.23,0.10,0.18,0.11,0.05,0.06,0.02,0.02), nrow = 9, byrow = TRUE)
-rownames(dat) <- r
-colnames(dat) <- c
-
-library(FactoMineR)
-#res.CA <- CA(dat)
-
-dat.f <- data.frame(country = r, dat) %>% 
-     pivot_longer(
-          cols = 2:11,
-          names_to = "topics",
-          values_to = "prop"
-     ) 
-
-library(ggplot2)
-p <- ggplot(aes(y = country, x = prop, fill = country, group = country), data = dat.f)
-p <- p + geom_bar(stat = "identity")
-p <- p + facet_wrap(~ topics, ncol = 5)
-p <- p + theme_minimal() + theme(legend.position = "none") + labs(y = "", x = "")
-p

### 2024-07-13 Functions/discard/corr.scatter.R

@@ -1,20 +1,0 @@
-corr.scatter <- function(x, s = 12) {
-     p <- ggscatter(data.frame(x), 
-                    x = "d1", 
-                    y = "d2", 
-                    point = FALSE,
-                    color = "cluster",
-                    palette = "uchicago",
-                    font.label =  s,
-                    label = "lab", 
-                    repel = TRUE)
-     p <- p + theme(legend.position = "none",
-                    axis.line.x = element_blank(),
-                    axis.line.y = element_blank())
-     p <- p + geom_vline(aes(xintercept = 0), 
-                         color = "gray", linetype = 2)
-     p <- p + geom_hline(aes(yintercept = 0), 
-                         color = "gray", linetype = 2)
-     p <- p + labs(x = "First Axis", y = "Second Axis")
-     return(p)
-     }

### 2024-07-13 Functions/discard/eigval.scatter.R

@@ -1,13 +1,0 @@
-eigval.scatter <- function(x, s = 2) {
-     p <- ggscatter(x, 
-                    x = "k", 
-                    y = "value",
-                    size = s,
-                    color = "red")
-     p <- p + labs(y = "", x = "k") 
-     p <- p + theme(axis.text.x = element_blank(),
-                    axis.text.y = element_text(size = 8),
-                    axis.line.y = element_blank(),
-                    axis.line.x = element_blank())
-     return(p)
-     }

### 2024-07-13 Functions/discard/eigvec.scatter.R

@@ -1,22 +1,0 @@
-eigvec.scatter <- function(x, s = 10) {
-     p <- ggscatter(x, 
-                    y = "rank", 
-                    x = "value", 
-                    point = FALSE,
-                    font.label =  s,
-                    label = "lab", 
-                    color = "cluster",
-                    palette = "Dark2",
-                    repel = TRUE
-                    #ellipse = TRUE,
-                    #ellipse.type = "convex",
-                    #ellipse.alpha = 0.15,
-                    #ellipse.border.remove = TRUE
-     )
-     p <- p + theme(axis.text.y = element_blank(),
-                    axis.line.x = element_blank(),
-                    axis.line.y = element_blank(),
-                    legend.position = "none")
-     p <- p + labs(y = "Rank", x = "First Axis")
-     return(p)
-     }

### 2024-07-13 Functions/discard/grs-plot-codechunk.R
current plotss

@@ -1,0 +1,46 @@
+
+source(here("Functions", "TwoModeGRS.R"))
+tmSR.res <- TwoModeGRS(A)
+a <- tmSR.res$eigvec.plot.r 
+b <- tmSR.res$eigvec.plot.c 
+c <- tmSR.res$eigval.plot.r
+d <- tmSR.res$eigval.plot.c
+a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
+e <- tmSR.res$A.plot
+
+dat.r <- data.frame(ca = as.numeric(tm$ca.eigvec.r[,1])*-1,
+                    sr = tmSR.res$ev.r1,
+                    lab = rownames(A))
+dat.c <- data.frame(ca = as.numeric(tm$ca.eigvec.c[,1])*-1,
+                    sr = tmSR.res$ev.c1,
+                    lab = colnames(A))
+cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
+g <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
+               label = "lab", repel = TRUE, 
+               add = "reg.line", 
+               add.params = list(color = "gray80", 
+                                 linewidth = 0.5, linetype = 2)) +
+     annotate("text", x = -.2, y = .3, 
+              label = paste("r = ", cor), size = 6) 
+cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
+h <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
+               label = "lab", repel = TRUE, 
+               add = "reg.line",
+               add.params = list(color = "gray80", 
+                                 linewidth = 0.5, linetype = 2)) +
+     annotate("text", x = -.2, y = .3, 
+              label = paste("r = ", cor), size = 6) 
+
+png(file = here("Plots", "grs-plot-eigen-p.png"), width=600, height=625)
+a
+dev.off()
+png(file = here("Plots", "grs-plot-eigen-g.png"), width=600, height=625)
+b
+dev.off()
+png(file = here("Plots", "grs-plot-reord.png"), width=600, height=625)
+e
+dev.off()
+png(file = here("Plots", "grs-corr-scatter.png"), width=350, height=625)
+g / h
+dev.off()

### 2024-07-13 main.tex
Updates from Overleaf

@@
 \subsection{Empirical Example} \label{subsec:ex}
-Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}.
+Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons in the Southern Women data \citep{davis1941}, using a bump chart to plot the rank order trajectory of persons across reflections.
 The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
-
-As the bottom panel of the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to.
+As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to.
 However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
@@
 
-The bottom panel of Figure~\ref{fig:pg-reflections} shows the corresponding bump chart for the events.
+Figure~\ref{fig:g-refs} shows the corresponding bump chart for the events.
 Just like for these actors, the equilibrium reflective centralities recover the ordering of the columns according to Doreian et al.
@@
 
-\begin{figure}
+\begin{figure}[ht!]
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
-    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
-    \label{fig:pg-reflections}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-reflections-sw.png}
+            \caption{Reflective centrality trajectories of persons.}
+            \label{fig:p-refs}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-reflections-sw.png}
+            \caption{Reflective centrality trajectories of groups.}
+            \label{fig:g-refs}
+    \end{subfigure}
+    \caption{Reflective centralities for persons and groups (even reflections)}
+    \label{fig:refs}
 \end{figure}
@@
 Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}.
-Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums around steady values after a few iterations.
+
+Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations.
 These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach.
@@
 The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
+
+\begin{figure}[ht!]
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
+    \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-reflections-sw.png}
+            \caption{Reflective centrality trajectories of persons.}
+            \label{fig:p-refs}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-reflections-sw.png}
+            \caption{Reflective centrality trajectories of groups.}
+            \label{fig:g-refs}
+    \end{subfigure}
+    \caption{Reflective centralities for persons and groups (even reflections)}
+    \label{fig:refs}
+\end{figure}
 

### 2024-07-13 analysis.qmd
updated tmSR function and plot

@@
 
-   png(file = here("Plots", "p-ca-eigvec-p.png"), width=500, height=500)
+   png(file = here("Plots", "p-ca-eigvec.png"), width=500, height=500)
    a
    dev.off() 
-   png(file = here("Plots", "g-ca-eigvec-g.png"), width=500, height=500)
+   png(file = here("Plots", "g-ca-eigvec.png"), width=500, height=500)
    b
@@
     e <- tmSR.res$A.plot
-
    dat.r <- data.frame(ca = as.numeric(tm$ca.eigvec.r[,1]),

### 2024-07-13 Functions/TwoModeSR.R

@@ -23,7 +23,7 @@ TwoModeSR <- function(A, k = 4, n.eig = 4) {
    ev.c2 <- e.c[, 2] #second main eigenvector of column similarity matrix
    
    # Affiliation matrix plot (SimRank ordering)
-   A.ord <- as.matrix(A[order(ev.r1), order(ev.c1)])
+   A.ord <- as.matrix(A[order(ev.r1), order(ev.c1*-1)])
    A.plot <- ggcorrplot(t(A.ord)) 
    A.plot <- A.plot + scale_x_discrete(position = "top")
    A.plot <- A.plot + theme(legend.position = "none",
@@ -50,15 +50,29 @@ TwoModeSR <- function(A, k = 4, n.eig = 4) {
                               cluster = as.factor(k.c$cluster))
    eigvec.plot.r <- eigvec.scatter(eigvec.dat.r, s = 14)
    eigvec.plot.c <- eigvec.scatter(eigvec.dat.c, s = 14)
+   
+   #k-means similarity-based clustering
+   km.r <- hkmeans(data.frame(e.r[, 1:n.eig]), n.eig)
+   km.c <- hkmeans(data.frame(e.c[, 1:n.eig]), n.eig)
+   
+   # Correspondence Plot (SimRank)
+   a <- rbind(e.r[, 1:2], e.c[, 1:2]*-1)
+   colnames(a) <- c("d1", "d2")
+   b <- c(rownames(A), colnames(A))
+   c <- factor(c(km.r$cluster, km.c$cluster))
+   d <- data.frame(a, cluster = c, lab = b)
+   corr.plot <- corr.scatter(d)
 
 return(list(
    ev.r1 = ev.r1,
-   ev.c1 = ev.c1,
+   ev.c1 = ev.c1, 
    A.plot = A.plot,
    eigval.plot.c = eigval.plot.c,
    eigval.plot.r = eigval.plot.r,
    eigvec.plot.c = eigvec.plot.c,
-   eigvec.plot.r = eigvec.plot.r))
+   eigvec.plot.r = eigvec.plot.r,
+   corr.plot = corr.plot)
+   )
    }
 
 

### 2024-07-13 main.tex
Updates from Overleaf

@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-reflections-sw.png}
+        \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
             \caption{Reflective centrality trajectories of persons.}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-reflections-sw.png}
+        \includegraphics[width=1.0\textwidth]{Plots/g-reflections.png}
             \caption{Reflective centrality trajectories of groups.}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-reflections-sw.png}
-            \caption{Reflective centrality trajectories of persons.}
-            \label{fig:p-refs}
+        \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
+            \caption{Re-ordered Affiliation matrix (CA).}
+            \label{fig:ca-reord}
     \end{subfigure}
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-reflections-sw.png}
-            \caption{Reflective centrality trajectories of groups.}
-            \label{fig:g-refs}
-    \end{subfigure}
-    \caption{Reflective centralities for persons and groups (even reflections)}
-    \label{fig:refs}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{Plots/ca-v-bon-reord.png}
+        \includegraphics[width=1.0\textwidth]{Plots/bon-reord.png}
+            \caption{Re-ordered Affiliation matrix (Bonacich).}
+            \label{fig:bon-reord}
+    \end{subfigure}
     \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score.}
@@
 \begin{figure}
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-    \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}
-    \caption{CA Eigenvector plot for persons (a) and groups (c).
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-ca-eigvec.png}
+            \caption{Eigenvector plot (Persons).}
+            \label{fig:ca-reord}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-ca-eigvec.png}
+            \caption{Eigenvector plot (Groups).}
+            \label{fig:bon-reord}
+    \end{subfigure}
+    \caption{CA Eigenvector plot for persons (a) and groups (b).
 The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis.
-Insets show the eigenvalue orderings on the x-axis ($k$) for persons (b) and groups (d).}
+Insets show the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
     \label{fig:ca-eigvec}
@@
 \begin{figure}
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-    \includegraphics[width=1.0\textwidth]{Plots/ca-sim.png}
+     \begin{subfigure}[b]{0.425\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-ca-sim.png}
+            \caption{CA Similarity plot (Persons).}
+            \label{fig:p-ca-sim}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-ca-sim.png}
+            \caption{CA Similarity plot (Groups).}
+            \label{fig:g-ca-sim}
+    \end{subfigure} \\
+     \begin{subfigure}[b]{0.65\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
+            \caption{CA correspondence plot.}
+            \label{fig:ca-corr}
+    \end{subfigure}
     \caption{CA similarity plots for persons (a) and groups (b) on the left and right top panel.
@@
 \begin{figure}
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-    \includegraphics[width=1.0\textwidth]{Plots/bon-sim.png}
+     \begin{subfigure}[b]{0.425\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-bon-sim.png}
+            \caption{Bonacich Similarity plot (Persons).}
+            \label{fig:p-bon-sim}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-bon-sim.png}
+            \caption{Bonacich Similarity plot (Groups).}
+            \label{fig:g-bon-sim}
+    \end{subfigure} \\
+     \begin{subfigure}[b]{0.65\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
+            \caption{Bonacich correspondence plot.}
+            \label{fig:bon-corr}
+    \end{subfigure}
     \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel.
@@
 \begin{figure}
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-    \includegraphics[width=0.8\textwidth]{Plots/sr-plot.png}
+     \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/sr-plot-reord.png}
+            \caption{Bonacich Similarity plot (Persons).}
+            \label{fig:p-bon-sim}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-sr-plot-eigen.png}
+            \caption{Bonacich Similarity plot (Groups).}
+            \label{fig:g-bon-sim}
+    \end{subfigure} 
+     \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-sr-plot-eigen.png}
+            \caption{Bonacich correspondence plot.}
+            \label{fig:bon-corr}
+    \end{subfigure} \\
+     \begin{subfigure}[b]{0.6\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/sr.png}
+            \caption{Bonacich Similarity plot (Groups).}
+            \label{fig:g-bon-sim}
+    \end{subfigure} 
+     \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-sr-plot-eigen.png}
+            \caption{SimRank correspondence plot.}
+            \label{fig:sr-corr}
+    \end{subfigure}
     \caption{Simrank versus CA comparison.

### 2024-07-13 main.tex
Updates from Overleaf

@@
 
+\begin{figure}
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
+    \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-ca-ref-corr.png}
+            \caption{CA/Reflection Correlation (Persons).}
+            \label{fig:p-ca-ref}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-ca-ref-corr.png}
+            \caption{CA/Reflection Correlation (Groups)}
+            \label{fig:g-ca-ref}
+    \end{subfigure}
+    \caption{Scatter plot of the scores computed by the method of reflections ($q = 26$) on the y-axis and the first dimension of the Correspondence Analysis of the two-mode affiliation matrix on the x-axis, for both persons (a) and groups (b).
+}
+    \label{fig:ca-ref-corr}
+\end{figure}
+
 Degree-pre-weighting does not radically alter the nature of the data.
@@
     \centering
-     \begin{subfigure}[b]{0.3\textwidth}
+     \begin{subfigure}[b]{0.35\textwidth}
         \includegraphics[width=1.0\textwidth]{Plots/sr-plot-reord.png}
-            \caption{Bonacich Similarity plot (Persons).}
+            \caption{Re-ordered Affiliation Matrix (SimRank).}
             \label{fig:p-bon-sim}
@@
         \includegraphics[width=1.0\textwidth]{Plots/p-sr-plot-eigen.png}
-            \caption{Bonacich Similarity plot (Groups).}
+            \caption{SimRank Eigenvector Plot (Persons).}
             \label{fig:g-bon-sim}
@@
         \includegraphics[width=1.0\textwidth]{Plots/g-sr-plot-eigen.png}
-            \caption{Bonacich correspondence plot.}
+            \caption{SimRank Eigenvector Plot (Groups).}
             \label{fig:bon-corr}
     \end{subfigure} \\
-     \begin{subfigure}[b]{0.6\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/sr.png}
-            \caption{Bonacich Similarity plot (Groups).}
+     \begin{subfigure}[b]{0.55\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/sr-corr-plot.png}
+            \caption{SimRank Correspondence Plot}
             \label{fig:g-bon-sim}
     \end{subfigure} 
-     \begin{subfigure}[b]{0.2\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-sr-plot-eigen.png}
-            \caption{SimRank correspondence plot.}
+     \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/sr-corr-scatter.png}
+            \caption{SimRank/CA Correlation.}
             \label{fig:sr-corr}
@@
 Panel (a) shows the affiliation matrix re-ordered according to the first eigenvector of the SVD of the SimRank similarity matrix.
-Panels (b) and (d) show the corresponding eigenvector and eigenvalue---panels (c) and (e)---plots for persons and groups, respectively, of the SimRank similarity matrix.
-Panel (f) shows the correspondence plot based on the eigenvector decomposition of the SimRank similarity matrix.
-Panels (g) and (h) show the Pearson correlation between the scores corresponding to the first eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis).}
+Panels (b) and (c) show the corresponding eigenvector and eigenvalue plots for persons and groups, respectively, of the SimRank similarity matrix.
+Panel (d) shows the correspondence plot based on the eigenvector decomposition of the first two non-trivial eigenvectors of the SimRank similarity matrix.
+Panel (e) shows the Pearson correlation between the scores corresponding to the first eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis) for persons (top) and groups (bottom).}
     \label{fig:simrank}

### 2024-07-13 analysis.qmd
updatded analysis

@@
    p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
-   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "18th Reflection (Standardized)")
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
    p <- p + theme(axis.text = element_text(size = 16),
@@
                             color = "blue", size = 6)
-   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "14th Reflection (Standardized)")
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
    p <- p + theme(axis.text = element_text(size = 16),
@@
    source(here("Functions", "TwoModeSR.R"))
-   tmSR.res <- TwoModeSR(A)
+   tmSR.res <- TwoModeSR(A, n.eig = 5)
     a <- tmSR.res$eigvec.plot.r 
@@
    dev.off()
+   png(file = here("Plots", "sr-corr-plot.png"), width=600, height=625)
+   tmSR.res$corr.plot
+   dev.off()
+
 ```

### 2025-01-26 Functions/bump.plot.R
yikes

@@ -1,0 +1,29 @@
+bump.plot <- function(x, p.name = "plot.png", w = 600, h = 625, c, rev = 1, n.r = 0.01) {
+     p <- ggplot(data = dplyr::filter(x, n.iter < c), 
+                 mapping = aes(x = n.iter, y = ref*rev, color = mode))
+     p <- p + geom_bump(linewidth = 1.1, smooth = 8)
+     p <- p + geom_point(size = 5)
+     p <- p + geom_text_repel(data = dplyr::filter(x, n.iter == 2), 
+                              aes(x = n.iter, label = mode), 
+                              size = 5, nudge_x = -3)
+     p <- p + geom_text_repel(data = dplyr::filter(x, n.iter == c),
+                              aes(x = n.iter, label = mode),
+                              size = 5, nudge_x = n.r)
+     p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
+     p <- p + scale_x_continuous(limits = c(-2, c + 2), breaks = seq(2, c-2, by = 2))
+     p <- p + labs(x = "Reflection")
+     p <- p + scale_y_reverse()
+     p <- p + theme(legend.position = "none",
+                    panel.grid.major = element_blank(),
+                    axis.text.y = element_blank(),
+                    axis.title.y = element_blank(),
+                    axis.text.x = element_text(size = 16)
+     )
+     set.seed(145)
+     c <- sample(cols25(25), length(unique(x$mode)), replace = FALSE)
+     p <- p + scale_color_manual(values=as.vector(c))
+     png(file = here("Plots", p.name), width = w, height = h)
+     p
+     dev.off()
+return(p)
+}

### 2025-01-26 Functions/cult.dat.R

@@ -1,0 +1,35 @@
+cult.dat <- function() {
+     library(dplyr)
+     library(haven)
+     taste.dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
+     taste.dat <- taste.dat %>% 
+          dplyr::select("id", ends_with(c("lis", "like")), -starts_with("none")) %>% 
+          dplyr::select(c(1:41)) %>% 
+          na.omit() %>% 
+          mutate(Classical = classicallike * classicallis,
+                 Opera = operalike * operalis,
+                 Jazz = jazzlike * jazzlis,
+                 Broadway = bwaystlike * bwaystlis,
+                 Easy = moodezlike * moodezlis, 
+                 Bigband = bbandlike * bbandlis,
+                 Classic_Rock = croldlike * croldlis,
+                 Country = countrylike * countrylis,
+                 Bluegrass = blueglike * blueglis,
+                 Folk = folklike * folklis,
+                 Gospel = hymgoslike * hymgoslis,
+                 Salsa = latlpsallike * latspsallis,
+                 Rap_Hip_Hop = raphiphoplike * raphiphoplis,
+                 Blues_RandB = blurblike * blurblis,
+                 Reggae = reggaelike * reggaelis,
+                 Pop = toppoplike * toppoplis,
+                 Contemp_Rock = controcklike * controcklis,
+                 Indie_Alt = indaltlike * indaltlis,
+                 Dance_Club = danclublike * danclublis,
+                 Metal = hvymtllike * hvymtllis
+          ) %>%  #people are linked to genres that the both like and listen to
+          dplyr::select(id, Classical:Metal) 
+          A <- as.matrix(taste.dat[, 2:ncol(taste.dat)])
+          rownames(A) <- taste.dat$id
+          A <- A[-which(rowSums(A) == 0), ]
+     return(A)
+     }

### 2025-01-26 Functions/discard/bump.plot.R

@@ -1,29 +1,0 @@
-bump.plot <- function(x, p.name = "plot.png", w = 600, h = 625, c, rev = 1, n.r = 0.01) {
-     p <- ggplot(data = dplyr::filter(x, n.iter < c), 
-                 mapping = aes(x = n.iter, y = ref*rev, color = mode))
-     p <- p + geom_bump(linewidth = 1.1, smooth = 8)
-     p <- p + geom_point(size = 5)
-     p <- p + geom_text_repel(data = dplyr::filter(x, n.iter == 2), 
-                              aes(x = n.iter, label = mode), 
-                              size = 5, nudge_x = -3)
-     p <- p + geom_text_repel(data = dplyr::filter(x, n.iter == c),
-                              aes(x = n.iter, label = mode),
-                              size = 5, nudge_x = n.r)
-     p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
-     p <- p + scale_x_continuous(limits = c(-2, c + 2), breaks = seq(2, c-2, by = 2))
-     p <- p + labs(x = "Reflection")
-     p <- p + scale_y_reverse()
-     p <- p + theme(legend.position = "none",
-                    panel.grid.major = element_blank(),
-                    axis.text.y = element_blank(),
-                    axis.title.y = element_blank(),
-                    axis.text.x = element_text(size = 16)
-     )
-     set.seed(145)
-     c <- sample(cols25(25), length(unique(x$mode)), replace = FALSE)
-     p <- p + scale_color_manual(values=as.vector(c))
-     png(file = here("Plots", p.name), width = w, height = h)
-     p
-     dev.off()
-return(p)
-}

### 2025-01-26 Functions/discard/cult.dat.R

@@ -1,35 +1,0 @@
-cult.dat <- function() {
-     library(dplyr)
-     library(haven)
-     taste.dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
-     taste.dat <- taste.dat %>% 
-          dplyr::select("id", ends_with(c("lis", "like")), -starts_with("none")) %>% 
-          dplyr::select(c(1:41)) %>% 
-          na.omit() %>% 
-          mutate(Classical = classicallike * classicallis,
-                 Opera = operalike * operalis,
-                 Jazz = jazzlike * jazzlis,
-                 Broadway = bwaystlike * bwaystlis,
-                 Easy = moodezlike * moodezlis, 
-                 Bigband = bbandlike * bbandlis,
-                 Classic_Rock = croldlike * croldlis,
-                 Country = countrylike * countrylis,
-                 Bluegrass = blueglike * blueglis,
-                 Folk = folklike * folklis,
-                 Gospel = hymgoslike * hymgoslis,
-                 Salsa = latlpsallike * latspsallis,
-                 Rap_Hip_Hop = raphiphoplike * raphiphoplis,
-                 Blues_RandB = blurblike * blurblis,
-                 Reggae = reggaelike * reggaelis,
-                 Pop = toppoplike * toppoplis,
-                 Contemp_Rock = controcklike * controcklis,
-                 Indie_Alt = indaltlike * indaltlis,
-                 Dance_Club = danclublike * danclublis,
-                 Metal = hvymtllike * hvymtllis
-          ) %>%  #people are linked to genres that the both like and listen to
-          dplyr::select(id, Classical:Metal) 
-          A <- as.matrix(taste.dat[, 2:ncol(taste.dat)])
-          rownames(A) <- taste.dat$id
-          A <- A[-which(rowSums(A) == 0), ]
-     return(A)
-     }

### 2025-01-26 Functions/tmCA.R

@@ -1,0 +1,25 @@
+tmCA <- function(x) {
+     # Initial matrices and vectors
+     A <- x[ , colSums(x) != 0, drop = FALSE] #dropping isolates
+     At <- t(A) #affiliation matrix transpose
+     r <- nrow(A) #number of rows
+     c <- ncol(A) #number of columns
+     rn <- rownames(A) #row name vector
+     cn <- colnames(A) #column name vector
+     iDr <- diag(1/rowSums(A)) #inverse of row diagonal degree matrix
+     iDc <- diag(1/colSums(A)) #inverse of column diagonal degree matrix
+     # Spectral matrices
+     eig.r <- eigen(iDr %*% A %*% iDc %*% At) #eigendecomposition of centrality weighted row similarity matrix
+     eig.c <- eigen(iDc %*% At %*% iDr %*% A) #eigendecomposition of centrality weighted column similarity matrix
+     eigvec.r <- round(Re(eig.r$vectors), 4) # Extracting real row eigenvectors and rounding
+     eigvec.c <- round(Re(eig.c$vectors), 4) # Extracting real column eigenvectors and rounding
+     eigval.r <- round(Re(eig.r$values), 4) # Extracting real row eigenvalues and rounding
+     eigval.c <- round(Re(eig.c$values), 4) # Extracting real column eigenvalues and rounding
+     rownames(eigvec.r) <- rn
+     rownames(eigvec.c) <- cn
+     return(list(eigvec.r = eigvec.r, 
+                 eigvec.c = eigvec.c, 
+                 eigval.r = eigval.r,
+                 eigval.c = eigval.c)
+     )
+}

### 2025-01-26 Reviews and Response/editor-letter.tex
Updates from Overleaf

\documentclass[]{letter}
\begin{document}

Dear Author,

Thank you for submitting your manuscript to the special issue of Social Networks on “Duality.” To assist with its evaluation, we enlisted two reviewers with significant expertise and a direct interest in the focal topic of your paper. We find their comments highly relevant and constructive. We have also read your paper independently, approaching it from the perspective of a general reader of Social Networks who may not have direct experience in your specific area of interest but would be drawn to innovative work that advances our understanding of network phenomena and – more specifically, network measurement and modeling.

Overall, we believe that incorporating the insightful feedback provided by the reviewers will significantly strengthen your manuscript and make it more compelling. Therefore, we are pleased to invite you to revise and resubmit your work for further consideration. Please note, however, that the opportunity to revise your manuscript does not guarantee its inclusion in the special issue.

We generally agree with the reviewers’ comments and strongly encourage you to take all of them into account, with a special focus on the key points I highlight below.

Reviewer 1 (R1) clearly appreciates the goals of your work and endorses its overall objectives. However, R1 raises concerns about potential errors in the matrix you analyze (specifically, the “Southern women data”). Since this is a canonical dataset, it is crucial that the version you use aligns with those employed in the studies you reference as benchmarks. If there are any discrepancies, please explain how and why the data you used differ. R1 offers some helpful advice on how to address this.

R1 also points out the need for further discussion on the odd and even iterations of the “reflective algorithm.” More generally, R1 finds that your "Discussion and Conclusions" section is much stronger and clearer than the Introduction. We share this view and encourage you to revisit your Introduction, clarifying the paper’s contribution and highlighting its novelty compared to existing work.

Reviewer 2 (R2) provides similar feedback, emphasizing the need for greater clarity around the distinct contribution of your research and its implications for future studies of social networks and beyond. To summarize R2's key point: What will readers learn from your paper that they did not know before? And what will they be able to do after reading your paper that they could not do before? I believe you can address these questions directly and convincingly. Both reviewers encourage you to make these points more explicit, especially in the Introduction and Motivation sections.

R2 also suggests broadening the context of your work by acknowledging additional relevant research on two-mode networks (2MN). For instance, consider reviewing the dual projection approach proposed by Everett and Borgatti in their 2013 paper published in a special issue of Social Networks on 2MN, as well as the hierarchy of dependence relations discussed by Robins and Pattison in the same special issue. Additionally, the more recent work of Lerner and Lomi (2022, Journal of Complex Networks), which uses the same “Southern Women” data, offers a framework that resonates with your own – albeit in a different modeling framework.

Finally, we agree with R2 that the link between your work and Breiger’s duality principle remains somewhat underdeveloped in the current manuscript. Strengthening this connection would greatly enhance the theoretical depth of your paper and align it more convincingly with the theme of the special issue.

We would like to bring to your attention a limited number of notational issues that may benefit from some clarification in the revised draft.

- p. 4 of "the sum of the q-1 centralities" is a little confusing. If we correctly understand your point, q-1 refers to the iteration rather than how many centralities are being summed. Perhaps this could be that could be reworded.
- Notation in eq. 19 and 20 seems slightly awkward - the left-hand side should somehow show that those are the transformed values, not the same as on the right-hand side (which currently is written the same way).
- For eq. 21 and 22, it might be helpful to note how the eigenanalysis of the two product matrices relates to the SVD of the original.

In conclusion, we believe the reviewers have provided several valuable and actionable recommendations that will help bring your manuscript to its next stage of development. We hope you agree that addressing their feedback will ultimately enhance the quality and impact of your work.

Thank you again for considering Social Networks as a potential outlet for your research. We look forward to reviewing your revised submission.

Sincerely,

Alessandro Lomi and Philippa Pattison
Executive Guest Editors, Social Networks
\end{document}

### 2025-01-26 Reviews and Response/response-memo.tex

\documentclass[]{letter}
\begin{document}

\textbf{Editors}

\textsc{Reviewer 1}

\underline{State key points and conclusions at the beginning}:

\textbf{Reviewer}: \textit{In terms of its overall organization, your paper is (probably, necessarily) complicated. For me, the discussion/conclusion (Section 5) is much more clear and powerfully written than is the abstract and Introduction (Section 1). As you think about improving the paper, I hope you will find ways to assert the main Section 5 points crisply at the beginning of the paper.}

\textbf{Response}:

\underline{Add distinction between Bonacich and CA to the highlights}:

\textbf{Reviewer}: \textit{(For example, I would like to see a ``Highlights" bullet point that says that Bonacich centrality is useful for revealing the position of the nodes in a center-periphery pattern, to the extent one exists, whereas CA is useful for revealing the presence of different groups or communities in the data, to the extent that the dual network is not homogeneous.)}

\textbf{Response}:

\underline{Data errors in the Southern Women two-mode matrix}:

\textbf{Reviewer}: \textit{Importantly, there are 4 errors in the matrix (``Southern Women data") that you analyze. For this dataset, you (in the note to Table 1) cite Doreian et al (2004, Table 4). Among the attendees at Event E4, you show Frances, Eleanor, and Ruth. However they did not attend this event. You also show Nora as absent from Event E8; however, Nora attended E8. These four errors are discrepant in comparison to your cited source (Doreian et al, 2004, Table 4), and they are discrepant as well in comparison to the original source (Davis, Gardener, and Gardener, 1941, p. 141, which has events in the same order as in your Table 1, and a slightly different order of women; this table is reproduced in Freeman, 2003, p. 42).}

\textbf{Response}:

\underline{Appropriateness of the term ``reflective centrality''}:

\textbf{Reviewer}: \textit{I have some resistance to your term "reflective centrality" (Section 2 title) to characterize the scores of the respective modes - precisely because, as you importantly show, in the limit the reflective scores are first dimensions of a CA, and (as you also importantly emphasize) the first CA dimension is about finding groups, not about ordering nodes. Hidalgo and Hausmann (2009) do not use the term "centrality" for the reflective scores they define and compute. I wonder whether ``reflective scores" might be a better term; I'm not sure, and simply ask you to think about it.}

\textbf{Response}:

\underline{Differences between odd and even iterations}:

\textbf{Reviewer}: \textit{What complicates things, as you know, is the difference between odd and even iterations of the reflective algorithm (see your comments in the first paragraph on p. 4, as well as comments of H \& H). One point you might consider adding is that, for the first even iterations (iterations 2, 4, 6, 8, and 10), the magnitude (absolute value) of the correlation of reflected scores with Bonacich centrality (for persons, and for groups) tends to increase with subsequent iterations (and, in the limit, is .682 for persons by iteration 40 for persons), whereas, for the first odd iterations (iterations 3, 5, 7, 9, 11), the magnitude of the correlation decreases (toward a limiting value, once again, of .682 by iteration 41). Can you relate this to the differences between odd and even iterations that you (and H\&H) discuss?}

\textbf{Response}:

\underline{Missed connection between Bonacich and CA}:

\textbf{Reviewer}: \textit{One of the high points of your paper - bringing great clarity, and of great usefulness to practitioners - is your Fig. 2, showing how the first (person and group) dimensions of the CA reveal ``groupiness" (my poor word choice), whereas the dual Bonacich centrality scores reveal the extent to which a center-periphery pattern obtains in the data. A point you miss here (however, I apologize if you do say this somewhere) is that the second dimension of Bonacich centrality (that is, the second dimension of the SVD of an affiliation matrix like Table 1) looks very much like the first CA dimension, emphasizing communities rather than center-periphery. This comment, like that of the previous paragraph, I think helps you to show how highly entangled are the concepts of reflective scores and centrality, and (therefore) how important it is to recognize when each is appropriate.}

\textbf{Response}:

\underline{Reversing column ordering on some figures}:

\textbf{Reviewer}: \textit{(In Fig. 2, both parts (a) and (b), nothing substantive or mathematical would change if you keep the row ordering you show, but reverse the column ordering. However, reversing the column ordering would help the reader to see, in panel (a), the groups, and, in panel (b), the center and the periphery. A similar comment applies to several of your other figures.)}

\textbf{Response}:

\underline{Confusing equations}:

\textbf{Reviewer}: \textit{What confused me about eqs. (23) and (24) is that I don't think the (respective) expressions in parentheses are symmetric matrices, and so I don't see how to take an eigenvector. (Each of the rows, but not the columns, sum to 1.)}

\textbf{Response}:

\underline{Confusing Figures}:

\textbf{Reviewer}: \textit{Figs 3, 4, 5, and 6 each have multiple panels (referred to as a, b, c, etc), but, in the current version of the paper, it is difficult or impossible to figure out which part of the figure has which label (a, b, c, …). Hopefully the formatting here can be improved.}

\textbf{Response}:

\underline{Missing Freeman citation}

\textbf{Response}:

\newpage
\textsc{Reviewer 2}

\underline{Insufficient engagement with the duality idea}

\textbf{Reviewer}: \textit{In my opinion, though, the paper does not take Breiger's argument in due consideration, even more as the piece is aimed at contributing to a special issue on ``Duality in social networks". The latter principle seems to remain in the background, while it should explicitly inform most of the analyses presented in this work.}

\textbf{Response}:

\underline{More extensive introduction and background regarding previous uses of CA and MCA}

\textbf{Reviewer}: \textit{I suggest extending the introduction, enriching the discussion of the background as for CA in the context of two-mode networks. Given the importance of the technique and its variants (like MCA), it is surprising that the authors devote such a limited room to the relevant framework. Consider that the discussion of CA in the context of two-mode networks seems not comprehensive as it misses the use of MCA in the same context \ldots}

\textbf{Response}:

\underline{More consideration of the visualization aspects of the CA/MCA}

\textbf{Reviewer}: \textit{Further, as the introduction states, there is an attempt at seemingly diminishing the power of visualization as a key feature of CA, as if it were a secondary concern in this domain. I think the authors should first consider carefully the actual utility of data visualization via CA/MCA and then also mind that, although their work aims to move ``beyond the focus on data visualization" (p. 2), they in fact make use of visual representation quite heavily in this paper. This seems to me rather misleading.}

\textbf{Response}:

\underline{Concerns over the interpretation of CA biplot distances}

\textbf{Reviewer}: \textit{As regards CA and its ``conventional" - so to speak - interpretation which the authors criticize and want to go beyond, I disagree on the idea that using CA in a way different than the authors' would be less useful/informative or, even worse, ``off the mark". Here I specifically refer to a passage on p. 12 \ldots While I would agree that one can interpret the distances between units - and, even more aptly, between the units and the axes origin - on a CA plot in terms of distances from an independence condition, I would not dismiss the importance of evaluating, in the above fashion, how ``people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups". This has clear substantive value for interpretation of the data thanks to use of CA.}

\textbf{Response}:

\underline{Insufficient discussion of positional equivalence}

\textbf{Reviewer}: \textit{As for the paper's argument, one may ask what's new in the current proposal and what's the point in it, beyond already known things. This issue appears to affect, for example, the opportunity to analyse the forms of structural or regular equivalence in two-mode networks. This theme is paid very little attention, though. And yet, it is very relevant to the paper - as the authors seem at least to know, as they mention it (one time only, on p. 14). In the course of the paper there are some passages where the point would be relevant, as is the case with the bottom of p. 10, when the authors recall the works of Doreian, Batagelj and Ferligoj (2004), Kovács (2010) and Lizardo (2024), and assert that their own results are comparable to the latter's \ldots This is clearly a matter of positional equivalence, which deserves more room in the paper, in my opinion.}

\textbf{Response}:

\underline{Insufficient discussion of blockmodeling approaches}

\textbf{Reviewer}: \textit{In addition, I would like to stress that the article sets forth an argument that, in the beginning, revolves around centrality, but ends with a more detailed discussion about how to assess relational similarity in two-mode networks. Here, there seems to be a lack of consideration regarding other relevant approaches. Blockmodeling, for instance, is not duly considered, although it is well known as a sort of gold standard in the analysis of equivalences - particularly generalized equivalences (I suggest seeing the book by Doreian, Batagelj, and Ferligoj, 2005, in addition to the Social Networks article by the same authors that is cited in the paper) \ldots I would like the authors to pay more attention to this absence in their discussion. More explicitly, what does the authors' proposal add to the analysis of structural/regular/generalized equivalences as it is already treated via blockmodeling (and also via MCA)?}

\textbf{Response}:

\underline{Other analytic strategies}

\textbf{Reviewer}: \textit{I also think about other analytic strategies like Cluster Correspondence Analysis (Cluster CA) applied to two-mode networks, for which I may suggest reading and referencing relevant works (such as D'Ambrosio, Serino and Ragozini, 2021).}

\textbf{Response}:

\underline{Better highlights section}

\textbf{Reviewer}: \textit{- The Highglights section should be improved in the sense that it looks now like a sort of abstract, while it needs to appear as a bullet-like summary of the paper that will make its argument better understood at the first glance}

\textbf{Response}:

\underline{Unclear figure}

\textbf{Reviewer}: \textit{- Figure 3 is not clear, in particular it should have a proper layout where the content of the current insets would be shown separately (now it is a bit confusing)}

\textbf{Response}:

\end{document}

### 2025-01-26 Reviews and Response/reviewer1.tex

\documentclass[]{letter}
\begin{document}
\textbf{Reviewer 1}

Given the proliferation of methods for ranking nodes in a network, along with the resultant confusion that confronts the analyst, there should be a special place in heaven for papers that show how seemingly different methods are related (and conversely). This paper resides in that special place. In economics a major paper published in PNAS by Hidalgo and Hausmann (2009), cited already over 3,700 times, develops and works with a measure of economic complexity - what they call the Method of Reflections. Their article (including its supplementary material) does not contain the words ``centrality," ``eigenvector," or ``correspondence analysis," and yet you (relying in part on your synthesis of a burgeoning multidisciplinary literature, and in part on your own novel insights) show how H\&H's reflective method is related to these other concepts, including how, in the limit, their method reproduces the first dimension of a correspondence analysis (CA), and therefore provides analysts with a ``new" way to understand what CA is doing, and how it works. A major theme in your paper is the similarities and (in particular) differences between Bonacich centrality scores (1991, Bonacich \& Lloyd 2001), a bedrock of network analysis, and correspondence analysis.

Among the insights in this paper that should be especially valuable to network analysts and that should affect research practice is that the CA plot captures, not that people who attend many of the same events will appear close in the plot (and conversely for similar groups), but rather that people who are surprisingly similar (from the point of view of a null model such as independence) will be close in the plot, after taking people's activity levels and sizes of the groups they belong to into account (and dually for groups; your p. 12). Another such insight concerns when (respectively) Bonacich centrality vs CA is most appropriate to be used.

Your paper provides an exploration of the concept of duality that is profound, in that the various methods that you bring together and relate to one another all are based on the idea that each "mode" of a two-mode network is constituted by the other mode. All the methods you relate are ways to model this co-constitution. Correspondence analysis and related techniques are core methods for the analysis of duality. In this sense, your paper is a very good fit indeed to this special issue.

In terms of its overall organization, your paper is (probably, necessarily) complicated. For me, the discussion / conclusion (Section 5) is much more clear and powerfully written than is the abstract and Introduction (Section 1). As you think about improving the paper, I hope you will find ways to assert the main Section 5 points crisply at the beginning of the paper. (For example, I would like to see a "Highlights" bullet point that says that Bonacich centrality is useful for revealing the position of the nodes in a center-periphery pattern, to the extent one exists, whereas CA is useful for revealing the presence of different groups or communities in the data, to the extent that the dual network is not homogeneous.)

Now I will turn to more specific comments. Importantly, there are 4 errors in the matrix ("Southern Women data") that you analyze. For this dataset, you (in the note to Table 1) cite Doreian et al (2004, Table 4). Among the attendees at Event E4, you show Frances, Eleanor, and Ruth. However they did not attend this event. You also show Nora as absent from Event E8; however, Nora attended E8. These four errors are discrepant in comparison to your cited source (Doreian et al, 2004, Table 4), and they are discrepant as well in comparison to the original source (Davis, Gardener, and Gardener, 1941, p. 141, which has events in the same order as in your Table 1, and a slightly different order of women; this table is reproduced in Freeman, 2003, p. 42). The data table in Kovacs (2010, p. 205) is correct (it is the same table as in Doreian et al), whereas the data table in Lizardo (2024, given in the qmd file in the supplementary material published for that article, lines 22-48 of the script) has the same 4 errors as in your manuscript.

So you will need to correct your Table 1 and, more importantly, all the subsequent analyses. In preparing this review, I have reproduced many of your analyses on the correct data, and it seems to me that, for the most part, the differences are not substantial and not substantive. Nonetheless, using the correct data is of course a requirement.

I have some resistance to your term "reflective centrality" (Section 2 title) to characterize the scores of the respective modes - precisely because, as you importantly show, in the limit the reflective scores are first dimensions of a CA, and (as you also importantly emphasize) the first CA dimension is about finding groups, not about ordering nodes. Hidalgo and Hausmann (2009) do not use the term "centrality" for the reflective scores they define and compute. I wonder whether "reflective scores" might be a better term; I'm not sure, and simply ask you to think about it.

What complicates things, as you know, is the difference between odd and even iterations of the reflective algorithm (see your comments in the first paragraph on p. 4, as well as comments of H \& H). One point you might consider adding is that, for the first even iterations (iterations 2, 4, 6, 8, and 10), the magnitude (absolute value) of the correlation of reflected scores with Bonacich centrality (for persons, and for groups) tends to increase with subsequent iterations (and, in the limit, is .682 for persons by iteration 40 for persons), whereas, for the first odd iterations (iterations 3, 5, 7, 9, 11), the magnitude of the correlation decreases (toward a limiting value, once again, of .682 by iteration 41). Can you relate this to the differences between odd and even iterations that you (and H\&H) discuss?

One of the high points of your paper - bringing great clarity, and of great usefulness to practitioners - is your Fig. 2, showing how the first (person and group) dimensions of the CA reveal ``groupi-ness" (my poor word choice), whereas the dual Bonacich centrality scores reveal the extent to which a center-periphery pattern obtains in the data. A point you miss here (however, I apologize if you do say this somewhere) is that the second dimension of Bonacich centrality (that is, the second dimension of the SVD of an affiliation matrix like Table 1) looks very much like the first CA dimension, emphasizing communities rather than center-periphery. This comment, like that of the previous paragraph, I think helps you to show how highly entangled are the concepts of reflective scores and centrality, and (therefore) how important it is to recognize when each is appropriate.

(In Fig. 2, both parts (a) and (b), nothing substantive or mathematical would change if you keep the row ordering you show, but reverse the column ordering. However, reversing the column ordering would help the reader to see, in panel (a), the groups, and, in panel (b), the center and the periphery. A similar comment applies to several of your other figures.)

What confused me about eqs. (23) and (24) is that I don't think the (respective) expressions in parentheses are symmetric matrices, and so I don't see how to take an eigenvector. (Each of the rows, but not the columns, sum to 1.)

Figs 3, 4, 5, and 6 each have multiple panels (referred to as a, b, c, etc), but, in the current version of the paper, it is difficult or impossible to figure out which part of the figure has which label (a, b, c, …). Hopefully the formatting here can be improved.

Reference not cited in your paper:

Freeman, L.C. (2003). "Finding Social Groups: A Meta-Analysis of the Southern Women Data." Pp. 39-77 in Dynamic Social Network Modeling and Analysis: Workshop Summary and Papers, edd. R. Breiger, K. Carley, P. Pattison. Washington, DC: National Academies Press.

\end{document}

### 2025-01-27 Reviews and Response/reviewer2.tex

\documentclass[]{letter}
\begin{document}
\textbf{Reviewer 2}

The paper titled ``The Correspondence Analysis of Two-Mode Networks Revisited" deals with a classic formal argument regarding centrality in two-mode networks, i.e. that of using information on the other mode to compute centrality for a given mode of the network. It proposes a revision of Bonacich's proposal for calculating two-mode centrality that deploys Correspondence Analysis (CA) to obtain dual centrality scores in a ``reflective" fashion. In addition, it provides an attempt to revise the use of CA for analysing the relational patterns of two-mode networks.

The paper is interesting and full of accurate formal reasoning around the above topics, recalling back Bonacich's original argument and comparing it with a version of reflective centrality, then furtherly developing the discussion by taking into account the ``generalized relational similarity" (GRS) approach to two-mode networks.

However, although the paper does much effort to illustrate the potentiality of CA to better understand the relational patterns of two-mode networks, with a rich and informative suite of commentary and graphical representations, I've got several reservations to make about this work - mainly issues of argument strength, clarity, and lack of reference to other relevant topics - and I shall describe them as follows.

The paper mentions the important legacy with Breiger's seminal work on duality in affiliation (two-mode) networks, as it provides a key grounding to the elaboration of centrality in such networks, stressing the bond between the two modes in analysing them. In my opinion, though, the paper does not take Breiger's argument in due consideration, even more as the piece is aimed at contributing to a special issue on ``Duality in social networks". The latter principle seems to remain in the background, while it should explicitly inform most of the analyses presented in this work.

I suggest extending the introduction, enriching the discussion of the background as for CA in the context of two-mode networks. Given the importance of the technique and its variants (like MCA), it is surprising that the authors devote such a limited room to the relevant framework. Consider that the discussion of CA in the context of two-mode networks seems not comprehensive as it misses the use of MCA in the same context - indeed, by merely citing it (one time) without discussing it properly. Both CA and MCA lead to appreciate visually (but also via the set of measures available in each tool) the similarities/differences between nodes and group of nodes in one mode by analysing their relational patterns with respect to the other mode.
Further, as the introduction states, there is an attempt at seemingly diminishing the power of visualization as a key feature of CA, as if it were a secondary concern in this domain. I think the authors should first consider carefully the actual utility of data visualization via CA/MCA and then also mind that, although their work aims to move ``beyond the focus on data visualization" (p. 2), they in fact make use of visual representation quite heavily in this paper. This seems to me rather misleading.

As regards CA and its ``conventional" - so to speak - interpretation which the authors criticize and want to go beyond, I disagree on the idea that using CA in a way different than the authors' would be less useful/informative or, even worse, ``off the mark". Here I specifically refer to a passage on p. 12:
"This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members [in common?] will appear close in the plot. But this (still common) interpretation is off the mark. What the CA correspondence plot distance captures is, instead, people and groups that are surprisingly similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account. Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups".
While I would agree that one can interpret the distances between units - and, even more aptly, between the units and the axes origin - on a CA plot in terms of distances from an independence condition, I would not dismiss the importance of evaluating, in the above fashion, how ``people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups". This has clear substantive value for interpretation of the data thanks to use of CA.

As for the paper's argument, one may ask what's new in the current proposal and what's the point in it, beyond already known things. This issue appears to affect, for example, the opportunity to analyse the forms of structural or regular equivalence in two-mode networks. This theme is paid very little attention, though. And yet, it is very relevant to the paper - as the authors seem at least to know, as they mention it (one time only, on p. 14). In the course of the paper there are some passages where the point would be relevant, as is the case with the bottom of p. 10, when the authors recall the works of Doreian, Batagelj and Ferligoj (2004), Kovács (2010) and Lizardo (2024), and assert that their own results are comparable to the latter's: "This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to similar groups, where group similarity is defined dually as having members in common who are themselves similar." This is clearly a matter of positional equivalence, which deserves more room in the paper, in my opinion.

In addition, I would like to stress that the article sets forth an argument that, in the beginning, revolves around centrality, but ends with a more detailed discussion about how to assess relational similarity in two-mode networks. Here, there seems to be a lack of consideration regarding other relevant approaches. Blockmodeling, for instance, is not duly considered, although it is well known as a sort of gold standard in the analysis of equivalences - particularly generalized equivalences (I suggest seeing the book by Doreian, Batagelj, and Ferligoj, 2005, in addition to the Social Networks article by the same authors that is cited in the paper).

Consider two paragraphs on p. 14. Above, it is highlighted that CA, as it is shown in the plot of Figure 5(c), namely in the authors' usage, can show that some ``core events […] are more inclusive of peripheral members" while others ``are more "cliquish," including only core members". Note that this way of assessing the partitions is akin to that provided by blockmodeling for two-mode networks when dealing with core-periphery structures. On the same p. 14, below (in section 4.6), when GRS and the ``SimRank" measure are considered, the theme of positional equivalences (mainly regular equivalence) is chiefly relevant (but not mentioned) to the assertions ``People are similar if they belong to similar groups" and ``Groups are similar if they share similar members". I would like the authors to pay more attention to this absence in their discussion. More explicitly, what does the authors' proposal add to the analysis of structural/regular/generalized equivalences as it is already treated via blockmodeling (and also via MCA)?

I also think about other analytic strategies like Cluster Correspondence Analysis (Cluster CA) applied to two-mode networks, for which I may suggest reading and referencing relevant works (such as D'Ambrosio, Serino and Ragozini, 2021).

As a more general comment, I think that the paper looks like a thorough discussion of different ways of calculating and representing recursive centrality measures for two-mode networks. This is fine, unless one considers that the attempt turns out to be a sort of juxtaposition of assessments of each approach to centrality and position in two-mode networks.

The problem is that the recurring comparison between the different approaches, albeit interesting, is poorly leveraged to sustain the article's argument. It seems that by considering each step of that process of comparing different approaches one misses the whole picture. (Consider also the fact that the discussion/conclusion section is still not sufficiently elaborate, and this does not help in this regard.)

In sum, my contention is that the authors' argument should be strengthened to show what worth is reading this piece beyond already available literature. I think the paper is instructive as it discusses analytically the relevant topics, but it should be clearer about how this work will have an impact on the debate regarding CA, two-mode networks and the search for ways of better assessing the relational patterns in such networks.

Other minor points are as follows:

- The Highglights section should be improved in the sense that it looks now like a sort of abstract, while it needs to appear as a bullet-like summary of the paper that will make its argument better understood at the first glance

- Figure 3 is not clear, in particular it should have a proper layout where the content of the current insets would be shown separately (now it is a bit confusing)

\end{document}

### 2025-01-27 main-R0.tex

--- 2024-07-13 main.tex
@@
 generalized blockmodel with events 1-6 separated from events 10-15 by events 7-9 (see also \citet{kovacs2010generalized}).
-Figure~\ref{fig:pg-reflections} shows that $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
+Clearly, $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
 Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering the average of the average number of memberships of its members (and so forth).
@@
 9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
+As we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix for both the row and column objects.
 

### 2025-01-27 main-R1.tex

--- 2025-01-27 main-R0.tex
@@
 The rest of the paper is organized as follows.
-In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
+In the next section, I motivate the use of CA to produce a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
 This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes.
@@
 I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers the core-periphery structure of the two-mode network instead.
+I contrast this ordering to that provided by the Bonacich dual centrality scores, which, in contrast to the CA ordering, reveals a dual community partition and recovers the core-periphery structure of the two-mode network instead.
 In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

### 2025-01-27 main-R1.tex
Updates from Overleaf

@@
 
-\begin{table}
-\caption{Southern Women Data.}
-\label{tab:southern}
-\begin{tabular}[]{lcccccccccccccc}
-\hline
- & E1 & E2 & E3 & E4 & E5 & E6 & E7 & E8 & E9 & E10 & E11 & E12 & E13 & E14 \\
- \hline
-EVELYN & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-LAURA & 1 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-THERESA & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-BRENDA & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-CHARLOTTE & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-FRANCES & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-ELEANOR & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-RUTH & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-VERNE & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 1 & 0 & 0 \\
-MYRA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 \\
-KATHERINE & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-SYLVIA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-NORA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 \\
-HELEN & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 \\
-OLIVIA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-FLORA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-PEARL & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-DOROTHY & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-\hline
-\end{tabular}
-\textsuperscript{Note: Rows ordered according to the generalized
-blockmodeling solution of Doreian et al. (2004, Table 4).} \\
-\end{table}
+\include{southern-women-corrected}
 

### 2025-01-27 Reviews and Response/response-memo.tex
changed paper

@@
-\documentclass[]{letter}
+\documentclass{article}
+\usepackage{graphicx}
 \begin{document}
@@
 
-\textbf{Response}:
+\begin{figure}[h!]
+    \includegraphics[width=1.0\textwidth]{Reviews and Response/sw-original.png}
+\end{figure}
+
+\textbf{Response}: Thanks for calling attention to this---rather embarrassing---error; it's definitely a cautionary tale against manually copying data.
+I have corrected the data, and as you can see in the analysis file included as supplementary material, I now extract the Southern Women data directly from the \texttt{networkdata} package.
+My table now matches those in the references you mentioned exactly.
+Note, however, that while you are correct that Frances, Eleanor, and Ruth did not attend E4, Nora did not attend E8, at least according to the published tables in Doreian et al (2004), Kovacs (2010), and the Figure 1 from Freeman (2003), shown above; in these papers, as you can see, Nora's event set is \{E6, E7, E9, E10, E11, E12, E13, E14\}.
+I think the confusion comes from the fact that Freeman used a non-chronological order of events (as is evident from the date labels in the figure) that emphasizes the group structure of the women.
+This non-chronological ordering leads to a different labeling of events than when events are arranged in chronological order; this is the ordering of events used by Doreian et al.
+2004 and Kovacs (2010).
+In this event labeling, Nora does not attend E8 (the event that took place on 9/16.) However, note that this is obviously \textit{not} the ordering of events in the original Breiger (1974)---nor for that matter the earlier Doreian (1979) piece in \textit{Social Networks} that used Q-Analysis.
+In this \textit{chronological} labeling of events, Nora does attend ``E8" (the event that took place on 5/19), as evident in Breiger (1974, p.
+186, fig.
+2a).
+In any case, to forestall any confusion, and given that the in the \texttt{networkdata} package, the \texttt{igraph} object containing the Southern Women data come ordered using the Freeman (2003) ordering, I'm sticking to the non-chronological event labeling.
+I include a very brief discussion in the paper clarifying this issue.
 

### 2025-01-27 main-R1.tex
changed paper

@@
 \subsection{Empirical Example} \label{subsec:ex}
-Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons in the Southern Women data \citep{davis1941}, using a bump chart to plot the rank order trajectory of persons across reflections.
+The following analyses use the classic Southern Women data \citep{davis1941} as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, wang2009exponential, everett2018measuring, batagelj2022analysis, lizardo2024two}.
+As we can see, novel insights can still be garnered from these old data.
+
+Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections.
 The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.

### 2025-01-27 analysis-R1.qmd
new stuff

---
title: "Correspondence Analysis of Two Mode Networks"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
---

# Setup
```{r setup}
   knitr::opts_chunk$set(include = TRUE, 
                       echo = FALSE, 
                       warning = FALSE, 
                       message = FALSE)
   library(FactoMineR)
   library(here)
   library(igraph)
   library(kableExtra)
   library(networkdata)
```

# Southern Women Data Table
```{r HTML data table}
   g <- southern_women
   A <- as.matrix(as_biadjacency_matrix(g))
   p.ord <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRNA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
   A <- A[p.ord, ]
   rownames(A) <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
   colnames(A) <- paste("E", 1:14, sep = "")
   kbl(A, align = "c", format = "latex", 
       booktabs = TRUE, linesep = "") %>% 
      kable_minimal(full_width = F, font_size = 10) %>% 
      save_kable(here("southern-women-corrected.tex"))
```

```{r}
   png(file = here("Plots", "ca-southern-women.png"), width=700, height=700)
   CA(A)
   dev.off()
```

### 2025-01-27 analysis.qmd
new stuff

@@
 ---
-title: "Two Mode Networks and Relational Similarity"
+title: "Correspondence Analysis of Two Mode Networks"
 author: "Omar Lizardo"
@@
    library(patchwork)
+   library(networkdata)
 ```
@@
 ```{r HTML data table}
-   source(here("Functions", "sw.dat.R"))
-   A <- sw.dat()
-   A <- A[sort(rownames(A)), ]
-   kbl(A, align = "c", format = "html") %>% 
-      kable_minimal(full_width = F, font_size = 26) %>% 
-      row_spec(1:18, extra_css = "border-bottom: 0px;line-height: 20px;") %>% 
-      save_kable(here("southern-women.html"))
+   g <- southern_women
+   A <- as.matrix(as_biadjacency_matrix(g))
+   p.ord <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRNA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
+   A <- A[p.ord, ]
+   colnames(A) <- paste("E", 1:14, sep = "")
+   kbl(A, align = "c", format = "latex", 
+       booktabs = TRUE, linesep = "") %>% 
+      kable_minimal(full_width = F, font_size = 10) %>% 
+      save_kable(here("southern-women-corrected.tex"))
 ```

### 2025-01-27 analysis-R1.qmd
make it so

@@
                        message = FALSE)
+   library(conflicted)
    library(FactoMineR)
@@
    library(networkdata)
+   conflicted::conflicts_prefer(dplyr::filter)
 ```
@@
 ```
+
+```{r Reflections Bump Chart (Persons)}
+   source(here("Functions", "reflections.R"))
+   source(here("Functions", "ref.long.dat.R"))
+   evens <- function(x) subset(x, x %% 2 == 0)
+   odds <- function(x) subset(x, x %% 2 != 0)
+   a <- reflections(A, iter = 26)
+   p.e.r <- a$p.r[, evens(1:20)]
+   g.e.r <- a$g.r[, evens(1:20)]
+   b <- ref.long.dat(a = p.e.r, b = g.e.r, max.iter = 26)
+   plot.dat <- b$person
+   p <- ggplot(data = filter(plot.dat, n.iter < 18),
+               mapping = aes(x = n.iter, y = ref, color = person))
+   # geoms
+   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
+   p <- p + geom_point(size = 5)
+   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
+                                    aes(x = n.iter, label = person), 
+                                    size = 5, nudge_x = -3)
+   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 18), 
+                              aes(x = n.iter, label = person), 
+                              size = 5, nudge_x = 0.01)
+
+   # fixings
+   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
+   p <- p + scale_x_continuous(limits = c(-2, 20), breaks = evens(2:16))
+   p <- p + labs(x = "Reflection")
+   p <- p + scale_y_reverse()
+   p <- p + theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 16)
+        )
+   set.seed(145)
+   c <- sample(cols25(25), 18, replace = FALSE)
+   p1 <- p + scale_color_manual(values=as.vector(c))
+   png(file = here("Plots", "p-reflections.png"), width=600, height=625)
+   p1
+   dev.off()
+```
+
+```{r Reflections Bump Chart (Groups)}
+   plot.dat <- b$group
+   p <- ggplot(data = filter(plot.dat, n.iter < 14),
+               mapping = aes(x = n.iter, y = ref, color = group))
+   #geoms
+   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
+   p <- p + geom_point(size = 5)
+   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
+                                    aes(x = n.iter, label = group), 
+                                    size = 6, nudge_x = -1)
+   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 14), 
+                              aes(x = n.iter, label = group), 
+                              size = 6, nudge_x = -0.01)
+   #fixings
+   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
+   p <- p + scale_x_continuous(limits = c(1, 14), breaks = evens(2:12))
+   p <- p + labs(x = "Reflection")
+   p <- p + scale_y_reverse()
+   p <- p + theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 16)
+        )
+   set.seed(145)
+   c <- sample(cols25(25), 18, replace = FALSE)
+   p2 <- p + scale_color_manual(values=as.vector(c))
+   png(file = here("Plots", "g-reflections.png"), width=600, height=625)
+   p2
+   dev.off()
+```

### 2025-01-27 main-R1.tex
Updates from Overleaf

--- 2025-01-27 main-R1.tex
@@
 \subsection{Empirical Example} \label{subsec:ex}
-Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons in the Southern Women data \citep{davis1941}, using a bump chart to plot the rank order trajectory of persons across reflections.
+The following analyses use the classic Southern Women data \citep{davis1941} as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, everett2018measuring, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}.
+As we can see, novel insights can still be garnered from these old data.
+
+Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections.
 The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.

### 2025-01-27 main-R1.tex

--- 2025-01-27 main-R1.tex
@@
 \subsection{Empirical Example} \label{subsec:ex}
-Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons in the Southern Women data \citep{davis1941}, using a bump chart to plot the rank order trajectory of persons across reflections.
+\subsubsection{Souther Women Data}
+The following analyses use the classic Southern Women data \citep{davis1941} (DGG) as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, everett2018measuring, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}.
+As we can see, novel insights can still be garnered from these old data.
+
+One key source of confusion in the mini-literature that exists around this dataset concern the order of presentation of the rows and columns.
+The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous.
+It \textit{could} be treated as nominal, but the original DGG table includes event \textit{dates}, which means that it could also be treated as an ordered variable.
+As \citet{freeman2003finding} notes, in the canonical original table both the row and column modes had be re-ordered in presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} included in chronological order.
+To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "woman 1" or woman 11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns, but in fact, as shown in~\ref{tab:southern}, it's the event that occurred on August 3rd, and there were three events that occurred after that one (E4, E8, and E13) in the DGG numbering.
+
+To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, with the events ordered chronologically.
+This means that the arbitrary ordinal labeling of the rows (e.g., "w1, "w2", "w3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding}.
+Different versions of the dataset reproduce different versions of the labeling.
+Of course, at the end of the day, the specific labeling does not really matter, since we are only interested in the structural signature of the rows and columns.
+The labeling does matter, however, when trying to reconcile claims across publications from the narrative, since "Woman 1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to correspond to "Woman 1" or "E8" in the other.
+To forestall confusion, in what follows I use the numeric labeling corresponding to the original DGG data table as reproduced in \citet[Figure 1]{freeman2003finding}, and shown in Table~\ref{tab:southern}, which include both the arbitrary ordered labels (e.g., "w1" or "E1") and both the woman's name and the event date,, with "Myrna" rebaptized to "Myra" to correspond to the original DGG table \citep{freeman2003finding}.
+
+\subsubsection{Reflective Score Trajectory}
+
+Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections.
 The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.

### 2025-01-27 main-R1.tex

--- 2025-01-27 main-R1.tex
@@
 
-\include{southern-women-corrected}
-
+\begin{figure}[ht!]
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
+    \centering
+    \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
+    \caption{Southern Women data with rows and columns ordered as in the original DGG table.}
+    \label{fig:sw}
+\end{figure}
 \subsection{Empirical Example} \label{subsec:ex}
-The following analyses use the classic Southern Women data \citep{davis1941} as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, wang2009exponential, everett2018measuring, batagelj2022analysis, lizardo2024two}.
+\subsubsection{Souther Women Data}
+The following analyses use the classic Southern Women data \citep{davis1941} (DGG) as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, borgatti2011analyzing, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}.
 As we can see, novel insights can still be garnered from these old data.
 
+One key source of confusion in the mini-literature that exists around this dataset concern the order of presentation of the rows and columns.
+The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous.
+It \textit{could} be treated as nominal, but the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.
+As \citet{freeman2003finding} notes, in the canonical original table both the row and column modes had been re-ordered in presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order in terms of column ordering.
+To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns, but in fact, as shown in~\ref{tab:southern}, it is the event that occurred on August 3rd, and there were three events that occurred after that one (E4, E8, and E13) in the DGG numbering.
+
+To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, but with the events ordered chronologically, which differs from the original DGG Table reproduced in \citet{freeman2003finding}.
+This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding}.
+Different versions of the dataset reproduce different versions of the labeling.
+Of course, at the end of the day, the specific labeling of the nodes does not really matter, since we are usually interested in the structural properties of the rows and columns.
+The even ordering however, does matter for analyses that try to exploit the temporal properties of the data.
+Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to correspond to "W1" or "E8" in the other (e.g., \citet{doreian1979evolution}).
+
+To forestall confusion, in what follows I use the numeric labeling corresponding to the original DGG data table as reproduced in \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{tab:southern}, which include both the arbitrary ordered labels (e.g., "W1" or "E1") and both the woman's name and the event date in parentheses, with "Myrna" rebaptized to her given name of "Myra" to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.
+
+\subsubsection{Reflective Score Trajectory}
 Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections.

### 2025-01-27 analysis-R1.qmd
Ay caramba

@@
      library(FactoMineR)
+     library(ggcorrplot)
+     library(ggplot2)
      library(here)
@@
      library(networkdata)
+     library(stringr)
      conflicted::conflicts_prefer(dplyr::filter)
@@
 ```{r HTML data table}
-   g <- southern_women
-   A <- as.matrix(as_biadjacency_matrix(g))
-   p.ord <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRNA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-   A <- A[p.ord, ]
-   rownames(A) <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-   colnames(A) <- paste("E", 1:14, sep = "")
-   kbl(A, align = "c", format = "latex", 
-       booktabs = TRUE, linesep = "") %>% 
-      kable_minimal(full_width = F, font_size = 10) %>% 
-      save_kable(here("southern-women-corrected.tex"))
+     A <- as.matrix(as_biadjacency_matrix(southern_women))
+     rownames(A) <- str_to_title(rownames(A))
+     colnames(A) <- str_replace(colnames(A), "/", "-")
+     wlab <- paste("W", 1:18, sep = "")
+     wlab <- paste(wlab, rownames(A), sep = " (")
+     wlab <- paste(wlab, ")", sep = "")
+     elab <- paste("E", 1:14, sep = "")
+     elab <- paste(elab, colnames(A), sep = " (")
+     elab <- paste(elab, ")", sep = "")
+     rownames(A) <- wlab
+     colnames(A) <- elab
+     p <- ggcorrplot(t(A), colors = c("white", "white", "red")) 
+     p <- p + theme(legend.position = "none", 
+                  axis.text.y = element_text(size = 16),
+                  axis.text.x = element_text(size = 16, angle = 45, hjust = -0.2),
+                  )
+     p <- p + scale_x_discrete(position = "top") 
+     p <- p + scale_y_discrete(limits = rev)
+     jpeg(file = here("Plots", "southern-women.jpg"), width=800, height=800)
+     p
+     dev.off()
 ```
 
-```{r}
-   png(file = here("Plots", "ca-southern-women.png"), width=700, height=700)
-   CA(A)
+```{r CA of affiliation matrix}
+   ca.res <- CA(A, graph = FALSE)
+   plot.dat <- data.frame(rbind(ca.res$row$coord, ca.res$col$coord)[, 1:2])
+   plot.dat <- cbind(plot.dat, mode = c(rep(1, 18), rep(2, 14)), names = rownames(plot.dat))
+   p <- ggplot(plot.dat, aes(y = Dim.2, x = Dim.1, 
+                             color = factor(mode), label = names))
+   p <- p + geom_vline(xintercept = 0, color = "gray")
+   p <- p + geom_hline(yintercept = 0, color = "gray")
+   p <- p + geom_text_repel(size = 4, max.overlaps = 50)
+   p <- p + theme_minimal() 
+   p <- p + theme(legend.position = "none",
+                  axis.text = element_text(size = 18),
+                  axis.title = element_text(size = 20)
+                  )
+   p <- p + labs(x = "Second Dimension", y = "First Dimension")
+   p <- p + scale_color_manual(values = c('red', 'blue'))
+   png(file = here("Plots", "ca-southern-women.png"), width=600, height=625)
+   p
    dev.off()
@@
 ```
+
+```{r CA versus Reflections Scatter}
+   source(here("Functions", "TwoModeCA.R"))
+   tm <- TwoModeCA(A)
+   p.plot.dat <- data.frame(c = tm$ca.eigvec.r[, 1]*-1, r = a$p.s[, 26])
+   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
+   p <- p + geom_point(size = 3, color = "blue") 
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
+   p
+   dev.off()
+
+   g.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1]*-1, 
+                                  r = a$g.s[, 26]))
+   p <- ggplot(data = g.plot.dat, aes(x = c, y = r))
+   p <- p + geom_point(size = 3, color = "red") 
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_text_repel(aes(label = rownames(g.plot.dat)), 
+                            color = "blue", size = 6)
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
+   p
+   dev.off()
+```

### 2025-01-29 analysis-R1.qmd
new stuff

--- 2025-01-27 analysis-R1.qmd
@@
       library(conflicted)
+      library(cowplot)
       library(FactoMineR)
+      library(ggcorrplot)
+      library(ggbump)
+      library(ggplot2)
+      library(ggpubr)
+      library(ggrepel)
       library(here)
+      library(factoextra)
       library(igraph)
@@
       library(networkdata)
+      library(pals)
+      library(stringr)
       conflicted::conflicts_prefer(dplyr::filter)
@@
 # Southern Women Data Table
-```{r HTML data table}
-   g <- southern_women
-   A <- as.matrix(as_biadjacency_matrix(g))
-   p.ord <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRNA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-   A <- A[p.ord, ]
-   rownames(A) <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-   colnames(A) <- paste("E", 1:14, sep = "")
-   kbl(A, align = "c", format = "latex", 
-       booktabs = TRUE, linesep = "") %>% 
-      kable_minimal(full_width = F, font_size = 10) %>% 
-      save_kable(here("southern-women-corrected.tex"))
+```{r Southern Women Data Matrix}
+     A <- as.matrix(as_biadjacency_matrix(southern_women))
+     p.names <- str_to_title(rownames(A))
+     p.names[which(p.names == "Myrna")] <- "Myra"
+     rownames(A) <- p.names
+     colnames(A) <- str_replace(colnames(A), "/", "-")
+     wlab <- paste("W", 1:18, sep = "")
+     wlab <- paste(wlab, rownames(A), sep = " (")
+     wlab <- paste(wlab, ")", sep = "")
+     elab <- paste("E", 1:14, sep = "")
+     elab <- paste(elab, colnames(A), sep = " (")
+     elab <- paste(elab, ")", sep = "")
+     rownames(A) <- wlab
+     colnames(A) <- elab
+     p <- ggcorrplot(t(A), colors = c("white", "white", "red")) 
+     p <- p + theme(legend.position = "none", 
+                  axis.text.y = element_text(size = 16),
+                  axis.text.x = element_text(size = 16, angle = 45, hjust = -0.2),
+                  )
+     p <- p + scale_x_discrete(position = "top") 
+     p <- p + scale_y_discrete(limits = rev)
+     jpeg(file = here("Plots", "southern-women.jpg"), width=800, height=800)
+     p
+     dev.off()
 ```
 
-```{r}
-   png(file = here("Plots", "ca-southern-women.png"), width=700, height=700)
-   CA(A)
+```{r CA of Aaffiliation Matrix}
+   ca.res <- CA(A, graph = FALSE)
+   plot.dat <- data.frame(rbind(ca.res$row$coord, ca.res$col$coord)[, 1:2])
+   plot.dat <- cbind(plot.dat, mode = c(rep(1, 18), rep(2, 14)), names = rownames(plot.dat))
+   p <- ggplot(plot.dat, aes(y = Dim.2, x = Dim.1, 
+                             color = factor(mode), label = names))
+   p <- p + geom_vline(xintercept = 0, color = "gray")
+   p <- p + geom_hline(yintercept = 0, color = "gray")
+   p <- p + geom_text_repel(size = 4, max.overlaps = 50)
+   p <- p + theme_minimal() 
+   p <- p + theme(legend.position = "none",
+                  axis.text = element_text(size = 18),
+                  axis.title = element_text(size = 20)
+                  )
+   p <- p + labs(x = "Second Dimension", y = "First Dimension")
+   p <- p + scale_color_manual(values = c('red', 'blue'))
+   png(file = here("Plots", "ca-southern-women.png"), width=600, height=625)
+   p
    dev.off()
@@
 ```
+
+```{r CA versus Reflections Scatter}
+   source(here("Functions", "TwoModeEig.R"))
+   tm <- TwoModeEig(A)
+   p.plot.dat <- data.frame(c = tm$ca.eigvec.r[, 1]*-1, r = a$p.s[, 26])
+   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
+   p <- p + geom_point(size = 3, color = "blue") 
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
+   p
+   dev.off()
+
+   g.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1]*-1, 
+                                  r = a$g.s[, 26]))
+   p <- ggplot(data = g.plot.dat, aes(x = c, y = r))
+   p <- p + geom_point(size = 3, color = "red") 
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_text_repel(aes(label = rownames(g.plot.dat)), 
+                            color = "blue", size = 6)
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
+   p
+   dev.off()
+```
+
+```{r CA versus Bonacich Re-ordered Affiliation Matrices}
+   a <- TwoModeCA(A)
+   b <- a$ca.A
+   c <- a$bon.A
+   png(file = here("Plots", "ca-reord.png"), width=500, height=500)
+   b
+   dev.off() 
+   png(file = here("Plots", "bon-reord.png"), width=500, height=500)
+   c
+   dev.off() 
+```

### 2025-01-29 Functions/TwoModeEig.R

@@ -1,0 +1,86 @@
+TwoModeEig <- function(x, 
+                      k.r = 2, 
+                      k.c = 2, 
+                      b.r = 2,
+                      b.c = 2,
+                      d.r = 2, 
+                      d.c = 2) {
+     # Initial matrices and vectors
+     A <- x[ , colSums(x) != 0, drop = FALSE] #dropping isolates
+     At <- t(A) #affiliation matrix transpose
+     r <- nrow(A) #number of rows
+     c <- ncol(A) #number of columns
+     rn <- rownames(A) #row name vector
+     cn <- colnames(A) #column name vector
+     u.r <- matrix(1, nrow = r, ncol = 1) #unit row vector
+     u.c <- matrix(1, nrow = c, ncol = 1) #unit column vector
+     
+     # Derived matrices and vectors
+     AAt <- A %*% At #row projection
+     AtA <- At %*% A #column projection
+     N <- as.numeric(t(u.r) %*% A %*% u.c) #affiliation matrix size
+     D.r <- diag(as.vector(A %*% u.c), r, r) #row diagonal degree matrix
+     D.c <- diag(as.vector(At %*% u.r), c, c) #column diagonal degree matrix
+     iDr <- solve(D.r) #inverse of row diagonal degree matrix
+     iDc <- solve(D.c)  #inverse of column diagonal degree matrix
+     S.r <- A %*% iDc %*% At #degree-normalized similarity matrix for rows
+     S.c <- At %*% iDr %*% A #degree-normalized similarity matrix for columns
+     
+     # Spectral matrices
+     CA.r <- eigen(iDr %*% A %*% iDc %*% At) #row CA
+     CA.c <- eigen(iDc %*% At %*% iDr %*% A) #column CA
+     S.r.eig <- eigen(S.r) #eigendecomposition of row similarity matrix 
+     S.c.eig <- eigen(S.c) #eigendecomposition of column similarity matrix 
+     L.r.eig <- eigen(D.r - S.r) #eigendecomposition of row similarity matrix Laplacian
+     L.c.eig <- eigen(D.c - S.c) #eigendecomposition of column similarity matrix Laplacian
+     eigvec.r <- Re(CA.r$vectors[, 2:r]) #row coordinates
[…]

### 2025-01-29 main-R1.tex

@@
     \centering
-    \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
+    \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
     \caption{Southern Women data with rows and columns ordered as in the original DGG table.}
@@
 \end{figure}
+
 \subsection{Empirical Example} \label{subsec:ex}
@@
 
-One key source of confusion in the mini-literature that exists around this dataset concern the order of presentation of the rows and columns.
+One key source of confusion in the mini-literature that exists around this dataset concerns the order of presentation of the rows and columns.
 The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous.
-It \textit{could} be treated as nominal, but the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.
-As \citet{freeman2003finding} notes, in the canonical original table both the row and column modes had been re-ordered in presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order in terms of column ordering.
-To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns, but in fact, as shown in~\ref{tab:southern}, it is the event that occurred on August 3rd, and there were three events that occurred after that one (E4, E8, and E13) in the DGG numbering.
-
-To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, but with the events ordered chronologically, which differs from the original DGG Table reproduced in \citet{freeman2003finding}.
-This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding}.
-Different versions of the dataset reproduce different versions of the labeling.
-Of course, at the end of the day, the specific labeling of the nodes does not really matter, since we are usually interested in the structural properties of the rows and columns.
-The even ordering however, does matter for analyses that try to exploit the temporal properties of the data.
-Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to correspond to "W1" or "E8" in the other (e.g., \citet{doreian1979evolution}).
+It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.
+As \citet{freeman2003finding} notes, in the canonical original table, both the row and column modes had been re-ordered in a presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns.
+To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns.
+In fact, as shown in~\ref{tab:southern}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.
+
+To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, but with the events ordered \textit{chronologically}, which differs from the original DGG Table reproduced in \citet{freeman2003finding}.
+This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; different versions of the dataset reproduce different versions of the labeling.
+Of course, the specific labeling of the nodes does not really matter since we are usually interested in the structural properties of the rows and columns.
+Event ordering, however, matters for analyses that try to exploit the temporal properties of the data.
+Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to "W1" or "E8" in the other (e.g., \citet{doreian1979evolution}).
 

### 2025-01-29 Reviews and Response/response-memo.tex

@@
 
-\textbf{Response}:
+\textbf{Response}: Thanks for this suggestion. I have now included a relevant highlight bulletpoint noting the relevant analytic contrast of Bonacich versus CA.
 
@@
 In any case, to forestall any confusion, and given that the in the \texttt{networkdata} package, the \texttt{igraph} object containing the Southern Women data come ordered using the Freeman (2003) ordering, I'm sticking to the non-chronological event labeling.
-I include a very brief discussion in the paper clarifying this issue.
+I now include a brief discussion in the paper clarifying this issue.
 
@@
 
-\underline{More extensive introduction and background regarding previous uses of CA and MCA}
+\underline{More extensive introduction and background}
 

### 2025-01-29 highlights.tex

@@
   \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
-  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
+  \item CA can extract latent positional information on people and groups, akin to recent work on ``generalized relational similarity" in two-mode networks. 
+  \item The Bonacich dual centrality is useful for highlighting core-periphery patterns in the two-mode case in contrast to CA scores which are more appropriate to revealing the latent group structure
 \end{itemize}

### 2025-01-29 main-R1.tex

@@
 I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I contrast this ordering to that provided by the Bonacich dual centrality scores, which, in contrast to the CA ordering, reveals a dual community partition and recovers the core-periphery structure of the two-mode network instead.
+I contrast this ordering to that provided by the Bonacich dual centrality scores, which, in contrast to the CA ordering, reveals a dual community partition and recovers the core-periphery structure of the two-mode network instead \citep{everett2013dual}.
 In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
+
+\section{Previous uses of CA and MCA in Two-Mode Network Data Analysis}
+
+\subsection{CA as a Data Visualization Tool}
+
+\citet{borgatti1997network} provided an early in-depth considerations of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
+The discussion was placed in a section on ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as visualization technique.
+According to \citet[246]{borgatti1997network}, CA ``can be seen as a method for representing points in a metric space such that distances between the points are meaningful.'' More specifically, \citet{borgatti1997network}, also using the Southern Women data as illustration,  claimed that in the CA map, points representing the women will appear closer if the women attend ``mostly the same events,'' and conversely points representing the events will be placed near one another if the events were attended by ``mostly the same women,'' and that women and events would appear closer in the plot if ``those women attended those events.'' \citet[247-249]{borgatti1997network} then went on to focus on various limitations of CA as visualization tool, noting that given the limiting range of dichotomous network data, two-dimensional representations ``will almost always be severely inaccurate or misleading.'' They also noted that since CA is a technique designed to deal with valued frequency data, its use with binary network data would lead to dimensional scores that would be either ``meaningless'' or ``difficult to interpret.'' Finally, they noted that since distances in the CA biplot are not Euclidean they would mislead interpreters who would try to impose such a metric to make sense of them.
+
+In a paper published shortly thereafter \citet{roberts2000correspondence}, responded to some of the criticisms that \citet{borgatti1997network} levied at the use of CA for analyzing two-mode networks in their original piece.
+First, \citet{roberts2000correspondence} noted that there were many examples in the standing of using CA to analyze binary non-frequency data, and that therefore the uses of CA are not constrained by the nature or provenance of the numerical quantities recorded in each cell of a two-mode network affiliation matrix.
+\citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and and the event points in the two-dimensional representation and one that used all thirteen dimensions.
+Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that even (and similarly for column profiles corresponding to events).
+Thus, the point distances are indeed interpretable as a kind of degree-weighted similarity between objects in each mode.
+
+\subsection{CA and Two-Mode Centrality}
+One of the earliest considerations of CA as a data-analytic tool for two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks.
+In that paper \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people buy the groups to which they belong,'' then it would seem that CA would be an appropriate analytic technique.
+According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (square quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation (in the least-squared error) sense of the similarities between row and column objects.
+Bonacich's discussion of CA came after his influential introduction of the Eigenvector centrality approach for two-mode network analysis.
+Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of a matrix---e.g., the Singular Value Decomposition (SVD).
+
+As we will see later, the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated.
+Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight ad describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques, and for events their likelihood of being attended by the most representative members of them
+\citep[164]{bonacich1991simultaneous}.
+
+In an influential paper on centrality in two-mode networks \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA.
+Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data.
+In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the even she attends, and for events their scores are given by the sum of the eigenvector centralities of its members.
+Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship obtained between the CA scores corresponding to persons and groups.
+The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends, and for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event.
+\citet{faust1997centrality} also specified the particular way that the affiliation matrix $\mathbf{A}$ had to be modified before applying the SVD (recall this is the step that creates the difference between CA and Eigenvector centrality), noting that each cell $a_{pg}$ is divided by the square root of the product of the number of events the row woman attends and the number of people attending the column event ($\sqrt{\sum_p a_{pg}\sum_g a_{pg}}$).
 
@@
     \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
-    \caption{Southern Women data with rows and columns ordered as in the original DGG table.}
+    \caption{Southern Women data matrix with rows and columns ordered as in the original DGG table.}
     \label{fig:sw}
@@
 \subsection{Empirical Example} \label{subsec:ex}
-\subsubsection{Souther Women Data}
-The following analyses use the classic Southern Women data \citep{davis1941} (DGG) as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, borgatti2011analyzing, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}.
-As we can see, novel insights can still be garnered from these old data.
-
-One key source of confusion in the mini-literature that exists around this dataset concerns the order of presentation of the rows and columns.
+\subsubsection{Southern Women Data}
+The following analyses use the classic ``Southern Women'' data \citep{davis1941} (DGG) as the running example.
+This data is doubly appropriate for showcasing the relevance of methods that exploit the duality between two sets of entities, as it was it used in Breiger's \citeyearpar{breiger1974duality} classic paper.
+Moreover, Southern Women is dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten much longer in the last two decades since Freeman's methodological ``meta-analysis'' \citep[e.g.,][among many others]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, kovacs2010generalized, brusco2011analysis, borgatti2014analyzing, everett2013dual, lerner2022dynamic, batagelj2022analysis, lizardo2024two}.
+Despite this long list of studies of the Southern Women data, we will see that novel insights can still be garnered from this old data source.
+
+Despite its undisputed pedigree as a methodological touchstone in studies of duality and dual constitution in two-mode networks, a key source of confusion in the mini-literature that exists around the Southern Women data concerns the order of presentation of the rows and columns.
 The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous.
@@
 As \citet{freeman2003finding} notes, in the canonical original table, both the row and column modes had been re-ordered in a presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns.
+
 To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns.
-In fact, as shown in~\ref{tab:southern}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.
-
-To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, but with the events ordered \textit{chronologically}, which differs from the original DGG Table reproduced in \citet{freeman2003finding}.
-This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; different versions of the dataset reproduce different versions of the labeling.
+In fact, as shown in Figure~\ref{fig:sw}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.\footnote{Using the DGG labeling, the chronologically ordered set of events are \{E11, E5, E2, E7, E12, E9, E3, E6, E10, E1, E14 E8, E4, E13\} \citep[p.
+68]{everett2018measuring}.}
+
+To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women---taken from \citet[83]{homans1950human}---from that of the original DGG table reproduced in \citet{freeman2003finding}, and ordered the events from left to right in the affiliation matrix \textit{chronologically}, which, as we have seen, also differs from the original DGG Table reproduced in \citet{freeman2003finding}.
+This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; in all, different versions of the dataset reproduce different versions of the labeling.
+
 Of course, the specific labeling of the nodes does not really matter since we are usually interested in the structural properties of the rows and columns.
-Event ordering, however, matters for analyses that try to exploit the temporal properties of the data.
-Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to "W1" or "E8" in the other (e.g., \citet{doreian1979evolution}).
-
-To forestall confusion, in what follows I use the numeric labeling corresponding to the original DGG data table as reproduced in \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{tab:southern}, which include both the arbitrary ordered labels (e.g., "W1" or "E1") and both the woman's name and the event date in parentheses, with "Myrna" rebaptized to her given name of "Myra" to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.
+Event ordering, however, matters for analyses that try to exploit the temporal properties of the data \citep[e.g.,][]{everett2018measuring, lerner2022dynamic}.
+Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication \citep[e.g.,][]{freeman2003finding} may not correspond to "W1" or "E8" in the other \citep[e.g.,][]{doreian1979evolution}.
+
+To forestall confusion, in what follows, I use the numeric labeling corresponding to the original DGG data table as reproduced in (among others) \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{fig:sw}.
+My presentation includes both the arbitrary ordered labels (e.g., "W1" or "E1") and both the woman's name and the event date in parentheses, with "Myrna" (W11) rebaptized to her given name of "Myra" to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.
 
@@
 As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
+Accordingly, the $C^R$ reflective scores reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
+
 The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side).
-The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
+In fact, as \citet[p.
+206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in a the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to direct optimization methods.
+The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
 
@@
 I showed that just like the reflective centralities that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums.
-This exercise clarifies the links between CA and the dual two-mode eigenvector centrality in a more coherent and systematic way, an issue that was left open and somewhat unclear in Bonacich's (1991) classic paper.
+This exercise clarifies the links between CA and the dual two-mode eigenvector centrality more coherently and systematically, an issue that was left open and somewhat unclear in Bonacich's \citeyearpar{bonacich1991simultaneous} classic paper.
 Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.
@@
 When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data.
-This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events.
+This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events \citep{everett2013dual}.
 
@@
 But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to?
-Here we argued that CA recovers a type of generalized relational similarity (GRS).
+Here, we show that CA recovers a type of generalized relational similarity (GRS).
 That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar.

### 2025-01-29 Reviews and Response/response-memo.tex
Updates from Overleaf

@@
 
-\textbf{Response}:
+\textbf{Response}: Thanks for this suggestion. I have re-plotted the figures using the suggested column ordering.
 
@@
 
-\textbf{Response}:
+\textbf{Response}: Thanks for this observation.
+You are correct that I did not adequately explain where the eigenvectors corresponding to the CA (and HH) scores come from.
+You are right that the equations result in row-stochastic matrices (row sums equal 1.0) that are square but not symmetric.
+You can still do an eigenvector decomposition of each of these matrices, but the first (dominant) eigenvalue will be 1.0, and the associated eigenvectors will be filled with constant values.
+So the relevant eigenvector we seek, which is the first CA dimension, is that associated with the \textit{second} largest eigenvalue (and the second CA dimension is that associated with the third, and so forth).
+Another way of saying it is that we can perform the usual eigendecomposition of the matrices produced by (23) and (24), discard the first trivial eigenvalue and eigenvector and treat the second as the first, the third as the second, etc., and that will be equivalent to the CA solution (the same scores that simple CA of the original rectangular Affiliation matrix would yield).
+The relevant section of the paper has been rewritten to clarify these issues.
 

### 2025-01-29 main-R1.tex
Updates from Overleaf

@@
 As we will see later, the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated.
-Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight ad describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques, and for events their likelihood of being attended by the most representative members of them
+Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight ad describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by the most representative members of them
 \citep[164]{bonacich1991simultaneous}.
@@
 Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data.
-In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the even she attends, and for events their scores are given by the sum of the eigenvector centralities of its members.
-Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship obtained between the CA scores corresponding to persons and groups.
+In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the event she attends; for events, their scores are given by the sum of the eigenvector centralities of its members.
+Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship was obtained between the CA scores corresponding to persons and groups.
 The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends, and for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event.
@@
 The following analyses use the classic ``Southern Women'' data \citep{davis1941} (DGG) as the running example.
-This data is doubly appropriate for showcasing the relevance of methods that exploit the duality between two sets of entities, as it was it used in Breiger's \citeyearpar{breiger1974duality} classic paper.
-Moreover, Southern Women is dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten much longer in the last two decades since Freeman's methodological ``meta-analysis'' \citep[e.g.,][among many others]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, kovacs2010generalized, brusco2011analysis, borgatti2014analyzing, everett2013dual, lerner2022dynamic, batagelj2022analysis, lizardo2024two}.
+This data is doubly appropriate for showcasing the relevance of methods that exploit the duality between two sets of entities, as it was used in Breiger's \citeyearpar{breiger1974duality} classic paper.
+Moreover, Southern Women is a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts.
+The list has only gotten much longer in the last two decades since Freeman's methodological ``meta-analysis'' \citep[e.g.,][among many others]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, kovacs2010generalized, brusco2011analysis, borgatti2014analyzing, everett2013dual, lerner2022dynamic, batagelj2022analysis, lizardo2024two}.
 Despite this long list of studies of the Southern Women data, we will see that novel insights can still be garnered from this old data source.
@@
 However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
-Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al.
-(2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
+Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the dominant partition between the women observed by \citet{freeman2003finding}, but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
 
@@
 Clearly, $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
-Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering the average of the average number of memberships of its members (and so forth).
+Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering (the average of) the average number of memberships of its members (and so forth).
 
@@
 In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings.
-Instead, Bonacich noted that, in analyzing the Southern Women data shown in Table~\ref{tab:southern}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
+Instead, Bonacich noted that, in analyzing the Southern Women data shown in Figure~\ref{fig:sw}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
 
@@
 
-Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
+Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\cite{bonacich1991simultaneous, faust2005using}.
+Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.
+
+To see this, consider the $|P| \times |P|$ matrix $Dp$, containing the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
 In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as the solution of the eigensystem:
+Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose (recall that a matrix is row stochastic if its rows sum to one).
+
+For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
+
+\begin{equation}
+    \mathbf{P}_{pg} = \mathbf{D}p^{-1}\mathbf{A}
+    \label{eq:ca1}
+\end{equation}
+
+We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
+
+\begin{equation}
+    \mathbf{P}_{gp} = \mathbf{D}_G^{-1}\mathbf{A}^T
+    \label{eq:ca2}
+\end{equation}
+
+Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step due to \citet{breiger1974duality} of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
+
+For the people, the degree-weighted projection is:
+
+\begin{equation}
+\mathbf{P}_{pp} = \mathbf{P}_{pg}\mathbf{P}_{gp}
+    \label{eq:ca3}
+\end{equation}
+
+And for the groups:
+
+\begin{equation}
+    \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
+    \label{eq:ca3}
+\end{equation}
+
+Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($P \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
+For the Southern Women data, they are show in Figure~\ref{fig:p-norm} and~\ref{fig:g-norm}.
+
+It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as a solution of the eigensystem\footnote{More specifically since both equations~\ref{eq:dam1} and~\ref{eq:dam2} produce row stochastic matrices (row sums are fixed to sum to 1.0), their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
+Accordingly, the scores we seek, which are the same as the CA scores and the HH reflective centrality scores, are those stored in the eigenvectors associated with the \textit{second} largest eigenvalue~\citep{van2021correspondence}.}:
 

### 2025-01-29 main-R1.tex
Updates from Overleaf

@@
 
-\citet{borgatti1997network} provided an early in-depth considerations of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
-The discussion was placed in a section on ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as visualization technique.
-According to \citet[246]{borgatti1997network}, CA ``can be seen as a method for representing points in a metric space such that distances between the points are meaningful.'' More specifically, \citet{borgatti1997network}, also using the Southern Women data as illustration,  claimed that in the CA map, points representing the women will appear closer if the women attend ``mostly the same events,'' and conversely points representing the events will be placed near one another if the events were attended by ``mostly the same women,'' and that women and events would appear closer in the plot if ``those women attended those events.'' \citet[247-249]{borgatti1997network} then went on to focus on various limitations of CA as visualization tool, noting that given the limiting range of dichotomous network data, two-dimensional representations ``will almost always be severely inaccurate or misleading.'' They also noted that since CA is a technique designed to deal with valued frequency data, its use with binary network data would lead to dimensional scores that would be either ``meaningless'' or ``difficult to interpret.'' Finally, they noted that since distances in the CA biplot are not Euclidean they would mislead interpreters who would try to impose such a metric to make sense of them.
+\citet{borgatti1997network} provided an early in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
+The discussion was placed in a section on ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as a visualization technique.
+According to \citet[246]{borgatti1997network}, CA ``can be seen as a method for representing points in a metric space such that distances between the points are meaningful.'' More specifically, \citet{borgatti1997network}, also using the Southern Women data as illustration,  claimed that in the CA map, points representing the women will appear closer if the women attend ``mostly the same events.'' Conversely, points representing the events will be placed near one another if the events were attended by ``mostly the same women,'' and that women and events would appear closer in the plot if ``those women attended those events.'' \citet[247-249]{borgatti1997network} then went on to focus on various limitations of CA as a visualization tool, noting that given the limiting range of dichotomous network data, two-dimensional representations ``will almost always be severely inaccurate or misleading.'' They also noted that since CA is a technique designed to deal with valued frequency data, its use with binary network data would lead to dimensional scores that would be either ``meaningless'' or ``difficult to interpret.'' Finally, they noted that since distances in the CA biplot are not Euclidean they would mislead interpreters who would try to impose such a metric to make sense of them.
 
@@
 
+\begin{figure}[ht!]
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
+    \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/p-norm.png}
+            \caption{Degree-weighted projection (People).}
+            \label{fig:ca-reord}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/g-norm.png}
+            \caption{Degree-weighted projection (Groups).}
+            \label{fig:bon-reord}
+    \end{subfigure}
+    \caption{}
+    \label{fig:dwproj}
+\end{figure}
+
 For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
@@
 Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($P \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
-For the Southern Women data, they are show in Figure~\ref{fig:p-norm} and~\ref{fig:g-norm}.
+For the Southern Women data, they are shown in Figure~\ref{fig:dwproj}.
 

### 2025-01-29 main-R1.tex

@@
 
-This just says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
-
-Of course, it is evident that these sum of sums would diverge to a bigger and bigger quantity at each step $q$.
+This says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
+
+Of course, these sums of sums would diverge to a larger and larger quantity at each step $q$.
 To prevent this and guarantee convergence, we normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:
@@
     \end{subfigure}
-    \caption{}
+    \caption{Degree-weighted projection matrices of the Southern Women data.}
     \label{fig:dwproj}
@@
 
-For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
+For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
 
@@
 
-We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
+We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
 
@@
     \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
-    \label{eq:ca3}
+    \label{eq:ca4}
 \end{equation}
@@
 For the Southern Women data, they are shown in Figure~\ref{fig:dwproj}.
-
-It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as a solution of the eigensystem\footnote{More specifically since both equations~\ref{eq:dam1} and~\ref{eq:dam2} produce row stochastic matrices (row sums are fixed to sum to 1.0), their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
-Accordingly, the scores we seek, which are the same as the CA scores and the HH reflective centrality scores, are those stored in the eigenvectors associated with the \textit{second} largest eigenvalue~\citep{van2021correspondence}.}:
-
-\begin{equation}
-    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
+Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person.
+Thus, higher values indicate an affinity or proximity between the people (and the groups in the corresponding matrix).
+
+To get to the main point of all of this, it can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as a solution of the eigensystem:
+
+\begin{equation}
+    \mathbf{P}_{pp}C^R_p = \lambda_2C^R_p
+    %\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
     \label{eq:dam1}
@@
 \begin{equation}
-    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
+    \mathbf{P}_{gg}C^R_g = \lambda_2C^R_g
+    %\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
     \label{eq:dam2}
-\end{equation}
-
-Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups.
-The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
-
-This can be clearly seen if we express $\left(Dg^{-1}A^TDp^{-1}A\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
+\end{equation}\
+
+Note that the $C^R$ scores are obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue.
+This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
+Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective centrality scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
+9.17]{fouss2016algorithms}.
+Note also that just like the eigenvector decomposition, the matrices implied by equations~\ref{eq:dam1} and~\ref{eq:dam2} have equivalent sets of (non-zero) eigenvalues.
+
+More importantly, because $\mathbf{P}_{pp} = Dp^{-1}ADg^{-1}A^T$ and $\mathbf{P}_{gg} = Dg^{-1}A^TDp^{-1}A$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
+Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ and $\mathbf{P}_{pp}$ for people and $A^TA$ and $\mathbf{P}_{gg}$ for groups.
+The key difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
+
+This can be clearly seen if we express $\left(Dg^{-1}\mathbf{A}^TDp^{-1}\mathbf{A}\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
 
@@
 
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
+Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}\mathbf{A}D_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
 As we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix for both the row and column objects.

### 2025-01-29 Reviews and Response/response-memo.tex

@@
 You are right that the equations result in row-stochastic matrices (row sums equal 1.0) that are square but not symmetric.
-You can still do an eigenvector decomposition of each of these matrices, but the first (dominant) eigenvalue will be 1.0, and the associated eigenvectors will be filled with constant values.
+You can still do a spectral decomposition of each of these matrices, but the first (dominant) eigenvalue will be 1.0, and the associated eigenvectors will be filled with constant values.
 So the relevant eigenvector we seek, which is the first CA dimension, is that associated with the \textit{second} largest eigenvalue (and the second CA dimension is that associated with the third, and so forth).
-Another way of saying it is that we can perform the usual eigendecomposition of the matrices produced by (23) and (24), discard the first trivial eigenvalue and eigenvector and treat the second as the first, the third as the second, etc., and that will be equivalent to the CA solution (the same scores that simple CA of the original rectangular Affiliation matrix would yield).
+Another way of saying it is that we can perform the usual eigendecomposition of the row-stochastic projection matrices, discard the first trivial eigenvalue and eigenvector and treat the second as the first, the third as the second, etc., and that will be equivalent to the CA solution (the same scores that simple CA of the original rectangular Affiliation matrix would yield).
 The relevant section of the paper has been rewritten to clarify these issues.

### 2025-01-29 main-R1.tex

@@
 
-Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($P \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
-For the Southern Women data, they are shown in Figure~\ref{fig:dwproj}.
+Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
+Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
 Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person.
 Thus, higher values indicate an affinity or proximity between the people (and the groups in the corresponding matrix).
+The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to exactly the same groups).
 

### 2025-01-29 main-R1.tex

@@
 Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich two-mode, the centrality of people is proportional to the centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
-These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
+These equations make clear that in the Bonacich two-mode, the centrality of people is equal to the sum of the centralities scores of the groups they join, and the centrality of groups is equivalent to the sum of the centralities of their members.
+This approach thus nicely captures the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
 Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
@@
 
-Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step due to \citet{breiger1974duality} of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
+Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step---due to \citet{breiger1974duality}---of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
 
@@
 Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person.
-Thus, higher values indicate an affinity or proximity between the people (and the groups in the corresponding matrix).
-The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to exactly the same groups).
+Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
+The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).
 
@@
 
-Note that the $C^R$ scores are obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue.
+With the $C^R$ scores obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the degree-weighted projection matrices.
 This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.

### 2025-01-31 Reviews and Response/response-memo.tex

@@
 \textbf{Reviewer}: \textit{I have some resistance to your term "reflective centrality" (Section 2 title) to characterize the scores of the respective modes - precisely because, as you importantly show, in the limit the reflective scores are first dimensions of a CA, and (as you also importantly emphasize) the first CA dimension is about finding groups, not about ordering nodes.
-Hidalgo and Hausmann (2009) do not use the term "centrality" for the reflective scores they define and compute.
+Hidalgo and Hausmann (2009) do not use the term ``centrality" for the reflective scores they define and compute.
 I wonder whether ``reflective scores" might be a better term; I'm not sure, and simply ask you to think about it.}
@@
 You can still do a spectral decomposition of each of these matrices, but the first (dominant) eigenvalue will be 1.0, and the associated eigenvectors will be filled with constant values.
-So the relevant eigenvector we seek, which is the first CA dimension, is that associated with the \textit{second} largest eigenvalue (and the second CA dimension is that associated with the third, and so forth).
-Another way of saying it is that we can perform the usual eigendecomposition of the row-stochastic projection matrices, discard the first trivial eigenvalue and eigenvector and treat the second as the first, the third as the second, etc., and that will be equivalent to the CA solution (the same scores that simple CA of the original rectangular Affiliation matrix would yield).
+So the relevant eigenvector we seek, which is the first CA dimension, is that associated with the \textit{second} largest eigenvalue, and the second CA dimension is that associated with the third, and so forth.
+Another way of saying it is that we can perform the usual eigendecomposition of the row-stochastic projection matrices, discard the first trivial eigenvalue and eigenvector and treat the second as the first, the third as the second, etc., and that will be equivalent to the CA solution, the same scores that simple CA of the original rectangular Affiliation matrix would yield (but note that we run into a similar situation when computing CA via SVD of the row and column mass normalized affiliation matrix, which results in a solution with a first eigenvalue of 1.0 and associated trivial eigenvectors with person and event scores proportional to the row and column masses, which need to be discarded with the focus being on the second, third, eigenvalues and eigenvectors as Faust 2005, p.
+126 notes).
 The relevant section of the paper has been rewritten to clarify these issues.

### 2025-01-31 main-R1.tex

@@
 
+In this section, I consider previous contributions regarding the usefulness and limitations of CA as a methodological tool for the analysis of two-mode network data.
+For reasons of space and relevance, therefore my discussion is limited in two ways.
+First, I focus attention on (1) pieces discussing primarily methodological aspects of CA and (2) with regard to the analysis of two-mode network data or affiliation networks.\footnote{Naturally, I also exclude the rich literature that has used methods other than CA to analyze affiliation networks.} There are, of course, various substantive applications of CA to the analysis of social networks, as well as substantive and methodological applications of CA to the analysis of one-mode network data; these will not be part of the discussion here.
+
+With these restrictions in mind, the literature on CA and two-mode networks can be divided into two streams.
+The first contributions focus on the usefulness of CA as a tool for the visualization or ``joint display'' of the two sets of entities in an affiliation network.
+Secondly, some contributions focus on the similarities and differences between CA and factoring---those based on a spectral decomposition of a suitable matrix---techniques for assigning scores to the nodes of a two-mode network, where the scores exploit the duality between the two sets of entities and could be, under some circumstances, interpreted as ``centrality'' scores.
+I discuss each in turn.
+
 \subsection{CA as a Data Visualization Tool}
 
-\citet{borgatti1997network} provided an early in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
-The discussion was placed in a section on ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as a visualization technique.
-According to \citet[246]{borgatti1997network}, CA ``can be seen as a method for representing points in a metric space such that distances between the points are meaningful.'' More specifically, \citet{borgatti1997network}, also using the Southern Women data as illustration,  claimed that in the CA map, points representing the women will appear closer if the women attend ``mostly the same events.'' Conversely, points representing the events will be placed near one another if the events were attended by ``mostly the same women,'' and that women and events would appear closer in the plot if ``those women attended those events.'' \citet[247-249]{borgatti1997network} then went on to focus on various limitations of CA as a visualization tool, noting that given the limiting range of dichotomous network data, two-dimensional representations ``will almost always be severely inaccurate or misleading.'' They also noted that since CA is a technique designed to deal with valued frequency data, its use with binary network data would lead to dimensional scores that would be either ``meaningless'' or ``difficult to interpret.'' Finally, they noted that since distances in the CA biplot are not Euclidean they would mislead interpreters who would try to impose such a metric to make sense of them.
+\citet{borgatti1997network} provided an early and influential in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
+The discussion was placed in a section of the paper titled ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as a visualization technique.
+According to \citet[246]{borgatti1997network}, CA ``can be seen as a method for representing points in a metric space such that distances between the points are meaningful.'' More specifically, \citet{borgatti1997network}, also using the Southern Women data as illustration,  claimed that in the CA map, points representing the women will appear closer if the women attend ``mostly the same events.'' Conversely, points representing the events will be placed near one another if the events were attended by ``mostly the same women,'' and that women and events would appear closer in the plot if ``those women attended those events.'' \citet[247-249]{borgatti1997network} then went on to focus on various limitations of CA as a visualization tool, noting that given the limiting range of dichotomous network data, two-dimensional representations ``will almost always be severely inaccurate or misleading.'' They also noted that since CA is a technique designed to deal with valued frequency data, its use with binary network data would lead to dimensional scores that would be either ``meaningless'' or ``difficult to interpret.'' Finally, they noted that since distances in the CA biplot are not Euclidean, they would mislead interpreters who would try to impose such a metric to make sense of them.
 
 In a paper published shortly thereafter \citet{roberts2000correspondence}, responded to some of the criticisms that \citet{borgatti1997network} levied at the use of CA for analyzing two-mode networks in their original piece.
-First, \citet{roberts2000correspondence} noted that there were many examples in the standing of using CA to analyze binary non-frequency data, and that therefore the uses of CA are not constrained by the nature or provenance of the numerical quantities recorded in each cell of a two-mode network affiliation matrix.
-\citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and and the event points in the two-dimensional representation and one that used all thirteen dimensions.
+First, \citet{roberts2000correspondence} noted that there were many examples in the standing of using CA to analyze binary non-frequency data and that, therefore, the uses of CA are not constrained by the nature or provenance of the numerical quantities recorded in each cell of a two-mode network affiliation matrix.
+\citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and the event points in the two-dimensional representation and one that used all thirteen dimensions.
 Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that even (and similarly for column profiles corresponding to events).
 Thus, the point distances are indeed interpretable as a kind of degree-weighted similarity between objects in each mode.
+
+Perhaps the most thorough discussion of the uses of CA as a tool for the ``joint display'' of affiliation networks is that provided by \citet{faust2005using}.
+According to \citet[118]{faust2005using}, previous debates regarding the usefulness---or lack thereof---of CA as a visualization tool were marred by imprecision and lack of ``formal specification'' regarding the exact substantive meaning of the scores obtained via the procedure and a related lack of sustained discussion as to the exact ``relationship between the model and the input data.'' Faust's basic point is that CA provides a way to both place actors and events in an affiliation network in a joint space such that the inter-point distances are interpretable and such that the link to the original data (e.g., the affiliation network matrix) is transparent.
+
+\citet{faust2005using} makes several contributions in this piece.
+First, Faust specifies the exact matrix that is subject to the singular value decomposition when CA is applied directly to the affiliation matrix of a two-mode network $\mathbf{A}$; namely, a scaled version where $\mathbf{A}$ is pre-multiplied by a diagonal matrix containing the square root of its row sums and post-multiplied by a diagonal matrix containing the square root of its column sums (equivalent to dividing every non-zero entry of $a_{pg}$ by the square root of the product of the number of memberships of person $p$ and the number of members of group $g$ before applying the SVD.
+Second, Faust specifies the exact relationship linking the resulting row and column CA scores and the original affiliation matrix input, providing the ``reconstruction equation" that takes us from the left and right singular vectors and eigenvalues to the former.
+Thirdly, Faust clarifies the duality linking the row and column scores, noting that the scores for the people are (activity) weighted sums of the scores that the groups they join, and the scores for the groups are (group size) weighted sums of the scores of their members.
+Finally, \citet[128ff]{faust2005using} clarifies the link between the CA scores---and different normalizations thereof---assigned to persons and events in an affiliation network on each dimension and (same node-set) inter-point distance in a joint space.
+Like \citet{roberts2000correspondence} before her, Faust clarifies that in the low (usually two) dimensional representation, inter-point distances are best (least squares) approximations of the chi-square distance between the membership profiles of pairs of people (or groups), where the profile of a given person-row or group-column is equal to the non-zero entries of the affiliation matrix divided by the corresponding row or column sum (people's activities and group sizes).\footnote{Note that because the chi-square distances are defined on these degree-weighted entries, they define a kind of degree-weighted similarity between pairs of people and pairs of groups, but not an exact similarity score as would be obtained, for instance, by using a criterion such as proportion matching \citep[208]{everett2013dual}.}
+
+Worth mentioning in this regard is an intriguing recent contribution by \citet{d2014use}, who convincingly advocate for using Multiple Correspondence Analysis (MCA) for the joint display of persons and groups in affiliation networks.
+The key difference between CA and MCA is that MCA works with a transformed version of the affiliation matrix, in which the rows are treated as ``cases'' and the columns as categorical attributes of the case, where each possible value represents a modality that the column object can take with respect to the cases.
+In the case of binary network data, the affiliation matrix is transformed into a new matrix $\mathbf{Z}$ with the same number of rows (persons) as the original and with twice the number of columns representing either membership ($z_{pg+} = 1$) or non-membership ($z_{pg-} = 1$) in that a group.
+The MCA scores are obtained via SVD of a centered and doubly normalized version of $\mathbf{Z}$.
+\citet{d2014use} discuss various advantages of the MCA graphical representation of two-mode networks, including the capacity to produce interpretable joint displays of actors and events and the ability to incorporate exogenous actor and event attributes as supplementary variables.
+However, because what follows bears more directly on the use of simple CA for two-mode network analysis, this discussion is less relevant.
 
@@
 As we will see later, the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated.
-Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight ad describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by the most representative members of them
+Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by the most representative members of them
 \citep[164]{bonacich1991simultaneous}.
 
-In an influential paper on centrality in two-mode networks \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA.
+In an influential paper on centrality in two-mode networks, \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA.
 Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data.
@@
 In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
-Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).
-Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
+Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
 As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
+This is consistent with the spirit of the original approach, recently described by \citet[92]{hidalgo2021economic} as centered on the ``\textit{duality} between economic inputs and outputs'' (italics mine).
 
@@
 
-\section{Illustrative Analysis of the Southern Women Data} \label{sec:anal}
 \subsection{Row and Column Re-ordered Affiliation Matrices}

### 2025-01-31 analysis-R1.qmd
new plots

--- 2025-01-27 analysis-R1.qmd
@@
 ---
-title: "Correspondence Analysis of Two Mode Networks"
+title: "The Correspondence Analysis of Two Mode Networks Revisited"
 author: "Omar Lizardo"
@@
       library(conflicted)
-   library(FactoMineR)
+      library(cowplot)
+      library(ggcorrplot)
+      library(ggbump)
+      library(ggplot2)
+      library(ggpubr)
+      library(ggrepel)
       library(here)
+      library(factoextra)
       library(igraph)
@@
       library(networkdata)
+      library(pals)
+      library(patchwork)
+      library(stringr)
       conflicted::conflicts_prefer(dplyr::filter)
@@
 
-# Southern Women Data Table
-```{r HTML data table}
-   g <- southern_women
-   A <- as.matrix(as_biadjacency_matrix(g))
-   p.ord <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRNA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-   A <- A[p.ord, ]
-   rownames(A) <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
-   colnames(A) <- paste("E", 1:14, sep = "")
-   kbl(A, align = "c", format = "latex", 
-       booktabs = TRUE, linesep = "") %>% 
-      kable_minimal(full_width = F, font_size = 10) %>% 
-      save_kable(here("southern-women-corrected.tex"))
-```
-
-```{r}
-   png(file = here("Plots", "ca-southern-women.png"), width=700, height=700)
-   CA(A)
+```{r Southern Women Data Matrix}
+     A <- as.matrix(as_biadjacency_matrix(southern_women))
+     p.names <- str_to_title(rownames(A))
+     p.names[which(p.names == "Myrna")] <- "Myra"
+     rownames(A) <- p.names
+     colnames(A) <- str_replace(colnames(A), "/", "-")
+     wlab <- paste("W", 1:18, sep = "")
+     wlab <- paste(wlab, rownames(A), sep = " (")
+     wlab <- paste(wlab, ")", sep = "")
+     elab <- paste("E", 1:14, sep = "")
+     elab <- paste(elab, colnames(A), sep = " (")
+     elab <- paste(elab, ")", sep = "")
+     rownames(A) <- wlab
+     colnames(A) <- elab
+     p <- ggcorrplot(t(A), colors = c("white", "white", "white"), 
+                     lab = TRUE, lab_size = 6.5) 
+     p <- p + theme(legend.position = "none", 
+                  axis.text.y = element_text(size = 16, family = "serif"),
+                  axis.text.x = element_text(size = 16, angle = 45, hjust = -0.2, family = "serif")
+                  )
+     p <- p + scale_x_discrete(position = "top") 
+     p <- p + scale_y_discrete(limits = rev)
+     jpeg(file = here("Plots", "southern-women.jpg"), width=800, height=800)
+     p
      dev.off()
@@
 
-```{r Reflections Bump Chart (Persons)}
+```{r Faust-Style CA}
+   rs <- rowSums(A)
+   cs <- colSums(A)
+   R <- diag(1/sqrt(rs))
+   C <- diag(1/sqrt(cs))
+   A.n <- R %*% A %*% C
+   ca.res <- svd(A.n)
+   N <- sum(A)
+   d1.r <- ca.res$d[2] * ca.res$u[, 2] * sqrt(N/rs)
+   d1.c <- ca.res$d[2] * ca.res$v[, 2] * sqrt(N/cs)
+   d1 <- c(d1.r, d1.c)
+   d2.r <- ca.res$d[3] * ca.res$u[, 3] * sqrt(N/rs)
+   d2.c <- ca.res$d[3] * ca.res$v[, 3] * sqrt(N/cs)
+   d2 <- c(d2.r, d2.c)
+```
+
+```{r Two Mode Eigenvector Scores}
+   eig.r <- eigen(A %*% t(A))$vectors
+   eig.c <- eigen(t(A) %*% A)$vectors
+```
+
+```{r Reflections Bump Charts}
    source(here("Functions", "reflections.R"))
@@
    dev.off()
-```
 
-```{r Reflections Bump Chart (Groups)}
    plot.dat <- b$group
@@
 ```
+
+```{r CA versus Reflections Scatter}
+   p.plot.dat <- data.frame(c = tm$ca.eigvec.r[, 1]*-1, r = a$p.s[, 26])
+   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_point(size = 3, color = "blue") 
+   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
+   p
+   dev.off()
+
+   g.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1]*-1, r = a$g.s[, 26]))
+   p <- ggplot(data = g.plot.dat, aes(x = c, y = r))
+   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
+   p <- p + geom_point(size = 3, color = "red") 
+   p <- p + geom_text_repel(aes(label = rownames(g.plot.dat)), 
+                            color = "blue", size = 6)
+   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
+   p <- p + theme(axis.text = element_text(size = 16),
+                  axis.title = element_text(size = 16))
+   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
+   p
+   dev.off()
+```
+
+```{r CA versus Bonacich Re-ordered Affiliation Matrices}
+      # Affiliation matrix plot (CA ordering)
+      A.ord <- as.matrix(A[order(d1.r*-1), order(d1.c)])
+      A.plot <- ggcorrplot(t(A.ord)) 
+      A.plot <- A.plot + scale_x_discrete(position = "top")
+      A.plot1 <- A.plot + theme(legend.position = "none",
+                               axis.text.x = element_text(hjust = -0.2))
+      # Affiliation matrix plot (Bonacich ordering)
+      A.ord <- as.matrix(A[order(eig.r[, 1]*-1), order(eig.c[, 1])])
+      A.plot <- ggcorrplot(t(A.ord)) 
+      A.plot <- A.plot + scale_x_discrete(position = "top")
+      A.plot2 <- A.plot + theme(legend.position = "none",
+                               axis.text.x = element_text(hjust = -0.2))
+
+     # Affiliation matrix plot (Bonacich ordering)
+      A.ord2 <- as.matrix(A[order(eig.r[, 2]*-1), order(eig.c[, 2])])
+      A.plot <- ggcorrplot(t(A.ord2)) 
+      A.plot <- A.plot + scale_x_discrete(position = "top")
+      A.plot3 <- A.plot + theme(legend.position = "none",
+                               axis.text.x = element_text(hjust = -0.2))
+
+      png(file = here("Plots", "ca-reord.png"), width=500, height=500)
+      A.plot1
+      dev.off() 
+      png(file = here("Plots", "bon-reord1.png"), width=500, height=500)
+      A.plot2
+      dev.off() 
+      png(file = here("Plots", "bon-reord2.png"), width=500, height=500)
+      A.plot3
+      dev.off()
+```
+
+```{r Row Schastic Projections Plot}
+   D.p <- diag(1/rowSums(A)) #inverse of degree matrix of persons
+   P.pg <- D.p %*% A
+   rownames(P.pg) <- rownames(A) 
+   D.g <- diag(1/colSums(A)) #inverse of degree matrix of groups
+   P.gp <- D.g %*% t(A)
+   rownames(P.gp) <- colnames(A)
+   P.pp <- P.pg %*% P.gp
+   P.gg <- P.gp %*% P.pg
+   rownames(P.pp) <- substr(colnames(P.pp), start = 1, stop = 3)
+   rownames(P.gg) <- substr(colnames(P.gg), start = 1, stop = 3)
+   P.p <- ggcorrplot(P.pp, lab = TRUE) +
+      theme(legend.position = "none", 
+            axis.text.x = element_text(hjust = -0.2)) + 
+      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
+      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
+   P.g <- ggcorrplot(P.gg, lab = TRUE) +
+      theme(legend.position = "none", 
+            axis.text.x = element_text(hjust = -0.2)) + 
+      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
+      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
+   png(file = here("Plots", "p-norm.png"), width=600, height=500)
+   P.p
+   dev.off() 
+   png(file = here("Plots", "g-norm.png"), width=600, height=500)
+   P.g
+   dev.off() 
+```
+
+```{r Normalized similarity Matrices}
+      S.r <- A %*% diag(1/cs) %*% t(A) #degree-normalized similarity matrix for rows
+      S.c <- t(A) %*% diag(1/rs) %*% A #degree-normalized similarity matrix for columns
+      eig.Sr <- eigen(S.r)
+      eig.Sc <- eigen(S.c)
+```
+
+```{r CA Eigenvector Plot} 
+      eigvec.dat.r <- data.frame(rank = rank(d1.r), 
+                             value = as.numeric(d1.r),
+                             lab = names(d1.r))
+      eigvec.dat.c <- data.frame(rank = rank(d1.c), 
+                             value = as.numeric(d1.c),
+                             lab = names(d1.c))
+      eigval.dat.r <- data.frame(k = 1:nrow(A), value = as.numeric(eig.Sr$values)) %>% 
+             mutate(value = value / max(value))
+      eigval.dat.c <- data.frame(k = 1:ncol(A), value = as.numeric(eig.Sc$values)) %>% 
+             mutate(value = value / max(value))
+
+      # Eigenvector plot for the people
+      p <- ggscatter(eigvec.dat.r, y = "rank", x = "value", point = FALSE,
+                 font.label =  10, label = "lab", repel = TRUE)
+      p <- p + theme(axis.text.y = element_blank(),
+                 axis.line.x = element_blank(),
+                 axis.line.y = element_blank(),
+                 axis.ticks = element_blank(),
+                 legend.position = "none")
+      p <- p + labs(y = "Rank", x = "First Axis") 
+      a <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")
+
+      # Eigenvector plot for the groups
+      p <- ggscatter(eigvec.dat.c, y = "rank", x = "value", point = FALSE,
+                 font.label =  10, label = "lab", repel = TRUE)
+      p <- p + theme(axis.text.y = element_blank(),
+                 axis.line.x = element_blank(),
+                 axis.line.y = element_blank(),
+                 axis.ticks = element_blank(),
+                 legend.position = "none")
+      p <- p + labs(y = "Rank", x = "First Axis") 
+      b <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")
+
+      # Eigenvalue plot for the people
+      p <- ggscatter(eigval.dat.r, x = "k", y = "value", size = 2, color = "black")
+      p <- p + labs(y = "", x = "k") 
+      c <- p + theme(axis.text.x = element_blank(),
+                    axis.text.y = element_text(size = 8),
+                    axis.line.y = element_blank(),
+                    axis.line.x = element_blank())
+
+      # Eigenvalue plot for the groups
+      p <- ggscatter(eigval.dat.c, x = "k", y = "value", size = 2, color = "black")
+      p <- p + labs(y = "", x = "k") 
+      d <- p + theme(axis.text.x = element_blank(),
+                    axis.text.y = element_text(size = 8),
+                    axis.line.y = element_blank(),
+                    axis.line.x = element_blank())
+      a <- a + inset_element(c, left = 0.7, right = 1, bottom = 0, top = 0.45)
+      b <- b + inset_element(d, left = 0.7, right = 1, bottom = 0, top = 0.45)
+      png(file = here("Plots", "p-ca-eigvec.png"), width=500, height=500)
+      a
+      dev.off() 
+      png(file = here("Plots", "g-ca-eigvec.png"), width=500, height=500)
+      b
+      dev.off() 
+```
+
+```{r CA Similarity and Correspondence Plot}
+   norm.sim.r <- as.matrix(S.r/max(S.r))
+   norm.sim.c <- as.matrix(S.c/max(S.c))
+   norm.sim.r <- norm.sim.r[order(d1.r), order(d1.r)]
+   norm.sim.c <- norm.sim.c[order(d1.c), order(d1.c)]
+
+   # CA Similarity Plot (People)
+   p <- ggcorrplot(norm.sim.r)
+   p <- p + scale_x_discrete(position = "top")
+   p <- p + scale_y_discrete(limits = rev)
+   a <- p + theme(legend.position = "none",
+                 axis.text.x = element_text(hjust = -0.2, size = 10))
+
+   # CA Similarity Plot (Groups)
+   p <- ggcorrplot(norm.sim.c)
+   p <- p + scale_x_discrete(position = "top")
+   p <- p + scale_y_discrete(limits = rev)
+   b <- p + theme(legend.position = "none",
+                 axis.text.x = element_text(hjust = -0.2, size = 10)) 
+   # CA Correspondence Plot
+   plot.dat <- data.frame(d1, d2, mode = c(rep(1, 18), rep(2, 14)))
+   p <- ggplot(plot.dat, aes(y = d2, x = d1, color = factor(mode), label = rownames(plot.dat)))
+   p <- p + geom_vline(xintercept = 0, color = "gray")
+   p <- p + geom_hline(yintercept = 0, color = "gray")
+   p <- p + geom_text_repel(size = 4, max.overlaps = 50)
+   p <- p + theme_minimal() 
+   p <- p + theme(legend.position = "none",
+                  axis.text = element_text(size = 18),
+                  axis.title = element_text(size = 20)
+                  )
+   p <- p + labs(x = "Second Dimension", y = "First Dimension")
+   c <- p + scale_color_manual(values = c('red', 'blue'))
+   png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=900)
+   (a + b) / c
+   dev.off()
+```
+
+```{r Bonacich Similarity and Correspondence Plot}
+   AAt <- A %*% t(A)
+   AtA <- t(A) %*% A
+   norm.sim.r <- as.matrix(AAt/max(AAt))
+   norm.sim.c <- as.matrix(AtA/max(AtA))
+   norm.sim.r <- norm.sim.r[order(eig.r[, 1]), order(eig.r[ ,1])]
+   norm.sim.c <- norm.sim.c[order(eig.c[, 1]), order(eig.c[ ,1])]
+
+   # Bonacich Similarity Plot (People)
+   p <- ggcorrplot(norm.sim.r)
+   p <- p + scale_x_discrete(position = "top")
+   p <- p + scale_y_discrete(limits = rev)
+   a <- p + theme(legend.position = "none",
+                 axis.text.x = element_text(hjust = -0.2, size = 10))
+
+   # Bonacich Similarity Plot (Groups)
+   p <- ggcorrplot(norm.sim.c)
+   p <- p + scale_x_discrete(position = "top")
+   p <- p + scale_y_discrete(limits = rev)
+   b <- p + theme(legend.position = "none",
+                 axis.text.x = element_text(hjust = -0.2, size = 10)) 
+   # Bonacich Correspondence Plot
+   eig.d1 <- c(eig.r[, 1], eig.c[, 1])
+   eig.d2 <- c(eig.r[, 2], eig.c[, 2])
+   plot.dat <- data.frame(eig.d1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
+   p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(mode), 
+                             label = c(rownames(A), colnames(A))))
+   p <- p + geom_vline(xintercept = 0, color = "gray")
+   p <- p + geom_hline(yintercept = 0, color = "gray")
+   p <- p + geom_text_repel(size = 4, max.overlaps = 50)
+   p <- p + theme_minimal() 
+   p <- p + theme(legend.position = "none",
+                  axis.text = element_text(size = 18),
+                  axis.title = element_text(size = 20)
+                  )
+   p <- p + labs(x = "Second Dimension", y = "First Dimension")
+   c <- p + scale_color_manual(values = c('red', 'blue'))
+   png(file = here("Plots", "bon-corr-plot.png"), width=1200, height=900)
+   (a + b) / c
+   dev.off()
+```

### 2025-01-31 Reviews and Response/response-memo.tex
Updates from Overleaf

@@
 
-\textbf{Response}:
+\textbf{Response}: Thanks very much for raising this point, which made me think about and consider issues that I missed in the first version of the paper.
+You are right that the second dimension of the Bonacich centrality reveals a dimension of ``groupiness'' in the data; however, this dimension is not quite the same as that of CA.
+In fact, it became clear to me that what it reveals are continuous approximations to structural equivalence in the data, which makes sense if, indeed, the CA scores are continuous approximations to a regular equivalence style criterion.
+We can check this claim by comparing the blocking obtained from the second Bonacich dimension and that produced by Everett and Borgatti (2013, Table 5) using a proportion-matching criterion for structural equivalence and see that they agree closely.
+This reinforces my previous claim, which was poorly argued for in the previous version of the paper, that the Bonacich decomposition comes closer to reproducing the original ``input data'' than the CA decomposition.
 

### 2025-02-01 main-R1.tex

@@
 Secondly, some contributions focus on the similarities and differences between CA and factoring---those based on a spectral decomposition of a suitable matrix---techniques for assigning scores to the nodes of a two-mode network, where the scores exploit the duality between the two sets of entities and could be, under some circumstances, interpreted as ``centrality'' scores.
-I discuss each in turn.
-
-\subsection{CA as a Data Visualization Tool}
-
-\citet{borgatti1997network} provided an early and influential in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
+
+\subsection{Previous Work on CA and Two-Mode Networks}
+
+One of the earliest considerations of CA as a data-analytic tool for two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks.
+In that paper, \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people by the groups to which they belong,'' then it would seem that CA would be an appropriate analytic technique.
+According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (square quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation (in the least-squared error) sense of the similarities between row and column objects.
+Bonacich's discussion of CA came after his influential introduction of the Eigenvector centrality approach for two-mode network analysis.
+Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of the affiliation matrix---e.g., the Singular Value Decomposition (SVD).
+As we will see later, both can also be thought of as based on the spectral decomposition of the dual projection matrices.
+
+As we will see later (section~\ref{subsec:refeigen}), the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated.
+Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by the most representative members of them
+\citep[164]{bonacich1991simultaneous}.
+
+In an influential paper on centrality in two-mode networks, \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA.
+Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data.
+In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the event she attends; for events, their scores are given by the sum of the eigenvector centralities of its members.
+Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship was obtained between the CA scores corresponding to persons and groups.
+The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends, and for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event.
+\citet{faust1997centrality} also specified the particular way that the affiliation matrix $\mathbf{A}$ had to be modified before applying the SVD (recall this is the step that creates the difference between CA and Eigenvector centrality), noting that each cell $a_{pg}$ is divided by the square root of the product of the number of events the row woman attends and the number of people attending the column event ($\sqrt{\sum_p a_{pg}\sum_g a_{pg}}$), something that she further clarify in a subsequent contribution \citep{faust2005using}.
+
+\citet{borgatti1997network} provided another early and influential in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
 The discussion was placed in a section of the paper titled ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as a visualization technique.
@@
 Thirdly, Faust clarifies the duality linking the row and column scores, noting that the scores for the people are (activity) weighted sums of the scores that the groups they join, and the scores for the groups are (group size) weighted sums of the scores of their members.
+
 Finally, \citet[128ff]{faust2005using} clarifies the link between the CA scores---and different normalizations thereof---assigned to persons and events in an affiliation network on each dimension and (same node-set) inter-point distance in a joint space.
-Like \citet{roberts2000correspondence} before her, Faust clarifies that in the low (usually two) dimensional representation, inter-point distances are best (least squares) approximations of the chi-square distance between the membership profiles of pairs of people (or groups), where the profile of a given person-row or group-column is equal to the non-zero entries of the affiliation matrix divided by the corresponding row or column sum (people's activities and group sizes).\footnote{Note that because the chi-square distances are defined on these degree-weighted entries, they define a kind of degree-weighted similarity between pairs of people and pairs of groups, but not an exact similarity score as would be obtained, for instance, by using a criterion such as proportion matching \citep[208]{everett2013dual}.}
+Like \citet{roberts2000correspondence} before her, Faust specifies that in the low (usually two) dimensional representation, inter-point distances are best (least squares) approximations of the chi-square distance between the membership profiles of pairs of people (or groups), where the profile of a given person-row or group-column is equal to the non-zero entries of the affiliation matrix divided by the corresponding row or column sum (people's activities and group sizes).\footnote{Note that because the chi-square distances are defined on these degree-weighted entries, they define a kind of degree-weighted similarity between pairs of people and pairs of groups, but not an exact similarity score as would be obtained, for instance, by using a criterion such as proportion matching \citep[208]{everett2013dual}.}
 
@@
 
-\subsection{CA and Two-Mode Centrality}
-One of the earliest considerations of CA as a data-analytic tool for two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks.
-In that paper \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people buy the groups to which they belong,'' then it would seem that CA would be an appropriate analytic technique.
-According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (square quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation (in the least-squared error) sense of the similarities between row and column objects.
-Bonacich's discussion of CA came after his influential introduction of the Eigenvector centrality approach for two-mode network analysis.
-Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of a matrix---e.g., the Singular Value Decomposition (SVD).
-
-As we will see later, the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated.
-Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by the most representative members of them
-\citep[164]{bonacich1991simultaneous}.
-
-In an influential paper on centrality in two-mode networks, \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA.
-Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data.
-In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the event she attends; for events, their scores are given by the sum of the eigenvector centralities of its members.
-Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship was obtained between the CA scores corresponding to persons and groups.
-The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends, and for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event.
-\citet{faust1997centrality} also specified the particular way that the affiliation matrix $\mathbf{A}$ had to be modified before applying the SVD (recall this is the step that creates the difference between CA and Eigenvector centrality), noting that each cell $a_{pg}$ is divided by the square root of the product of the number of events the row woman attends and the number of people attending the column event ($\sqrt{\sum_p a_{pg}\sum_g a_{pg}}$).
-
-\section{Reflective Centralities in Two-Mode Networks} \label{sec:ref2mode}
-In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
+\section{Reflective Node Scores in Two-Mode Networks} \label{sec:ref2mode}
+In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of assigning meaningful scores to nodes in a two-mode network using what they called at the time a ``reflective'' approach.
 Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
@@
 
-If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}.
+If we are going to assign substantively meaningful scores to nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}.
 A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $G$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.
@@
 That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
-As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.
+
+As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order'' quantities $C^R(2)$ for both people and groups using the \textit{averaged} first-order centralities of the entities in the other mode they are connected to.
 
@@
 
-Equation \ref{eq:R2_p} says ``people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
-Equation \ref{eq:R2_g} says ``groups are more central when the average activity of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
-Of course, we can keep on going and define third-order reflections.
+While Equation~\ref{eq:R1_p} assigns a high value to people who belong to a lot of groups, Equation~\ref{eq:R2_p} assigns a high value to people who, on average, belong to large groups (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
+In the same way, while Equation~\ref{eq:R1_g} assigns a high value to groups that have lots of members, Equation~\ref{eq:R1_g} assigns a high value to groups that, on average, have members who themselves have lots of memberships (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
+Of course, we can keep on going and define ``third-order'' reflections.
 
@@
 
-Equation \ref{eq:R3_p} says something like ``people are more central when the average sum of the average activity of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
-Equation \ref{eq:R3_g} says, ''groups are more central when the average sum of the average size of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
+As we saw, while for people, Equation~\ref{eq:R2_p} measured the averaged size of the groups they join, Equation~\ref{eq:R3_p} assigns a high value to people who, on average, belong to groups who are themselves attended by highly active members (e.g., whenever $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
+In the same way, while Equation~\ref{eq:R2_g} assigns a high value to groups whose members have lots of memberships, Equation~\ref{eq:R3_g} assigns a high value to groups that, on average, have members who themselves (also on average) belong to large groups (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
 
@@
 It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.
-As \citet{freeman2003finding} notes, in the canonical original table, both the row and column modes had been re-ordered in a presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns.
+As \citet{freeman2003finding} notes, in the canonical original table included in \citet{davis1941}, both the row and column modes had been re-ordered in a presentation to reveal what DGG ascertained (using intuitive methods and in-depth knowledge of the case) was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns.
 
@@
 In fact, as shown in Figure~\ref{fig:sw}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.\footnote{Using the DGG labeling, the chronologically ordered set of events are \{E11, E5, E2, E7, E12, E9, E3, E6, E10, E1, E14 E8, E4, E13\} \citep[p.
-68]{everett2018measuring}.}
-
-To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women---taken from \citet[83]{homans1950human}---from that of the original DGG table reproduced in \citet{freeman2003finding}, and ordered the events from left to right in the affiliation matrix \textit{chronologically}, which, as we have seen, also differs from the original DGG Table reproduced in \citet{freeman2003finding}.
+68]{everett2018measuring}.} To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women---taken from \citet[83]{homans1950human}---from that of the original DGG table reproduced in \citet{freeman2003finding}, and ordered the events from left to right in the affiliation matrix \textit{chronologically}, which, as we have seen, also differs from the original DGG Table reproduced in \citet{freeman2003finding}.
 This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; in all, different versions of the dataset reproduce different versions of the labeling.
@@
 Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication \citep[e.g.,][]{freeman2003finding} may not correspond to "W1" or "E8" in the other \citep[e.g.,][]{doreian1979evolution}.
-
 To forestall confusion, in what follows, I use the numeric labeling corresponding to the original DGG data table as reproduced in (among others) \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{fig:sw}.
@@
 However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
-Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the dominant partition between the women observed by \citet{freeman2003finding}, but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
-
+
+In the exhaustive methodological meta-analysis mentioned earlier, \citet{freeman2003finding} noted that ``the consensus of the analytic procedures is to assign women 1 through 9 to one group and women 10 through 18 to the other.'' Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover this same dominant partition reading from top to bottom.
 Figure~\ref{fig:g-refs} shows the corresponding bump chart for the events.
-Just like for these actors, the equilibrium reflective centralities recover the ordering of the columns according to Doreian et al.
-generalized blockmodel with events 1-6 separated from events 10-15 by events 7-9 (see also \citet{kovacs2010generalized}).
 Clearly, $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
 Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering (the average of) the average number of memberships of its members (and so forth).
+Notably, the equilibrium reflective centralities recover and ordering of groups columns that re-appears across many articles that have analyzed these data with events 1-6 separated from events 10-15 by events 7-9 \citep[see e.g.,][]{doreian2004generalized, kovacs2010generalized, lizardo2024two, everett2013dual}.
 
@@
 
-\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks} \label{sec:ca}
-The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
-In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings.
-Instead, Bonacich noted that, in analyzing the Southern Women data shown in Figure~\ref{fig:sw}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
-
-As we will see, Bonacich was only half right.
-Indeed, the CA scores capture a version of ``clique'' (more accurately, community) membership, but they---particularly the first dimension---also capture a version of centrality in the most general sense of ranking nodes based on a meaningful criterion \citep{borgatti2006graph}.
-In fact, they capture the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
+\section{CA and Dual Centrality in Two-Mode Networks} \label{sec:ca}
+The reader may ask what the point of going through all these reflective scoring, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
+Recall that, as noted earlier in that paper, Bonacich also considered CA in passing but dismissed it as a tool for assigning scores to nodes in two-mode networks, noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.
+While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network.
+In fact, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
 
@@
 Where $A$ is the original affiliation matrix.
-Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, obtained from the singular value decomposition (SVD) of the rectangular affiliation matrix $A$.
+Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$.
 Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
@@
 Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich two-mode, the centrality of people is equal to the sum of the centralities scores of the groups they join, and the centrality of groups is equivalent to the sum of the centralities of their members.
+These equations make clear that in the Bonacich dual-scoring approach, the centrality of people is equal to the sum of the centralities scores of the groups they join, and the centrality of groups is equivalent to the sum of the centralities of their members.
 This approach thus nicely captures the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
@@
 This says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
-
 Of course, these sums of sums would diverge to a larger and larger quantity at each step $q$.
-To prevent this and guarantee convergence, we normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:
+To prevent this and guarantee convergence, we can normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:
 
@@
 Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}.
-
 Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations.
@@
 This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' centralities and the Bonacich two-mode centralities.
-In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.
-
-\subsection{HH Reflective Centrality as an Eigenvector Centrality} \label{subsec:refeigen}
+In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.\footnote{Note that this is precisely the difference between the Bonacich centrality in one-mode networks \citep{bonacich1972factoring} and the PageRank centrality \citep{brin1998anatomy}.}
+
+\subsection{HH Reflections and Correspondence Analysis} \label{subsec:refeigen}
 Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem.
@@
 \begin{equation}
-    \left(AA^T\right)C^B_p = \lambda^2 C^B_p
+    \left(AA^T\right)C^B_p = \lambda C^B_p
     \label{eq:bon3}
@@
 \begin{equation}
-    \left(A^TA\right)C^B_g = \lambda^2 C^B_g
+    \left(A^TA\right)C^B_g = \lambda C^B_g
     \label{eq:bon4}
@@
 Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose (recall that a matrix is row stochastic if its rows sum to one).
+
+For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
+
+\begin{equation}
+    \mathbf{P}_{pg} = \mathbf{D}p^{-1}\mathbf{A}
+    \label{eq:ca1}
+\end{equation}
+
+We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
+
+\begin{equation}
+    \mathbf{P}_{gp} = \mathbf{D}_G^{-1}\mathbf{A}^T
+    \label{eq:ca2}
+\end{equation}
+
+Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step---due to \citet{breiger1974duality}---of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
+
+For the people, the degree-weighted projection is:
+
+\begin{equation}
+\mathbf{P}_{pp} = \mathbf{P}_{pg}\mathbf{P}_{gp}
+    \label{eq:ca3}
+\end{equation}
+
+And for the groups:
+
+\begin{equation}
+    \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
+    \label{eq:ca4}
+\end{equation}
+
+Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
+Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
+Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person.
+Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
+The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).
 
@@
 
-For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
-
-\begin{equation}
-    \mathbf{P}_{pg} = \mathbf{D}p^{-1}\mathbf{A}
-    \label{eq:ca1}
-\end{equation}
-
-We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
-
-\begin{equation}
-    \mathbf{P}_{gp} = \mathbf{D}_G^{-1}\mathbf{A}^T
-    \label{eq:ca2}
-\end{equation}
-
-Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step---due to \citet{breiger1974duality}---of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
-
-For the people, the degree-weighted projection is:
-
-\begin{equation}
-\mathbf{P}_{pp} = \mathbf{P}_{pg}\mathbf{P}_{gp}
-    \label{eq:ca3}
-\end{equation}
-
-And for the groups:
-
-\begin{equation}
-    \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
-    \label{eq:ca4}
-\end{equation}
-
-Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
-Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
-Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person.
-Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
-The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).
-
 To get to the main point of all of this, it can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as a solution of the eigensystem:
@@
 Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective centrality scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms}.
-Note also that just like the eigenvector decomposition, the matrices implied by equations~\ref{eq:dam1} and~\ref{eq:dam2} have equivalent sets of (non-zero) eigenvalues.
+9.17]{fouss2016algorithms}.\footnote{Note also that just like the eigenvector decomposition, the matrices implied by equations~\ref{eq:dam1} and~\ref{eq:dam2} have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix.}
 
@@
         \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
-            \caption{Re-ordered Affiliation matrix (CA).}
+            \caption{Re-ordered Affiliation matrix (CA First Dimension).}
             \label{fig:ca-reord}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/bon-reord.png}
-            \caption{Re-ordered Affiliation matrix (Bonacich).}
+        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
+            \caption{Re-ordered Affiliation matrix (Bonacich Centrality).}
+            \label{fig:bon-reord`}
+    \end{subfigure}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
+            \caption{Re-ordered Affiliation matrix (Bonacich Second Dimension).}
             \label{fig:bon-reord}
     \end{subfigure}
-    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score.}
+    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score, (c) row and columns re-ordered according to person and group score on the second Bonacich dimension.}
     \label{fig:ca-v-bon}
@@
 
-\subsection{Row and Column Re-ordered Affiliation Matrices}
+\subsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
 Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
@@
 In fact, as \citet[p.
-206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in a the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to direct optimization methods.
+206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to direct optimization methods.
 The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
@@
     \centering
-     \begin{subfigure}[b]{0.425\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-ca-sim.png}
-            \caption{CA Similarity plot (Persons).}
-            \label{fig:p-ca-sim}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-ca-sim.png}
-            \caption{CA Similarity plot (Groups).}
-            \label{fig:g-ca-sim}
-    \end{subfigure} \\
-     \begin{subfigure}[b]{0.65\textwidth}
         \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
@@
             \label{fig:ca-corr}
-    \end{subfigure}
     \caption{CA similarity plots for persons (a) and groups (b) on the left and right top panel.
@@
     \centering
-     \begin{subfigure}[b]{0.425\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-bon-sim.png}
-            \caption{Bonacich Similarity plot (Persons).}
-            \label{fig:p-bon-sim}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-bon-sim.png}
-            \caption{Bonacich Similarity plot (Groups).}
-            \label{fig:g-bon-sim}
-    \end{subfigure} \\
-     \begin{subfigure}[b]{0.65\textwidth}
         \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
@@
             \label{fig:bon-corr}
-    \end{subfigure}
     \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel.
@@
 
-\subsection{One-Mode Projection Matrix and Eigevenctor Plot}
+\subsection{One-Mode Projection Matrix and Eigenvector Plot} \label{sec:eigplot}
 This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
@@
     \centering
-     \begin{subfigure}[b]{0.35\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/sr-plot-reord.png}
-            \caption{Re-ordered Affiliation Matrix (SimRank).}
-            \label{fig:p-bon-sim}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-sr-plot-eigen.png}
-            \caption{SimRank Eigenvector Plot (Persons).}
-            \label{fig:g-bon-sim}
-    \end{subfigure} 
-     \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-sr-plot-eigen.png}
-            \caption{SimRank Eigenvector Plot (Groups).}
-            \label{fig:bon-corr}
-    \end{subfigure} \\
-     \begin{subfigure}[b]{0.55\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/sr-corr-plot.png}
-            \caption{SimRank Correspondence Plot}
-            \label{fig:g-bon-sim}
-    \end{subfigure} 
-     \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/sr-corr-scatter.png}
-            \caption{SimRank/CA Correlation.}
-            \label{fig:sr-corr}
-    \end{subfigure}
+        \includegraphics[width=0.9\textwidth]{Plots/simrank-v-ca.png}
+            \label{fig:sim-v-ca}
     \caption{Simrank versus CA comparison.
-Panel (a) shows the affiliation matrix re-ordered according to the first eigenvector of the SVD of the SimRank similarity matrix.
-Panels (b) and (c) show the corresponding eigenvector and eigenvalue plots for persons and groups, respectively, of the SimRank similarity matrix.
-Panel (d) shows the correspondence plot based on the eigenvector decomposition of the first two non-trivial eigenvectors of the SimRank similarity matrix.
-Panel (e) shows the Pearson correlation between the scores corresponding to the first eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis) for persons (top) and groups (bottom).}
+Panels (a) and (c) show the final SimRank similarity matrices for persons and groups (respectively) with rows and columns ordered according to the first CA dimension scores.
+Panel (b) shows the Pearson correlation between the scores corresponding to the second eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis) for persons, panel (d) shows the same correlation for groups.}
     \label{fig:simrank}

### 2025-02-04 analysis-R1.qmd
plots

@@
 
-# Setup
-```{r setup}
+```{r Setup}
       knitr::opts_chunk$set(include = TRUE, 
@@
       conflicted::conflicts_prefer(dplyr::filter)
+      source(here("Functions", "reflections.R"))
+      source(here("Functions", "ref.long.dat.R"))
+      source(here("Functions", "SimRank.R"))
 ```
@@
 
-```{r Faust-Style CA}
-   rs <- rowSums(A)
-   cs <- colSums(A)
-   R <- diag(1/sqrt(rs))
-   C <- diag(1/sqrt(cs))
-   A.n <- R %*% A %*% C
-   ca.res <- svd(A.n)
-   N <- sum(A)
-   d1.r <- ca.res$d[2] * ca.res$u[, 2] * sqrt(N/rs)
-   d1.c <- ca.res$d[2] * ca.res$v[, 2] * sqrt(N/cs)
+```{r Faust-Style Direct CA}
+   ca.res <- svd(diag(1/sqrt(rowSums(A))) %*% A %*% diag(1/sqrt(colSums(A))))
+   d1.r <- ca.res$d[2] * ca.res$u[, 2] * sqrt(N/rowSums(A))
+   d1.c <- ca.res$d[2] * ca.res$v[, 2] * sqrt(N/colSums(A))
    d1 <- c(d1.r, d1.c)
-   d2.r <- ca.res$d[3] * ca.res$u[, 3] * sqrt(N/rs)
-   d2.c <- ca.res$d[3] * ca.res$v[, 3] * sqrt(N/cs)
+   d2.r <- ca.res$d[3] * ca.res$u[, 3] * sqrt(N/rowSums(A))
+   d2.c <- ca.res$d[3] * ca.res$v[, 3] * sqrt(N/colSums(A))
    d2 <- c(d2.r, d2.c)
+   d3.r <- ca.res$d[4] * ca.res$u[, 4] * sqrt(N/rowSums(A))
+   d3.c <- ca.res$d[4] * ca.res$v[, 4] * sqrt(N/colSums(A))
+   d3 <- c(d3.r, d3.c)
+   d4.r <- ca.res$d[5] * ca.res$u[, 5] * sqrt(N/rowSums(A))
+   d4.c <- ca.res$d[5] * ca.res$v[, 5] * sqrt(N/colSums(A))
+   d4 <- c(d4.r, d4.c)
 ```
 
-```{r Two Mode Eigenvector Scores}
+```{r Dual Projection CA}
+   P.pg <- diag(1/rowSums(A)) %*% A
+   rownames(P.pg) <- rownames(A) 
+   P.gp <- diag(1/colSums(A)) %*% t(A)
+   rownames(P.gp) <- colnames(A)
+   P.pp <- P.pg %*% P.gp
+   P.gg <- P.gp %*% P.pg
+   CA.r <- eigen(P.pp)
+   CA.c <- eigen(P.gg)
+   rownames(CA.r$vectors) <- rownames(A)
+   rownames(CA.c$vectors) <- colnames(A)
+```
+
+```{r Weighted Similarity CA}
+   S.r <- A %*% diag(1/colSums(A)) %*% t(A)
+   S.c <- t(A) %*% diag(1/rowSums(A)) %*% A
+   SCA.r <- eigen(diag(nrow(A)) - (diag(1/rowSums(A)) %*% S.r))
+   SCA.c <- eigen(diag(ncol(A)) - (diag(1/colSums(A)) %*% S.c))
+   rownames(SCA.r$vectors) <- rownames(A)
+   rownames(SCA.c$vectors) <- colnames(A)
+```
+
+```{Two Dual Projection Eigenvector Scores}
    eig.r <- eigen(A %*% t(A))$vectors
    eig.c <- eigen(t(A) %*% A)$vectors
+   rownames(eig.r) <- rownames(A)
+   rownames(eig.c) <- colnames(A)
 ```
@@
 ```{r Reflections Bump Charts}
-   source(here("Functions", "reflections.R"))
-   source(here("Functions", "ref.long.dat.R"))
    evens <- function(x) subset(x, x %% 2 == 0)
@@
 
-```{r Row Schastic Projections Plot}
+```{r Row Stochastic Matrices for CA Projections Plot}
    D.p <- diag(1/rowSums(A)) #inverse of degree matrix of persons
@@
 
-```{r Normalized similarity Matrices}
-      S.r <- A %*% diag(1/cs) %*% t(A) #degree-normalized similarity matrix for rows
-      S.c <- t(A) %*% diag(1/rs) %*% A #degree-normalized similarity matrix for columns
-      eig.Sr <- eigen(S.r)
-      eig.Sc <- eigen(S.c)
-```
-
 ```{r CA Eigenvector Plot} 
@@
                              lab = names(d1.c))
-      eigval.dat.r <- data.frame(k = 1:nrow(A), value = as.numeric(eig.Sr$values)) %>% 
+      eigval.dat.r <- data.frame(k = 1:nrow(A)-1, value = as.numeric(CA.r$values)) %>% 
              mutate(value = value / max(value))
-      eigval.dat.c <- data.frame(k = 1:ncol(A), value = as.numeric(eig.Sc$values)) %>% 
+      eigval.dat.c <- data.frame(k = 1:ncol(A)-1, value = as.numeric(CA.c$values)) %>% 
              mutate(value = value / max(value))
@@
       p <- p + scale_y_discrete(limits = rev)
-   b <- p + theme(legend.position = "none",
+      b <- p + theme(legend.position = "none", plot.margin = margin(l = 100), 
                     axis.text.x = element_text(hjust = -0.2, size = 10))
+
       # CA Correspondence Plot
-   plot.dat <- data.frame(d1, d2, mode = c(rep(1, 18), rep(2, 14)))
-   p <- ggplot(plot.dat, aes(y = d2, x = d1, color = factor(mode), label = rownames(plot.dat)))
+      plot.dat <- data.frame(a = d3, 
+                             b = d1, 
+                             mode = c(rep(1, 18), rep(2, 14)))
+      p <- ggplot(plot.dat, aes(y = a, x = b, color = factor(mode), label = rownames(plot.dat)))
       p <- p + geom_vline(xintercept = 0, color = "gray")
@@
       png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=900)
-   (a + b) / c
+      (a + b) / c + 
+         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
+         theme(plot.tag = element_text(size = 26))
       dev.off()
@@
       p <- p + scale_y_discrete(limits = rev)
-   b <- p + theme(legend.position = "none",
+      b <- p + theme(legend.position = "none", plot.margin = margin(l = 100),
                     axis.text.x = element_text(hjust = -0.2, size = 10))
+
       # Bonacich Correspondence Plot
@@
       png(file = here("Plots", "bon-corr-plot.png"), width=1200, height=900)
-   (a + b) / c
+      (a + b) / c + 
+         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
+         theme(plot.tag = element_text(size = 26))
       dev.off()
 ```
+
+```{r CA versus SimRank}
+      res <- SimRank(A)
+      Sr <- res$Sr
+      Sc <- res$Sc
+      s1.r <- eigen(Sr)$vectors[, 2]
+      s1.c <- eigen(Sc)$vectors[, 2]
+      diag(Sr) <- 0
+      diag(Sc) <- 0
+      Sr <- Sr[order(d1.r), order(d1.r)]
+      Sc <- Sc[order(d1.c), order(d1.c)]
+      rownames(Sr) <- substr(colnames(P.pp), start = 1, stop = 3)
+      rownames(Sc) <- substr(colnames(P.gg), start = 1, stop = 3)
+      a <- ggcorrplot(Sr, lab = TRUE, show.diag = FALSE, lab_size = 3) +
+         theme(legend.position = "none", 
+               axis.text.x = element_text(hjust = -0.2)) + 
+         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
+         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
+                              low = "white", mid = "yellow", high = "purple")
+
+      b <- ggcorrplot(Sc, lab = TRUE, show.diag = FALSE, lab_size = 3) +
+         theme(legend.position = "none", 
+               axis.text.x = element_text(hjust = -0.2)) + 
+         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
+         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
+                              low = "white", mid = "yellow", high = "purple")
+
+      dat.r <- data.frame(ca = d1.r,
+                          sr = s1.r,
+                          lab = rownames(A))
+      dat.c <- data.frame(ca = d1.c,
+                          sr = s1.c,
+                          lab = colnames(A))
+      cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
+
+      c <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
+                     label = "lab", repel = TRUE, 
+                     add = "reg.line", 
+                     add.params = list(color = "gray80", 
+                                       linewidth = 0.5, linetype = 2)) +
+                     annotate("text", x = -.2, y = .3, 
+                              label = paste("r = ", cor), size = 6) 
+      cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
+
+      d <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
+                     label = "lab", repel = TRUE, 
+                     add = "reg.line",
+                     add.params = list(color = "gray80", 
+                                       linewidth = 0.5, linetype = 2)) +
+                     annotate("text", x = -.2, y = .3, 
+                              label = paste("r = ", cor), size = 6) 
+      png(file = here("Plots", "simrank-v-ca.png"), width=1000, height=900)
+      (a + c + plot_layout(widths = c(2, 1))) / (b + d + plot_layout(widths = c(2, 1))) + 
+         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
+         theme(plot.tag = element_text(size = 26))
+      dev.off() 
+```

### 2025-02-04 main-R1.tex

@@
 
-\section{CA and Dual Centrality in Two-Mode Networks} \label{sec:ca}
-The reader may ask what the point of going through all these reflective scoring, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
-Recall that, as noted earlier in that paper, Bonacich also considered CA in passing but dismissed it as a tool for assigning scores to nodes in two-mode networks, noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.
-While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network.
-In fact, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
-
-To see this, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:
+\section{CA and Dual Scoring in Two-Mode Networks} \label{sec:ca}
+The reader may ask what the point of going through all these reflective scoring since \citet{bonacich1991simultaneous} already developed a dual conception of scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected.
+Recall that, as noted earlier in that paper, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.
+
+While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures like degree).
+In fact, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
+This makes the connection between the Bonacich and the CA scoring approaches more intimate than Bonacich thought.
+
+To begin to see the linkages between CA and the Bonacich approach, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:
 
@@
 Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$.
-Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
+Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific scores.
 
@@
 Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich dual-scoring approach, the centrality of people is equal to the sum of the centralities scores of the groups they join, and the centrality of groups is equivalent to the sum of the centralities of their members.
-This approach thus nicely captures the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
-Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
-The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.
+These equations make clear that in the Bonacich dual-scoring approach, the scores assigned to people are equal to the sum of the scores of the groups they join, and the scores assigned to groups are equivalent to the sum of the scores simultaneously assigned to their members.
+This approach thus nicely captures the duality property since the scores of nodes in each mode are given by aggregating their connections to nodes in the other mode.
+
+However, Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
+The main difference is that in HH's reflective equations, a person's centrality is proportional to the \textit{group-degree-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{person-degree-weighted} centrality of the people who are members.
 We will return to this crucial point later.
@@
 \subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
-To see the connection between the reflective and Bonacich two-mode centralities more clearly, we can motivate the Bonacich two-mode centrality using the same ``reflective'' approach we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}.
-Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centrality for two-mode networks~\citep{bonacich1991simultaneous}, but it will help clarify the similarities and differences between the HH and the Bonacich approaches.
+To see the connection between the CA and Bonacich dual scores more clearly, we can motivate the Bonacich eigenvector-based approach using the same ``reflective'' exposition we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}.
+Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centralities for two-mode networks; however, it will help clarify the similarities and differences between the HH and the Bonacich approaches.
+
 Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:
@@
 These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach.
-Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power'' method of \citet{mises1929praktische}).
-
-This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' centralities and the Bonacich two-mode centralities.
-In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.\footnote{Note that this is precisely the difference between the Bonacich centrality in one-mode networks \citep{bonacich1972factoring} and the PageRank centrality \citep{brin1998anatomy}.}
-
-\subsection{HH Reflections and Correspondence Analysis} \label{subsec:refeigen}
+Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power method'' of \citet{mises1929praktische}).
+
+This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' scores and the Bonacich two-mode eigenvector centralities.
+In fact, both can be seen as instantiating an underlying ``prismatic'' model---in the sense of \citet{podolny2001networks}---of how centrality is distributed in the two-mode network.
+The Bonacich approach, based on sums of sums, favors overall \textit{volume}---or two-mode ``total effects'' in the sense of \citet{friedkin1991theoretical}---thus, groups receive more centrality points from their more central members and fewer centrality points from their less central members.
+Likewise, people receive more centrality points from their more central memberships and fewer centrality points from their less central memberships, and so on through every iteration.
+The HH reflective approach (and CA by implication) favors \textit{averages} rather than volume; thus, a person with a few memberships can contribute as much centrality to the group they belong to as a person with many memberships if the person with fewer memberships belongs to (on average) more central groups and the person with more memberships belongs to (on average) less central ones.
+In the same way, a person can receive as much centrality from a small group they belong to compared to a large group, as long as the small group (on average) has more central members in it, and the large group (on average) has less central members.
+
+\subsection{HH Reflections as Correspondence Analysis} \label{subsec:refeigen}
 Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem.
@@
 \begin{equation}
-    \left(AA^T\right)C^B_p = \lambda C^B_p
+    \left(\mathbf{AA}^T\right)C^B_p = \lambda C^B_p
     \label{eq:bon3}
@@
 \begin{equation}
-    \left(A^TA\right)C^B_g = \lambda C^B_g
+    \left(\mathbf{A}^T\mathbf{A}\right)C^B_g = \lambda C^B_g
     \label{eq:bon4}
@@
 
-This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.
-
-Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\cite{bonacich1991simultaneous, faust2005using}.
-Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.
-
-To see this, consider the $|P| \times |P|$ matrix $Dp$, containing the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
-In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
+This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.
+
+We can motivate CA using the same dual projection strategy.
+Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\cite{bonacich1991simultaneous, faust2005using}, which gives the impression that CA is exclusively a ``direct'' rather than a dual projection method.
+Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection approach just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.
+
+To see this, consider the $|P| \times |P|$ matrix $\mathbf{D}p$, containing the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
+In the same way, consider the $|G| \times |G|$ matrix $\mathbf{D}g$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
 Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose (recall that a matrix is row stochastic if its rows sum to one).
@@
 
-We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
+We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the row-stochastic matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
 
@@
 Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
-Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
+Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), positive definite, but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
 Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person.
@@
 
-With the $C^R$ scores obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the degree-weighted projection matrices.
-This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
+With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices.\footnote{Note also that the degree normalized projection matrices have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix, with values between zero and one \citep{van2021correspondence}.}  This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
 Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective centrality scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms}.\footnote{Note also that just like the eigenvector decomposition, the matrices implied by equations~\ref{eq:dam1} and~\ref{eq:dam2} have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix.}
-
-More importantly, because $\mathbf{P}_{pp} = Dp^{-1}ADg^{-1}A^T$ and $\mathbf{P}_{gg} = Dg^{-1}A^TDp^{-1}A$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ and $\mathbf{P}_{pp}$ for people and $A^TA$ and $\mathbf{P}_{gg}$ for groups.
+9.17]{fouss2016algorithms} or via the eigenvalue decomposition of the weighted projection matrices as in~\ref{eq:dam1} and~\ref{eq:dam2}.
+And as we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix.
+
+More importantly, because $\mathbf{P}_{pp} = \mathbf{D}p^{-1}\mathbf{A}\mathbf{D}g^{-1}\mathbf{A}^T$ and $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
+Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups.
 The key difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
 
-This can be clearly seen if we express $\left(Dg^{-1}\mathbf{A}^TDp^{-1}\mathbf{A}\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
+This can be clearly seen if we express $\mathbf{P}_{PP}= \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
 
@@
 
-The same reasoning applies to groups in~\ref{eq:dam2}, whose entries are given by:
+The same reasoning applies to groups in the matrix $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$, whose entries are given by:
 
@@
 
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}\mathbf{A}D_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
-As we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix for both the row and column objects.
+Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}; accordingly, the first non-trivial CA dimension for persons and groups given by the right and left eigenvectors corresponding to the second-largest eigenvalue from the SVD of the matrix $\mathbf{D}_p^{-1/2}\mathbf{A}\mathbf{D}_g^{-1/2}$ \citep[126]{faust2005using}.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
 
@@
             \caption{Re-ordered Affiliation matrix (Bonacich Centrality).}
-            \label{fig:bon-reord`}
+            \label{fig:bon-reord1}
     \end{subfigure}
@@
             \caption{Re-ordered Affiliation matrix (Bonacich Second Dimension).}
-            \label{fig:bon-reord}
+            \label{fig:bon-reord2}
     \end{subfigure}
@@
 
-\subsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
-Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
+\subsection{Comparing Bonacich and CA Dual Scoring in the Southern Women Data}
+\subsubsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
+Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
 Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
@@
 In fact, as \citet[p.
-206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to direct optimization methods.
+206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods.
 The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
 
-\subsection{Eigenvector Plot} \label{subsec:eigplot}
+\subsubsection{Eigenvector Plot} \label{subsec:eigplot}
 We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
@@
 Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
-Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach.
+Note that this is the canonical consensus partitioning (for the people) highlighted by \citet{freeman2003finding} across twenty-one different methods and subsequently reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity (GRS) approach.
 This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.
@@
             \caption{Eigenvector plot (Persons).}
-            \label{fig:ca-reord}
+            \label{fig:p-eigvec}
     \end{subfigure}
@@
             \caption{Eigenvector plot (Groups).}
-            \label{fig:bon-reord}
+            \label{fig:g-eigvec}
     \end{subfigure}
@@
 
-\subsection{Normalized Similarity Plot} \label{subsec:normsim}
-To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one.
-Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
+\subsubsection{Normalized Similarity Plot} \label{subsec:normsim}
+To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = \mathbf{AD}g^{-1}\mathbf{A}^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors in a paper---minus one.
+Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = \mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
 As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
@@
 
-\subsection{CA Correspondence Plot} \label{subsec:caplot}
+\subsubsection{CA Correspondence Plot} \label{subsec:caplot}
 The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions.
@@
 People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm).
-The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart.
+The same goes for groups; those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart.
 The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.
@@
 
-\subsection{One-Mode Projection Matrix and Eigenvector Plot} \label{sec:eigplot}
+\subsubsection{One-Mode Projection Matrix and Eigenvector Plot} \label{sec:eigplot}
 This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.

### 2025-02-20 analysis-R1.qmd
plots!

@@
 ```{r Faust-Style Direct CA}
-   ca.res <- svd(diag(1/sqrt(rowSums(A))) %*% A %*% diag(1/sqrt(colSums(A))))
-   d1.r <- ca.res$d[2] * ca.res$u[, 2] * sqrt(N/rowSums(A))
-   d1.c <- ca.res$d[2] * ca.res$v[, 2] * sqrt(N/colSums(A))
-   d1 <- c(d1.r, d1.c)
-   d2.r <- ca.res$d[3] * ca.res$u[, 3] * sqrt(N/rowSums(A))
-   d2.c <- ca.res$d[3] * ca.res$v[, 3] * sqrt(N/colSums(A))
-   d2 <- c(d2.r, d2.c)
-   d3.r <- ca.res$d[4] * ca.res$u[, 4] * sqrt(N/rowSums(A))
-   d3.c <- ca.res$d[4] * ca.res$v[, 4] * sqrt(N/colSums(A))
-   d3 <- c(d3.r, d3.c)
-   d4.r <- ca.res$d[5] * ca.res$u[, 5] * sqrt(N/rowSums(A))
-   d4.c <- ca.res$d[5] * ca.res$v[, 5] * sqrt(N/colSums(A))
-   d4 <- c(d4.r, d4.c)
+   FCA <- svd(diag(1/sqrt(rowSums(A))) %*% A %*% diag(1/sqrt(colSums(A))))
+   w.r <- sqrt(sum(A)/rowSums(A))
+   w.c <- sqrt(sum(A)/colSums(A))
+   FCA$d <- FCA$d[2:14]
+   FCA$u <- FCA$u[, 2:14]
+   FCA$v <- FCA$v[, 2:14]
+   #FCA$u <- FCA$u * FCA$d * w.r
+   #FCA$v <- FCA$v * FCA$d * w.c
+   FCA.d <- rbind(FCA$u, FCA$v)
+   rownames(FCA.d) <- c(rownames(A), colnames(A))
+   round(FCA.d, 3)
 ```
@@
 
-```{Two Dual Projection Eigenvector Scores}
+```{r Dual Projection Eigenvector Scores}
    eig.r <- eigen(A %*% t(A))$vectors
@@
 ```{r CA versus Reflections Scatter}
-   p.plot.dat <- data.frame(c = tm$ca.eigvec.r[, 1]*-1, r = a$p.s[, 26])
+   p.plot.dat <- data.frame(c = CA.r$vectors[, 2]*-1, 
+                            r = a$p.s[, 26])
    p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
@@
 
-   g.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1]*-1, r = a$g.s[, 26]))
+   g.plot.dat <- data.frame(cbind(c = CA.c$vectors[, 2]*-1, 
+                                  r = a$g.s[, 26]))
    p <- ggplot(data = g.plot.dat, aes(x = c, y = r))
@@
 ```{r Row Stochastic Matrices for CA Projections Plot}
-   D.p <- diag(1/rowSums(A)) #inverse of degree matrix of persons
-   P.pg <- D.p %*% A
+   P.pg <- diag(1/rowSums(A)) %*% A
    rownames(P.pg) <- rownames(A) 
-   D.g <- diag(1/colSums(A)) #inverse of degree matrix of groups
-   P.gp <- D.g %*% t(A)
+   P.gp <- diag(1/colSums(A)) %*% t(A)
    rownames(P.gp) <- colnames(A)
@@
    rownames(P.gg) <- substr(colnames(P.gg), start = 1, stop = 3)
-   P.p <- ggcorrplot(P.pp, lab = TRUE) +
+   P.p <- ggcorrplot(P.pp, lab = TRUE, lab_size = 5,
+                     colors = c(rep("white", 3))) +
       theme(legend.position = "none", 
-            axis.text.x = element_text(hjust = -0.2)) + 
-      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
-      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
-   P.g <- ggcorrplot(P.gg, lab = TRUE) +
+            axis.text.x = element_text(hjust = -0.2, size = 16),
+            axis.text.y = element_text(size = 16)) + 
+      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
+   P.g <- ggcorrplot(P.gg, lab = TRUE, lab_size = 5,
+                     colors = c(rep("white", 3))) +
       theme(legend.position = "none", 
-            axis.text.x = element_text(hjust = -0.2)) + 
-      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
-      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
-   png(file = here("Plots", "p-norm.png"), width=600, height=500)
+            axis.text.x = element_text(hjust = -0.2, size = 16),
+            axis.text.y = element_text(size = 16)) + 
+      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
+   png(file = here("Plots", "p-norm.png"), width=900, height=900)
    P.p
    dev.off() 
-   png(file = here("Plots", "g-norm.png"), width=600, height=500)
+   png(file = here("Plots", "g-norm.png"), width=900, height=900)
    P.g
@@
 ```{r CA Eigenvector Plot} 
-      eigvec.dat.r <- data.frame(rank = rank(d1.r), 
-                             value = as.numeric(d1.r),
-                             lab = names(d1.r))
-      eigvec.dat.c <- data.frame(rank = rank(d1.c), 
-                             value = as.numeric(d1.c),
-                             lab = names(d1.c))
-      eigval.dat.r <- data.frame(k = 1:nrow(A)-1, value = as.numeric(CA.r$values)) %>% 
-             mutate(value = value / max(value))
-      eigval.dat.c <- data.frame(k = 1:ncol(A)-1, value = as.numeric(CA.c$values)) %>% 
+      eigvec.dat.r <- data.frame(rank = rank(CA.r$vectors[,2]), 
+                             value = as.numeric(CA.r$vectors[,2]),
+                             lab = names(CA.r$vectors[,2]))
+      eigvec.dat.c <- data.frame(rank = rank(CA.c$vectors[,2]), 
+                             value = as.numeric(CA.c$vectors[,2]),
+                             lab = names(CA.c$vectors[,2]))
+      eigval.dat.c <- data.frame(k = 1:ncol(A)-1, 
+                                 value = as.numeric(CA.c$values)) %>% 
              mutate(value = value / max(value))
@@
                  axis.ticks = element_blank(),
+                 plot.margin = margin(l = 100),
                  legend.position = "none")
@@
 
-      # Eigenvalue plot for the people
-      p <- ggscatter(eigval.dat.r, x = "k", y = "value", size = 2, color = "black")
+      # Eigenvalue plot for the groups
+      p <- ggscatter(eigval.dat.c, x = "k", y = "value", size = 2, color = "black")
       p <- p + labs(y = "", x = "k") 
@@
                     axis.line.x = element_blank())
-
-      # Eigenvalue plot for the groups
-      p <- ggscatter(eigval.dat.c, x = "k", y = "value", size = 2, color = "black")
-      p <- p + labs(y = "", x = "k") 
-      d <- p + theme(axis.text.x = element_blank(),
-                    axis.text.y = element_text(size = 8),
-                    axis.line.y = element_blank(),
-                    axis.line.x = element_blank())
-      a <- a + inset_element(c, left = 0.7, right = 1, bottom = 0, top = 0.45)
-      b <- b + inset_element(d, left = 0.7, right = 1, bottom = 0, top = 0.45)
-      png(file = here("Plots", "p-ca-eigvec.png"), width=500, height=500)
-      a
-      dev.off() 
-      png(file = here("Plots", "g-ca-eigvec.png"), width=500, height=500)
-      b
+      b <- b + inset_element(c, left = 0.6, right = 1, bottom = 0, top = 0.45)
+      png(file = here("Plots", "ca-eigvec.png"), width=1000, height=600)
+      a + b + plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
+         theme(plot.tag = element_text(size = 20))
       dev.off() 
@@
 ```{r CA Similarity and Correspondence Plot}
-      norm.sim.r <- as.matrix(S.r/max(S.r))
-      norm.sim.c <- as.matrix(S.c/max(S.c))
-      norm.sim.r <- norm.sim.r[order(d1.r), order(d1.r)]
-      norm.sim.c <- norm.sim.c[order(d1.c), order(d1.c)]
+      norm.sim.r <- as.matrix(P.pp/max(P.pp))
+      norm.sim.c <- as.matrix(P.gg/max(P.gg))
+      norm.sim.r <- norm.sim.r[order(CA.r$vectors[,2]), 
+                               order(CA.r$vectors[,2])]
+      norm.sim.c <- norm.sim.c[order(CA.c$vectors[,2]), 
+                               order(CA.c$vectors[,2])]
 
@@
       p <- p + scale_y_discrete(limits = rev)
-      b <- p + theme(legend.position = "none", plot.margin = margin(l = 100), 
+      b <- p + theme(legend.position = "none", 
+                     plot.margin = margin(l = 100), 
                     axis.text.x = element_text(hjust = -0.2, size = 10))
@@
       # CA Correspondence Plot
-      plot.dat <- data.frame(a = d3, 
-                             b = d1, 
+      plot.dat <- data.frame(a = FCA.d[, 2], 
+                             b = FCA.d[, 1], 
                              mode = c(rep(1, 18), rep(2, 14)))
@@
                      )
-      p <- p + labs(x = "Second Dimension", y = "First Dimension")
+      p <- p + labs(x = "First Dimension", y = "Second Dimension")
       c <- p + scale_color_manual(values = c('red', 'blue'))
-      png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=900)
-      (a + b) / c + 
+
+      plot.dat <- data.frame(a = FCA.d[, 3], 
+                             b = FCA.d[, 1], 
+                             mode = c(rep(1, 18), rep(2, 14)))
+      p <- ggplot(plot.dat, aes(y = a, x = b, color = factor(mode), label = rownames(plot.dat)))
+      p <- p + geom_vline(xintercept = 0, color = "gray")
+      p <- p + geom_hline(yintercept = 0, color = "gray")
+      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
+      p <- p + theme_minimal() 
+      p <- p + theme(legend.position = "none",
+                     axis.text = element_text(size = 18),
+                     axis.title = element_text(size = 20)
+                     )
+      p <- p + labs(x = "First Dimension", y = "Third Dimension")
+      d <- p + scale_color_manual(values = c('red', 'blue'))
+
+      png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=1200)
+      (a + b) / c / d + 
          plot_annotation(tag_levels = 'a', tag_suffix = ')') & 

### 2025-02-20 main-R1.tex

@@
     \centering
-     \begin{subfigure}[b]{0.45\textwidth}
+     \begin{subfigure}[b]{0.55\textwidth}
         \includegraphics[width=1.0\textwidth]{Plots/p-norm.png}
-            \caption{Degree-weighted projection (People).}
-            \label{fig:ca-reord}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
+            %\caption{Degree-weighted projection (People).}
+            %\label{fig:ca-reord}
+    \end{subfigure} \\
+     \begin{subfigure}[b]{0.55\textwidth}
         \includegraphics[width=1.0\textwidth]{Plots/g-norm.png}
-            \caption{Degree-weighted projection (Groups).}
-            \label{fig:bon-reord}
+            %\caption{Degree-weighted projection (Groups).}
+            %\label{fig:bon-reord}
     \end{subfigure}
@@
         \includegraphics[width=1.0\textwidth]{Plots/p-ca-ref-corr.png}
-            \caption{CA/Reflection Correlation (Persons).}
+            \caption{}
             \label{fig:p-ca-ref}
@@
         \includegraphics[width=1.0\textwidth]{Plots/g-ca-ref-corr.png}
-            \caption{CA/Reflection Correlation (Groups)}
+            \caption{}
             \label{fig:g-ca-ref}
@@
         \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
-            \caption{Re-ordered Affiliation matrix (CA First Dimension).}
+            \caption{}
             \label{fig:ca-reord}
@@
         \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
-            \caption{Re-ordered Affiliation matrix (Bonacich Centrality).}
+            \caption{}
             \label{fig:bon-reord1}
@@
         \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
-            \caption{Re-ordered Affiliation matrix (Bonacich Second Dimension).}
+            \caption{}
             \label{fig:bon-reord2}
@@
 \begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-ca-eigvec.png}
-            \caption{Eigenvector plot (Persons).}
-            \label{fig:p-eigvec}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-ca-eigvec.png}
-            \caption{Eigenvector plot (Groups).}
-            \label{fig:g-eigvec}
-    \end{subfigure}
+        \includegraphics[width=1.0\textwidth]{Plots/ca-eigvec.png}            
+        \label{fig:eigvec}
     \caption{CA Eigenvector plot for persons (a) and groups (b).
 The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis.
-Insets show the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
+The inset plot (c) shows the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
     \label{fig:ca-eigvec}

### 2025-02-20 main-R1.tex
Updates from Overleaf

@@
 
-\section{Previous uses of CA and MCA in Two-Mode Network Data Analysis}
-
-In this section, I consider previous contributions regarding the usefulness and limitations of CA as a methodological tool for the analysis of two-mode network data.
-For reasons of space and relevance, therefore my discussion is limited in two ways.
-First, I focus attention on (1) pieces discussing primarily methodological aspects of CA and (2) with regard to the analysis of two-mode network data or affiliation networks.\footnote{Naturally, I also exclude the rich literature that has used methods other than CA to analyze affiliation networks.} There are, of course, various substantive applications of CA to the analysis of social networks, as well as substantive and methodological applications of CA to the analysis of one-mode network data; these will not be part of the discussion here.
-
-With these restrictions in mind, the literature on CA and two-mode networks can be divided into two streams.
+\section{Previous uses of CA in Two-Mode Network Data Analysis}
+
+In this section, I consider previous contributions regarding the usefulness and limitations of simple CA as a methodological tool for the analysis of two-mode network data as applied directly to the affiliation matrix recording ties between actors and events.
+For reasons of space and relevance, therefore, my discussion is circumscribed in four ways: I focus attention on (1) pieces discussing primarily methodological or interpretive aspects of CA, (2) with regard to the analysis of two-mode network data or affiliation networks,\footnote{There are, of course, various substantive applications of CA for the analysis of two-mode networks~\citep[e.g.,][]{breiger2000tool, faust2002scaling, schweizer1991power, serino2024mapping, ragozini2018analysis} as well as substantive and methodological applications of CA to the analysis of one-mode network data~\citep[e.g.,][]{noma1985scaling, kumbasar1994systematic, lizardo2020correspondence}; these will not be part of the discussion here.} (3)  dealing with the usual binary affiliation matrix measuring a single relation between persons and groups at one point in time, where links are observed only between persons and groups.\footnote{As such, recent work dealing with the application of CA or other factoring approaches to time-varying, multiplex, multilevel, or valued two-mode network data \citep{ragozini2014correspondence, ragozini2015multiple, zhu2016correspondence} will also be outside the scope of my discussion.} Finally, (4) there has been a spate of recent work focusing on the uses and advantages of using Multiple Correspondence Analysis---a variant of CA applied to a disjunctive doubly centered version of the original affiliation matrix) to analyze and visualize two-mode networks \citep[e.g.,][]{dramalidis2016subset, desposito2014use, ragozini2014correspondence}.
+Because much of what I have to say deals with the uses and interpretation of the scores obtained from simple CA applied directly to the affiliation matrix (or its one-mode projections), my discussion will not be directly relevant to this line of work, except for a recent paper directly comparing the substantive interpretation of the $\chi^2$ distances in simple CA and MCA with regard to their substantive interpretation as indexes of structural similarity, which bears more directly to some of the subsequent discussion \citep{desposito2014comparison}.
+
+%With these restrictions in mind, the literature on CA, duality, and two-mode networks can be divided into two streams.
 The first contributions focus on the usefulness of CA as a tool for the visualization or ``joint display'' of the two sets of entities in an affiliation network.
@@
 \subsection{Previous Work on CA and Two-Mode Networks}
-
-One of the earliest considerations of CA as a data-analytic tool for two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks.
+One of the earliest considerations of CA as a data-analytic tool for examining duality relations in two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks.
 In that paper, \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people by the groups to which they belong,'' then it would seem that CA would be an appropriate analytic technique.
-According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (square quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation (in the least-squared error) sense of the similarities between row and column objects.
+According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (scare quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation---in the least-squared error sense---of the similarities between row and column objects.
 Bonacich's discussion of CA came after his influential introduction of the Eigenvector centrality approach for two-mode network analysis.
-Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of the affiliation matrix---e.g., the Singular Value Decomposition (SVD).
-As we will see later, both can also be thought of as based on the spectral decomposition of the dual projection matrices.
-
-As we will see later (section~\ref{subsec:refeigen}), the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated.
-Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by the most representative members of them
-\citep[164]{bonacich1991simultaneous}.
-
-In an influential paper on centrality in two-mode networks, \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA.
-Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data.
+Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of the affiliation matrix---namely, the Singular Value Decomposition (SVD).
+As we will see in section~\ref{subsec:refeigen}, both CA and centrality analysis are mathematically linked in a second way, as both are based on the spectral decomposition of the dual \textit{projections} of the affiliation matrix.
+In this particular respect, the mathematics connecting the Eigenvector approach to centrality analysis and CA are more similar than even Bonacich appreciated.
+
+For Bonacich, the main differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality,'' but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also do later in the paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by their most representative members  \citep[164]{bonacich1991simultaneous}.
+
+In an influential paper on centrality in two-mode networks, \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality, CA, and dual relations between actors and events in two-mode networks.
+Faust noted that one thing both techniques had in common was that they both produced scores that truly exploited the duality of the two modes in the data.
 In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the event she attends; for events, their scores are given by the sum of the eigenvector centralities of its members.
 Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship was obtained between the CA scores corresponding to persons and groups.
-The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends, and for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event.
-\citet{faust1997centrality} also specified the particular way that the affiliation matrix $\mathbf{A}$ had to be modified before applying the SVD (recall this is the step that creates the difference between CA and Eigenvector centrality), noting that each cell $a_{pg}$ is divided by the square root of the product of the number of events the row woman attends and the number of people attending the column event ($\sqrt{\sum_p a_{pg}\sum_g a_{pg}}$), something that she further clarify in a subsequent contribution \citep{faust2005using}.
-
-\citet{borgatti1997network} provided another early and influential in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect.
+The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends.
+Similarly, for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event, a classic case of duality via mutual definition.
+%\citet{faust1997centrality} also specified the particular way that the affiliation matrix $\mathbf{A}$ had to be modified before applying the SVD (recall this is the step that creates the difference between CA and Eigenvector centrality), noting that each cell of the affiliation matrix is divided by the square root of the product of the number of events the row woman attends and the number of people attending the column event.
+
+\citet{borgatti1997network} provided another early and influential in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect of actors and events in a single space.
 The discussion was placed in a section of the paper titled ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as a visualization technique.
@@
 
-In a paper published shortly thereafter \citet{roberts2000correspondence}, responded to some of the criticisms that \citet{borgatti1997network} levied at the use of CA for analyzing two-mode networks in their original piece.
+In a paper published shortly thereafter, \citet{roberts2000correspondence} responded to some of the criticisms \citet{borgatti1997network} levied at the use of CA for analyzing two-mode networks.
 First, \citet{roberts2000correspondence} noted that there were many examples in the standing of using CA to analyze binary non-frequency data and that, therefore, the uses of CA are not constrained by the nature or provenance of the numerical quantities recorded in each cell of a two-mode network affiliation matrix.
-\citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and the event points in the two-dimensional representation and one that used all thirteen dimensions.
-Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that even (and similarly for column profiles corresponding to events).
-Thus, the point distances are indeed interpretable as a kind of degree-weighted similarity between objects in each mode.
-
-Perhaps the most thorough discussion of the uses of CA as a tool for the ``joint display'' of affiliation networks is that provided by \citet{faust2005using}.
+\citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and the event points in the two-dimensional representation and one that used all thirteen dimensions.\footnote{In CA, the maximum number of dimensions is one minus the rank of the affiliation matrix, which in this case is fourteen, corresponding to the number of events.} Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that even (and similarly for column profiles corresponding to events).
+Thus, the point distances are indeed interpretable as a kind of \textit{degree-weighted similarity} between objects in each mode.
+This is a point that we will return to later.
+
+Perhaps the most thorough discussion of the uses of CA as a tool for the dual ``joint display'' of affiliation networks is that provided by \citet{faust2005using}.
 According to \citet[118]{faust2005using}, previous debates regarding the usefulness---or lack thereof---of CA as a visualization tool were marred by imprecision and lack of ``formal specification'' regarding the exact substantive meaning of the scores obtained via the procedure and a related lack of sustained discussion as to the exact ``relationship between the model and the input data.'' Faust's basic point is that CA provides a way to both place actors and events in an affiliation network in a joint space such that the inter-point distances are interpretable and such that the link to the original data (e.g., the affiliation network matrix) is transparent.
-
-\citet{faust2005using} makes several contributions in this piece.
-First, Faust specifies the exact matrix that is subject to the singular value decomposition when CA is applied directly to the affiliation matrix of a two-mode network $\mathbf{A}$; namely, a scaled version where $\mathbf{A}$ is pre-multiplied by a diagonal matrix containing the square root of its row sums and post-multiplied by a diagonal matrix containing the square root of its column sums (equivalent to dividing every non-zero entry of $a_{pg}$ by the square root of the product of the number of memberships of person $p$ and the number of members of group $g$ before applying the SVD.
-Second, Faust specifies the exact relationship linking the resulting row and column CA scores and the original affiliation matrix input, providing the ``reconstruction equation" that takes us from the left and right singular vectors and eigenvalues to the former.
-Thirdly, Faust clarifies the duality linking the row and column scores, noting that the scores for the people are (activity) weighted sums of the scores that the groups they join, and the scores for the groups are (group size) weighted sums of the scores of their members.
-
+Along the way, \citet{faust2005using} established several important points.
+First, Faust specifies the exact matrix that is subject to the singular value decomposition when CA is applied directly to the affiliation matrix of a two-mode network $\mathbf{A}$; namely, a scaled version where $\mathbf{A}$ is pre-multiplied by a diagonal matrix containing the square root of its row sums and post-multiplied by a diagonal matrix containing the square root of its column sums.
+This is equivalent, as both \citet{faust1997centrality} and \citet{roberts2000correspondence} had noted before, to dividing every non-zero entry of $a_{pg}$ by the square root of the product of the number of memberships of person $p$ and the number of members of group $g$ before applying the SVD.
+Second, Faust specifies the exact relationship linking the resulting row and column CA scores and the original affiliation matrix input, providing the ``reconstruction equation'' that takes us from the left and right singular vectors and eigenvalues to the former.
+Thirdly, as she did in the 1997 two-mode centrality paper, Faust clarifies the duality linking the row and column scores, noting that the scores for the people are (activity) weighted sums of the scores that the groups they join, and the scores for the groups are (group size) weighted sums of the scores of their members.
 Finally, \citet[128ff]{faust2005using} clarifies the link between the CA scores---and different normalizations thereof---assigned to persons and events in an affiliation network on each dimension and (same node-set) inter-point distance in a joint space.
-Like \citet{roberts2000correspondence} before her, Faust specifies that in the low (usually two) dimensional representation, inter-point distances are best (least squares) approximations of the chi-square distance between the membership profiles of pairs of people (or groups), where the profile of a given person-row or group-column is equal to the non-zero entries of the affiliation matrix divided by the corresponding row or column sum (people's activities and group sizes).\footnote{Note that because the chi-square distances are defined on these degree-weighted entries, they define a kind of degree-weighted similarity between pairs of people and pairs of groups, but not an exact similarity score as would be obtained, for instance, by using a criterion such as proportion matching \citep[208]{everett2013dual}.}
-
-Worth mentioning in this regard is an intriguing recent contribution by \citet{d2014use}, who convincingly advocate for using Multiple Correspondence Analysis (MCA) for the joint display of persons and groups in affiliation networks.
-The key difference between CA and MCA is that MCA works with a transformed version of the affiliation matrix, in which the rows are treated as ``cases'' and the columns as categorical attributes of the case, where each possible value represents a modality that the column object can take with respect to the cases.
-In the case of binary network data, the affiliation matrix is transformed into a new matrix $\mathbf{Z}$ with the same number of rows (persons) as the original and with twice the number of columns representing either membership ($z_{pg+} = 1$) or non-membership ($z_{pg-} = 1$) in that a group.
-The MCA scores are obtained via SVD of a centered and doubly normalized version of $\mathbf{Z}$.
-\citet{d2014use} discuss various advantages of the MCA graphical representation of two-mode networks, including the capacity to produce interpretable joint displays of actors and events and the ability to incorporate exogenous actor and event attributes as supplementary variables.
-However, because what follows bears more directly on the use of simple CA for two-mode network analysis, this discussion is less relevant.
+Like \citet{roberts2000correspondence} before her, Faust specifies that in the low (usually two) dimensional representation, inter-point distances are best (least squares) approximations of the chi-square distance between the membership profiles of pairs of people (or groups), where the profile of a given person-row or group-column is equal to the non-zero entries of the affiliation matrix divided by the corresponding row or column sum (people's activities and group sizes).\footnote{Note that because the chi-square distances are defined on these degree-weighted entries, they define a kind of \textit{degree-weighted similarity} between pairs of people and pairs of groups, but not an exact similarity score as would be obtained, for instance, by using a criterion such as the proportion of matches in the affiliation matrix \citep[208]{everett2013dual}.}
+
+In a recent paper, \citet{desposito2014comparison} compare CA and MCA as tools for representing the structural similarity between nodes in a two-mode network (actors in their case).
+Specifically, they analyze the significance of the typical $\chi^2$ distances defined from the scores obtained from the simple CA of the affiliation matrix as compared to the MCA of the indicator matrix of persons by groups with events treated as disjunctively coded binary variables measured over the actors (as just described).
+The focus of the analysis is the extent to which the $\chi^2$ distance of actor profiles obtained from the affiliation matrix $\mathbf{A}$ or the disjunctively coded indicator matrix $\mathbf{Z}$ best approaches a criterion for computing approximate structural equivalence between nodes (such as the Euclidean distance computed from the binary entries in the affiliation matrix).\footnote{In the case of MCA applied to the binary affiliation matrix, the latter is transformed into a new matrix $\mathbf{Z}$ with the same number of rows (persons) as the original and with twice the number of columns representing either membership ($z_{pg+} = 1$) or non-membership ($z_{pg-} = 1$) in that a group (disjunctive coding).
+The MCA scores are obtained via SVD of a centered and doubly normalized---with respect to person activity and group sizes---version of $\mathbf{Z}$ \citep{desposito2014use}.}  That is, \citet{desposito2014comparison} seek to ascertain whether CA or MCA is the best analytic tool to identify structurally equivalent actors in a two-mode network.
+Interestingly, \citet[115]{desposito2014comparison} note that just by examining the respective formulas defining the profile distances in CA versus MCA, it can be seen the CA is \textit{not} a good tool for revealing clusters of structurally equivalent actors.
+As the authors note, the reason is that the CA profile distance ``adopts a peculiar and more complex weighting system that could lead to a distorted representation of the actual relational patterns'' because the CA ``distance between actors (events) depends, not only on the pattern of participation (attendance), but also on the actor degree" and the event size.
+In addition, they note that in CA, small size events...are associated to larger weights...[and therefore] differences...related to such small size events have [a] stronger impact on the overall...distance than differences related to larger size events.'' \citet{desposito2014comparison} also note that, in contrast, the MCA definition of inter-profile distance between actors based on $\mathbf{Z}$ is not affected by the degree-weighting aspect, and that in this respect, ``[I]t turns out that MCA [distances] closely resembles the Euclidean distance." A conclusion that is supported by computational experiments conducted later in the paper, and which lead the authors to conclude that the MCA defined  $\chi^2$ distances ``better reproduces the \textit{actual} relational pattern embedded in affiliation networks'' \citep[122, italics added]{desposito2014comparison}.
+
+\subsection{Summary and Open Questions}
+As we can see from this brief and necessarily selective review, considerations of CA for two-mode network analysis have generally intertwined two basic themes.
+One beginning with \citet{bonacich1991simultaneous} and continuing through \citet{faust1997centrality, faust2005using} have commented on the formal commonalities and contrasts between ``centrality analysis'' and ``correspondence analysis,'' noticing striking similarities in the mathematical formulations of the traditional dual (eigenvector) score and the scores obtained from CA (as both come from the eigendecomposition of the affiliation matrix), and both result in scores that meet the strict criterion for duality (the scores of nodes in one mode are usually additive functions of the scores of nodes in the other mode).
+The other stream, which began with characterizations of CA as a visualization tool, focuses on the substantive meaning of the (low-dimensional representations of) the distances defined over the multidimensional space returned by the CA procedure.
+Here, multiple interpretations have been offered, from the idea that CA returns clique-like structures \citep{bonacich1991simultaneous}, to intimations that the closeness in the CA dimensions implies a more structured similarity closer to the idea of structural equivalence (e.g., such that actors that attend the same events and events that the same actors attend will be closer in the space).
+This approach sanctions the idea that the main goal of CA is to provide a parsimonious representation of the original ``input data'' \citep{faust2005using}, where this is understood to mean the original affiliation matrix.
+Yet, more recently, this line of interpretation has been controverted by those who have noticed that whatever the CA similarities are designed to do, they are a less-than-optimal representation of structural equivalence, given the specific weighting by the degree of nodes in each mode that is built into their construction \citep{desposito2014comparison}.
+
+In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral science was of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper, I show how a recent re-invention of CA for the analysis of ``economic complexity'' (two-mode networks of countries/regions by the products they produce) allows us to revisit controversies and open questions as to the meaning of CA scores (section ~\ref{sec:ref2mode}), and the relationship between correspondence analysis, centrality analysis, and the discovery of structurally similar clusters of actors in two-mode networks (section ~\ref{sec:ca}).
+Along the way, we will be able to further clarify the link between CA scores and traditional eigenvector centrality scores for two-mode networks, as well as ascertain the exact kind of similarity that is learned by estimating CA (and traditional eigenvector centrality) scores for nodes in two-mode networks along multiple dimensions.
 
@@
 In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of assigning meaningful scores to nodes in a two-mode network using what they called at the time a ``reflective'' approach.
-Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
+Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks.
+Hence, the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
 As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
@@
 
-As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order'' quantities $C^R(2)$ for both people and groups using the \textit{averaged} first-order centralities of the entities in the other mode they are connected to.
+As noted by \citet{hidalgo2009building}, the key to the reflective approach to computing the ``complexity'' of countries and products is the observation that, once we have these first-order quantities, it is possible to compute ``second-order'' quantities $C^R(2)$ for both people and groups using the \textit{averaged} first-order centralities of the entities in the other mode they are connected to.
 
@@
 
-Equation~\ref{eq:Rq_p} says that the reflective centrality of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
-Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
+Equation~\ref{eq:Rq_p} says that the reflective score of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
+Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective score is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
 
@@
 
-To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns.
+To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to ``W1'' or ``W11'' and ``E1'' or ``E13'' based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because ``E14'' in the original DGG data is not the last chronological even among the fourteen columns.
 In fact, as shown in Figure~\ref{fig:sw}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.\footnote{Using the DGG labeling, the chronologically ordered set of events are \{E11, E5, E2, E7, E12, E9, E3, E6, E10, E1, E14 E8, E4, E13\} \citep[p.
 68]{everett2018measuring}.} To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women---taken from \citet[83]{homans1950human}---from that of the original DGG table reproduced in \citet{freeman2003finding}, and ordered the events from left to right in the affiliation matrix \textit{chronologically}, which, as we have seen, also differs from the original DGG Table reproduced in \citet{freeman2003finding}.
-This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; in all, different versions of the dataset reproduce different versions of the labeling.
+This means that the arbitrary ordinal labeling of the rows (e.g., ``W1,'' ``W2,'' ``W3,'' etc.) and the columns (e.g., ``E1,'' ``E2,'' ``E3,'' etc.) in these (and other) publications---see, e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; in all, different versions of the dataset reproduce different versions of the labeling.
 
@@
 Event ordering, however, matters for analyses that try to exploit the temporal properties of the data \citep[e.g.,][]{everett2018measuring, lerner2022dynamic}.
-Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication \citep[e.g.,][]{freeman2003finding} may not correspond to "W1" or "E8" in the other \citep[e.g.,][]{doreian1979evolution}.
+Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since ``W1'' or ``E8'' in one publication \citep[e.g.,][]{freeman2003finding} may not correspond to ``W1'' or ``E8'' in the other \citep[e.g.,][]{doreian1979evolution}.
 To forestall confusion, in what follows, I use the numeric labeling corresponding to the original DGG data table as reproduced in (among others) \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{fig:sw}.
-My presentation includes both the arbitrary ordered labels (e.g., "W1" or "E1") and both the woman's name and the event date in parentheses, with "Myrna" (W11) rebaptized to her given name of "Myra" to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.
+My presentation includes both the arbitrarily ordered labels (e.g., ``W1'' or ``E1'') and both the woman's name and the event date in parentheses, with ``Myrna'' (W11) rebaptized to her given name of ``Myra'' to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.
 
 \subsubsection{Reflective Score Trajectory}
-Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections.
+Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective scores for persons, using a bump chart to plot the rank order trajectory of persons across reflections.
 The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
@@
 
-In the exhaustive methodological meta-analysis mentioned earlier, \citet{freeman2003finding} noted that ``the consensus of the analytic procedures is to assign women 1 through 9 to one group and women 10 through 18 to the other.'' Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover this same dominant partition reading from top to bottom.
+In the exhaustive methodological meta-analysis mentioned earlier, \citet{freeman2003finding} noted that ``the consensus of the analytic procedures is to assign women 1 through 9 to one group and women 10 through 18 to the other.'' Notably, the reflective score rankings after freezing ($q \geq 18$) recover this same dominant partition reading from top to bottom.
 Figure~\ref{fig:g-refs} shows the corresponding bump chart for the events.
@@
 Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering (the average of) the average number of memberships of its members (and so forth).
-Notably, the equilibrium reflective centralities recover and ordering of groups columns that re-appears across many articles that have analyzed these data with events 1-6 separated from events 10-15 by events 7-9 \citep[see e.g.,][]{doreian2004generalized, kovacs2010generalized, lizardo2024two, everett2013dual}.
+Notably, the equilibrium reflective scores recover and ordering of groups columns that re-appears across many articles that have analyzed these data with events 1-6 separated from events 10-15 by events 7-9 \citep[see e.g.,][]{doreian2004generalized, kovacs2010generalized, lizardo2024two, everett2013dual}.
 
@@
         \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
-            \caption{Reflective centrality trajectories of persons.}
+            \caption{Reflective score trajectories of persons.}
             \label{fig:p-refs}
@@
         \includegraphics[width=1.0\textwidth]{Plots/g-reflections.png}
-            \caption{Reflective centrality trajectories of groups.}
+            \caption{Reflective score trajectories of groups.}
             \label{fig:g-refs}
     \end{subfigure}
-    \caption{Reflective centralities for persons and groups (even reflections)}
+    \caption{Reflective scores for persons and groups (even reflections)}
     \label{fig:refs}
@@
 
-\section{CA and Dual Scoring in Two-Mode Networks} \label{sec:ca}
-The reader may ask what the point of going through all these reflective scoring since \citet{bonacich1991simultaneous} already developed a dual conception of scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected.
+\section{CA and Dual Centrality Scoring in Two-Mode Networks} \label{sec:ca}
+The reader may ask what the point of going through all this reflective scoring is since \citet{bonacich1991simultaneous} already developed a dual conception of scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected.
 Recall that, as noted earlier in that paper, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.
 
-While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures like degree).
-In fact, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
+While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures).
+In fact, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective scores discussed in the previous section \citep{mealy2019interpreting}.
 This makes the connection between the Bonacich and the CA scoring approaches more intimate than Bonacich thought.
@@
 
-\subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
-To see the connection between the CA and Bonacich dual scores more clearly, we can motivate the Bonacich eigenvector-based approach using the same ``reflective'' exposition we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}.
+\subsection{Bonacich Reflections} \label{subsec:bonref}
+To see the connection between the CA and Bonacich dual scores more clearly, we can motivate the Bonacich eigenvector-based approach using the same ``reflective'' exposition we used to introduce the HH reflective approach in Section~\ref{sec:ref2mode}.
 Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centralities for two-mode networks; however, it will help clarify the similarities and differences between the HH and the Bonacich approaches.
@@
 
-This says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
+This says that the Bonacich reflection at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
 Of course, these sums of sums would diverge to a larger and larger quantity at each step $q$.
@@
 Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem.
-As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.
+As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective scores can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.
 
@@
 We can motivate CA using the same dual projection strategy.
-Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\cite{bonacich1991simultaneous, faust2005using}, which gives the impression that CA is exclusively a ``direct'' rather than a dual projection method.
+Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\citep{borgatti1997network, bonacich1991simultaneous, faust1997centrality, faust2005using}.
+This gives the impression that CA is exclusively a ``direct'' rather than a dual projection method.
 Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection approach just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.
 
-To see this, consider the $|P| \times |P|$ matrix $\mathbf{D}p$, containing the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
+To see this, consider the $|P| \times |P|$ matrix $\mathbf{D}p$, containing the ``first order'' reflective scores of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
 In the same way, consider the $|G| \times |G|$ matrix $\mathbf{D}g$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose (recall that a matrix is row stochastic if its rows sum to one).
-
-For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
+Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose.\footnote{Recall that a matrix is row stochastic if its rows sum to one} For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
 
@@
 Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), positive definite, but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
-Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person.
+Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row person(group) and, following any Markovian sequence of $person-group-person'-group'$ hops, will reach the column person(group).
 Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
 The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).
+Figure~\ref{fig:dwproj} shows the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices for the Southern Women Data.
 
@@
 
-To get to the main point of all of this, it can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as a solution of the eigensystem:
+To get to the main point of all of this, it can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective scores can be obtained as a solution of the eigensystem:
 
@@
 With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices.\footnote{Note also that the degree normalized projection matrices have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix, with values between zero and one \citep{van2021correspondence}.}  This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
-Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective centrality scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms} or via the eigenvalue decomposition of the weighted projection matrices as in~\ref{eq:dam1} and~\ref{eq:dam2}.
+Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix via SVD or the eigenvalue decomposition of the weighted projection matrices as in~\ref{eq:dam1} and~\ref{eq:dam2} \citep[398, eq.
+9.17]{fouss2016algorithms}.
 And as we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix.
@@
 More importantly, because $\mathbf{P}_{pp} = \mathbf{D}p^{-1}\mathbf{A}\mathbf{D}g^{-1}\mathbf{A}^T$ and $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups.
-The key difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
+Both extract individual and group scores as eigenvectors of the one-mode projection of the original affiliation matrix: $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups.
+The key difference is that in CA, we pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
 
@@
 As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
-The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
+The HH reflective scores, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
 
@@
 
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}; accordingly, the first non-trivial CA dimension for persons and groups given by the right and left eigenvectors corresponding to the second-largest eigenvalue from the SVD of the matrix $\mathbf{D}_p^{-1/2}\mathbf{A}\mathbf{D}_g^{-1/2}$ \citep[126]{faust2005using}.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
+Now, as we saw earlier \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}; accordingly, the first non-trivial CA dimension for persons and groups given by the right and left eigenvectors corresponding to the second-largest eigenvalue from the SVD of the matrix $\mathbf{D}_p^{-1/2}\mathbf{A}\mathbf{D}_g^{-1/2}$ \citep[126]{faust2005using}.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of computing CA scores via reciprocal averaging \citep{hill1973reciprocal} \citep{mealy2019interpreting}.
 
@@
 Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
-This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large.
-On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
-
-Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
+This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large, thus adjusting similarity by group size so that co-memberships in groups everyone belongs to counts for less \citep{ragozini2014correspondence}.
+In the same way, on the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
+
+%Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
 In the Bonacich case, an eigendecomposition of the one-mode projection of the two-mode network; in the reflective case, an eigendecomposition of the degree-weighted one-mode projection of the same network.
@@
 \subsubsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
-Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
+Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$.\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic} using scores from CA, as they did in that paper for one-mode networks.} Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
 In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
-As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-Accordingly, the $C^R$ reflective scores reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
-
+As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure \citep[34]{wasserman1990correspondence}, and the $C^B$ re-ordered affiliation matrix having a triangular structure.
+
+Accordingly, the $C^R$ reflective scores reveal a dual partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
 The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side).
@@
 206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods.
-The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
-
-\subsubsection{Eigenvector Plot} \label{subsec:eigplot}
-We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
-If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
-The Southern Women data clearly belong to the second category.
-In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other.
-Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
-Note that this is the canonical consensus partitioning (for the people) highlighted by \citet{freeman2003finding} across twenty-one different methods and subsequently reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity (GRS) approach.
-This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.
-
-\begin{figure}
+The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward a similarity-based partitioning---with the exact kind of similarity revealed by CA yet to be determined---and $C^B$ more focused on finding a core of well-connected people and groups, separated from a less well-connected periphery.
+
+\begin{figure}[ht!]
     \centering
-        \includegraphics[width=1.0\textwidth]{Plots/ca-eigvec.png}            
+        \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}            
         \label{fig:eigvec}
@@
 
-\subsubsection{Normalized Similarity Plot} \label{subsec:normsim}
-To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = \mathbf{AD}g^{-1}\mathbf{A}^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors in a paper---minus one.
-Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = \mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
-As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
-This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition.
-These are shown as inset plots in Figure~\ref{fig:ca-eigvec} for both people and groups.
-Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.
-
-\begin{figure}
+\subsubsection{Eigenvector Plot} \label{subsec:eigplot}
+We can confirm that the first CA axis points toward the underlying similarity-based partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
+If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
+The Southern Women Data clearly belong in the second category.
+In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other.
+Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
+Note that this is the canonical consensus partitioning (for the people) highlighted by \citet{freeman2003finding} across twenty-one different methods and subsequently reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity (GRS) approach.
+This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar; we will return to this point shortly.
+
+\begin{figure}[ht!]
     \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
@@
         \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
-            \caption{CA correspondence plot.}
-            \label{fig:ca-corr}
-    \caption{CA similarity plots for persons (a) and groups (b) on the left and right top panel.
-The bottom panel (c) shows the correspondence plot for persons and groups with scores corresponding to the first dimension on the horizontal axis and scores corresponding to the second dimension on the vertical axis.}
-    \label{fig:ca-sim}
+    \caption{CA weighted projection plots for persons and groups on (a) the left and (b) the right top panel.
+The bottom panels show the correspondence plot for persons and groups with scores corresponding to (c) the first dimension on the horizontal axis and scores corresponding to the second dimension on the vertical axis, and (d) the first dimension on the horizontal axis and scores corresponding to the third dimension on the vertical axis.}
+    \label{fig:ca}
 \end{figure}
 
-To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
-In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity.
-Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups.
-As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.
-
-\subsubsection{CA Correspondence Plot} \label{subsec:caplot}
-The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions.
-People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices.
-The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups.
-People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm).
-The same goes for groups; those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart.
-The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.
-
-Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups).
-This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot.
+\subsubsection{CA Projection and Correspondence Plot} \label{subsec:caplot}
+Panels (a) and (b) of Figure~\ref{fig:ca} show the same weighted projection matrices from Figure~\ref{fig:dwproj} but this time with the rows and columns re-ordered according to the scores corresponding to the first CA dimension.
+The bottom panel of the same figure shows the usual correspondence plot of the first three CA dimensions.
+This figure illustrates a couple of important points.
+First, as panels (a) and (b) show, distances along the first CA axis correspond to degree-weighted similarities between persons and groups recorded in the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices.
+Pairs of people and groups with the largest entries in the matrix have minimal differences in the first CA dimension score, appearing close to one another when we reshuffle the rows and columns of the weighted projection matrix.
+This also means that the distances between pairs of people and groups in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the respective \textit{degree-normalized one-mode projections}.
+Pairs of people with large values in the projection matrix $P_{pp}$---such as \{Olivia, Flora\},  \{Katherine, Sylvia, Nora\}, \{Dorothy, Verne\}, and \{Brenda, Laura, Evelyn\}--- appear closer in the correspondence plot.
+The same goes for pairs of groups; those with large values in $P_{gg}$ matrix---such as \{E13, E14\} and \{E1, E2, E3, E4\}---appear closer in the correspondence plot, while those with small weighted similarity values are placed far apart.
+Distances between same-mode entities in the correspondence plot are thus a function of their (inverse) degree-normalized similarity.
+
+Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{ragozini2014correspondence}.
+This interpretation implies (for instance) that two people who attend many of the same events or groups with many members in common will appear close in the plot.
 But this (still common) interpretation is off the mark.
-What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs.
+What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.
+This type of degree-normalization can be the preferred approach in some applications since people with many memberships and groups with many members will appear to be similar to many other people and groups if we go with the raw number of other-mode objects shared.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs.
 2.13 and 2.14]{fouss2016algorithms}.
-As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
+As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} This implies that the similarities between pairs of people and groups will be weighted by the degrees of the other-mode objects that are shared between them.
+Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
 In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.
 
+\textit{}Another thing to note is that since the CA row and column scores in each dimension are obtained from the eigendecomposition of the normalized similarity matrices $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$, each dimension provides independent information on the similarity clustering of persons and groups.
+Thus, as shown in panels (a) and (b) of Figure~\ref{fig:ca}, while the first horizontal dimension recovers the main weighted-similarity partition separating the two dominant clusters of persons and groups, the second and third dimensions recover secondary and tertiary similarity clusters of people and groups net of the main two-block distinction.\footnote{We know from inset plot (c) of Figure~\ref{fig:ca-eigvec}, that the eigenvalues corresponding to the first three dimensions exhibit strong separation from the rest, suggesting that these provide a good low-rank approximation of the original degree-weighted projection matrix for persons and groups.} Thus, the second dimension (the y-axis in panel (c) of Figure~\ref{fig:ca}) separates \{Flora, Nora\} and \{E9, E11\} from the rest of persons and groups, consistent with the strong similarity between these nodes in the original weighted projection matrix.
+We, of course, know from the original affiliation matrix in Figure~\ref{fig:sw} that \{Flora, Nora\} are maximally similar---they are structurally equivalent---and \textit{only} attend events \{E9, E11\}.
+Accordingly, these two events end up being similar in this second dimension because they are attended by the people who are also most similar in the same dimension.
+The third CA dimension (the y-axis in panel (d) of Figure~\ref{fig:ca}), by contrast, separates \{Pearl Ruth, Dorothy, Verne, Myra\} and \{E8, E9\} from the rest; E8, along with E7, is one of the most popularly attended events, and thus contributes little to differentiating between the main partition of people and events on the first correspondence plot in panel (c).
+The pattern of attendance of \{Pearl Ruth, Dorothy, Verne, Myra\} is one characterized by high attendance at these popular events but sparser attendance at the events that serve to differentiate the two primary clusters.
+It is precisely this affinity of similar people for similar events that is picked up in the third dimension of the CA plot.
 \begin{figure}
@@
 Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}.
-I showed that just like the reflective centralities that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums.
+I showed that just like the reflective scores that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums.
 This exercise clarifies the links between CA and the dual two-mode eigenvector centrality more coherently and systematically, an issue that was left open and somewhat unclear in Bonacich's \citeyearpar{bonacich1991simultaneous} classic paper.
@@
 
-A better approach, similar to the one exemplified here, would be to present a comparison of the reflective centralities obtained by CA and Bonacich side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering.
+A better approach, similar to the one exemplified here, would be to present a comparison of the reflective scores obtained by CA and Bonacich side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering.
 Of course, suppose the analyst is more interested in such a ``subgroup'' analytic approach.

### 2025-02-21 analysis-R1.qmd
foo

@@
       # Affiliation matrix plot (CA ordering)
-      A.ord <- as.matrix(A[order(d1.r*-1), order(d1.c)])
-      A.plot <- ggcorrplot(t(A.ord)) 
-      A.plot <- A.plot + scale_x_discrete(position = "top")
-      A.plot1 <- A.plot + theme(legend.position = "none",
+      A.ord <- as.matrix(A[order(CA.r$vectors[, 2]), order(CA.c$vectors[, 2])])
+      p <- ggcorrplot(t(A.ord)) 
+      p <- p + scale_x_discrete(position = "top")
+      p <- p + scale_y_discrete(limits = rev)
+      p1 <- p + theme(legend.position = "none",
                                axis.text.x = element_text(hjust = -0.2))
       # Affiliation matrix plot (Bonacich ordering)
-      A.ord <- as.matrix(A[order(eig.r[, 1]*-1), order(eig.c[, 1])])
-      A.plot <- ggcorrplot(t(A.ord)) 
-      A.plot <- A.plot + scale_x_discrete(position = "top")
-      A.plot2 <- A.plot + theme(legend.position = "none",
+      A.ord <- as.matrix(A[order(eig.r[, 1]), order(eig.c[, 1])])
+      p <- ggcorrplot(t(A.ord)) 
+      p <- p + scale_x_discrete(position = "top")
+      p <- p + scale_y_discrete(limits = rev)
+      p2 <- p + theme(legend.position = "none",
                                axis.text.x = element_text(hjust = -0.2)) 
-
      # Affiliation matrix plot (Bonacich ordering)
-      A.ord2 <- as.matrix(A[order(eig.r[, 2]*-1), order(eig.c[, 2])])
-      A.plot <- ggcorrplot(t(A.ord2)) 
-      A.plot <- A.plot + scale_x_discrete(position = "top")
-      A.plot3 <- A.plot + theme(legend.position = "none",
+      A.ord <- as.matrix(A[order(eig.r[, 2]), order(eig.c[, 2])])
+      p <- ggcorrplot(t(A.ord)) 
+      p <- p + scale_x_discrete(position = "top")
+      p <- p + scale_y_discrete(limits = rev)
+      p3 <- p + theme(legend.position = "none",
                                axis.text.x = element_text(hjust = -0.2)) 
-
       png(file = here("Plots", "ca-reord.png"), width=500, height=500)
-      A.plot1
+      p1
       dev.off() 
       png(file = here("Plots", "bon-reord1.png"), width=500, height=500)
-      A.plot2
+      p2
       dev.off() 
       png(file = here("Plots", "bon-reord2.png"), width=500, height=500)
-      A.plot3
+      p3
       dev.off()
@@
                              lab = names(CA.c$vectors[,2]))
-      eigval.dat.c <- data.frame(k = 1:ncol(A)-1, 
-                                 value = as.numeric(CA.c$values)) %>% 
+      eigval.dat.c <- data.frame(k = 1:13, 
+                             value = CA.c$values[2:14]/max(CA.c$values[2:14])) %>% 
              mutate(value = value / max(value))
-
       # Eigenvector plot for the people
@@
 ```{r CA Similarity and Correspondence Plot}
-      norm.sim.r <- as.matrix(P.pp/max(P.pp))
-      norm.sim.c <- as.matrix(P.gg/max(P.gg))
+      norm.sim.r <- P.pp
+      norm.sim.c <- P.gg
+      diag(norm.sim.r) <- 0
+      diag(norm.sim.c) <- 0
       norm.sim.r <- norm.sim.r[order(CA.r$vectors[,2]), 
@@
       p <- p + scale_y_discrete(limits = rev)
+      p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.r+0.02)), 
+                                    low = "lightblue", 
+                                    high =  "darkred", 
+                                    mid = "red", 
+                                    midpoint = max(norm.sim.r)/2)
       a <- p + theme(legend.position = "none", 
                     axis.text.x = element_text(hjust = -0.2, size = 10)) 
-
       # CA Similarity Plot (Groups)
@@
       p <- p + scale_y_discrete(limits = rev)
+      p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.c+0.02)), 
+                                    low = "lightblue", 
+                                    high =  "darkred", 
+                                    mid = "red", 
+                                    midpoint = max(norm.sim.c)/2)
       b <- p + theme(legend.position = "none", 
@@
 ```{r Bonacich Similarity and Correspondence Plot}
-      AAt <- A %*% t(A)
-      AtA <- t(A) %*% A
-      norm.sim.r <- as.matrix(AAt/max(AAt))
-      norm.sim.c <- as.matrix(AtA/max(AtA))
+      norm.sim.r <- A %*% t(A)
+      norm.sim.c <- t(A) %*% A
+      diag(norm.sim.r) <- 0
+      diag(norm.sim.c) <- 0
       norm.sim.r <- norm.sim.r[order(eig.r[, 1]), order(eig.r[ ,1])]
       norm.sim.c <- norm.sim.c[order(eig.c[, 1]), order(eig.c[ ,1])]
-
       # Bonacich Similarity Plot (People)
@@
       p <- p + scale_y_discrete(limits = rev)
+      p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.r+1)), 
+                                    low = "lightblue", 
+                                    high =  "darkred", 
+                                    mid = "red",  
+                                    midpoint = max(norm.sim.r)/2)
       a <- p + theme(legend.position = "none",
@@
       p <- p + scale_y_discrete(limits = rev)
+      p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.c+1)), 
+                                    low = "lightblue", 
+                                    high =  "darkred", 
+                                    mid = "red",  
+                                    midpoint = max(norm.sim.c)/2)
       b <- p + theme(legend.position = "none", plot.margin = margin(l = 100),
@@
       eig.d2 <- c(eig.r[, 2], eig.c[, 2])
-      plot.dat <- data.frame(eig.d1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
+      plot.dat <- data.frame(eig.d1*-1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
       p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(mode), 
@@
                      )
-      p <- p + labs(x = "Second Dimension", y = "First Dimension")
+      p <- p + labs(x = "First Dimension", y = "Second Dimension")
       c <- p + scale_color_manual(values = c('red', 'blue'))

### 2025-02-21 old-similarity-verbiage.tex

--- 2024-07-13 main.tex
@@
-%% 
-%% Copyright 2019-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'CAS Bundle'.
-%% --------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'CAS Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for cas-dc documentclass for 
-%% double column output.
-
-%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
-\documentclass[a4paper,fleqn]{cas-sc}
-
-%\usepackage[numbers]{natbib}
-%\usepackage[authoryear]{natbib}
-\usepackage[authoryear,longnamesfirst]{natbib}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage{longtable}
-
-%%%Author definitions
-\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
-\tsc{WGM}
-\tsc{QE}
-\tsc{EP}
-\tsc{PMS}
-\tsc{BEC}
-\tsc{DE}
-%%%
-
-% Uncomment and use as if needed
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdefinition{rmk}{Remark}
-%\newproof{pf}{Proof}
-%\newproof{pot}{Proof of Theorem \ref{thm}}
-
-\begin{document}
-\let\WriteBookmarks\relax
-\def\floatpagepagefraction{1}
-\def\textpagefraction{.001}
-
-% Short title
-\shorttitle{Correspondence Analysis of Two-Mode Networks}
-
-% Short author
-%\shortauthors{O. Lizardo}
-
-% Main title of the paper
-\title [mode = title]{The Correspondence Analysis of Two-Mode Networks Revisited}                      
-% Title footnote mark
-% eg: \tnotemark[1]
-%\tnotemark[1,2]
-
-% Title footnote 1.
-% eg: \tnotetext[1]{Title footnote text}
-% \tnotetext[<tnote number>]{<tnote text>} 
-%\tnotetext[1]{}
-
-%\tnotetext[2]{.}
-
-% First author
-%
-% Options: Use if required
-% eg: \author[1,3]{Author Name}[type=editor,
-%       style=chinese,
-%       auid=000,
-%       bioid=1,
-%       prefix=Sir,
-       %orcid=0000-0002-5405-3007,
-%       facebook=<facebook id>,
-%       twitter=<twitter id>,
-%       linkedin=<linkedin id>,
-%       gplus=<gplus id>]
-%\author[1]{Omar Lizardo}[orcid=0000-0002-5405-3007]
-
-% Corresponding author indication
-%\cormark[1]
-
-% Footnote of the first author
-%\fnmark[1]
-
-% Email id of the first author
-%\ead{olizardo@soc.ucla.edu}
-
-% URL of the first author
-%\ead[url]{http://olizardo.bol.ucla.edu/}
-
-%  Credit authorship
-%\credit{Conceptualization of this study, Methodology, Software}
-
-%Address/affiliation
-%\affiliation[1]{organization={University of California, Los Angeles},
-   % addressline={264 Haines Hall}, 
-   % city={Los Angeles},
-    %citysep={},  Uncomment if no comma needed between city and postcode
-    % postcode={90095}, 
-    % state={CA},
-    % country={USA}}
-
- %Corresponding author text
-%\cortext[cor1]{Corresponding author}
-%\cortext[cor2]{Principal corresponding author}
-
-%Footnote text
-%\fntext[fn1]{This paper was written while the author was partially supported by National Science Foundation grant HNDS-R \#2214215.}
- % author as well.}
-%\fntext[fn2]{Another author footnote, this is a very long footnote and
- % it should be a really long footnote. But this footnote is not yet
- % sufficiently long enough to make two lines of footnote text.}
-
-% For a title note without a number/mark
-%\nonumnote{This note has no numbers. In this work we demonstrate $a_b$
-%  the formation Y\_1 of a new type of polariton on the interface
- % between a cuprous oxide slab and a polystyrene micro-sphere placed
- % on the slab.
- % }
-% Here goes the abstract
-\begin{abstract}
-This paper reconsiders the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting aspects that have not been previously emphasized, going beyond the use of CA as a visualization tool.
-It argues that CA can be used to compute a dual centrality score on both modes, related to but not mathematically equivalent to the \citet{bonacich1991simultaneous} two-mode centrality.
-This ``reflective'' centrality score connects the use of CA in two-mode network analysis to its use in other disciplines as a method of ordination.
-Additionally, I show CA can extract latent positional information on people and groups, with similarities to recent work on ``generalized relational similarity" in two-mode networks.
-Thus, CA can be used for indirect community or subgroup identification in two-mode networks, connecting to its use in some disciplines as a clustering method.
-The paper demonstrates these applications of CA, comparing them with the Bonacich dual centrality scores, and provides guidance on using CA for structural analysis in network analysis.
-\end{abstract}
-
-% Use if graphical abstract is present
-% \begin{graphicalabstract}
-% \includegraphics{figs/grabs.pdf}
-% \end{graphicalabstract}
-
-%Research highlights
-\begin{highlights}
-  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
-  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
-  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
-\end{highlights}
-
-% Keywords
-% Each keyword is separated by \sep
-\begin{keywords}
-Correspondence Analysis \sep Generalized Relational Similarity \sep Duality \sep Two-mode networks \sep Centrality 
-\end{keywords}
-
-\maketitle
-\newpage
-\section{Introduction} \label{sec:intro}
-While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
-Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}.
-This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Nevertheless, it departs from the already-mentioned previous efforts in highlighting aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before, moving beyond the focus on data visualization.
-
-Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
-With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general centrality score connecting the use of CA in two-mode network analysis to way CA features in some disciplines also concerned with two-mode data (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
-In social network analysis, this applies to conceptualizing the venerable duality between people and the groups they join \citep{breiger1974duality}.
-
-In addition to its capacity to reveal latent ordinal structure in two-mode data, I also argue that CA can also be used to extract latent \textit{positional} information on people and groups.
-Particularly, there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
-As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups).
-Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks connecting to how CA is used in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}, namely, the discovery of sociometrically similar subsets of entities in affiliation networks.
-In this way, CA emerges as a tool central to the usual network-analytic tasks and not just as a visualization or data-summarization tool.
-
-\subsection{Organization of the Paper} \label{subsec:org}
-The rest of the paper is organized as follows.
-In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
-This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes.
-This approach is computationally and mathematically related but \textit{not} equivalent to the eigenvalue decomposition of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores.
-I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers the core-periphery structure of the two-mode network instead.
-In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
-
-\section{Reflective Centralities in Two-Mode Networks} \label{sec:ref2mode}
-In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach.
-Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).
-Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
-As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
-
-If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}.
-A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $G$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.
-
-Given this, the degree-centrality of people is given by:
-
-\begin{equation}
-    C^R_p(1) = \sum_g a_{pg}
-    \label{eq:R1_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(1) = \sum_p a_{pg}
-  \label{eq:R1_g}
-\end{equation}
-
-That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
-As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.
-
-For people, these are given by:
-
-\begin{equation}
-    C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
-    \label{eq:R2_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
-    \label{eq:R2_g}
-\end{equation}
-
-Equation \ref{eq:R2_p} says ``people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
-Equation \ref{eq:R2_g} says ``groups are more central when the average activity of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
-Of course, we can keep on going and define third-order reflections.
-
-For the people, these are given by:
-
-\begin{equation}
-   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
-   \label{eq:R3_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
-   \label{eq:R3_g}
-\end{equation}
-
-Equation \ref{eq:R3_p} says something like ``people are more central when the average sum of the average activity of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
-Equation \ref{eq:R3_g} says, ''groups are more central when the average sum of the average size of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
-
-Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes).
-On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to).
-In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity).
-On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size).
-While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).
-
-More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases).
-
-For people, these are given by:
-
-\begin{equation}   
-    C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
-   \label{eq:Rq_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
-   \label{eq:Rq_g}
-\end{equation}
-
-Equation~\ref{eq:Rq_p} says that the reflective centrality of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
-Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
-
-\begin{table}
-\caption{Southern Women Data.}
-\label{tab:southern}
-\begin{tabular}[]{lcccccccccccccc}
-\hline
- & E1 & E2 & E3 & E4 & E5 & E6 & E7 & E8 & E9 & E10 & E11 & E12 & E13 & E14 \\
- \hline
-EVELYN & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-LAURA & 1 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-THERESA & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-BRENDA & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-CHARLOTTE & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-FRANCES & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-ELEANOR & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-RUTH & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-VERNE & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 1 & 0 & 0 \\
-MYRA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 \\
-KATHERINE & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-SYLVIA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 1 & 1 & 1 \\
-NORA & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 1 \\
-HELEN & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 & 0 \\
-OLIVIA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-FLORA & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 0 \\
-PEARL & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-DOROTHY & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 \\
-\hline
-\end{tabular}
-\textsuperscript{Note: Rows ordered according to the generalized
-blockmodeling solution of Doreian et al. (2004, Table 4).} \\
-\end{table}
-
-\subsection{Empirical Example} \label{subsec:ex}
-Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons in the Southern Women data \citep{davis1941}, using a bump chart to plot the rank order trajectory of persons across reflections.
-The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
-As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to.
-However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
-Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al.
-(2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
-
-Figure~\ref{fig:g-refs} shows the corresponding bump chart for the events.
-Just like for these actors, the equilibrium reflective centralities recover the ordering of the columns according to Doreian et al.
-generalized blockmodel with events 1-6 separated from events 10-15 by events 7-9 (see also \citet{kovacs2010generalized}).
-Figure~\ref{fig:pg-reflections} shows that $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
-Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering the average of the average number of memberships of its members (and so forth).
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
-            \caption{Reflective centrality trajectories of persons.}
-            \label{fig:p-refs}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-reflections.png}
-            \caption{Reflective centrality trajectories of groups.}
-            \label{fig:g-refs}
-    \end{subfigure}
-    \caption{Reflective centralities for persons and groups (even reflections)}
-    \label{fig:refs}
-\end{figure}
-
-\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks} \label{sec:ca}
-The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected.
-In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings.
-Instead, Bonacich noted that, in analyzing the Southern Women data shown in Table~\ref{tab:southern}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.
-
-As we will see, Bonacich was only half right.
-Indeed, the CA scores capture a version of ``clique'' (more accurately, community) membership, but they---particularly the first dimension---also capture a version of centrality in the most general sense of ranking nodes based on a meaningful criterion \citep{borgatti2006graph}.
-In fact, they capture the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.
-
-To see this, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:
-
-\begin{equation}
-    AC^B_p = \lambda C^B_g 
-    \label{eq:bon1}
-\end{equation}
-
-\begin{equation}
-    A^TC^B_g = \lambda C^B_p 
-    \label{eq:bon2}
-\end{equation}
-
-Where $A$ is the original affiliation matrix.
-Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, obtained from the singular value decomposition (SVD) of the rectangular affiliation matrix $A$.
-Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.
-
-For people, these are given by:
-
-\begin{equation}
-    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
-    \label{eq:faust1}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
-    \label{eq:faust2}
-\end{equation}
-
-Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich two-mode, the centrality of people is proportional to the centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members.
-These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode.
-Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
-The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.
-We will return to this crucial point later.
-
-\subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
-To see the connection between the reflective and Bonacich two-mode centralities more clearly, we can motivate the Bonacich two-mode centrality using the same ``reflective'' approach we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}.
-Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centrality for two-mode networks~\citep{bonacich1991simultaneous}, but it will help clarify the similarities and differences between the HH and the Bonacich approaches.
-Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:
-
-\begin{equation}
-   C^B_p(2) = \sum_g a_{pg}C^R_g(1)
-   \label{eq:BR2_p}
-\end{equation}
-
-\begin{equation}
-   C^B_g(2) = \sum_p a_{pg}C^R_p(1)
-   \label{eq:BR2_g}
-\end{equation}
-
-Equation~\ref{eq:BR2_p} says that people are central when the sum of the number of members of the groups they belong to is large.
-Equation~\ref{eq:BR2_g} says groups are central when the sum of the number of memberships of the people who belong to them is large.
-
-As before, we can keep on going and define a third-order reflection using the formulas:
-
-\begin{equation}
-   C^B_p(3) = \sum_g a_{pg}C^R_g(2)
-   \label{eq:BR3_p}
-\end{equation}
-
-\begin{equation}
-   C^B_g(3) = \sum_p a_{pg}C^R_p(2)
-   \label{eq:BR3_g}
-\end{equation}
-
-Equation~\ref{eq:BR3_p} says that people are central when the sum of the sum of the number of memberships held by the people who belong to the groups they belong to is large.
-Equation~\ref{eq:BR3_g} says that groups are central when the sum of the sum of the number of members of the groups their members belong to is large.
-Once again, we can keep going and define fourth order, fifth order, and higher reflections $C^R_p(4), C^R_p(4) \ldots C^R_p(q)$, where the centralities of nodes in one mode are based on the sums, of the sums, of the sums, of the centralities of nodes in the other mode.
-
-More generally, and in parallel with equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}, the reflective Bonacich centralities for persons and groups are given by:
-
-\begin{equation}
-   C^B_p(q) = \sum_g a_{pg}C^R_g(q-1)
-   \label{eq:BRq_p}
-\end{equation}
-
-\begin{equation}
-   C^B_g(q) = \sum_p a_{pg}C^R_p(q-1)
-   \label{eq:BRq_g}
-\end{equation}
-
-This just says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
-
-Of course, it is evident that these sum of sums would diverge to a bigger and bigger quantity at each step $q$.
-To prevent this and guarantee convergence, we normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:
-
-\begin{equation}
-   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
-   \label{eq:BRq_pn}
-\end{equation}
-
-\begin{equation}
-   C^B_g(q) = \frac{C^B_g(q)}{||C^B_g(q)||_2}
-   \label{eq:BRq_gn}
-\end{equation}
-
-Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}.
-
-Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations.
-These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach.
-Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power'' method of \citet{mises1929praktische}).
-
-This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' centralities and the Bonacich two-mode centralities.
-In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.
-
-\subsection{HH Reflective Centrality as an Eigenvector Centrality} \label{subsec:refeigen}
-Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem.
-As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.
-
-To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
-Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:
-
-\begin{equation}
-    \left(AA^T\right)C^B_p = \lambda^2 C^B_p
-    \label{eq:bon3}
-\end{equation}
-
-\begin{equation}
-    \left(A^TA\right)C^B_g = \lambda^2 C^B_g
-    \label{eq:bon4}
-\end{equation}
-
-This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.
-
-Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
-In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as the solution of the eigensystem:
-
-\begin{equation}
-    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
-    \label{eq:dam1}
-\end{equation}
-
-\begin{equation}
-    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
-    \label{eq:dam2}
-\end{equation}
-
-Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups.
-The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
-
-This can be clearly seen if we express $\left(Dg^{-1}A^TDp^{-1}A\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
-
-\begin{equation}
-    a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
-    \label{eq:mealy1}
-\end{equation}
-
-In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$.
-Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
-As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
-The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
-
-The same reasoning applies to groups in~\ref{eq:dam2}, whose entries are given by:
-
-\begin{equation}
-    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
-    \label{eq:mealy2}
-\end{equation}
-
-Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq.
-9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.
-
-Degree-pre-weighting does not radically alter the nature of the data.
-Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
-This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large.
-On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
-
-Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
-In the Bonacich case, an eigendecomposition of the one-mode projection of the two-mode network; in the reflective case, an eigendecomposition of the degree-weighted one-mode projection of the same network.
-Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
-Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to.
-The (average) centrality of groups, in turn, is a function of their members' (average) centrality.
-The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
-            \caption{Re-ordered Affiliation matrix (CA).}
-            \label{fig:ca-reord}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/bon-reord.png}
-            \caption{Re-ordered Affiliation matrix (Bonacich).}
-            \label{fig:bon-reord}
-    \end{subfigure}
-    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score.}
-    \label{fig:ca-v-bon}
-\end{figure}
-
-\section{Illustrative Analysis of the Southern Women Data} \label{sec:anal}
-\subsection{Row and Column Re-ordered Affiliation Matrices}
-Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
-In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
-As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
-The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side).
-The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
-
-\subsection{Eigenvector Plot} \label{subsec:eigplot}
-We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
-If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
-The Southern Women data clearly belong to the second category.
-In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other.
-Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
-Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach.
-This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.
-
-\begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-ca-eigvec.png}
-            \caption{Eigenvector plot (Persons).}
-            \label{fig:ca-reord}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-ca-eigvec.png}
-            \caption{Eigenvector plot (Groups).}
-            \label{fig:bon-reord}
-    \end{subfigure}
-    \caption{CA Eigenvector plot for persons (a) and groups (b).
-The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis.
-Insets show the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
-    \label{fig:ca-eigvec}
-\end{figure}
-
-\subsection{Normalized Similarity Plot} \label{subsec:normsim}
-To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one.
-Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
+To further appreciate the links between CA, HH reflective scoring, and community partitioning in two-mode networks, note that $S_p = \mathbf{AD}g^{-1}\mathbf{A}^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors in a paper---minus one.
+Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = \mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups).
 As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$.
@@
 Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.
-
-\begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.425\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-ca-sim.png}
-            \caption{CA Similarity plot (Persons).}
-            \label{fig:p-ca-sim}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-ca-sim.png}
-            \caption{CA Similarity plot (Groups).}
-            \label{fig:g-ca-sim}
-    \end{subfigure} \\
-     \begin{subfigure}[b]{0.65\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
-            \caption{CA correspondence plot.}
-            \label{fig:ca-corr}
-    \end{subfigure}
-    \caption{CA similarity plots for persons (a) and groups (b) on the left and right top panel.
-The bottom panel (c) shows the correspondence plot for persons and groups with scores corresponding to the first dimension on the horizontal axis and scores corresponding to the second dimension on the vertical axis.}
-    \label{fig:ca-sim}
-\end{figure}
-
-To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups).
-In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity.
-Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups.
-As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.
-
-\subsection{CA Correspondence Plot} \label{subsec:caplot}
-The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions.
-People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices.
-The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups.
-People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm).
-The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart.
-The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.
-
-Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups).
-This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot.
-But this (still common) interpretation is off the mark.
-What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs.
-2.13 and 2.14]{fouss2016algorithms}.
-As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
-In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.
-
-\begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.425\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-bon-sim.png}
-            \caption{Bonacich Similarity plot (Persons).}
-            \label{fig:p-bon-sim}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-bon-sim.png}
-            \caption{Bonacich Similarity plot (Groups).}
-            \label{fig:g-bon-sim}
-    \end{subfigure} \\
-     \begin{subfigure}[b]{0.65\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
-            \caption{Bonacich correspondence plot.}
-            \label{fig:bon-corr}
-    \end{subfigure}
-    \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel.
-The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis.}
-    \label{fig:bon-sim}
-\end{figure}
-
-\subsection{One-Mode Projection Matrix and Eigevenctor Plot}
-This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
-But as we have seen, this is \textit{not} what CA is designed to do.
-Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes.
-Notably, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two eigenvectors of the usual one-mode projections of the matrix (where the first is, of course, the standard Bonacich dual centrality score).
-
-To illustrate, Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by other-mode degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score.
-Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
-Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}  We can see that the plot of the first two eigenvectors does a good job of recovering the raw connectivity structure of the Southern Women affiliation network, partitioning the core persons and groups (on the upper-right) from the more peripheral ones (on the lower-left).
-
-Moreover, if all we are interested in is capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then Figure~\ref{fig:bon-sim}(c) does a better job of that than the usual CA correspondence plot in Figure~\ref{fig:ca-sim}(c).
-For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine, Sylvia\right]$ for $\left[E10, E12\right]$.
-In the same way, core events like $\left[E8, E9, E10\right]$---shown on the lower half of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the upper half.
-The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.
-
-This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's Bonacich eigenvector centrality.
-For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot.
-Similarly, the main difference between $\left[Ruth\right]$ and $\left[Sylvia\right]$, despite their comparable $C^B$ scores, is that the former's (closer to the upper half of the plot) event profile is composed mainly of core events.
-In contrast, peripheral events dominate the latter's attendance profile (shown in the bottom half of the plot), accordingly, $\left[Ruth\right]$ and $\left[Sylvia\right]$ are assigned to distinct clusters by the k-means algorithm.
-
-\subsection{Correspondence Analysis and Generalized Relational Similarity} \label{subsec:cagensim}
-As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.
-Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}, where objects' similarities are given by their pattern of connections to objects in the other mode.
-This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence \citep{everett1994regular}.
-In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.\footnote{\citet{lizardo2024two} shows the connection between Kovacs's idea of generalized relations similarity and the two-mode network projection.}
-
-An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}.
-In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong to, and the similarity of groups is a function of the people who belong to them, making the similarity of persons and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
-Thus,
-
-\begin{itemize}
-    \item People are \textit{similar} if they belong to \textit{similar} groups.
-    \item Groups are \textit{similar} if they share \textit{similar} members.
-\end{itemize}
-
-Which is consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
-To accomplish this, \citet[540, eq.
-2 and eq.
-3]{jeh2002simrank} propose that we define the similarity of each pair of people $S(p, p')$  as given by:
-
-\begin{equation}
-    S(p, p') = \frac{\alpha}{C^R(1)_pC^R(1)_{p'}}
-    \sum_{i = 1}^{C^R(1)_p} \sum_{j = 1}^{C^R(1)_{p'}} 
-    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
-    \label{eq:simrank1}
-\end{equation}
-
-Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to, and $\alpha$ is a free parameter obeying the restriction: $0 > \alpha < 1$.
-By construction, $S(p, p) = 1$, for all $p$.
-Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of each unordered pair of groups they both belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.
-
-Likewise, for groups, the SimRank similarities are given by:
-
-\begin{equation}
-    S(g, g') = \frac{\alpha}{C^R(1)_gC^R(1)_{g'}}
-    \sum_{i = 1}^{C^R(1)_g} \sum_{j = 1}^{C^R(1)_{g'}} 
-    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
-    \label{eq:simrank2}
-\end{equation}
-
-Where $p(i)_{i \in N(g)}$ is the $i^{th}$ person in the set of members of group $g$, and $p(j)_{j \in N(g')}$ is the $j^{th}$ person in the set of members of group $g'$, and $S(g, g) = 1$.
-Thus, the SimRank similarity between two groups is a function of the sum of the similarities of each pair of people who belong to both groups, weighted by the reciprocal of the products of the number of members of each group multiplied by $\alpha$.
-
-\begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/sr-plot-reord.png}
-            \caption{Bonacich Similarity plot (Persons).}
-            \label{fig:p-bon-sim}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-sr-plot-eigen.png}
-            \caption{Bonacich Similarity plot (Groups).}
-            \label{fig:g-bon-sim}
-    \end{subfigure} 
-     \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-sr-plot-eigen.png}
-            \caption{Bonacich correspondence plot.}
-            \label{fig:bon-corr}
-    \end{subfigure} \\
-     \begin{subfigure}[b]{0.6\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/sr.png}
-            \caption{Bonacich Similarity plot (Groups).}
-            \label{fig:g-bon-sim}
-    \end{subfigure} 
-     \begin{subfigure}[b]{0.2\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-sr-plot-eigen.png}
-            \caption{SimRank correspondence plot.}
-            \label{fig:sr-corr}
-    \end{subfigure}
-    \caption{Simrank versus CA comparison.
-Panel (a) shows the affiliation matrix re-ordered according to the first eigenvector of the SVD of the SimRank similarity matrix.
-Panels (b) and (d) show the corresponding eigenvector and eigenvalue---panels (c) and (e)---plots for persons and groups, respectively, of the SimRank similarity matrix.
-Panel (f) shows the correspondence plot based on the eigenvector decomposition of the SimRank similarity matrix.
-Panels (g) and (h) show the Pearson correlation between the scores corresponding to the first eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis).}
-    \label{fig:simrank}
-\end{figure}
-
-In two-mode networks, SimRank scores for each pair of nodes across the two sets can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
-Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
-We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}, which is confirmed here for the Southern Women data.
-
-Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
-The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
-Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections \citep{mealy2019interpreting, van2021correspondence}.
-
-And indeed, they are not.
-Figure~\ref{fig:simrank}(a) shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices---with $\alpha$ set to $0.8$---for the row and column nodes.
-Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two communities of persons and groups revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
-
-Figure~\ref{fig:simrank}(b) and Figure~\ref{fig:simrank}(d) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
-Looking at the plots from left to right, we can see that the partitioning of the node sets on the first informative eigenvector is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
-
-Figure~\ref{fig:simrank}(f) shows the correspondence plot of the first and second eigenvectors of the SimRank similarity matrices for both persons and groups.
-Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c), suggesting that the first-CA dimensions of the two-mode network recover clusters of ``similar'' entities in the two node sets, where internode similarity is a generalized relational similarity as defined earlier.
-Thus, [$Flora, Olivia$] appear close because they choose similar groups, as do [$Sylvia, Katherine, Helen$].
-In the same way, groups [$E9, E11$] are close in the plot because they are chosen by similar people, as are events [$E10, E12, E13, E14$].
-Figures~\ref{fig:simrank}(g) and (h) show a regression plot of the first CA eigenvector for persons (top) and groups (bottom) on the x-axis and the first eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
-As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that these first eigenvectors capture the same (GSR) information across the two approaches.
-
-\section{Discussion and Concluding Remarks} \label{sec:disc}
-This paper reconsiders the role of CA in the analysis of two-mode network data.
-We began with an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to simple CA of the affiliation matrix.
-Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}.
-I showed that just like the reflective centralities that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums.
-This exercise clarifies the links between CA and the dual two-mode eigenvector centrality in a more coherent and systematic way, an issue that was left open and somewhat unclear in Bonacich's (1991) classic paper.
-Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.
-
-But CA does more than reveal a latent dual ordering of nodes in the two-mode network.
-In addition to doing this, CA can be shown to reveal \textit{groupings} of nodes based on some conception of the similarity of their connections to nodes in the other mode.
-This is evident once we use the scores of the first CA dimension to re-order the rows and columns of the original affiliation matrix.
-When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data.
-This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events.
-
-This is another way in which CA and the Bonacich dual centrality approach systematically differ; one is geared to uncovering blocks of actors with similar connections to events in the other mode (and vice versa), while the other reveals blocks of actors who are the most active and who attend the largest events.
-Both are, of course, legitimate ways of analyzing the structure of a two-mode network.
-Still, they differ in terms of the structural patterns that they are sensitive to, with the CA approach closer to a community partition where nodes that are surprisingly similar end up in the same clusters---where ``surprising'' means similarity based on deviations from a suitable null model based on random mixing given their first-order degrees.
-
-But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to?
-Here we argued that CA recovers a type of generalized relational similarity (GRS).
-That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar.
-Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, we saw that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach.
-Thus, we can clarify that two-mode network CA reveals latent groupings of nodes based on a GRS criterion.
-
-Overall, the preceding has shown that CA can be upgraded from a method designed to generate joint plots and visualization of two-mode data to one that can be seen as more ``central'' to the usual social-network-analytic tasks, like ordering the nodes in the two sets according to some substantively meaningful rank---centrality analysis---or finding sets of similar actors in the network (subgroup or community detection).
-At the very least, it seems like the Bonacich style dual centrality based on the eigenvector decomposition of the raw affiliation matrix should not be the default ``reflective'' centrality approach unless the analyst has the explicit analytic goal of exploring center-periphery partitioning in the network.
-
-A better approach, similar to the one exemplified here, would be to present a comparison of the reflective centralities obtained by CA and Bonacich side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering.
-Of course, suppose the analyst is more interested in such a ``subgroup'' analytic approach.
-In that case, the Bonacich dual centrality approach is less helpful (even if multiple dimensions are considered), and CA should be the first line of attack.
-
-%% Loading bibliography style file]
-%\bibliographystyle{model1-num-names}
-\bibliographystyle{cas-model2-names}
-% Loading bibliography database
-\bibliography{ca.bib}
-\end{document}
-
-The Elsevier cas-dc class is based on the
-standard article class and supports almost all of the functionality of
-that class. In addition, it features commands and options to format the
-\begin{itemize} \item document style \item baselineskip \item front
-matter \item keywords and MSC codes \item theorems, definitions and
-proofs \item labels of enumerations \item citation style and labeling.
-\end{itemize}
-
-This class depends on the following packages
-for its proper functioning:
-
-\begin{enumerate}
-\itemsep=0pt
-\item {natbib.sty} for citation processing;
-\item {geometry.sty} for margin settings;
-\item {fleqn.clo} for left aligned equations;
-\item {graphicx.sty} for graphics inclusion;
-\item {hyperref.sty} optional packages if hyperlinking is
-  required in the document;
-\end{enumerate}
-
-All the above packages are part of any
-standard \LaTeX{} installation.
-Therefore, the users need not be
-bothered about downloading any extra packages.
-
-\section{Installation}
-
-The package is available at author resources page at Elsevier
-(\url{http://www.elsevier.com/locate/latex}).
-The class may be moved or copied to a place, usually,\linebreak
-\verb+$TEXMF/tex/latex/elsevier/+, %$%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-or a folder which will be read                   
-by \LaTeX{} during document compilation.  The \TeX{} file
-database needs updation after moving/copying class file.  Usually,
-we use commands like \verb+mktexlsr+ or \verb+texhash+ depending
-upon the distribution and operating system.
-
-\section{Front matter}
-
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. 
-You are recommended to conform your choice to the journal you 
-are submitting to.
-
-\section{Bibliography styles}
-
-There are various bibliography styles available. You can select the
-style of your choice in the preamble of this document. These styles are
-Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use Bib\TeX\ to generate your bibliography and include DOIs
-whenever available.
-
-Here are two sample references: 
-\cite{Fortunato2010}
-\cite{Fortunato2010,NewmanGirvan2004}
-\cite{Fortunato2010,Vehlowetal2013}
-
-\section{Floats}
-{Figures} may be included using the command,\linebreak 
-\verb+\includegraphics+ in
-combination with or without its several options to further control
-graphic. \verb+\includegraphics+ is provided by {graphic[s,x].sty}
-which is part of any standard \LaTeX{} distribution.
-{graphicx.sty} is loaded by default. \LaTeX{} accepts figures in
-the postscript format while pdf\LaTeX{} accepts {*.pdf},
-{*.mps} (metapost), {*.jpg} and {*.png} formats. 
-pdf\LaTeX{} does not accept graphic files in the postscript format.
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=.75]{figs/Fig1.pdf}
-	\caption{The evanescent light - $1S$ quadrupole coupling
-	($g_{1,l}$) scaled to the bulk exciton-photon coupling
-	($g_{1,2}$). The size parameter $kr_{0}$ is denoted as $x$ and
-	the \PMS is placed directly on the cuprous oxide sample ($\delta
-	r=0$, See also Table \protect\ref{tbl1}).}
-	\label{FIG:1}
-\end{figure}
-
-The \verb+table+ environment is handy for marking up tabular
-material. If users want to use {multirow.sty},
-{array.sty}, etc., to fine control/enhance the tables, they
-are welcome to load any package of their choice and
-{cas-dc.cls} will work in combination with all loaded
-packages.
-
-\begin{table}[width=.9\linewidth,cols=4,pos=h]
-\caption{This is a test caption. This is a test caption. This is a test
-caption. This is a test caption.}\label{tbl1}
-\begin{tabular*}{\tblwidth}{@{} LLLL@{} }
-\toprule
-Col 1 & Col 2 & Col 3 & Col4\\
-\midrule
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-12345 & 12345 & 123 & 12345 \\
-\bottomrule
-\end{tabular*}
-\end{table}
-
-\section[Theorem and ...]{Theorem and theorem like environments}
-
-{cas-dc.cls} provides a few shortcuts to format theorems and
-theorem-like environments with ease. In all commands the options that
-are used with the \verb+\newtheorem+ command will work exactly in the same
-manner. {cas-dc.cls} provides three commands to format theorem or
-theorem-like environments:
-
-\begin{verbatim}
- \newtheorem{theorem}{Theorem}
- \newtheorem{lemma}[theorem]{Lemma}
- \newdefinition{rmk}{Remark}
- \newproof{pf}{Proof}
- \newproof{pot}{Proof of Theorem \ref{thm2}}
-\end{verbatim}
-
-The \verb+\newtheorem+ command formats a
-theorem in \LaTeX's default style with italicized font, bold font
-for theorem heading and theorem number at the right hand side of the
-theorem heading.  It also optionally accepts an argument which
-will be printed as an extra heading in parentheses.
-
-\begin{verbatim}
-  \begin{theorem} 
-   For system (8), consensus can be achieved with 
-   $\|T_{\omega z}$ ...
-     \begin{eqnarray}\label{10}
-     ....
-     \end{eqnarray}
-  \end{theorem}
-\end{verbatim}
-
-\newtheorem{theorem}{Theorem}
-
-\begin{theorem}
-For system (8), consensus can be achieved with 
-$\|T_{\omega z}$ ...
-\begin{eqnarray}\label{10}
-....
-\end{eqnarray}
-\end{theorem}
-
-The \verb+\newdefinition+ command is the same in
-all respects as its \verb+\newtheorem+ counterpart except that
-the font shape is roman instead of italic.  Both
-\verb+\newdefinition+ and \verb+\newtheorem+ commands
-automatically define counters for the environments defined.
-
-The \verb+\newproof+ command defines proof environments with
-upright font shape.  No counters are defined.
-
-\section[Enumerated ...]{Enumerated and Itemized Lists}
-{cas-dc.cls} provides an extended list processing macros
-which makes the usage a bit more user friendly than the default
-\LaTeX{} list macros.   With an optional argument to the
-\verb+\begin{enumerate}+ command, you can change the list counter
-type and its attributes.
-
-\begin{verbatim}
- \begin{enumerate}[1.]
- \item The enumerate environment starts with an optional
-   argument `1.', so that the item counter will be suffixed
-   by a period.
- \item You can use `a)' for alphabetical counter and '(i)' 
-  for roman counter.
-  \begin{enumerate}[a)]
-    \item Another level of list with alphabetical counter.
-    \item One more item before we start another.
-    \item One more item before we start another.
-    \item One more item before we start another.
-    \item One more item before we start another.
-\end{verbatim}
-
-Further, the enhanced list environment allows one to prefix a
-string like `step' to all the item numbers.
-
-\begin{verbatim}
- \begin{enumerate}[Step 1.]
-  \item This is the first step of the example list.
-  \item Obviously this is the second step.
-  \item The final step to wind up this example.
- \end{enumerate}
-\end{verbatim}
-
-\section{Cross-references}
-In electronic publications, articles may be internally
-hyperlinked. Hyperlinks are generated from proper
-cross-references in the article.  For example, the words
-\textcolor{black!80}{Fig.~1} will never be more than simple text,
-whereas the proper cross-reference \verb+\ref{tiger}+ may be
-turned into a hyperlink to the figure itself:
-\textcolor{blue}{Fig.~1}.  In the same way,
-the words \textcolor{blue}{Ref.~[1]} will fail to turn into a
-hyperlink; the proper cross-reference is \verb+\cite{Knuth96}+.
-Cross-referencing is possible in \LaTeX{} for sections,
-subsections, formulae, figures, tables, and literature
-references.
-
-\section{Bibliography}
-
-Two bibliographic style files (\verb+*.bst+) are provided ---
-{model1-num-names.bst} and {model2-names.bst} --- the first one can be
-used for the numbered scheme. This can also be used for the numbered
-with new options of {natbib.sty}. The second one is for the author year
-scheme. When  you use model2-names.bst, the citation commands will be
-like \verb+\citep+,  \verb+\citet+, \verb+\citealt+ etc. However when
-you use model1-num-names.bst, you may use only \verb+\cite+ command.
-
-\verb+thebibliography+ environment.  Each reference is a\linebreak
-\verb+\bibitem+ and each \verb+\bibitem+ is identified by a label,
-by which it can be cited in the text:
-
-\noindent In connection with cross-referencing and
-possible future hyperlinking it is not a good idea to collect
-more that one literature item in one \verb+\bibitem+.  The
-so-called Harvard or author-year style of referencing is enabled
-by the \LaTeX{} package {natbib}. With this package the
-literature can be cited as follows:
-
-\begin{enumerate}[\textbullet]
-\item Parenthetical: \verb+\citep{WB96}+ produces (Wettig \& Brown, 1996).
-\item Textual: \verb+\citet{ESG96}+ produces Elson et al. (1996).
-\item An affix and part of a reference:\break
-\verb+\citep[e.g.][Ch. 2]{Gea97}+ produces (e.g. Governato et
-al., 1997, Ch. 2).
-\end{enumerate}
-
-In the numbered scheme of citation, \verb+\cite{<label>}+ is used,
-since \verb+\citep+ or \verb+\citet+ has no relevance in the numbered
-scheme.  {natbib} package is loaded by {cas-dc} with
-\verb+numbers+ as default option.  You can change this to author-year
-or harvard scheme by adding option \verb+authoryear+ in the class
-loading command.  If you want to use more options of the {natbib}
-package, you can do so with the \verb+\biboptions+ command.  For
-details of various options of the {natbib} package, please take a
-look at the {natbib} documentation, which is part of any standard
-\LaTeX{} installation.
-
-\appendix
-\section{My Appendix}
-Appendix sections are coded under \verb+\appendix+.
-
-\verb+\printcredits+ command is used after appendix sections to list 
-author credit taxonomy contribution roles tagged using \verb+\credit+ 
-in frontmatter.
-
-\printcredits
-
-%\vskip3pt
-
-\bio{}
-Author biography without author photo.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-\endbio
-
-\bio{figs/pic1}
-Author biography with author photo.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-\endbio
-
-\bio{figs/pic1}
-Author biography with author photo.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-Author biography. Author biography. Author biography.
-\endbio

### 2025-03-24 analysis-R1.qmd
do it

@@
       p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.r+0.02)), 
-                                    low = "lightblue", 
+                                    low = "white", 
                                     high =  "darkred", 
@@
       p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.c+0.02)), 
-                                    low = "lightblue", 
+                                    low = "white", 
                                     high =  "darkred", 
@@
       p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.r+1)), 
-                                    low = "lightblue", 
+                                    low = "white", 
                                     high =  "darkred", 
@@
       p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.c+1)), 
-                                    low = "lightblue", 
+                                    low = "white", 
                                     high =  "darkred", 
@@
       # Bonacich Correspondence Plot
-      eig.d1 <- c(eig.r[, 1], eig.c[, 1])
-      eig.d2 <- c(eig.r[, 2], eig.c[, 2])
-      plot.dat <- data.frame(eig.d1*-1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
+      eig.d1 <- c(eig.r[, 1], eig.c[, 1])*-1
+      eig.d2 <- c(eig.r[, 2], eig.c[, 2])*-1
+      plot.dat <- data.frame(eig.d1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
       p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(mode), 
@@
       c <- p + scale_color_manual(values = c('red', 'blue'))
-      png(file = here("Plots", "bon-corr-plot.png"), width=1200, height=900)
-      (a + b) / c + 
+      png(file = here("Plots", "bon-corr-plot.png"), width=1500, height=1000)
+      (a + b) / (c) + 
          plot_annotation(tag_levels = 'a', tag_suffix = ')') & 

### 2025-03-24 main-R1.tex
Updates from Overleaf

@@
 \begin{abstract}
-This paper reconsiders the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting aspects that have not been previously emphasized, going beyond the use of CA as a visualization tool.
-It argues that CA can be used to compute a dual centrality score on both modes, related to but not mathematically equivalent to the \citet{bonacich1991simultaneous} two-mode centrality.
-This ``reflective'' centrality score connects the use of CA in two-mode network analysis to its use in other disciplines as a method of ordination.
-Additionally, I show CA can extract latent positional information on people and groups, with similarities to recent work on ``generalized relational similarity" in two-mode networks.
-Thus, CA can be used for indirect community or subgroup identification in two-mode networks, connecting to its use in some disciplines as a clustering method.
-The paper demonstrates these applications of CA, comparing them with the Bonacich dual centrality scores, and provides guidance on using CA for structural analysis in network analysis.
+This paper reconsiders the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting aspects that have not been previously emphasized and clarifying its linkages to two-mode centrality.
+I show that using CA to compute a dual score on both sets of nodes in a two-mode network is related to but not mathematically equivalent to the \citet{bonacich1991simultaneous} two-mode centrality.
+This ``reflective'' approach to computing dual scores connects the use of CA in two-mode network analysis to its use in other disciplines as a method of ordination.
+Additionally, I show that CA can extract latent positional information on people and groups, connecting it to recent work on extracting ``generalized relational similarities'' in two-mode networks.
 \end{abstract}
@@
 While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
-Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}.
+Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{desposito2014use}.
+
 This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Nevertheless, it departs from the already-mentioned previous efforts in highlighting aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before, moving beyond the focus on data visualization.
-
-Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}.
-With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general centrality score connecting the use of CA in two-mode network analysis to way CA features in some disciplines also concerned with two-mode data (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities.
-In social network analysis, this applies to conceptualizing the venerable duality between people and the groups they join \citep{breiger1974duality}.
-
-In addition to its capacity to reveal latent ordinal structure in two-mode data, I also argue that CA can also be used to extract latent \textit{positional} information on people and groups.
+Nevertheless, it departs from previous efforts to highlight aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before.
+Particularly, I clarify the similarities and differences between the dual scores computed by CA on both modes and those obtained from the eigenvector-based approach proposed by \citet{bonacich1991simultaneous} for obtaining dual centralities in two-mode networks.
+These linkages are made clearer by recasting both CA and the Bonacich centralities as variations of a ``reflective'' strategy for analyzing two-mode network data via iterative dual scoring \citep{yang2022comparing, yang2020birank}, which proceeds by assigning substantively meaningful scores to the nodes in each set of a two-mode network using an iterative procedure equivalent to a spectral decomposition of a proximity matrix based on the one-mode projection.
+
+In this vein, I show how the CA approach to two-mode network analysis---re-invented in some corners of Network Science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting, van2021correspondence}--- is part of a more general toolkit for the analysis of two-mode networks.
+While the ``economic complexity'' interpretation of the CA scores on the first dimension has treated the resulting scores as ordinal rank measures---in the style of a centrality index---I review recent work that shows that they are better sense as partitioning the two sets of nodes into \textit{similarity classes} based on their patterns of connectivity to the \textit{most similar} nodes in the other set \citep{kovacs2010generalized, lizardo2024two}.
+
+Rather than ordering nodes on a unidimensional axis based on connectivity---like Bonacich's \citet{bonacich1991simultaneous} eigenvector-based approach---CA extracts latent \textit{positional} information on people and groups.
 Particularly, there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
-As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups).
-Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks connecting to how CA is used in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}, namely, the discovery of sociometrically similar subsets of entities in affiliation networks.
-In this way, CA emerges as a tool central to the usual network-analytic tasks and not just as a visualization or data-summarization tool.
+As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to \textit{similar} entities in the other mode (e.g., groups), thus following a clustering logic of ``regular equivalence.'' In this way, CA can be used as an approach to indirect blockmodeling in two-mode networks.
+This is similar to the way CA has been developed and used in some disciplines (e.g., computer and information science), namely, as a \textit{spectral clustering} method for detecting most similar node subsets in bipartite graphs \citep{wu2022spectral}.
+Overall, CA emerges as a tool central to the usual social-network-analytic task of extracting underlying structure from two-mode (and multi-mode) relational data.
 
@@
 The rest of the paper is organized as follows.
-In the next section, I motivate the use of CA to produce a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix.
-This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes.
-This approach is computationally and mathematically related but \textit{not} equivalent to the eigenvalue decomposition of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores.
-I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes.
-I contrast this ordering to that provided by the Bonacich dual centrality scores, which, in contrast to the CA ordering, reveals a dual community partition and recovers the core-periphery structure of the two-mode network instead \citep{everett2013dual}.
-In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.
-
-\section{Previous uses of CA in Two-Mode Network Data Analysis}
-
-In this section, I consider previous contributions regarding the usefulness and limitations of simple CA as a methodological tool for the analysis of two-mode network data as applied directly to the affiliation matrix recording ties between actors and events.
-For reasons of space and relevance, therefore, my discussion is circumscribed in four ways: I focus attention on (1) pieces discussing primarily methodological or interpretive aspects of CA, (2) with regard to the analysis of two-mode network data or affiliation networks,\footnote{There are, of course, various substantive applications of CA for the analysis of two-mode networks~\citep[e.g.,][]{breiger2000tool, faust2002scaling, schweizer1991power, serino2024mapping, ragozini2018analysis} as well as substantive and methodological applications of CA to the analysis of one-mode network data~\citep[e.g.,][]{noma1985scaling, kumbasar1994systematic, lizardo2020correspondence}; these will not be part of the discussion here.} (3)  dealing with the usual binary affiliation matrix measuring a single relation between persons and groups at one point in time, where links are observed only between persons and groups.\footnote{As such, recent work dealing with the application of CA or other factoring approaches to time-varying, multiplex, multilevel, or valued two-mode network data \citep{ragozini2014correspondence, ragozini2015multiple, zhu2016correspondence} will also be outside the scope of my discussion.} Finally, (4) there has been a spate of recent work focusing on the uses and advantages of using Multiple Correspondence Analysis---a variant of CA applied to a disjunctive doubly centered version of the original affiliation matrix) to analyze and visualize two-mode networks \citep[e.g.,][]{dramalidis2016subset, desposito2014use, ragozini2014correspondence}.
-Because much of what I have to say deals with the uses and interpretation of the scores obtained from simple CA applied directly to the affiliation matrix (or its one-mode projections), my discussion will not be directly relevant to this line of work, except for a recent paper directly comparing the substantive interpretation of the $\chi^2$ distances in simple CA and MCA with regard to their substantive interpretation as indexes of structural similarity, which bears more directly to some of the subsequent discussion \citep{desposito2014comparison}.
+In Section~\ref{sec:prevuses}, I review previous SNA work dealing with the uses of CA to analyze two-mode network data.
+Then, in Section~\ref{sec:ref2mode}, I consider a recent re-invention of CA (the economic complexity index) used to produce a type of Bonacich-style dual scoring via iterative scoring across the two modes of the affiliation matrix.
+This leads naturally to an abbreviated way of computing the same scores---CA and Bonacich---via spectral decomposition of the one-mode projection of the affiliation matrix for each set of nodes (degree-weighted in the case of CA and unweighted in the case of Bonacich).
+Section~\ref{sec:comparing} uses the classic Southern Women dataset to show the payoff of this approach, focusing on the distinct substantive insights extracted by CA and Bonacich scoring in a venerable dataset, showing that the CA-blocked affiliation matrix reveals a dual block partition between the two node sets (similar to that recovered by previous analyses of the same data using other methods), while the Bonacich-blocked affiliation recovers a core-periphery structure instead.
+Section~\ref{sec:cagensim} considers the hanging question of the exact substantive interpretation of the node groupings revealed by CA-style blocking, which are shown to be fairly close to those obtained by a computationally unrelated approach based on the idea of ``generalized relational similarity,'' suggesting that the CA blocks obey a logic distinct from structural equivalence.
+Finally, Section~\ref{sec:disc} closes with a summary of key points in the argument and suggestions for future development of the linkages between CA and positional and centrality analysis in two-mode networks.
+
+\section{Previous uses of CA in Two-Mode Network Data Analysis} \label{sec:prevuses}
+
+For reasons of space and relevance, my discussion of previous work is circumscribed in four ways, focusing on (1) pieces discussing primarily \textit{methodological} or interpretive aspects of CA, (2) regarding the analysis of \textit{two-mode network} data or affiliation networks,\footnote{There are, of course, various substantive applications of CA for the analysis of two-mode networks~\citep[e.g.,][]{breiger2000tool, faust2002scaling, schweizer1991power, serino2024mapping, ragozini2018analysis} as well as substantive and methodological applications of CA to the analysis of one-mode network data~\citep[e.g.,][]{noma1985scaling, kumbasar1994systematic, lizardo2020correspondence}; these will not be part of the discussion here.} (3) analyzing the usual \textit{binary} affiliation matrix measuring a single relation between persons and groups at one point in time, where links are observed only between persons and groups,\footnote{As such, recent work dealing with the application of CA or other factoring approaches to time-varying, multiplex, multilevel, or valued two-mode network data \citep{ragozini2014correspondence, ragozini2015multiple, zhu2016correspondence} will also be outside the scope of my discussion.} using with the \textit{simple CA} of the affiliation matrix.\footnote{There has been a spate of recent work focusing on the uses and advantages of using Multiple Correspondence Analysis or MCA---a variant of CA applied to a disjunctively-coded, doubly-centered version of the original affiliation matrix) to analyze and visualize two-mode networks \citep[e.g.,][]{dramalidis2016subset, desposito2014use, ragozini2014correspondence}.
+Because much of what I have to say deals with the uses and interpretation of the scores obtained from simple CA applied directly to the affiliation matrix (or its one-mode projections), my discussion will not be directly relevant to this line of work, except for a recent paper directly comparing the substantive interpretation of the $\chi^2$ distances in simple CA and MCA with regard to their substantive interpretation as indexes of structural similarity, which bears more directly to some of the subsequent discussion \citep{desposito2014comparison}.}
 
@@
 One of the earliest considerations of CA as a data-analytic tool for examining duality relations in two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks.
-In that paper, \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people by the groups to which they belong,'' then it would seem that CA would be an appropriate analytic technique.
+In that paper, \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people by the groups to which they belong,'' then it would seem that CA would be a relevant analytic technique.
 According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (scare quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation---in the least-squared error sense---of the similarities between row and column objects.
 Bonacich's discussion of CA came after his influential introduction of the Eigenvector centrality approach for two-mode network analysis.
-Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of the affiliation matrix---namely, the Singular Value Decomposition (SVD).
-As we will see in section~\ref{subsec:refeigen}, both CA and centrality analysis are mathematically linked in a second way, as both are based on the spectral decomposition of the dual \textit{projections} of the affiliation matrix.
+Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both rely on a version of the spectral decomposition of the affiliation matrix---namely, the Singular Value Decomposition (SVD) to obtain scores for the row and column objects in the two-mode network.
+As we will see in section~\ref{subsec:refeigen}, CA and centrality analysis are mathematically linked in a second way not noted by Bonacich, as they are also based on the spectral decomposition of the dual one-mode \textit{projections} of the original rectangular affiliation matrix.
 In this particular respect, the mathematics connecting the Eigenvector approach to centrality analysis and CA are more similar than even Bonacich appreciated.
 
-For Bonacich, the main differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality,'' but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also do later in the paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by their most representative members  \citep[164]{bonacich1991simultaneous}.
+For Bonacich, the main differences between CA and Eigenvector centrality analysis stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]efore the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality,'' but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also do later in this paper---noting that ``[r]ather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events,'' with the magnitude of each woman and event score in the main dimension seeming to indicate the ``purity'' of their membership in each of the two cliques and for events; that is, the likelihood of being attended by their most representative members  \citep[164]{bonacich1991simultaneous}.
 
@@
 First, \citet{roberts2000correspondence} noted that there were many examples in the standing of using CA to analyze binary non-frequency data and that, therefore, the uses of CA are not constrained by the nature or provenance of the numerical quantities recorded in each cell of a two-mode network affiliation matrix.
-\citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and the event points in the two-dimensional representation and one that used all thirteen dimensions.\footnote{In CA, the maximum number of dimensions is one minus the rank of the affiliation matrix, which in this case is fourteen, corresponding to the number of events.} Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that even (and similarly for column profiles corresponding to events).
+\citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and the event points in the two-dimensional representation and one that used all thirteen dimensions.\footnote{In CA, the maximum number of dimensions is one minus the rank of the affiliation matrix, which in this case is fourteen, corresponding to the number of events.} Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that event (and similarly for column profiles corresponding to events).
 Thus, the point distances are indeed interpretable as a kind of \textit{degree-weighted similarity} between objects in each mode.
@@
 The focus of the analysis is the extent to which the $\chi^2$ distance of actor profiles obtained from the affiliation matrix $\mathbf{A}$ or the disjunctively coded indicator matrix $\mathbf{Z}$ best approaches a criterion for computing approximate structural equivalence between nodes (such as the Euclidean distance computed from the binary entries in the affiliation matrix).\footnote{In the case of MCA applied to the binary affiliation matrix, the latter is transformed into a new matrix $\mathbf{Z}$ with the same number of rows (persons) as the original and with twice the number of columns representing either membership ($z_{pg+} = 1$) or non-membership ($z_{pg-} = 1$) in that a group (disjunctive coding).
-The MCA scores are obtained via SVD of a centered and doubly normalized---with respect to person activity and group sizes---version of $\mathbf{Z}$ \citep{desposito2014use}.}  That is, \citet{desposito2014comparison} seek to ascertain whether CA or MCA is the best analytic tool to identify structurally equivalent actors in a two-mode network.
-Interestingly, \citet[115]{desposito2014comparison} note that just by examining the respective formulas defining the profile distances in CA versus MCA, it can be seen the CA is \textit{not} a good tool for revealing clusters of structurally equivalent actors.
-As the authors note, the reason is that the CA profile distance ``adopts a peculiar and more complex weighting system that could lead to a distorted representation of the actual relational patterns'' because the CA ``distance between actors (events) depends, not only on the pattern of participation (attendance), but also on the actor degree" and the event size.
-In addition, they note that in CA, small size events...are associated to larger weights...[and therefore] differences...related to such small size events have [a] stronger impact on the overall...distance than differences related to larger size events.'' \citet{desposito2014comparison} also note that, in contrast, the MCA definition of inter-profile distance between actors based on $\mathbf{Z}$ is not affected by the degree-weighting aspect, and that in this respect, ``[I]t turns out that MCA [distances] closely resembles the Euclidean distance." A conclusion that is supported by computational experiments conducted later in the paper, and which lead the authors to conclude that the MCA defined  $\chi^2$ distances ``better reproduces the \textit{actual} relational pattern embedded in affiliation networks'' \citep[122, italics added]{desposito2014comparison}.
+The MCA scores are obtained via SVD of a centered and doubly normalized---with respect to person activity and group sizes---version of $\mathbf{Z}$ \citep{desposito2014use}.}  That is, \citet{desposito2014comparison} seek to ascertain whether CA or MCA is the best analytic tool to identify sets of \textit{structurally equivalent} actors in a two-mode network.
+
+Interestingly, \citet[115]{desposito2014comparison} note that just by examining the respective formulas defining the profile distances in CA versus MCA, we can verify that CA is \textit{not} a good tool for revealing clusters of structurally equivalent actors.
+As \citet{desposito2014comparison} put it, the reason is that the CA profile distance ``adopts a peculiar and more complex weighting system that could lead to a distorted representation of the actual relational patterns'' because the CA ``distance between actors (events) depends, not only on the pattern of participation (attendance)'' but also on the actor's \textit{degree} and the event \textit{sizes}.
+In addition, they note that because of this weighting scheme, in CA, ``small size events...are associated to larger weights...[and therefore] differences...related to such small size events have [a] stronger impact on the overall...distance than differences related to larger size events.'' \citet{desposito2014comparison} also point out that, in contrast, the MCA definition of inter-profile distance between actors based on $\mathbf{Z}$ is not affected by the degree-weighting aspect, and that in this respect, ``[I]t turns out that MCA [distances] closely resembles the Euclidean distance,'' which is a straightforward criterion for structural equivalence.
+This is a conclusion that is supported by computational experiments conducted later in the paper, and which leads \citet{desposito2014comparison} to conclude that the MCA defined  $\chi^2$ distances ``better reproduces the \textit{actual} relational pattern embedded in affiliation networks'' (122, italics added).
 
@@
 As we can see from this brief and necessarily selective review, considerations of CA for two-mode network analysis have generally intertwined two basic themes.
-One beginning with \citet{bonacich1991simultaneous} and continuing through \citet{faust1997centrality, faust2005using} have commented on the formal commonalities and contrasts between ``centrality analysis'' and ``correspondence analysis,'' noticing striking similarities in the mathematical formulations of the traditional dual (eigenvector) score and the scores obtained from CA (as both come from the eigendecomposition of the affiliation matrix), and both result in scores that meet the strict criterion for duality (the scores of nodes in one mode are usually additive functions of the scores of nodes in the other mode).
+One beginning with \citet{bonacich1991simultaneous} and continuing through \citet{faust1997centrality, faust2005using} have commented on the formal commonalities and contrasts between ``centrality analysis'' and ``correspondence analysis,'' noticing striking similarities in the mathematical formulations of the traditional dual (eigenvector) score and the scores obtained from CA, as both come from the eigendecomposition of the affiliation matrix, resulting in scores that meet the strict criterion for duality, as the scores of nodes in one mode are usually additive functions of the scores of nodes in the other mode.
 The other stream, which began with characterizations of CA as a visualization tool, focuses on the substantive meaning of the (low-dimensional representations of) the distances defined over the multidimensional space returned by the CA procedure.
@@
 This approach sanctions the idea that the main goal of CA is to provide a parsimonious representation of the original ``input data'' \citep{faust2005using}, where this is understood to mean the original affiliation matrix.
-Yet, more recently, this line of interpretation has been controverted by those who have noticed that whatever the CA similarities are designed to do, they are a less-than-optimal representation of structural equivalence, given the specific weighting by the degree of nodes in each mode that is built into their construction \citep{desposito2014comparison}.
-
-In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral science was of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper, I show how a recent re-invention of CA for the analysis of ``economic complexity'' (two-mode networks of countries/regions by the products they produce) allows us to revisit controversies and open questions as to the meaning of CA scores (section ~\ref{sec:ref2mode}), and the relationship between correspondence analysis, centrality analysis, and the discovery of structurally similar clusters of actors in two-mode networks (section ~\ref{sec:ca}).
-Along the way, we will be able to further clarify the link between CA scores and traditional eigenvector centrality scores for two-mode networks, as well as ascertain the exact kind of similarity that is learned by estimating CA (and traditional eigenvector centrality) scores for nodes in two-mode networks along multiple dimensions.
+Yet, more recently, this line of interpretation has been controverted by  \citep{desposito2014comparison} who have noticed that whatever the CA similarities are designed to do, they are a less-than-optimal representation of structural equivalence (and thus presumably the original input data), given the specific weighting by the degree of nodes in each mode that is built into their construction.
+
+In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral science was of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper and starting in the next section, I show how a recent re-invention of CA for the analysis of ``economic complexity''---two-mode networks of countries/regions by the products they are most adept at producing---in the now well-established field of ``Network Science'' in the twenty-first century, allows us to revisit controversies and open questions as to the meaning of CA scores in two-mode network data, and the relationship between correspondence analysis, centrality analysis, and positional analysis (e.g., the discovery of structurally similar clusters of actors in two-mode networks).
+Along the way, we will be able to shed light on the link between CA scores and traditional eigenvector centrality scores for two-mode networks, as well as ascertain the exact kind of similarity that is learned by estimating CA (and traditional eigenvector centrality) scores for nodes in two-mode networks along multiple dimensions.
 
@@
 
-That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.
+That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $\mathbf{A}$, and the column sums of the same matrix give the first-order centrality of groups.
 
@@
 In the same way, while Equation~\ref{eq:R1_g} assigns a high value to groups that have lots of members, Equation~\ref{eq:R1_g} assigns a high value to groups that, on average, have members who themselves have lots of memberships (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
-Of course, we can keep on going and define ``third-order'' reflections.
-
-For the people, these are given by:
+
+Of course, we can keep on going and define ``third-order'' reflections; for the people, these are given by:
 
@@
 
-As we saw, while for people, Equation~\ref{eq:R2_p} measured the averaged size of the groups they join, Equation~\ref{eq:R3_p} assigns a high value to people who, on average, belong to groups who are themselves attended by highly active members (e.g., whenever $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
+As we saw, while for people, Equation~\ref{eq:R2_p} measured the average size of the groups they join, Equation~\ref{eq:R3_p} assigns a high value to people who, on average, belong to groups who are themselves attended by highly active members (e.g., whenever $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
 In the same way, while Equation~\ref{eq:R2_g} assigns a high value to groups whose members have lots of memberships, Equation~\ref{eq:R3_g} assigns a high value to groups that, on average, have members who themselves (also on average) belong to large groups (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
@@
 
-\section{CA and Dual Centrality Scoring in Two-Mode Networks} \label{sec:ca}
+\subsection{CA and Dual Centrality Scoring in Two-Mode Networks} \label{sec:ca}
 The reader may ask what the point of going through all this reflective scoring is since \citet{bonacich1991simultaneous} already developed a dual conception of scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected.
@@
 \begin{equation}
-    AC^B_p = \lambda C^B_g 
+    \mathbf{A}C^B_p = \lambda C^B_g 
     \label{eq:bon1}
@@
 \begin{equation}
-    A^TC^B_g = \lambda C^B_p 
+    \mathbf{A}^TC^B_g = \lambda C^B_p 
     \label{eq:bon2}
@@
 Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$.
-Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific scores.
-
-For people, these are given by:
+
+Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific scores; for people, these are given by:
 
@@
 
-However, Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
+Note, however, the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
 The main difference is that in HH's reflective equations, a person's centrality is proportional to the \textit{group-degree-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{person-degree-weighted} centrality of the people who are members.
@@
 \subsection{Bonacich Reflections} \label{subsec:bonref}
-To see the connection between the CA and Bonacich dual scores more clearly, we can motivate the Bonacich eigenvector-based approach using the same ``reflective'' exposition we used to introduce the HH reflective approach in Section~\ref{sec:ref2mode}.
-Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centralities for two-mode networks; however, it will help clarify the similarities and differences between the HH and the Bonacich approaches.
+We can motivate the Bonacich eigenvector-based approach using the same ``reflective'' exposition we used to introduce the HH reflective approach in Section~\ref{sec:ref2mode}.
+Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centralities for two-mode networks; but it will help clarify the similarities and differences between the HH and the Bonacich approaches.
 
@@
 Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), positive definite, but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
-Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row person(group) and, following any Markovian sequence of $person-group-person'-group'$ hops, will reach the column person(group).
+Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row person (group) and, following any Markovian sequence of $person-group-person'-group'$ hops, will reach the column person (group), and where the probability of jumping from any person to a group they belong is given by the inverse of the number of groups they belong to (and similarly for groups) \citep[240]{deng2009generalized}.
 Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
@@
         \includegraphics[width=1.0\textwidth]{Plots/p-norm.png}
-            %\caption{Degree-weighted projection (People).}
-            %\label{fig:ca-reord}
+            \caption{People.}
+            \label{fig:p-dwproj}
     \end{subfigure} \\
@@
         \includegraphics[width=1.0\textwidth]{Plots/g-norm.png}
-            %\caption{Degree-weighted projection (Groups).}
-            %\label{fig:bon-reord}
+            \caption{Groups.}
+            \label{fig:g-dwproj}
     \end{subfigure}
@@
 
-To get to the main point of all of this, it can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective scores can be obtained as a solution of the eigensystem:
+It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective scores can be obtained as a solution of the eigensystem:
 
@@
     \label{eq:dam2}
-\end{equation}\
-
-With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices.\footnote{Note also that the degree normalized projection matrices have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix, with values between zero and one \citep{van2021correspondence}.}  This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
-Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix via SVD or the eigenvalue decomposition of the weighted projection matrices as in~\ref{eq:dam1} and~\ref{eq:dam2} \citep[398, eq.
+\end{equation}
+
+With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices, which are the same as the CA scores and the limit HH reflective scores \citep{van2021correspondence}.\footnote{As we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
+Note also that the degree normalized projection matrices have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix, with values between zero and one \citep{van2021correspondence}.} These scores will also be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups, respectively) obtained from a simple CA of the original affiliation matrix via SVD \citep[398, eq.
 9.17]{fouss2016algorithms}.
-And as we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix.
-
-More importantly, because $\mathbf{P}_{pp} = \mathbf{D}p^{-1}\mathbf{A}\mathbf{D}g^{-1}\mathbf{A}^T$ and $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
+This exact correspondence can be verified in Figure~\ref{fig:ca-ref-corr}, which shows a scatterplot and associated linear regression line of the standardized scores computed via the iterative method of reflections ($q = 26$) on the y-axis against those obtained from the second eigenvector of the degree-weighted projection matrices for persons and groups on the x-axis ($r = 1.0$).\footnote{Formally, the eigendecomposition of the degree-weighted projection (for people) is given by the solution to the equation $\mathbf{P}_{pp} = \mathbf{UVU}^{-1}$, where $\mathbf{U}$ is a $|P| \times |P|$ matrix whose columns are the eigenvectors of $\mathbf{P}_{pp}$ and $\mathbf{V}$ is a $|P| \times |P|$ matrix whose diagonals are the eigenvalues of $\mathbf{P}_{pp}$ (with zeros in every other cell).
+Columns two through $|P|$ of matrix $\mathbf{U}$ contain the standard CA scores for the people along the complete set of $|P| - 1$ dimensions; the same goes for $\mathbf{P}_{gg}$ in the case of groups.}
+
+Because $\mathbf{P}_{pp} = \mathbf{D}p^{-1}\mathbf{A}\mathbf{D}g^{-1}\mathbf{A}^T$ and $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
 Both extract individual and group scores as eigenvectors of the one-mode projection of the original affiliation matrix: $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups.
@@
 \begin{equation}
-    a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
+    p_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
     \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
@@
 \begin{equation}
-    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
+    p_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
     \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
@@
 
-Now, as we saw earlier \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}; accordingly, the first non-trivial CA dimension for persons and groups given by the right and left eigenvectors corresponding to the second-largest eigenvalue from the SVD of the matrix $\mathbf{D}_p^{-1/2}\mathbf{A}\mathbf{D}_g^{-1/2}$ \citep[126]{faust2005using}.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of computing CA scores via reciprocal averaging \citep{hill1973reciprocal} \citep{mealy2019interpreting}.
+Now, as we saw earlier \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}; accordingly, the first non-trivial CA dimension for persons and groups given by the right and left eigenvectors corresponding to the second-largest eigenvalue from the SVD of the matrix $\mathbf{D}_p^{-1/2}\mathbf{A}\mathbf{D}_g^{-1/2}$ \citep[126]{faust2005using}.} In fact, as \citet{mealy2019interpreting} have noted, the method of ``reflections'' is just a re-discovery of the older idea of computing CA scores via reciprocal averaging \citep{hill1973reciprocal}.
 
@@
 
-Degree-pre-weighting does not radically alter the nature of the data.
+As \citet{desposito2014comparison} have recently noted, degree-pre-weighting does alter the original input data, which means that CA is \textit{not} meant to reproduce a low dimensional approximation of the original affiliation matrix.
 Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
@@
 In the same way, on the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
-
-%Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical.
-In the Bonacich case, an eigendecomposition of the one-mode projection of the two-mode network; in the reflective case, an eigendecomposition of the degree-weighted one-mode projection of the same network.
-Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks.
-Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to.
-The (average) centrality of groups, in turn, is a function of their members' (average) centrality.
-The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.
 
@@
 
-\subsection{Comparing Bonacich and CA Dual Scoring in the Southern Women Data}
-\subsubsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
-Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$.\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic} using scores from CA, as they did in that paper for one-mode networks.} Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).
-In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
+\section{Comparing Bonacich and CA Dual Scoring in the Southern Women Data} \label{sec:comparing}
+\subsection{Row and Column Re-ordered Affiliation Matrices} 
+Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$.
+Each panel shows the original Southern Women affiliation matrix but with rows and columns blocked according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic}, using scores from CA to re-order the original matrix, as they did in that paper for one-mode networks.} In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
 As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure \citep[34]{wasserman1990correspondence}, and the $C^B$ re-ordered affiliation matrix having a triangular structure.
 
-Accordingly, the $C^R$ reflective scores reveal a dual partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
-The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side).
+Accordingly, the $C^R$ reflective scores reveal a dual-block partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
+The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic \textit{core-periphery} partitioning \citep{borgatti2000models}, with a group of highly active women (on the top half) who selectively co-participate in highly attended events (on the left-hand side) and less active women (on the bottom half) who go to less well-attended events (on the right-hand side).
 In fact, as \citet[p.
-206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods.
-The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward a similarity-based partitioning---with the exact kind of similarity revealed by CA yet to be determined---and $C^B$ more focused on finding a core of well-connected people and groups, separated from a less well-connected periphery.
+206]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic \citet{breiger1974duality} projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods.
+In fact, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is identical to that shown in~\citet[206, Table 2]{everett2013dual}.
+
+The larger point is that the standard eigenvector and CA scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward a similarity-based partitioning---with the exact kind of similarity revealed by CA yet to be determined---and $C^B$ more focused on finding a core of well-connected people and groups, separated from a less well-connected periphery.
 
@@
 
-\subsubsection{Eigenvector Plot} \label{subsec:eigplot}
+\subsection{Eigenvector Plot} \label{subsec:eigplot}
 We can confirm that the first CA axis points toward the underlying similarity-based partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
-If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
+If a two-mode network has no discernible block structure, the first dimension CA scores would be distributed as a continuous logistic curve; when block structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
 The Southern Women Data clearly belong in the second category.
-In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other.
-Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position.
-Note that this is the canonical consensus partitioning (for the people) highlighted by \citet{freeman2003finding} across twenty-one different methods and subsequently reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity (GRS) approach.
-This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar; we will return to this point shortly.
+In the people mode, we have \{Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\} on one side, \{Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\} on the other.
+Among events, we have \{E1, E2, E3, E4, E5, E6\} on side \{E9, E10, E11, E12, E13, E14\} on the other and \{E7, E8\} in an ambiguous middle position.
+Note that this is the canonical consensus partitioning (for the people) highlighted by \citet{freeman2003finding} across twenty-one different methods and subsequently reproduced (for both persons and groups) by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity (GRS) approach.
+This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar; we will return to this point in Section~\ref{sec:cagensim}.
 
@@
         \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
-    \caption{CA weighted projection plots for persons and groups on (a) the left and (b) the right top panel.
-The bottom panels show the correspondence plot for persons and groups with scores corresponding to (c) the first dimension on the horizontal axis and scores corresponding to the second dimension on the vertical axis, and (d) the first dimension on the horizontal axis and scores corresponding to the third dimension on the vertical axis.}
+    \caption{The top panels show the degree-weighted one-mode projection heatmap plots for (a) persons and (b) groups in the Southern Women Data with rows and columns re-ordered according to the first CA dimension.
+Darker red cells correspond to high entry values (normalized by the maximum) in the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices.
+The bottom panels show the correspondence plot for persons and groups with scores corresponding to (c) the first dimension on the horizontal axis with the second dimension on the vertical axis, and (d) the first dimension on the horizontal axis with the third dimension on the vertical axis.}
     \label{fig:ca}
@@
 
-\subsubsection{CA Projection and Correspondence Plot} \label{subsec:caplot}
-Panels (a) and (b) of Figure~\ref{fig:ca} show the same weighted projection matrices from Figure~\ref{fig:dwproj} but this time with the rows and columns re-ordered according to the scores corresponding to the first CA dimension.
-The bottom panel of the same figure shows the usual correspondence plot of the first three CA dimensions.
-This figure illustrates a couple of important points.
-First, as panels (a) and (b) show, distances along the first CA axis correspond to degree-weighted similarities between persons and groups recorded in the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices.
+\subsection{Degree-Weighted Projections and CA Correspondence Plot} \label{subsec:caplot}
+Panels (a) and (b) of Figure~\ref{fig:ca} shows a heatmap plot of the same weighted projection matrices from Figures~\ref{fig:p-dwproj} and~\ref{fig:g-dwproj}, but this time with the rows and columns re-ordered according to the scores corresponding to the first CA dimension.
+In the plot, darker red indicates greater similarity between pairs of persons and groups, and gray indicates less similarity.
+The bottom panels of the same figure show the usual correspondence plot of the first three CA dimensions, with the first dimension on the x-axis against the second dimension on the y-axis in panel (c) and the first dimension plotted against the third dimension on the y-axis in panel (d).
+
+This figure illustrates several important points.
+First, as the heatmap plots in panels (a) and (b) show, distances along the first CA axis correspond to degree-weighted similarities between persons and groups recorded in the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices.
 Pairs of people and groups with the largest entries in the matrix have minimal differences in the first CA dimension score, appearing close to one another when we reshuffle the rows and columns of the weighted projection matrix.
@@
 
-Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{ragozini2014correspondence}.
+Note that this differs from the usual interpretation of the CA correspondence plot, which is typically taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{desposito2014comparison}, or ``first order'' similarity \citet{kovacs2010generalized}.
 This interpretation implies (for instance) that two people who attend many of the same events or groups with many members in common will appear close in the plot.
@@
 Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
-In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.
-
-\textit{}Another thing to note is that since the CA row and column scores in each dimension are obtained from the eigendecomposition of the normalized similarity matrices $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$, each dimension provides independent information on the similarity clustering of persons and groups.
-Thus, as shown in panels (a) and (b) of Figure~\ref{fig:ca}, while the first horizontal dimension recovers the main weighted-similarity partition separating the two dominant clusters of persons and groups, the second and third dimensions recover secondary and tertiary similarity clusters of people and groups net of the main two-block distinction.\footnote{We know from inset plot (c) of Figure~\ref{fig:ca-eigvec}, that the eigenvalues corresponding to the first three dimensions exhibit strong separation from the rest, suggesting that these provide a good low-rank approximation of the original degree-weighted projection matrix for persons and groups.} Thus, the second dimension (the y-axis in panel (c) of Figure~\ref{fig:ca}) separates \{Flora, Nora\} and \{E9, E11\} from the rest of persons and groups, consistent with the strong similarity between these nodes in the original weighted projection matrix.
-We, of course, know from the original affiliation matrix in Figure~\ref{fig:sw} that \{Flora, Nora\} are maximally similar---they are structurally equivalent---and \textit{only} attend events \{E9, E11\}.
-Accordingly, these two events end up being similar in this second dimension because they are attended by the people who are also most similar in the same dimension.
-The third CA dimension (the y-axis in panel (d) of Figure~\ref{fig:ca}), by contrast, separates \{Pearl Ruth, Dorothy, Verne, Myra\} and \{E8, E9\} from the rest; E8, along with E7, is one of the most popularly attended events, and thus contributes little to differentiating between the main partition of people and events on the first correspondence plot in panel (c).
-The pattern of attendance of \{Pearl Ruth, Dorothy, Verne, Myra\} is one characterized by high attendance at these popular events but sparser attendance at the events that serve to differentiate the two primary clusters.
-It is precisely this affinity of similar people for similar events that is picked up in the third dimension of the CA plot.
+In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships \citep{desposito2014comparison}.
+
+Another thing to note is that since the CA row and column scores in each dimension are obtained from the eigendecomposition and subsequent low-rank approximation of the degree-normalized similarity matrices $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$, each dimension provides independent information on the similarity clustering of persons and groups.\footnote{We know from inset plot (c) of Figure~\ref{fig:ca-eigvec} that the eigenvalues corresponding to the first three dimensions exhibit strong separation from the rest, suggesting that these provide a good low-rank approximation of the original degree-weighted projection matrix for persons and groups.}  Thus, as shown in panels (c) and (d) of Figure~\ref{fig:ca}, as we have already seen, the first horizontal dimension recovers the main partition separating the two dominant clusters of persons and groups~\citep{freeman2003finding}.
+The second and third dimensions, by contrast, recover secondary and tertiary similarity clusters of people and groups independent of the main two-block partition.
+Thus, the second dimension (the y-axis in panel (c) of Figure~\ref{fig:ca}) separates \{Flora, Nora\} and \{E9, E11\} from the rest of persons and groups, consistent with the strong similarity between these nodes in the original weighted projection matrix.
+We know from the original affiliation matrix in Figure~\ref{fig:sw} that \{Flora, Nora\} are maximally similar---they are structurally equivalent---and \textit{only} attend events \{E9, E11\}, which makes these events similar in the second dimension given they are attended by the two people who are also most similar in the same dimension (this also accounts for the high similarity between the two events shown in the heatmap plot of the degree-weighted projection in Figure~\ref{fig:ca}(b)).
+
+The third CA dimension, shown in the y-axis in panel (d) of Figure~\ref{fig:ca} against the first dimension separation on the x-axis, by contrast, separates \{Eleanor, Pearl, Ruth, Dorothy, Verne, Myra\} and \{E8, E9\} from the rest.
+We know from Figure~\ref{fig:bon-reord1} that event E8, along with E7, is one of the most popularly attended events and thus contributes little to differentiate between the main partition of people and events on the first correspondence plot in panel (c), hence its location near the plot's origin.
+The pattern of attendance of these six women is one characterized by affiliation with these popular events combined with sparser attendance at the events that most serve to differentiate the two primary clusters.
+In his original paper, \citet[164]{bonacich1991simultaneous} noted that this third dimension ``corresponded closely to the centrality analysis pattern,'' an observation likely driven by the affinity between the actors who load highly on this dimensions and two of the most central events by raw popularity, \{E8, E9\}.
+What we can see instead is that the third CA dimension separates actors that are most ``peripheral'' to the two main clusters according to \citet{freeman2003finding}; namely, \{Eleanor, Pearl Ruth\} on one side and \{Dorothy, Verne, Myra\} on the other, from core members of the two main blocks, who are placed in the bottom of the correspondence plot in panel (c)---low scores in the third dimension---and toward the extreme left and right ends of the plot, indicating very high absolute value scores in the first dimension.
+
 \begin{figure}
@@
 
-\subsubsection{One-Mode Projection Matrix and Eigenvector Plot} \label{sec:eigplot}
-This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
+\subsection{Breiger Projections and the Bonacich Correspondence Plot} \label{sec:eigplot}
+Earlier, we discussed how CA is sometimes considered a technique useful to generate a correspondence plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
 But as we have seen, this is \textit{not} what CA is designed to do.
-Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes.
-Notably, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two eigenvectors of the usual one-mode projections of the matrix (where the first is, of course, the standard Bonacich dual centrality score).
-
-To illustrate, Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by other-mode degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score.
+Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes \citep{desposito2014comparison}.
+In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.
+
+Here, once again, the contrast with the traditional dual Bonacich scores can prove instructive.
+Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$)---the standard \citet{breiger1974duality} projections---with the rows and columns re-ordered according by the first eigenvector of the matrix, which is equivalent to the usual Bonacich eigenvector dual centrality score for two-mode networks (see equations~\ref{eq:bon3} and~\ref{eq:bon4}).
 Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
-Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}  We can see that the plot of the first two eigenvectors does a good job of recovering the raw connectivity structure of the Southern Women affiliation network, partitioning the core persons and groups (on the upper-right) from the more peripheral ones (on the lower-left).
-
-Moreover, if all we are interested in is capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then Figure~\ref{fig:bon-sim}(c) does a better job of that than the usual CA correspondence plot in Figure~\ref{fig:ca-sim}(c).
-For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine, Sylvia\right]$ for $\left[E10, E12\right]$.
-In the same way, core events like $\left[E8, E9, E10\right]$---shown on the lower half of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the upper half.
-The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.
-
-This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's Bonacich eigenvector centrality.
-For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot.
-Similarly, the main difference between $\left[Ruth\right]$ and $\left[Sylvia\right]$, despite their comparable $C^B$ scores, is that the former's (closer to the upper half of the plot) event profile is composed mainly of core events.
-In contrast, peripheral events dominate the latter's attendance profile (shown in the bottom half of the plot), accordingly, $\left[Ruth\right]$ and $\left[Sylvia\right]$ are assigned to distinct clusters by the k-means algorithm.
-
-\subsection{Correspondence Analysis and Generalized Relational Similarity} \label{subsec:cagensim}
+Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix.
+
+This suggests that the eigenvector centralities do encode a kind of similarity between pairs of people and pairs of groups; it just happens to be the unweighted similarity scores given by their raw number of shared memberships (for people) and shared members (for groups).
+This also means that the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of members will end up being similar to lots of other people (and the same for large groups).
+This also means that higher order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode other forms of similarity between pairs of people and pairs of groups, independently of the main axis dominated by overall person activity and group size.
+
+While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice---\citep[see][for a discussion in the one-mode network case]{iacobucci2017eigenvector}---the exact meaning of this mysterious second eigenvector remains unclear, especially in the two-mode case.
+We can use the same ``a posteriori blockmodeling'' \citep{wasserman1990correspondence} approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue.
+Figure~\ref{fig:bon-reord2} shows the affiliation matrix blocked according to the \textit{second} Bonacich dimension, and Figure~\ref{fig:bon-sim} (c) shows the ``Bonacich Correspondence Plot'' with the usual eigenvector centrality score on the x-axis plotted against the second eigenvector of the Breiger projection matrix for both persons and groups on the y-axis.
+
+We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension reveals a \textit{dual triangular} structure, revealing two secondary core and peripheries that are not encoded in either the first Bonacich or CA dimensions.
+Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering of a version of the projection matrices where the proximities were designed to detect structural equivalence (e.g., the cell entry in each projection matrix was the proportion of matches between each pair of people and groups).\footnote{As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that
+attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is the position of Myrna/Myra's, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
+
+This analysis points to a new way to interpret the second Bonacich dimension, namely, as a way to reveal secondary core-periphery partitions net of a main core-periphery partition in the two-mode network (if any exists).
+Thus, in the Bonacich Correspondence Plot shown in Figure~\ref{fig:bon-sim}(c), we can see that persons and groups are ordered by their ``main'' core-periphery scores from left-to-right, while the vertical segregation reveals
+
+\section{Correspondence Analysis and Generalized Relational Similarity} \label{sec:cagensim}
 As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.

### 2025-03-25 main-R1.tex
Updates from Overleaf

@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
+        \includegraphics[width=1.0\textwidth]{Plots/ca-reord1.png}
             \caption{}
-            \label{fig:ca-reord}
+            \label{fig:ca-reord1}
     \end{subfigure}
@@
 
-The larger point is that the standard eigenvector and CA scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward a similarity-based partitioning---with the exact kind of similarity revealed by CA yet to be determined---and $C^B$ more focused on finding a core of well-connected people and groups, separated from a less well-connected periphery.
-
 \begin{figure}[ht!]
@@
 This suggests that the eigenvector centralities do encode a kind of similarity between pairs of people and pairs of groups; it just happens to be the unweighted similarity scores given by their raw number of shared memberships (for people) and shared members (for groups).
-This also means that the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of members will end up being similar to lots of other people (and the same for large groups).
-This also means that higher order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode other forms of similarity between pairs of people and pairs of groups, independently of the main axis dominated by overall person activity and group size.
-
-While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice---\citep[see][for a discussion in the one-mode network case]{iacobucci2017eigenvector}---the exact meaning of this mysterious second eigenvector remains unclear, especially in the two-mode case.
-We can use the same ``a posteriori blockmodeling'' \citep{wasserman1990correspondence} approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue.
+In this respect, the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of memberships will end up being similar to lots of other people (and the same for large groups).
+Also, just like with CA and the degree-weighted projections, higher order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode forms of similarity between pairs of people and pairs of groups, independently of the main axis dominated by overall person activity and group size.
+
+While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice \citep[see][for a discussion in the one-mode network case]{iacobucci2017eigenvector} the exact meaning of this mysterious second eigenvector of the Breiger projection matrices remains unclear, especially in the two-mode case.
+We can use the same ``a posteriori blockmodeling'' approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue \citep{wasserman1990correspondence}.
 Figure~\ref{fig:bon-reord2} shows the affiliation matrix blocked according to the \textit{second} Bonacich dimension, and Figure~\ref{fig:bon-sim} (c) shows the ``Bonacich Correspondence Plot'' with the usual eigenvector centrality score on the x-axis plotted against the second eigenvector of the Breiger projection matrix for both persons and groups on the y-axis.
@@
 We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension reveals a \textit{dual triangular} structure, revealing two secondary core and peripheries that are not encoded in either the first Bonacich or CA dimensions.
-Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering of a version of the projection matrices where the proximities were designed to detect structural equivalence (e.g., the cell entry in each projection matrix was the proportion of matches between each pair of people and groups).\footnote{As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that
-attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is the position of Myrna/Myra's, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
+Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering a version of the projection matrices of the Southern Women data where the proximities were designed to detect \textit{structural equivalence} (e.g., the cell entry in each projection matrix was the proportion of matches between each pair of people and groups).\footnote{As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is the position of Myrna/Myra's, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
+This may be partly due to the discrepant position of ``core event'' E9 (according to the main eigenvector dimension) which is grouped with the other peripheral events \{E10, E11, E12, E13, E14\} in Figure~\ref{fig:bon-reord2}.
 
 This analysis points to a new way to interpret the second Bonacich dimension, namely, as a way to reveal secondary core-periphery partitions net of a main core-periphery partition in the two-mode network (if any exists).
-Thus, in the Bonacich Correspondence Plot shown in Figure~\ref{fig:bon-sim}(c), we can see that persons and groups are ordered by their ``main'' core-periphery scores from left-to-right, while the vertical segregation reveals
+Furthermore, the similarities between persons and groups defined over the two-dimensional Bonacich space, shown in Figure~\ref{fig:bon-sim} (c), can thus be used to reveal structurally equivalent clusters of persons and events similar to those obtained by \citet[Table 5]{everett2013dual} using UCINET's combinatorial optimization routine specifying a three cluster solution of the (proportion matching) projection matrix for persons and groups.
+Figure~\ref{fig:bonclust} shows the results of a k-means analysis using the first two Bonacich dimensions to define between-person and groups distances, revealing a three cluster solution equivalent to that obtained by Borgatti and Everett, with the exception of Myrna/Myra's position as discussed earlier.
+This suggests that the Bonacich approach to similarity clustering is distinct from CA in being useful for revealing structural equivalence, thus making the Bonacich dimensions a more faithful basis for a low-rank reconstruction of the original ``input data'' than CA \citep{desposito2014comparison}.
 

### 2025-03-25 main-R1.tex

--- 2025-03-24 main-R1.tex
@@
 
-This paper contributes to the stream of previous work applying CA to analyze two-mode data.
-Nevertheless, it departs from previous efforts to highlight aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before.
-Particularly, I clarify the similarities and differences between the dual scores computed by CA on both modes and those obtained from the eigenvector-based approach proposed by \citet{bonacich1991simultaneous} for obtaining dual centralities in two-mode networks.
-These linkages are made clearer by recasting both CA and the Bonacich centralities as variations of a ``reflective'' strategy for analyzing two-mode network data via iterative dual scoring \citep{yang2022comparing, yang2020birank}, which proceeds by assigning substantively meaningful scores to the nodes in each set of a two-mode network using an iterative procedure equivalent to a spectral decomposition of a proximity matrix based on the one-mode projection.
-
-In this vein, I show how the CA approach to two-mode network analysis---re-invented in some corners of Network Science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting, van2021correspondence}--- is part of a more general toolkit for the analysis of two-mode networks.
-While the ``economic complexity'' interpretation of the CA scores on the first dimension has treated the resulting scores as ordinal rank measures---in the style of a centrality index---I review recent work that shows that they are better sense as partitioning the two sets of nodes into \textit{similarity classes} based on their patterns of connectivity to the \textit{most similar} nodes in the other set \citep{kovacs2010generalized, lizardo2024two}.
-
-Rather than ordering nodes on a unidimensional axis based on connectivity---like Bonacich's \citet{bonacich1991simultaneous} eigenvector-based approach---CA extracts latent \textit{positional} information on people and groups.
-Particularly, there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}.
-As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to \textit{similar} entities in the other mode (e.g., groups), thus following a clustering logic of ``regular equivalence.'' In this way, CA can be used as an approach to indirect blockmodeling in two-mode networks.
-This is similar to the way CA has been developed and used in some disciplines (e.g., computer and information science), namely, as a \textit{spectral clustering} method for detecting most similar node subsets in bipartite graphs \citep{wu2022spectral}.
-Overall, CA emerges as a tool central to the usual social-network-analytic task of extracting underlying structure from two-mode (and multi-mode) relational data.
-
 \subsection{Organization of the Paper} \label{subsec:org}
@@
 
-In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral science was of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper and starting in the next section, I show how a recent re-invention of CA for the analysis of ``economic complexity''---two-mode networks of countries/regions by the products they are most adept at producing---in the now well-established field of ``Network Science'' in the twenty-first century, allows us to revisit controversies and open questions as to the meaning of CA scores in two-mode network data, and the relationship between correspondence analysis, centrality analysis, and positional analysis (e.g., the discovery of structurally similar clusters of actors in two-mode networks).
+In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral sciences was one of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation at roughly contemporaneous intervals throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper and starting in the next section, I show how a recent re-invention of CA for the analysis of ``economic complexity''---two-mode networks of countries/regions by the products they are most adept at producing---in the now well-established field of ``Network Science'' in the twenty-first century, allows us to revisit controversies and open questions as to the meaning of CA scores in two-mode network data, and the relationship between correspondence analysis, centrality analysis, and positional analysis (e.g., the discovery of structurally similar clusters of actors in two-mode networks).
 Along the way, we will be able to shed light on the link between CA scores and traditional eigenvector centrality scores for two-mode networks, as well as ascertain the exact kind of similarity that is learned by estimating CA (and traditional eigenvector centrality) scores for nodes in two-mode networks along multiple dimensions.
@@
 \subsection{CA and Dual Centrality Scoring in Two-Mode Networks} \label{sec:ca}
-The reader may ask what the point of going through all this reflective scoring is since \citet{bonacich1991simultaneous} already developed a dual conception of scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected.
-Recall that, as noted earlier in that paper, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.
-
-While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures).
-In fact, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective scores discussed in the previous section \citep{mealy2019interpreting}.
+The reader may ask what the point of going through all this reflective scoring is since \citet{bonacich1991simultaneous} already developed a dual approach to scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected.
+As we have already noted, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.
+
+While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although, as we will see, the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures).
+In fact, as I will show in a bit, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective scores discussed in the previous section \citep{mealy2019interpreting}.
 This makes the connection between the Bonacich and the CA scoring approaches more intimate than Bonacich thought.
@@
 
-Where $A$ is the original affiliation matrix.
-Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$.
+Where $\mathbf{A}$ is the original affiliation matrix.
+Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda_1$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $\mathbf{A}$.
 
@@
 \begin{equation}
-    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
+    C^B_p = \frac{1}{\lambda_1}\sum_{g}a_{pg}C^B_g
     \label{eq:faust1}
@@
 \begin{equation}
-    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
+    C^B_g = \frac{1}{\lambda_1}\sum_{p}a_{pg}C^B_p
     \label{eq:faust2}
@@
 
-Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
+Where $\lambda_1$ is the largest (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
 These equations make clear that in the Bonacich dual-scoring approach, the scores assigned to people are equal to the sum of the scores of the groups they join, and the scores assigned to groups are equivalent to the sum of the scores simultaneously assigned to their members.
@@
 
-This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.
+This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode ``Breiger'' \citeyearpar{breiger1974duality} projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.
 
@@
 \subsection{Row and Column Re-ordered Affiliation Matrices} 
-Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns blocked according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic}, using scores from CA to re-order the original matrix, as they did in that paper for one-mode networks.} In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
+Figures~\ref{fig:ca-reord1} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$.
+Each panel shows the original Southern Women affiliation matrix but with rows and columns blocked according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic}, using scores from CA and Bonacich scoring to re-order the original matrix, as they did in that paper for one-mode networks.} In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
 As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure \citep[34]{wasserman1990correspondence}, and the $C^B$ re-ordered affiliation matrix having a triangular structure.
@@
 206]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic \citet{breiger1974duality} projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods.
-In fact, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is identical to that shown in~\citet[206, Table 2]{everett2013dual}.
-
-The larger point is that the standard eigenvector and CA scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward a similarity-based partitioning---with the exact kind of similarity revealed by CA yet to be determined---and $C^B$ more focused on finding a core of well-connected people and groups, separated from a less well-connected periphery.
+In fact, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is substantively identical to that shown in~\citet[206, Table 2]{everett2013dual}, who previously extracted a core-periphery partitioning of the same data.
 
@@
 
-Note that this differs from the usual interpretation of the CA correspondence plot, which is typically taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{desposito2014comparison}, or ``first order'' similarity \citet{kovacs2010generalized}.
+Note that this differs from the usual interpretation of the CA correspondence plot, which is typically taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{desposito2014comparison}, or ``first order'' similarity \citep{kovacs2010generalized}.
 This interpretation implies (for instance) that two people who attend many of the same events or groups with many members in common will appear close in the plot.
@@
 The pattern of attendance of these six women is one characterized by affiliation with these popular events combined with sparser attendance at the events that most serve to differentiate the two primary clusters.
-In his original paper, \citet[164]{bonacich1991simultaneous} noted that this third dimension ``corresponded closely to the centrality analysis pattern,'' an observation likely driven by the affinity between the actors who load highly on this dimensions and two of the most central events by raw popularity, \{E8, E9\}.
+In his original paper, \citet[164]{bonacich1991simultaneous} noted that this third dimension ``corresponded closely to the centrality analysis pattern,'' an observation likely driven by the affinity between the actors who load highly on this dimension and two of the most central events by raw popularity, \{E8, E9\}.
 What we can see instead is that the third CA dimension separates actors that are most ``peripheral'' to the two main clusters according to \citet{freeman2003finding}; namely, \{Eleanor, Pearl Ruth\} on one side and \{Dorothy, Verne, Myra\} on the other, from core members of the two main blocks, who are placed in the bottom of the correspondence plot in panel (c)---low scores in the third dimension---and toward the extreme left and right ends of the plot, indicating very high absolute value scores in the first dimension.
 
-\begin{figure}
+\subsection{Breiger Projections and the Bonacich Correspondence Plot} \label{sec:eigplot}
+Earlier, we discussed how CA is sometimes considered a technique useful to generate a correspondence plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
+But as we have seen, this is \textit{not} what CA is designed to do.
+Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes \citep{desposito2014comparison}.
+In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix---or a Markov transition matrix under a different interpretation---between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.
+
+Here, once again, the contrast with the traditional dual Bonacich scores can prove instructive.
+Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$)---the standard \citet{breiger1974duality} projections---with the rows and columns re-ordered according by the first eigenvector of the matrix, which is equivalent to the usual Bonacich eigenvector dual centrality score for two-mode networks (see equations~\ref{eq:bon3} and~\ref{eq:bon4}).
+Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
+Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix.
+
+This suggests that the eigenvector centralities do encode a kind of similarity between pairs of people and pairs of groups; it just happens to be based on the unweighted similarity scores given by the raw number of shared memberships (for people) and shared members (for groups).
+In this respect, the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of memberships will end up being ``similar`` to more people on the standard projection matrix by the simple fact of having more memberships (and the same for large groups with respect to members).
+Persons and groups high in eigenvector centrality can be thought of as being the most similar to most others, condensing the information in the Breiger projection matrix into a single ordinal rank.
+Also, just like with CA and the degree-weighted projections, higher-order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode forms of similarity between pairs of people and pairs of groups, net of the main axis dominated by overall person activity and group size.
+
+\begin{figure}[ht!]
     \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
@@
     \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel.
-The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis.}
+The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis, points are colored according a k-means three-cluster solution.}
     \label{fig:bon-sim}
@@
 
-\subsection{Breiger Projections and the Bonacich Correspondence Plot} \label{sec:eigplot}
-Earlier, we discussed how CA is sometimes considered a technique useful to generate a correspondence plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
-But as we have seen, this is \textit{not} what CA is designed to do.
-Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes \citep{desposito2014comparison}.
-In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.
-
-Here, once again, the contrast with the traditional dual Bonacich scores can prove instructive.
-Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$)---the standard \citet{breiger1974duality} projections---with the rows and columns re-ordered according by the first eigenvector of the matrix, which is equivalent to the usual Bonacich eigenvector dual centrality score for two-mode networks (see equations~\ref{eq:bon3} and~\ref{eq:bon4}).
-Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
-Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix.
-
-This suggests that the eigenvector centralities do encode a kind of similarity between pairs of people and pairs of groups; it just happens to be the unweighted similarity scores given by their raw number of shared memberships (for people) and shared members (for groups).
-This also means that the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of members will end up being similar to lots of other people (and the same for large groups).
-This also means that higher order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode other forms of similarity between pairs of people and pairs of groups, independently of the main axis dominated by overall person activity and group size.
-
-While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice---\citep[see][for a discussion in the one-mode network case]{iacobucci2017eigenvector}---the exact meaning of this mysterious second eigenvector remains unclear, especially in the two-mode case.
-We can use the same ``a posteriori blockmodeling'' \citep{wasserman1990correspondence} approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue.
-Figure~\ref{fig:bon-reord2} shows the affiliation matrix blocked according to the \textit{second} Bonacich dimension, and Figure~\ref{fig:bon-sim} (c) shows the ``Bonacich Correspondence Plot'' with the usual eigenvector centrality score on the x-axis plotted against the second eigenvector of the Breiger projection matrix for both persons and groups on the y-axis.
-
-We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension reveals a \textit{dual triangular} structure, revealing two secondary core and peripheries that are not encoded in either the first Bonacich or CA dimensions.
-Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering of a version of the projection matrices where the proximities were designed to detect structural equivalence (e.g., the cell entry in each projection matrix was the proportion of matches between each pair of people and groups).\footnote{As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that
-attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is the position of Myrna/Myra's, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
+While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice \citep[see][for discussion in the one-mode network case]{iacobucci2017eigenvector} the exact meaning of the mysterious second eigenvector remains unclear in the two-mode case.
+We can use the same ``a posteriori blockmodeling'' approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue \citep{wasserman1987stochastic}.
+Figure~\ref{fig:bon-reord2} shows the affiliation matrix blocked according to the \textit{second} Bonacich dimension, and Figure~\ref{fig:bon-sim} (c) shows the ``Bonacich Correspondence Plot'' with the usual eigenvector centrality score on the horizontal x-axis plotted against the second eigenvector of the Breiger projection matrix for both persons and groups on the vertical y-axis.
+
+We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension diplays a \textit{dual triangular} structure, revealing two secondary core and peripheries not evident in the blocking according to either the first CA or standard eigenvector scores.
+Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering a version of the projection matrices of the Southern Women data where the proximities were designed to detect \textit{structural equivalence}.\footnote{In that application, the cell entry in each projection matrix was the \textit{proportion of matches} between each pair of people and groups.
+As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is Myrna/Myra's position, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
+This may be partly due to the discrepant position of ``core event'' E9 along the columns of the blocked matrix which is grouped with the second group of ``peripheral'' events \{E10, E11, E12, E13, E14\} in Figure~\ref{fig:bon-reord2}.
 
 This analysis points to a new way to interpret the second Bonacich dimension, namely, as a way to reveal secondary core-periphery partitions net of a main core-periphery partition in the two-mode network (if any exists).
-Thus, in the Bonacich Correspondence Plot shown in Figure~\ref{fig:bon-sim}(c), we can see that persons and groups are ordered by their ``main'' core-periphery scores from left-to-right, while the vertical segregation reveals
+Furthermore, the similarities between persons and groups defined over the two-dimensional Bonacich space, shown in Figure~\ref{fig:bon-sim} (c), can thus be used to reveal \textit{structurally equivalent} clusters of persons and events similar to those obtained by \citet[Table 5]{everett2013dual}.
+This is confirmed by the coloring of the persons and groups in Figure~\ref{fig:bon-reord2}) which comes from a k-means cluster analysis using the first two Bonacich dimensions to define initial cluster centroids, revealing a three-cluster solution almost equivalent to that obtained by Everett and Borgatti, with the exception of Myrna/Myra's and event E9's assingment as discussed earlier.
+This suggests that the Bonacich approach to similarity clustering is distinct from CA in being useful for revealing structural equivalence, thus making the Bonacich dimensions a more faithful basis for a low-rank reconstruction of the original ``input data'' than CA if that is the goal of the analysis \citep{desposito2014comparison}.
 
@@
 
-Which is consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
+This definition consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
 To accomplish this, \citet[540, eq.

### 2025-03-25 main-R1.tex

@@
 Section~\ref{sec:comparing} uses the classic Southern Women dataset to show the payoff of this approach, focusing on the distinct substantive insights extracted by CA and Bonacich scoring in a venerable dataset, showing that the CA-blocked affiliation matrix reveals a dual block partition between the two node sets (similar to that recovered by previous analyses of the same data using other methods), while the Bonacich-blocked affiliation recovers a core-periphery structure instead.
-Section~\ref{sec:cagensim} considers the hanging question of the exact substantive interpretation of the node groupings revealed by CA-style blocking, which are shown to be fairly close to those obtained by a computationally unrelated approach based on the idea of ``generalized relational similarity,'' suggesting that the CA blocks obey a logic distinct from structural equivalence.
+In addition I show that considering higher dimensions of the Bonacich-style decomposition allows us to discover secondary core-periphery structures in the two-mode network, while also leading to a meaningful clustering of the nodes based on structural equivalence.
+Section~\ref{sec:cagensim} considers the hanging question of the exact substantive interpretation of the node clusters revealed by CA-style blocking, which are shown to be fairly close to those obtained by a computationally unrelated approach based on the idea of ``generalized relational similarity,'' suggesting that the CA blocks are distinct from those obtained following the logic of structural equivalence.
 Finally, Section~\ref{sec:disc} closes with a summary of key points in the argument and suggestions for future development of the linkages between CA and positional and centrality analysis in two-mode networks.
@@
         \includegraphics[width=1.0\textwidth]{Plots/ca-reord1.png}
-            \caption{}
+            \caption{First CA Dimension}
             \label{fig:ca-reord1}
@@
         \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
-            \caption{}
+            \caption{First Bonacich Dimension}
             \label{fig:bon-reord1}
@@
         \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
-            \caption{}
+            \caption{Second Bonacich Dimension}
             \label{fig:bon-reord2}
     \end{subfigure}
-    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score, (c) row and columns re-ordered according to person and group score on the second Bonacich dimension.}
+     \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=1.0\textwidth]{Plots/SR-reord1.png}
+            \caption{Second SimRank Eigenvector}
+            \label{fig:sr-reord1}
+    \end{subfigure}
+    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) rows and columns re-ordered according to the person and group score on the first CA dimension, (b) rows and columns re-ordered according to person and group Eigenvector centrality score, (c) rows and columns re-ordered according to person and group score on the second Bonacich dimension, (d) rows and columns re-ordered according to the second eigenvector of the SimRank similarity matrix.}
     \label{fig:ca-v-bon}
@@
 Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections \citep{mealy2019interpreting, van2021correspondence}.
-
-And indeed, they are not.
-Figure~\ref{fig:simrank}(a) shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices---with $\alpha$ set to $0.8$---for the row and column nodes.
-Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two communities of persons and groups revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
-
-Figure~\ref{fig:simrank}(b) and Figure~\ref{fig:simrank}(d) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
-Looking at the plots from left to right, we can see that the partitioning of the node sets on the first informative eigenvector is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
-
-Figure~\ref{fig:simrank}(f) shows the correspondence plot of the first and second eigenvectors of the SimRank similarity matrices for both persons and groups.
-Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c), suggesting that the first-CA dimensions of the two-mode network recover clusters of ``similar'' entities in the two node sets, where internode similarity is a generalized relational similarity as defined earlier.
+Beyond that, the SimRank approach can be thought of as yet another way of obtaining one-mode projections from an original two-mode network, in line with Everett and Borgatti's ``dual projection'' approach \citep{everett2013dual}.
+
+Figure~\ref{fig:sr-reord1} shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices---with $\alpha$ set to $0.8$---for the row and column nodes.
+Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling displays a block-diagonal structure separating the two communities of persons and groups as that revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
+Figure~\ref{fig:simrank}(a) and Figure~\ref{fig:simrank}(b) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
+Looking at the plots from left to right, we can see that the partitioning of the node sets on the most informative eigenvector of the SimRank similarity matrix is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
+
+Figure~\ref{fig:simrank}(f) shows the correspondence plot of the two main non-trivial eigenvectors of the SimRank similarity matrices (corresponding to the second and third eigevectors) for both persons and groups.
+Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c), suggesting that the first-CA dimensions of the two-mode network recover clusters of ``similar'' entities in the two node sets, where internode similarity is a generalized relational similarity.
 Thus, [$Flora, Olivia$] appear close because they choose similar groups, as do [$Sylvia, Katherine, Helen$].

### 2025-03-25 main-R1.tex
Updates from Overleaf

@@
     \caption{CA Eigenvector plot for persons (a) and groups (b).
-The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis.
-The inset plot (c) shows the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
+The first dimension CA score is on the x-axis, and the rank order of persons and groups on the same score is on the y-axis.
+The inset plot (c) shows the eigenvalue orderings on the x-axis ($k$) for persons and groups, showing a separation between the first three eigenvalues and the rest.}
     \label{fig:ca-eigvec}
@@
 Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling displays a block-diagonal structure separating the two communities of persons and groups as that revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
-Figure~\ref{fig:simrank}(a) and Figure~\ref{fig:simrank}(b) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
+Figure~\ref{fig:simrank-v-ca}(a) and Figure~\ref{fig:simrank-v-ca}(b) show an eigenvector plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
 Looking at the plots from left to right, we can see that the partitioning of the node sets on the most informative eigenvector of the SimRank similarity matrix is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
-
-Figure~\ref{fig:simrank}(f) shows the correspondence plot of the two main non-trivial eigenvectors of the SimRank similarity matrices (corresponding to the second and third eigevectors) for both persons and groups.
-Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c), suggesting that the first-CA dimensions of the two-mode network recover clusters of ``similar'' entities in the two node sets, where internode similarity is a generalized relational similarity.
-Thus, [$Flora, Olivia$] appear close because they choose similar groups, as do [$Sylvia, Katherine, Helen$].
-In the same way, groups [$E9, E11$] are close in the plot because they are chosen by similar people, as are events [$E10, E12, E13, E14$].
-Figures~\ref{fig:simrank}(g) and (h) show a regression plot of the first CA eigenvector for persons (top) and groups (bottom) on the x-axis and the first eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
+Figures~\ref{fig:simrank-v-ca}(c) and (d) show a regression plot of the first CA dimension scores for persons and groups (respectively) on the x-axis and the SimRank scores extracted from the main non-trivial eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
 As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that these first eigenvectors capture the same (GSR) information across the two approaches.

### 2025-03-30 analysis-R1.qmd
do it

@@
 
-```{r CA versus Bonacich Re-ordered Affiliation Matrices}
-      # Affiliation matrix plot (CA ordering)
-      A.ord <- as.matrix(A[order(CA.r$vectors[, 2]), order(CA.c$vectors[, 2])])
-      p <- ggcorrplot(t(A.ord)) 
-      p <- p + scale_x_discrete(position = "top")
-      p <- p + scale_y_discrete(limits = rev)
-      p1 <- p + theme(legend.position = "none",
-                               axis.text.x = element_text(hjust = -0.2))
-      # Affiliation matrix plot (Bonacich ordering)
-      A.ord <- as.matrix(A[order(eig.r[, 1]), order(eig.c[, 1])])
-      p <- ggcorrplot(t(A.ord)) 
-      p <- p + scale_x_discrete(position = "top")
-      p <- p + scale_y_discrete(limits = rev)
-      p2 <- p + theme(legend.position = "none",
-                               axis.text.x = element_text(hjust = -0.2)) 
-     # Affiliation matrix plot (Bonacich ordering)
-      A.ord <- as.matrix(A[order(eig.r[, 2]), order(eig.c[, 2])])
-      p <- ggcorrplot(t(A.ord)) 
-      p <- p + scale_x_discrete(position = "top")
-      p <- p + scale_y_discrete(limits = rev)
-      p3 <- p + theme(legend.position = "none",
-                               axis.text.x = element_text(hjust = -0.2)) 
-      png(file = here("Plots", "ca-reord.png"), width=500, height=500)
-      p1
-      dev.off() 
-      png(file = here("Plots", "bon-reord1.png"), width=500, height=500)
-      p2
-      dev.off() 
-      png(file = here("Plots", "bon-reord2.png"), width=500, height=500)
-      p3
-      dev.off()
-```
-
 ```{r Row Stochastic Matrices for CA Projections Plot}
@@
              mutate(value = value / max(value))
+
       # Eigenvector plot for the people
@@
                                order(CA.c$vectors[,2])]
+      # CA Spectral Clustering
+      ca.dat <- data.frame(rbind(CA.r$vectors[,2:4], CA.c$vectors[,2:4]))
+      ca.clus.p <- kmeans(ca.dat[1:18, ], centers = 3, nstart = 25)
+      ca.clus.g <- kmeans(ca.dat[19:32, ], centers = 3, nstart = 25)
+      ca.clus <- kmeans(ca.dat, centers = 3, nstart = 25)
 
@@
 
+      # K-means clustering of Bonacich solution
+      eig.dat <- data.frame(d1 = eig.d1, d2 = eig.d2)
+      eig.clus.p <- kmeans(eig.dat[1:18,], centers = 3, nstart = 25)$cluster
+      eig.clus.g <- kmeans(eig.dat[19:32,], centers = 3, nstart = 25)$cluster
+
       # Bonacich Correspondence Plot
@@
       eig.d2 <- c(eig.r[, 2], eig.c[, 2])*-1
-      plot.dat <- data.frame(eig.d1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
-      p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(mode), 
+      plot.dat <- data.frame(eig.d1, eig.d2, 
+                             mode = c(rep(1, 18), rep(2, 14)),
+                             clus = c(eig.clus.p, eig.clus.g+3)
+                             )
+      p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(clus), 
                                 label = c(rownames(A), colnames(A))))
@@
       p <- p + geom_hline(yintercept = 0, color = "gray")
-      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
+      p <- p + geom_text_repel(size = 4, max.overlaps = 50, fontface = "bold")
       p <- p + theme_minimal() 
@@
                      )
-      p <- p + labs(x = "First Dimension", y = "Second Dimension")
-      c <- p + scale_color_manual(values = c('red', 'blue'))
+      c <- p + labs(x = "First Dimension", y = "Second Dimension")
+      #c <- p + scale_color_manual(values = c('red', 'blue'))
       png(file = here("Plots", "bon-corr-plot.png"), width=1500, height=1000)
@@
 
+```{r SimRank Similarity Matrices}
+   SR.res <- SimRank(A)
+   SR.p <- SR.res$Sr
+   SR.g <- SR.res$Sc
+   rownames(SR.p) <- substr(colnames(SR.p), start = 1, stop = 3)
+   rownames(SR.g) <- substr(colnames(SR.g), start = 1, stop = 3)
+   a <- ggcorrplot(SR.p, lab = TRUE, lab_size = 5,
+                     colors = c(rep("white", 3))) +
+      theme(legend.position = "none", 
+            axis.text.x = element_text(hjust = -0.2, size = 16),
+            axis.text.y = element_text(size = 16)) + 
+      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
+   b <- ggcorrplot(SR.g, lab = TRUE, lab_size = 5,
+                     colors = c(rep("white", 3))) +
+      theme(legend.position = "none", 
+            axis.text.x = element_text(hjust = -0.2, size = 16),
+            axis.text.y = element_text(size = 16)) + 
+      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
+   png(file = here("Plots", "p-SimRank.png"), width=900, height=900)
+   P.p
+   dev.off() 
+   png(file = here("Plots", "g-SimRank.png"), width=900, height=900)
+   P.g
+   dev.off() 
+```
+
 ```{r CA versus SimRank}
-      res <- SimRank(A)
-      Sr <- res$Sr
-      Sc <- res$Sc
-      s1.r <- eigen(Sr)$vectors[, 2]
-      s1.c <- eigen(Sc)$vectors[, 2]
-      diag(Sr) <- 0
-      diag(Sc) <- 0
-      Sr <- Sr[order(d1.r), order(d1.r)]
-      Sc <- Sc[order(d1.c), order(d1.c)]
-      rownames(Sr) <- substr(colnames(P.pp), start = 1, stop = 3)
-      rownames(Sc) <- substr(colnames(P.gg), start = 1, stop = 3)
-      a <- ggcorrplot(Sr, lab = TRUE, show.diag = FALSE, lab_size = 3) +
-         theme(legend.position = "none", 
-               axis.text.x = element_text(hjust = -0.2)) + 
-         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
-         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
-                              low = "white", mid = "yellow", high = "purple")
+      # SimRank Eigenvector plot data
+      SR.eig.p <- eigen(SR.res$Sr)$vectors
+      SR.eig.g <- eigen(SR.res$Sc)$vectors
+      eigvec.dat.r <- data.frame(rank = rank(SR.eig.p[, 2]), 
+                             value = as.numeric(SR.eig.p[, 2]),
+                             lab = rownames(A))
+      eigvec.dat.c <- data.frame(rank = rank(SR.eig.g[, 2]), 
+                             value = as.numeric(SR.eig.g[, 2]),
+                             lab = colnames(A))
 
-      b <- ggcorrplot(Sc, lab = TRUE, show.diag = FALSE, lab_size = 3) +
-         theme(legend.position = "none", 
-               axis.text.x = element_text(hjust = -0.2)) + 
-         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
-         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
-                              low = "white", mid = "yellow", high = "purple")
+      # Eigenvector plot for the people
+      p <- ggscatter(eigvec.dat.r, y = "rank", x = "value", point = FALSE,
+                 font.label =  10, label = "lab", repel = TRUE)
+      p <- p + theme(axis.text.y = element_blank(),
+                 axis.line.x = element_blank(),
+                 axis.line.y = element_blank(),
+                 axis.ticks = element_blank(),
+                 legend.position = "none")
+      p <- p + labs(y = "Rank", x = "First Axis") 
+      a <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")
 
-      dat.r <- data.frame(ca = d1.r,
-                          sr = s1.r,
+      # Eigenvector plot for the groups
+      p <- ggscatter(eigvec.dat.c, y = "rank", x = "value", point = FALSE,
+                 font.label =  10, label = "lab", repel = TRUE)
+      p <- p + theme(axis.text.y = element_blank(),
+                 axis.line.x = element_blank(),
+                 axis.line.y = element_blank(),
+                 axis.ticks = element_blank(),
+                 plot.margin = margin(l = 100),
+                 legend.position = "none")
+      p <- p + labs(y = "Rank", x = "First Axis") 
+      b <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")
+
+      # Data for scatterplot
+      dat.p <- data.frame(ca = CA.r$vectors[, 2],
+                          sr = SR.eig.p[, 2],
                           lab = rownames(A))
-      dat.c <- data.frame(ca = d1.c,
-                          sr = s1.c,
+      dat.g <- data.frame(ca = CA.c$vectors[, 2],
+                          sr = SR.eig.g[, 2],
                           lab = colnames(A))
-      cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
 
-      c <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
+      # SimRank versus CA (People)
+      cor <- round(cor(dat.p[,c("ca", "sr")])[1, 2], 2)
+      c <- ggscatter(dat.p, x = "ca", y = "sr", font.label= 8, 
                      label = "lab", repel = TRUE, 
@@
                      annotate("text", x = -.2, y = .3, 
-                              label = paste("r = ", cor), size = 6) 
-      cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
+                              label = paste("r = ", cor), size = 6) +
+                     labs(x = "CA Scores", y = "SimRank Scores")
 
-      d <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
+      # SimRank versus CA (Groups)
+      cor <- round(cor(dat.g[,c("ca", "sr")])[1, 2], 2)
+      d <- ggscatter(dat.g, x = "ca", y = "sr", font.label= 8, 
                      label = "lab", repel = TRUE, 
@@
                      annotate("text", x = -.2, y = .3, 
-                              label = paste("r = ", cor), size = 6) 
+                              label = paste("r = ", cor), size = 6) +
+                     labs(x = "CA Scores", y = "SimRank Scores")
+
+      # Composite Plot
       png(file = here("Plots", "simrank-v-ca.png"), width=1000, height=900)
-      (a + c + plot_layout(widths = c(2, 1))) / (b + d + plot_layout(widths = c(2, 1))) + 
-         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
+         (a + b)/(c + d) + plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
          theme(plot.tag = element_text(size = 26))
@@
 ```
+
+```{r Blocked Affiliation Matrices}
+      # First CA Dimension
+      A.ord <- as.matrix(A[order(CA.r$vectors[, 2]), order(CA.c$vectors[, 2])])
+      p <- ggcorrplot(t(A.ord)) 
+      p <- p + scale_x_discrete(position = "top")
+      p <- p + scale_y_discrete(limits = rev)
+      p1 <- p + theme(legend.position = "none",
+                               axis.text.x = element_text(hjust = -0.2))
+      # First Bonacich Dimension
+      A.ord <- as.matrix(A[order(eig.r[, 1]), order(eig.c[, 1])])
+      p <- ggcorrplot(t(A.ord)) 
+      p <- p + scale_x_discrete(position = "top")
+      p <- p + scale_y_discrete(limits = rev)
+      p2 <- p + theme(legend.position = "none",
+                               axis.text.x = element_text(hjust = -0.2)) 
+     # Second Bonacich Dimension
+      A.ord <- as.matrix(A[order(eig.r[, 2]), order(eig.c[, 2])])
+      p <- ggcorrplot(t(A.ord)) 
+      p <- p + scale_x_discrete(position = "top")
+      p <- p + scale_y_discrete(limits = rev)
+      p3 <- p + theme(legend.position = "none",
+                               axis.text.x = element_text(hjust = -0.2)) 
+     # Second SimRank Eigenvector
+      A.ord <- as.matrix(A[order(s2.r), order(s2.c)])
+      p <- ggcorrplot(t(A.ord)) 
+      p <- p + scale_x_discrete(position = "top")
+      p <- p + scale_y_discrete(limits = rev)
+      p4 <- p + theme(legend.position = "none",
+                               axis.text.x = element_text(hjust = -0.2))
+      # Saving plots
+      png(file = here("Plots", "ca-reord1.png"), width=500, height=500)
+      p1
+      dev.off() 
+      png(file = here("Plots", "bon-reord1.png"), width=500, height=500)
+      p2
+      dev.off() 
+      png(file = here("Plots", "bon-reord2.png"), width=500, height=500)
+      p3
+      dev.off()
+      png(file = here("Plots", "SR-reord1.png"), width=500, height=500)
+      p4
+      dev.off()
+```

### 2025-03-30 analysis-R1.qmd
do it

@@
    p2
+   dev.off()
+```
+
+```{r Correlation between odds and even reflections and Bonacich Score}
+   a <- reflections(A, iter = 42)
+   p.e.r <- a$p.r[, evens(1:40)]
+   g.e.r <- a$g.r[, evens(1:40)]
+   p.o.r <- a$p.r[, odds(3:41)]
+   g.o.r <- a$g.r[, odds(3:41)]
+   plot.dat.p <- data.frame(p.o = as.vector(abs(cor(eig.r[, 1], p.o.r))),
+                            p.e = as.vector(abs(cor(eig.r[, 1], p.e.r))),
+                            ref.o = seq(3, 41, by = 2),
+                            ref.e = seq(2, 40, by = 2)
+                            )
+   p <- ggplot(plot.dat.p, aes(x = ref.e, y = p.e))
+   p <- p + geom_point(size = 2) + geom_line()
+   p <- p + ylim(0, 1) + labs(x = "Even Reflection", y = "Bonacich Correlation")
+   p1 <- p + scale_x_continuous(breaks = seq(2, 40, by = 4))
+   p <- ggplot(plot.dat.p, aes(x = ref.o, y = p.o))
+   p <- p + geom_point(size = 2) + geom_line()
+   p <- p + ylim(0, 1) + labs(x = "Odd Reflection", y = "Bonacich Correlation")
+   p2 <- p + scale_x_continuous(breaks = seq(3, 41, by = 4))
+   png(file = here("Plots", "bon-ref-corr-p.png"), width=800, height=400)
+      p1 + p2 & theme_minimal()
+   dev.off()
+   plot.dat.g <- data.frame(p.o = as.vector(abs(cor(eig.c[, 1], g.o.r))),
+                            p.e = as.vector(abs(cor(eig.c[, 1], g.e.r))),
+                            ref.o = seq(3, 41, by = 2),
+                            ref.e = seq(2, 40, by = 2)
+                            )
+   p <- ggplot(plot.dat.g, aes(x = ref.e, y = p.e))
+   p <- p + geom_point(size = 2) + geom_line()
+   p <- p + ylim(0, 1) + labs(x = "Even Reflection", y = "Bonacich Correlation")
+   p1 <- p + scale_x_continuous(breaks = seq(2, 40, by = 4))
+   p <- ggplot(plot.dat.g, aes(x = ref.o, y = p.o))
+   p <- p + geom_point(size = 2) + geom_line()
+   p <- p + ylim(0, 1) + labs(x = "Odd Reflection", y = "Bonacich Correlation")
+   p2 <- p + scale_x_continuous(breaks = seq(3, 41, by = 4))
+   png(file = here("Plots", "bon-ref-corr-g.png"), width=800, height=400)
+      p1 + p2 & theme_minimal()
    dev.off()
@@
 ```{r Bonacich Similarity and Correspondence Plot}
-      norm.sim.r <- A %*% t(A)
-      norm.sim.c <- t(A) %*% A
+      sim.p <- A %*% t(A)
+      sim.g <- t(A) %*% A
       diag(norm.sim.r) <- 0
       diag(norm.sim.c) <- 0
-      norm.sim.r <- norm.sim.r[order(eig.r[, 1]), order(eig.r[ ,1])]
-      norm.sim.c <- norm.sim.c[order(eig.c[, 1]), order(eig.c[ ,1])]
+      sim.p <- sim.p[order(eig.r[, 1]), order(eig.r[ ,1])]
+      sim.g <- sim.g[order(eig.c[, 1]), order(eig.c[ ,1])]
       # Bonacich Similarity Plot (People)
-      p <- ggcorrplot(norm.sim.r)
+      p <- ggcorrplot(sim.p)
       p <- p + scale_x_discrete(position = "top")
       p <- p + scale_y_discrete(limits = rev)
-      p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.r+1)), 
+      p <- p + scale_fill_gradient2(limit = c(0, max(sim.p+1)), 
                                     low = "white", 
@@
                                     mid = "red",  
-                                    midpoint = max(norm.sim.r)/2)
+                                    midpoint = max(sim.p)/2)
       a <- p + theme(legend.position = "none",
@@
       # Bonacich Similarity Plot (Groups)
-      p <- ggcorrplot(norm.sim.c)
+      p <- ggcorrplot(sim.g)
       p <- p + scale_x_discrete(position = "top")
       p <- p + scale_y_discrete(limits = rev)
-      p <- p + scale_fill_gradient2(limit = c(0, max(norm.sim.c+1)), 
+      p <- p + scale_fill_gradient2(limit = c(0, max(sim.g+1)), 
                                     low = "white", 
@@
                                     mid = "red",  
-                                    midpoint = max(norm.sim.c)/2)
+                                    midpoint = max(sim.g)/2)
       b <- p + theme(legend.position = "none", plot.margin = margin(l = 100),
@@
       # K-means clustering of Bonacich solution
-      eig.dat <- data.frame(d1 = eig.d1, d2 = eig.d2)
+      eig.dat <- data.frame(d1 = c(eig.r[, 1], eig.c[, 1]), 
+                            d2 = c(eig.r[, 2], eig.c[, 2]))
       eig.clus.p <- kmeans(eig.dat[1:18,], centers = 3, nstart = 25)$cluster
@@
       # Bonacich Correspondence Plot
-      eig.d1 <- c(eig.r[, 1], eig.c[, 1])*-1
-      eig.d2 <- c(eig.r[, 2], eig.c[, 2])*-1
-      plot.dat <- data.frame(eig.d1, eig.d2, 
+      plot.dat <- data.frame(cbind(eig.dat), 
                              mode = c(rep(1, 18), rep(2, 14)),
@@
                              )
-      p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(clus), 
+      p <- ggplot(plot.dat, aes(y = d2, x = d1*-1, 
+                                color = factor(clus), 
                                 label = c(rownames(A), colnames(A))))

### 2025-03-30 main-R1-short.tex
Updates from Overleaf

--- 2025-03-25 main-R1.tex
@@
 \begin{abstract}
-This paper reconsiders the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting aspects that have not been previously emphasized and clarifying its linkages to two-mode centrality.
-I show that using CA to compute a dual score on both sets of nodes in a two-mode network is related to but not mathematically equivalent to the \citet{bonacich1991simultaneous} two-mode centrality.
-This ``reflective'' approach to computing dual scores connects the use of CA in two-mode network analysis to its use in other disciplines as a method of ordination.
-Additionally, I show that CA can extract latent positional information on people and groups, connecting it to recent work on extracting ``generalized relational similarities'' in two-mode networks.
+This paper reconsiders the status of Correspondence Analysis (CA) as a tool for analyzing two-mode networks, comparing it with the Bonacich dual centrality approach and revealing their mathematical linkages as eigenvector-based methods.
+While Bonacich centrality identifies core-periphery structures and is helpful for clustering nodes based on the criterion of similarity via structural equivalence, CA is best at detecting subsets of actors and events based on a generalized relational similarity criterion, thus coming closer to clustering via regular equivalence.
+In the end, both CA and Bonacich centrality emerge as valuable, but distinct strategies for the dual projection analysis of two-mode networks, emphasizing the duality between actors and events.
 \end{abstract}
@@
 \begin{highlights}
-  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
-  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
-  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
+    \item Correspondence Analysis (CA) and Bonacich centrality are mathematically related, both being eigenvector-based methods applied to one-mode projections of two-mode networks.
+    \item Bonacich centrality is better suited for identifying core-periphery structures and structural equivalence-based clustering of nodes in the two-mode network.
+    \item CA excels at revealing clusters of nodes in both modes based on regular equivalence or ``generalized relational similarity.''
+    \item Both CA and Bonacich centrality are valuable, but distinct, tools for analyzing two-mode networks through a ``dual projection'' approach.
 \end{highlights}
@@
 \begin{keywords}
-Correspondence Analysis \sep Generalized Relational Similarity \sep Duality \sep Two-mode networks \sep Centrality 
+Correspondence Analysis \sep Bonacich Centrality \sep Duality \sep Two-mode networks \sep Clustering 
 \end{keywords}
@@
 \maketitle
+
 \newpage
-\section{Introduction} \label{sec:intro}
-While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature.
+\section{Introduction} \label{sec:1intro}
+Despite not being a frequently used tool for analyzing two-mode network data, Correspondence Analysis---hereafter CA---has had a somewhat rocky and distinctive career in the social networks literature.
 Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{desposito2014use}.
+
+This paper reconsiders the uses of CA as a tool in the analysis of the two-mode networks, clarifying its similarities and differences to other widely used approaches, such as Bonacich's \citeyearpar{bonacich1991simultaneous} dual centrality scoring.
+I show that both CA and Bonacich centrality analysis are close mathematical cousins, and can be thought of as eigenvector-based decomposition strategies of proximity matrices extracted from the one-mode projection of the original two-mode affiliation matrix.
+The key differences between CA and the Bonacich approach concern the type of underlying structure in the two-mode network that they are sensitive two, with the Bonacich approach more useful for detecting primary and secondary core-periphery partitions---and when used as a blocking tool revealing clustering of nodes based on structural equivalence---and CA more useful for revealing clusters of similar nodes, where similarity relations obey a logic closer to that of regular equivalence, or ``generalized relational similarity'' \citep{kovacs2010generalized}.
+In this respect, both Correspondence Analysis and Bonacich dual centrality analysis emerge as equally useful but distinct strategies in the extended toolkit of options useful for analyzing two-mode networks from a ``dual projection'' perspective \citep{everett2013dual}.
 
@@
 Section~\ref{sec:comparing} uses the classic Southern Women dataset to show the payoff of this approach, focusing on the distinct substantive insights extracted by CA and Bonacich scoring in a venerable dataset, showing that the CA-blocked affiliation matrix reveals a dual block partition between the two node sets (similar to that recovered by previous analyses of the same data using other methods), while the Bonacich-blocked affiliation recovers a core-periphery structure instead.
-In addition I show that considering higher dimensions of the Bonacich-style decomposition allows us to discover secondary core-periphery structures in the two-mode network, while also leading to a meaningful clustering of the nodes based on structural equivalence.
+In addition, I show that considering higher dimensions of the Bonacich-style decomposition allows us to discover secondary core-periphery structures in the two-mode network, while also leading to a meaningful clustering of the nodes based on structural equivalence.
 Section~\ref{sec:cagensim} considers the hanging question of the exact substantive interpretation of the node clusters revealed by CA-style blocking, which are shown to be fairly close to those obtained by a computationally unrelated approach based on the idea of ``generalized relational similarity,'' suggesting that the CA blocks are distinct from those obtained following the logic of structural equivalence.
@@
 
-For reasons of space and relevance, my discussion of previous work is circumscribed in four ways, focusing on (1) pieces discussing primarily \textit{methodological} or interpretive aspects of CA, (2) regarding the analysis of \textit{two-mode network} data or affiliation networks,\footnote{There are, of course, various substantive applications of CA for the analysis of two-mode networks~\citep[e.g.,][]{breiger2000tool, faust2002scaling, schweizer1991power, serino2024mapping, ragozini2018analysis} as well as substantive and methodological applications of CA to the analysis of one-mode network data~\citep[e.g.,][]{noma1985scaling, kumbasar1994systematic, lizardo2020correspondence}; these will not be part of the discussion here.} (3) analyzing the usual \textit{binary} affiliation matrix measuring a single relation between persons and groups at one point in time, where links are observed only between persons and groups,\footnote{As such, recent work dealing with the application of CA or other factoring approaches to time-varying, multiplex, multilevel, or valued two-mode network data \citep{ragozini2014correspondence, ragozini2015multiple, zhu2016correspondence} will also be outside the scope of my discussion.} using with the \textit{simple CA} of the affiliation matrix.\footnote{There has been a spate of recent work focusing on the uses and advantages of using Multiple Correspondence Analysis or MCA---a variant of CA applied to a disjunctively-coded, doubly-centered version of the original affiliation matrix) to analyze and visualize two-mode networks \citep[e.g.,][]{dramalidis2016subset, desposito2014use, ragozini2014correspondence}.
+For reasons of space and relevance, my discussion of previous work is circumscribed in four ways, focusing on (1) pieces discussing primarily \textit{methodological} or interpretive aspects of CA, (2) regarding the analysis of \textit{two-mode network} data or affiliation networks,\footnote{There are, of course, various substantive applications of CA for the analysis of two-mode networks~\citep[e.g.,][]{breiger2000tool, faust2002scaling, schweizer1991power, serino2024mapping, ragozini2018analysis} as well as substantive and methodological applications of CA to the analysis of one-mode network data~\citep[e.g.,][]{noma1985scaling, kumbasar1994systematic, lizardo2020correspondence}; these will not be part of the discussion here.} (3) analyzing the usual \textit{binary} affiliation matrix measuring a single relation between persons and groups at one point in time, where links are observed only between persons and groups,\footnote{As such, recent work dealing with the application of CA or other factoring approaches to time-varying, multiplex, multilevel, or valued two-mode network data \citep{ragozini2014correspondence, ragozini2015multiple, zhu2016correspondence} will also be outside the scope of my discussion.} using \textit{simple CA} of the affiliation matrix.\footnote{There has been a spate of recent work focusing on the uses and advantages of using Multiple Correspondence Analysis or MCA---a variant of CA applied to a disjunctively-coded, doubly-centered version of the original affiliation matrix) to analyze and visualize two-mode networks \citep[e.g.,][]{dramalidis2016subset, desposito2014use, ragozini2014correspondence}.
 Because much of what I have to say deals with the uses and interpretation of the scores obtained from simple CA applied directly to the affiliation matrix (or its one-mode projections), my discussion will not be directly relevant to this line of work, except for a recent paper directly comparing the substantive interpretation of the $\chi^2$ distances in simple CA and MCA with regard to their substantive interpretation as indexes of structural similarity, which bears more directly to some of the subsequent discussion \citep{desposito2014comparison}.}
@@
 
-\subsection{Previous Work on CA and Two-Mode Networks}
 One of the earliest considerations of CA as a data-analytic tool for examining duality relations in two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks.
@@
 As we will see in section~\ref{subsec:refeigen}, CA and centrality analysis are mathematically linked in a second way not noted by Bonacich, as they are also based on the spectral decomposition of the dual one-mode \textit{projections} of the original rectangular affiliation matrix.
-In this particular respect, the mathematics connecting the Eigenvector approach to centrality analysis and CA are more similar than even Bonacich appreciated.
+In this particular respect, the mathematical connection between the Eigenvector approach to centrality analysis and CA is more similar than even Bonacich appreciated.
 
@@
 Faust noted that one thing both techniques had in common was that they both produced scores that truly exploited the duality of the two modes in the data.
-In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the event she attends; for events, their scores are given by the sum of the eigenvector centralities of its members.
+In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the event she attends; for events, their scores are given by the sum of the eigenvector centralities of their members.
 Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship was obtained between the CA scores corresponding to persons and groups.
@@
 Second, Faust specifies the exact relationship linking the resulting row and column CA scores and the original affiliation matrix input, providing the ``reconstruction equation'' that takes us from the left and right singular vectors and eigenvalues to the former.
-Thirdly, as she did in the 1997 two-mode centrality paper, Faust clarifies the duality linking the row and column scores, noting that the scores for the people are (activity) weighted sums of the scores that the groups they join, and the scores for the groups are (group size) weighted sums of the scores of their members.
+Thirdly, as she did in the 1997 paper, Faust clarifies the duality linking the row and column scores, noting that the scores for the people are (activity) weighted sums of the scores that the groups they join, and the scores for the groups are (group size) weighted sums of the scores of their members.
 Finally, \citet[128ff]{faust2005using} clarifies the link between the CA scores---and different normalizations thereof---assigned to persons and events in an affiliation network on each dimension and (same node-set) inter-point distance in a joint space.
@@
 
-\subsection{Summary and Open Questions}
-As we can see from this brief and necessarily selective review, considerations of CA for two-mode network analysis have generally intertwined two basic themes.
-One beginning with \citet{bonacich1991simultaneous} and continuing through \citet{faust1997centrality, faust2005using} have commented on the formal commonalities and contrasts between ``centrality analysis'' and ``correspondence analysis,'' noticing striking similarities in the mathematical formulations of the traditional dual (eigenvector) score and the scores obtained from CA, as both come from the eigendecomposition of the affiliation matrix, resulting in scores that meet the strict criterion for duality, as the scores of nodes in one mode are usually additive functions of the scores of nodes in the other mode.
-The other stream, which began with characterizations of CA as a visualization tool, focuses on the substantive meaning of the (low-dimensional representations of) the distances defined over the multidimensional space returned by the CA procedure.
-Here, multiple interpretations have been offered, from the idea that CA returns clique-like structures \citep{bonacich1991simultaneous}, to intimations that the closeness in the CA dimensions implies a more structured similarity closer to the idea of structural equivalence (e.g., such that actors that attend the same events and events that the same actors attend will be closer in the space).
-This approach sanctions the idea that the main goal of CA is to provide a parsimonious representation of the original ``input data'' \citep{faust2005using}, where this is understood to mean the original affiliation matrix.
-Yet, more recently, this line of interpretation has been controverted by  \citep{desposito2014comparison} who have noticed that whatever the CA similarities are designed to do, they are a less-than-optimal representation of structural equivalence (and thus presumably the original input data), given the specific weighting by the degree of nodes in each mode that is built into their construction.
-
-In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral sciences was one of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation at roughly contemporaneous intervals throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper and starting in the next section, I show how a recent re-invention of CA for the analysis of ``economic complexity''---two-mode networks of countries/regions by the products they are most adept at producing---in the now well-established field of ``Network Science'' in the twenty-first century, allows us to revisit controversies and open questions as to the meaning of CA scores in two-mode network data, and the relationship between correspondence analysis, centrality analysis, and positional analysis (e.g., the discovery of structurally similar clusters of actors in two-mode networks).
-Along the way, we will be able to shed light on the link between CA scores and traditional eigenvector centrality scores for two-mode networks, as well as ascertain the exact kind of similarity that is learned by estimating CA (and traditional eigenvector centrality) scores for nodes in two-mode networks along multiple dimensions.
-
-\section{Reflective Node Scores in Two-Mode Networks} \label{sec:ref2mode}
-In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of assigning meaningful scores to nodes in a two-mode network using what they called at the time a ``reflective'' approach.
-Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks.
-Hence, the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
-As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
-This is consistent with the spirit of the original approach, recently described by \citet[92]{hidalgo2021economic} as centered on the ``\textit{duality} between economic inputs and outputs'' (italics mine).
-
-If we are going to assign substantively meaningful scores to nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}.
-A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $G$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.
-
-Given this, the degree-centrality of people is given by:
-
-\begin{equation}
-    C^R_p(1) = \sum_g a_{pg}
-    \label{eq:R1_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(1) = \sum_p a_{pg}
-  \label{eq:R1_g}
-\end{equation}
-
-That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $\mathbf{A}$, and the column sums of the same matrix give the first-order centrality of groups.
-
-As noted by \citet{hidalgo2009building}, the key to the reflective approach to computing the ``complexity'' of countries and products is the observation that, once we have these first-order quantities, it is possible to compute ``second-order'' quantities $C^R(2)$ for both people and groups using the \textit{averaged} first-order centralities of the entities in the other mode they are connected to.
-
-For people, these are given by:
-
-\begin{equation}
-    C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
-    \label{eq:R2_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
-    \label{eq:R2_g}
-\end{equation}
-
-While Equation~\ref{eq:R1_p} assigns a high value to people who belong to a lot of groups, Equation~\ref{eq:R2_p} assigns a high value to people who, on average, belong to large groups (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number).
-In the same way, while Equation~\ref{eq:R1_g} assigns a high value to groups that have lots of members, Equation~\ref{eq:R1_g} assigns a high value to groups that, on average, have members who themselves have lots of memberships (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
-
-Of course, we can keep on going and define ``third-order'' reflections; for the people, these are given by:
-
-\begin{equation}
-   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
-   \label{eq:R3_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
-   \label{eq:R3_g}
-\end{equation}
-
-As we saw, while for people, Equation~\ref{eq:R2_p} measured the average size of the groups they join, Equation~\ref{eq:R3_p} assigns a high value to people who, on average, belong to groups who are themselves attended by highly active members (e.g., whenever $a_{pg} = 1$ and $C^R_g(2)$ is a big number).
-In the same way, while Equation~\ref{eq:R2_g} assigns a high value to groups whose members have lots of memberships, Equation~\ref{eq:R3_g} assigns a high value to groups that, on average, have members who themselves (also on average) belong to large groups (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
-
-Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes).
-On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to).
-In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity).
-On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size).
-While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).
-
-More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases).
-
-For people, these are given by:
-
-\begin{equation}   
-    C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
-   \label{eq:Rq_p}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-   C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
-   \label{eq:Rq_g}
-\end{equation}
-
-Equation~\ref{eq:Rq_p} says that the reflective score of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
-Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective score is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-    \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
-    \caption{Southern Women data matrix with rows and columns ordered as in the original DGG table.}
-    \label{fig:sw}
-\end{figure}
-
-\subsection{Empirical Example} \label{subsec:ex}
-\subsubsection{Southern Women Data}
-The following analyses use the classic ``Southern Women'' data \citep{davis1941} (DGG) as the running example.
-This data is doubly appropriate for showcasing the relevance of methods that exploit the duality between two sets of entities, as it was used in Breiger's \citeyearpar{breiger1974duality} classic paper.
-Moreover, Southern Women is a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts.
-The list has only gotten much longer in the last two decades since Freeman's methodological ``meta-analysis'' \citep[e.g.,][among many others]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, kovacs2010generalized, brusco2011analysis, borgatti2014analyzing, everett2013dual, lerner2022dynamic, batagelj2022analysis, lizardo2024two}.
-Despite this long list of studies of the Southern Women data, we will see that novel insights can still be garnered from this old data source.
-
-Despite its undisputed pedigree as a methodological touchstone in studies of duality and dual constitution in two-mode networks, a key source of confusion in the mini-literature that exists around the Southern Women data concerns the order of presentation of the rows and columns.
-The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous.
-It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.
-As \citet{freeman2003finding} notes, in the canonical original table included in \citet{davis1941}, both the row and column modes had been re-ordered in a presentation to reveal what DGG ascertained (using intuitive methods and in-depth knowledge of the case) was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns.
-
-To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to ``W1'' or ``W11'' and ``E1'' or ``E13'' based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because ``E14'' in the original DGG data is not the last chronological even among the fourteen columns.
-In fact, as shown in Figure~\ref{fig:sw}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.\footnote{Using the DGG labeling, the chronologically ordered set of events are \{E11, E5, E2, E7, E12, E9, E3, E6, E10, E1, E14 E8, E4, E13\} \citep[p.
-68]{everett2018measuring}.} To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women---taken from \citet[83]{homans1950human}---from that of the original DGG table reproduced in \citet{freeman2003finding}, and ordered the events from left to right in the affiliation matrix \textit{chronologically}, which, as we have seen, also differs from the original DGG Table reproduced in \citet{freeman2003finding}.
-This means that the arbitrary ordinal labeling of the rows (e.g., ``W1,'' ``W2,'' ``W3,'' etc.) and the columns (e.g., ``E1,'' ``E2,'' ``E3,'' etc.) in these (and other) publications---see, e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; in all, different versions of the dataset reproduce different versions of the labeling.
-
-Of course, the specific labeling of the nodes does not really matter since we are usually interested in the structural properties of the rows and columns.
-Event ordering, however, matters for analyses that try to exploit the temporal properties of the data \citep[e.g.,][]{everett2018measuring, lerner2022dynamic}.
-Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since ``W1'' or ``E8'' in one publication \citep[e.g.,][]{freeman2003finding} may not correspond to ``W1'' or ``E8'' in the other \citep[e.g.,][]{doreian1979evolution}.
-To forestall confusion, in what follows, I use the numeric labeling corresponding to the original DGG data table as reproduced in (among others) \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{fig:sw}.
-My presentation includes both the arbitrarily ordered labels (e.g., ``W1'' or ``E1'') and both the woman's name and the event date in parentheses, with ``Myrna'' (W11) rebaptized to her given name of ``Myra'' to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.
-
-\subsubsection{Reflective Score Trajectory}
-Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective scores for persons, using a bump chart to plot the rank order trajectory of persons across reflections.
-The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.
-As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to.
-However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).
-
-In the exhaustive methodological meta-analysis mentioned earlier, \citet{freeman2003finding} noted that ``the consensus of the analytic procedures is to assign women 1 through 9 to one group and women 10 through 18 to the other.'' Notably, the reflective score rankings after freezing ($q \geq 18$) recover this same dominant partition reading from top to bottom.
-Figure~\ref{fig:g-refs} shows the corresponding bump chart for the events.
-Clearly, $E14$ experiences the most dramatic improvement in status as we move to higher reflections.
-Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering (the average of) the average number of memberships of its members (and so forth).
-Notably, the equilibrium reflective scores recover and ordering of groups columns that re-appears across many articles that have analyzed these data with events 1-6 separated from events 10-15 by events 7-9 \citep[see e.g.,][]{doreian2004generalized, kovacs2010generalized, lizardo2024two, everett2013dual}.
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
-            \caption{Reflective score trajectories of persons.}
-            \label{fig:p-refs}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-reflections.png}
-            \caption{Reflective score trajectories of groups.}
-            \label{fig:g-refs}
-    \end{subfigure}
-    \caption{Reflective scores for persons and groups (even reflections)}
-    \label{fig:refs}
-\end{figure}
-
-\subsection{CA and Dual Centrality Scoring in Two-Mode Networks} \label{sec:ca}
-The reader may ask what the point of going through all this reflective scoring is since \citet{bonacich1991simultaneous} already developed a dual approach to scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected.
-As we have already noted, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.
-
-While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although, as we will see, the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures).
-In fact, as I will show in a bit, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective scores discussed in the previous section \citep{mealy2019interpreting}.
-This makes the connection between the Bonacich and the CA scoring approaches more intimate than Bonacich thought.
-
-To begin to see the linkages between CA and the Bonacich approach, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:
-
-\begin{equation}
-    \mathbf{A}C^B_p = \lambda C^B_g 
-    \label{eq:bon1}
-\end{equation}
-
-\begin{equation}
-    \mathbf{A}^TC^B_g = \lambda C^B_p 
-    \label{eq:bon2}
-\end{equation}
-
-Where $\mathbf{A}$ is the original affiliation matrix.
-Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda_1$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $\mathbf{A}$.
-
-Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific scores; for people, these are given by:
-
-\begin{equation}
-    C^B_p = \frac{1}{\lambda_1}\sum_{g}a_{pg}C^B_g
-    \label{eq:faust1}
-\end{equation}
-
-And for groups:
-
-\begin{equation}
-    C^B_g = \frac{1}{\lambda_1}\sum_{p}a_{pg}C^B_p
-    \label{eq:faust2}
-\end{equation}
-
-Where $\lambda_1$ is the largest (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores.
-These equations make clear that in the Bonacich dual-scoring approach, the scores assigned to people are equal to the sum of the scores of the groups they join, and the scores assigned to groups are equivalent to the sum of the scores simultaneously assigned to their members.
-This approach thus nicely captures the duality property since the scores of nodes in each mode are given by aggregating their connections to nodes in the other mode.
-
-Note, however, the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}.
-The main difference is that in HH's reflective equations, a person's centrality is proportional to the \textit{group-degree-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{person-degree-weighted} centrality of the people who are members.
-We will return to this crucial point later.
-
-\subsection{Bonacich Reflections} \label{subsec:bonref}
-We can motivate the Bonacich eigenvector-based approach using the same ``reflective'' exposition we used to introduce the HH reflective approach in Section~\ref{sec:ref2mode}.
-Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centralities for two-mode networks; but it will help clarify the similarities and differences between the HH and the Bonacich approaches.
-
-Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:
-
-\begin{equation}
-   C^B_p(2) = \sum_g a_{pg}C^R_g(1)
-   \label{eq:BR2_p}
-\end{equation}
-
-\begin{equation}
-   C^B_g(2) = \sum_p a_{pg}C^R_p(1)
-   \label{eq:BR2_g}
-\end{equation}
-
-Equation~\ref{eq:BR2_p} says that people are central when the sum of the number of members of the groups they belong to is large.
-Equation~\ref{eq:BR2_g} says groups are central when the sum of the number of memberships of the people who belong to them is large.
-
-As before, we can keep on going and define a third-order reflection using the formulas:
-
-\begin{equation}
-   C^B_p(3) = \sum_g a_{pg}C^R_g(2)
-   \label{eq:BR3_p}
-\end{equation}
-
-\begin{equation}
-   C^B_g(3) = \sum_p a_{pg}C^R_p(2)
-   \label{eq:BR3_g}
-\end{equation}
-
-Equation~\ref{eq:BR3_p} says that people are central when the sum of the sum of the number of memberships held by the people who belong to the groups they belong to is large.
-Equation~\ref{eq:BR3_g} says that groups are central when the sum of the sum of the number of members of the groups their members belong to is large.
-Once again, we can keep going and define fourth order, fifth order, and higher reflections $C^R_p(4), C^R_p(4) \ldots C^R_p(q)$, where the centralities of nodes in one mode are based on the sums, of the sums, of the sums, of the centralities of nodes in the other mode.
-
-More generally, and in parallel with equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}, the reflective Bonacich centralities for persons and groups are given by:
-
-\begin{equation}
-   C^B_p(q) = \sum_g a_{pg}C^R_g(q-1)
-   \label{eq:BRq_p}
-\end{equation}
-
-\begin{equation}
-   C^B_g(q) = \sum_p a_{pg}C^R_p(q-1)
-   \label{eq:BRq_g}
-\end{equation}
-
-This says that the Bonacich reflection at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).
-Of course, these sums of sums would diverge to a larger and larger quantity at each step $q$.
-To prevent this and guarantee convergence, we can normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:
-
-\begin{equation}
-   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
-   \label{eq:BRq_pn}
-\end{equation}
-
-\begin{equation}
-   C^B_g(q) = \frac{C^B_g(q)}{||C^B_g(q)||_2}
-   \label{eq:BRq_gn}
-\end{equation}
-
-Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}.
-Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations.
-These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach.
-Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power method'' of \citet{mises1929praktische}).
-
-This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' scores and the Bonacich two-mode eigenvector centralities.
-In fact, both can be seen as instantiating an underlying ``prismatic'' model---in the sense of \citet{podolny2001networks}---of how centrality is distributed in the two-mode network.
-The Bonacich approach, based on sums of sums, favors overall \textit{volume}---or two-mode ``total effects'' in the sense of \citet{friedkin1991theoretical}---thus, groups receive more centrality points from their more central members and fewer centrality points from their less central members.
-Likewise, people receive more centrality points from their more central memberships and fewer centrality points from their less central memberships, and so on through every iteration.
-The HH reflective approach (and CA by implication) favors \textit{averages} rather than volume; thus, a person with a few memberships can contribute as much centrality to the group they belong to as a person with many memberships if the person with fewer memberships belongs to (on average) more central groups and the person with more memberships belongs to (on average) less central ones.
-In the same way, a person can receive as much centrality from a small group they belong to compared to a large group, as long as the small group (on average) has more central members in it, and the large group (on average) has less central members.
-
-\subsection{HH Reflections as Correspondence Analysis} \label{subsec:refeigen}
-Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem.
-As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective scores can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.
-
-To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}.
-Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:
-
-\begin{equation}
-    \left(\mathbf{AA}^T\right)C^B_p = \lambda C^B_p
-    \label{eq:bon3}
-\end{equation}
-
-\begin{equation}
-    \left(\mathbf{A}^T\mathbf{A}\right)C^B_g = \lambda C^B_g
-    \label{eq:bon4}
-\end{equation}
-
-This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode ``Breiger'' \citeyearpar{breiger1974duality} projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.
-
-We can motivate CA using the same dual projection strategy.
-Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\citep{borgatti1997network, bonacich1991simultaneous, faust1997centrality, faust2005using}.
-This gives the impression that CA is exclusively a ``direct'' rather than a dual projection method.
-Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection approach just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.
-
-To see this, consider the $|P| \times |P|$ matrix $\mathbf{D}p$, containing the ``first order'' reflective scores of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
-In the same way, consider the $|G| \times |G|$ matrix $\mathbf{D}g$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose.\footnote{Recall that a matrix is row stochastic if its rows sum to one} For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
-
-\begin{equation}
-    \mathbf{P}_{pg} = \mathbf{D}p^{-1}\mathbf{A}
-    \label{eq:ca1}
-\end{equation}
-
-We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the row-stochastic matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
-
-\begin{equation}
-    \mathbf{P}_{gp} = \mathbf{D}_G^{-1}\mathbf{A}^T
-    \label{eq:ca2}
-\end{equation}
-
-Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step---due to \citet{breiger1974duality}---of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
-
-For the people, the degree-weighted projection is:
-
-\begin{equation}
-\mathbf{P}_{pp} = \mathbf{P}_{pg}\mathbf{P}_{gp}
-    \label{eq:ca3}
-\end{equation}
-
-And for the groups:
-
-\begin{equation}
-    \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
-    \label{eq:ca4}
-\end{equation}
-
-Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
-Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), positive definite, but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
-Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row person (group) and, following any Markovian sequence of $person-group-person'-group'$ hops, will reach the column person (group), and where the probability of jumping from any person to a group they belong is given by the inverse of the number of groups they belong to (and similarly for groups) \citep[240]{deng2009generalized}.
-Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
-The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).
-Figure~\ref{fig:dwproj} shows the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices for the Southern Women Data.
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.55\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-norm.png}
-            \caption{People.}
-            \label{fig:p-dwproj}
-    \end{subfigure} \\
-     \begin{subfigure}[b]{0.55\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-norm.png}
-            \caption{Groups.}
-            \label{fig:g-dwproj}
-    \end{subfigure}
-    \caption{Degree-weighted projection matrices of the Southern Women data.}
-    \label{fig:dwproj}
-\end{figure}
-
-It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective scores can be obtained as a solution of the eigensystem:
-
-\begin{equation}
-    \mathbf{P}_{pp}C^R_p = \lambda_2C^R_p
-    %\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
-    \label{eq:dam1}
-\end{equation}
-
-\begin{equation}
-    \mathbf{P}_{gg}C^R_g = \lambda_2C^R_g
-    %\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
-    \label{eq:dam2}
-\end{equation}
-
-With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices, which are the same as the CA scores and the limit HH reflective scores \citep{van2021correspondence}.\footnote{As we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
-Note also that the degree normalized projection matrices have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix, with values between zero and one \citep{van2021correspondence}.} These scores will also be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups, respectively) obtained from a simple CA of the original affiliation matrix via SVD \citep[398, eq.
-9.17]{fouss2016algorithms}.
-This exact correspondence can be verified in Figure~\ref{fig:ca-ref-corr}, which shows a scatterplot and associated linear regression line of the standardized scores computed via the iterative method of reflections ($q = 26$) on the y-axis against those obtained from the second eigenvector of the degree-weighted projection matrices for persons and groups on the x-axis ($r = 1.0$).\footnote{Formally, the eigendecomposition of the degree-weighted projection (for people) is given by the solution to the equation $\mathbf{P}_{pp} = \mathbf{UVU}^{-1}$, where $\mathbf{U}$ is a $|P| \times |P|$ matrix whose columns are the eigenvectors of $\mathbf{P}_{pp}$ and $\mathbf{V}$ is a $|P| \times |P|$ matrix whose diagonals are the eigenvalues of $\mathbf{P}_{pp}$ (with zeros in every other cell).
-Columns two through $|P|$ of matrix $\mathbf{U}$ contain the standard CA scores for the people along the complete set of $|P| - 1$ dimensions; the same goes for $\mathbf{P}_{gg}$ in the case of groups.}
-
-Because $\mathbf{P}_{pp} = \mathbf{D}p^{-1}\mathbf{A}\mathbf{D}g^{-1}\mathbf{A}^T$ and $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group scores as eigenvectors of the one-mode projection of the original affiliation matrix: $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups.
-The key difference is that in CA, we pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
-
-This can be clearly seen if we express $\mathbf{P}_{PP}= \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:
-
-\begin{equation}
-    p_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
-    \label{eq:mealy1}
-\end{equation}
-
-In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$.
-Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}.
-As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups).
-The HH reflective scores, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.
-
-The same reasoning applies to groups in the matrix $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$, whose entries are given by:
-
-\begin{equation}
-    p_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
-    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
-    \label{eq:mealy2}
-\end{equation}
-
-Now, as we saw earlier \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}; accordingly, the first non-trivial CA dimension for persons and groups given by the right and left eigenvectors corresponding to the second-largest eigenvalue from the SVD of the matrix $\mathbf{D}_p^{-1/2}\mathbf{A}\mathbf{D}_g^{-1/2}$ \citep[126]{faust2005using}.} In fact, as \citet{mealy2019interpreting} have noted, the method of ``reflections'' is just a re-discovery of the older idea of computing CA scores via reciprocal averaging \citep{hill1973reciprocal}.
-
-\begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-ca-ref-corr.png}
-            \caption{}
-            \label{fig:p-ca-ref}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-ca-ref-corr.png}
-            \caption{}
-            \label{fig:g-ca-ref}
-    \end{subfigure}
-    \caption{Scatter plot of the scores computed by the method of reflections ($q = 26$) on the y-axis and the first dimension of the Correspondence Analysis of the two-mode affiliation matrix on the x-axis, for both persons (a) and groups (b).
-}
-    \label{fig:ca-ref-corr}
-\end{figure}
-
-As \citet{desposito2014comparison} have recently noted, degree-pre-weighting does alter the original input data, which means that CA is \textit{not} meant to reproduce a low dimensional approximation of the original affiliation matrix.
-Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}.
-This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large, thus adjusting similarity by group size so that co-memberships in groups everyone belongs to counts for less \citep{ragozini2014correspondence}.
-In the same way, on the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/ca-reord1.png}
-            \caption{First CA Dimension}
-            \label{fig:ca-reord1}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
-            \caption{First Bonacich Dimension}
-            \label{fig:bon-reord1}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
-            \caption{Second Bonacich Dimension}
-            \label{fig:bon-reord2}
-    \end{subfigure}
-     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/SR-reord1.png}
-            \caption{Second SimRank Eigenvector}
-            \label{fig:sr-reord1}
-    \end{subfigure}
-    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) rows and columns re-ordered according to the person and group score on the first CA dimension, (b) rows and columns re-ordered according to person and group Eigenvector centrality score, (c) rows and columns re-ordered according to person and group score on the second Bonacich dimension, (d) rows and columns re-ordered according to the second eigenvector of the SimRank similarity matrix.}
-    \label{fig:ca-v-bon}
-\end{figure}
-
-\section{Comparing Bonacich and CA Dual Scoring in the Southern Women Data} \label{sec:comparing}
-\subsection{Row and Column Re-ordered Affiliation Matrices} 
-Figures~\ref{fig:ca-reord1} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$.
-Each panel shows the original Southern Women affiliation matrix but with rows and columns blocked according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic}, using scores from CA and Bonacich scoring to re-order the original matrix, as they did in that paper for one-mode networks.} In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero.
-As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure \citep[34]{wasserman1990correspondence}, and the $C^B$ re-ordered affiliation matrix having a triangular structure.
-
-Accordingly, the $C^R$ reflective scores reveal a dual-block partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).
-The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic \textit{core-periphery} partitioning \citep{borgatti2000models}, with a group of highly active women (on the top half) who selectively co-participate in highly attended events (on the left-hand side) and less active women (on the bottom half) who go to less well-attended events (on the right-hand side).
-In fact, as \citet[p.
-206]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic \citet{breiger1974duality} projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods.
-In fact, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is substantively identical to that shown in~\citet[206, Table 2]{everett2013dual}, who previously extracted a core-periphery partitioning of the same data.
-
-\begin{figure}[ht!]
-    \centering
-        \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}            
-        \label{fig:eigvec}
-    \caption{CA Eigenvector plot for persons (a) and groups (b).
-The first dimension CA score is on the x-axis, and the rank order of persons and groups on the same score is on the y-axis.
-The inset plot (c) shows the eigenvalue orderings on the x-axis ($k$) for persons and groups, showing a separation between the first three eigenvalues and the rest.}
-    \label{fig:ca-eigvec}
-\end{figure}
-
-\subsection{Eigenvector Plot} \label{subsec:eigplot}
-We can confirm that the first CA axis points toward the underlying similarity-based partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis.
-If a two-mode network has no discernible block structure, the first dimension CA scores would be distributed as a continuous logistic curve; when block structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}.
-The Southern Women Data clearly belong in the second category.
-In the people mode, we have \{Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\} on one side, \{Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\} on the other.
-Among events, we have \{E1, E2, E3, E4, E5, E6\} on side \{E9, E10, E11, E12, E13, E14\} on the other and \{E7, E8\} in an ambiguous middle position.
-Note that this is the canonical consensus partitioning (for the people) highlighted by \citet{freeman2003finding} across twenty-one different methods and subsequently reproduced (for both persons and groups) by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity (GRS) approach.
-This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar; we will return to this point in Section~\ref{sec:cagensim}.
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-        \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
-    \caption{The top panels show the degree-weighted one-mode projection heatmap plots for (a) persons and (b) groups in the Southern Women Data with rows and columns re-ordered according to the first CA dimension.
-Darker red cells correspond to high entry values (normalized by the maximum) in the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices.
-The bottom panels show the correspondence plot for persons and groups with scores corresponding to (c) the first dimension on the horizontal axis with the second dimension on the vertical axis, and (d) the first dimension on the horizontal axis with the third dimension on the vertical axis.}
-    \label{fig:ca}
-\end{figure}
-
-\subsection{Degree-Weighted Projections and CA Correspondence Plot} \label{subsec:caplot}
-Panels (a) and (b) of Figure~\ref{fig:ca} shows a heatmap plot of the same weighted projection matrices from Figures~\ref{fig:p-dwproj} and~\ref{fig:g-dwproj}, but this time with the rows and columns re-ordered according to the scores corresponding to the first CA dimension.
-In the plot, darker red indicates greater similarity between pairs of persons and groups, and gray indicates less similarity.
-The bottom panels of the same figure show the usual correspondence plot of the first three CA dimensions, with the first dimension on the x-axis against the second dimension on the y-axis in panel (c) and the first dimension plotted against the third dimension on the y-axis in panel (d).
-
-This figure illustrates several important points.
-First, as the heatmap plots in panels (a) and (b) show, distances along the first CA axis correspond to degree-weighted similarities between persons and groups recorded in the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices.
-Pairs of people and groups with the largest entries in the matrix have minimal differences in the first CA dimension score, appearing close to one another when we reshuffle the rows and columns of the weighted projection matrix.
-This also means that the distances between pairs of people and groups in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the respective \textit{degree-normalized one-mode projections}.
-Pairs of people with large values in the projection matrix $P_{pp}$---such as \{Olivia, Flora\},  \{Katherine, Sylvia, Nora\}, \{Dorothy, Verne\}, and \{Brenda, Laura, Evelyn\}--- appear closer in the correspondence plot.
-The same goes for pairs of groups; those with large values in $P_{gg}$ matrix---such as \{E13, E14\} and \{E1, E2, E3, E4\}---appear closer in the correspondence plot, while those with small weighted similarity values are placed far apart.
-Distances between same-mode entities in the correspondence plot are thus a function of their (inverse) degree-normalized similarity.
-
-Note that this differs from the usual interpretation of the CA correspondence plot, which is typically taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{desposito2014comparison}, or ``first order'' similarity \citep{kovacs2010generalized}.
-This interpretation implies (for instance) that two people who attend many of the same events or groups with many members in common will appear close in the plot.
-But this (still common) interpretation is off the mark.
-What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.
-This type of degree-normalization can be the preferred approach in some applications since people with many memberships and groups with many members will appear to be similar to many other people and groups if we go with the raw number of other-mode objects shared.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs.
-2.13 and 2.14]{fouss2016algorithms}.
-As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} This implies that the similarities between pairs of people and groups will be weighted by the degrees of the other-mode objects that are shared between them.
-Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups.
-In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships \citep{desposito2014comparison}.
-
-Another thing to note is that since the CA row and column scores in each dimension are obtained from the eigendecomposition and subsequent low-rank approximation of the degree-normalized similarity matrices $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$, each dimension provides independent information on the similarity clustering of persons and groups.\footnote{We know from inset plot (c) of Figure~\ref{fig:ca-eigvec} that the eigenvalues corresponding to the first three dimensions exhibit strong separation from the rest, suggesting that these provide a good low-rank approximation of the original degree-weighted projection matrix for persons and groups.}  Thus, as shown in panels (c) and (d) of Figure~\ref{fig:ca}, as we have already seen, the first horizontal dimension recovers the main partition separating the two dominant clusters of persons and groups~\citep{freeman2003finding}.
-The second and third dimensions, by contrast, recover secondary and tertiary similarity clusters of people and groups independent of the main two-block partition.
-Thus, the second dimension (the y-axis in panel (c) of Figure~\ref{fig:ca}) separates \{Flora, Nora\} and \{E9, E11\} from the rest of persons and groups, consistent with the strong similarity between these nodes in the original weighted projection matrix.
-We know from the original affiliation matrix in Figure~\ref{fig:sw} that \{Flora, Nora\} are maximally similar---they are structurally equivalent---and \textit{only} attend events \{E9, E11\}, which makes these events similar in the second dimension given they are attended by the two people who are also most similar in the same dimension (this also accounts for the high similarity between the two events shown in the heatmap plot of the degree-weighted projection in Figure~\ref{fig:ca}(b)).
-
-The third CA dimension, shown in the y-axis in panel (d) of Figure~\ref{fig:ca} against the first dimension separation on the x-axis, by contrast, separates \{Eleanor, Pearl, Ruth, Dorothy, Verne, Myra\} and \{E8, E9\} from the rest.
-We know from Figure~\ref{fig:bon-reord1} that event E8, along with E7, is one of the most popularly attended events and thus contributes little to differentiate between the main partition of people and events on the first correspondence plot in panel (c), hence its location near the plot's origin.
-The pattern of attendance of these six women is one characterized by affiliation with these popular events combined with sparser attendance at the events that most serve to differentiate the two primary clusters.
-In his original paper, \citet[164]{bonacich1991simultaneous} noted that this third dimension ``corresponded closely to the centrality analysis pattern,'' an observation likely driven by the affinity between the actors who load highly on this dimension and two of the most central events by raw popularity, \{E8, E9\}.
-What we can see instead is that the third CA dimension separates actors that are most ``peripheral'' to the two main clusters according to \citet{freeman2003finding}; namely, \{Eleanor, Pearl Ruth\} on one side and \{Dorothy, Verne, Myra\} on the other, from core members of the two main blocks, who are placed in the bottom of the correspondence plot in panel (c)---low scores in the third dimension---and toward the extreme left and right ends of the plot, indicating very high absolute value scores in the first dimension.
-
-\subsection{Breiger Projections and the Bonacich Correspondence Plot} \label{sec:eigplot}
-Earlier, we discussed how CA is sometimes considered a technique useful to generate a correspondence plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$.
-But as we have seen, this is \textit{not} what CA is designed to do.
-Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes \citep{desposito2014comparison}.
-In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix---or a Markov transition matrix under a different interpretation---between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.
-
-Here, once again, the contrast with the traditional dual Bonacich scores can prove instructive.
-Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$)---the standard \citet{breiger1974duality} projections---with the rows and columns re-ordered according by the first eigenvector of the matrix, which is equivalent to the usual Bonacich eigenvector dual centrality score for two-mode networks (see equations~\ref{eq:bon3} and~\ref{eq:bon4}).
-Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).
-Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix.
-
-This suggests that the eigenvector centralities do encode a kind of similarity between pairs of people and pairs of groups; it just happens to be based on the unweighted similarity scores given by the raw number of shared memberships (for people) and shared members (for groups).
-In this respect, the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of memberships will end up being ``similar`` to more people on the standard projection matrix by the simple fact of having more memberships (and the same for large groups with respect to members).
-Persons and groups high in eigenvector centrality can be thought of as being the most similar to most others, condensing the information in the Breiger projection matrix into a single ordinal rank.
-Also, just like with CA and the degree-weighted projections, higher-order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode forms of similarity between pairs of people and pairs of groups, net of the main axis dominated by overall person activity and group size.
-
-\begin{figure}[ht!]
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-        \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
-            \caption{Bonacich correspondence plot.}
-            \label{fig:bon-corr}
-    \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel.
-The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis, points are colored according a k-means three-cluster solution.}
-    \label{fig:bon-sim}
-\end{figure}
-
-While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice \citep[see][for discussion in the one-mode network case]{iacobucci2017eigenvector} the exact meaning of the mysterious second eigenvector remains unclear in the two-mode case.
-We can use the same ``a posteriori blockmodeling'' approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue \citep{wasserman1987stochastic}.
-Figure~\ref{fig:bon-reord2} shows the affiliation matrix blocked according to the \textit{second} Bonacich dimension, and Figure~\ref{fig:bon-sim} (c) shows the ``Bonacich Correspondence Plot'' with the usual eigenvector centrality score on the horizontal x-axis plotted against the second eigenvector of the Breiger projection matrix for both persons and groups on the vertical y-axis.
-
-We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension diplays a \textit{dual triangular} structure, revealing two secondary core and peripheries not evident in the blocking according to either the first CA or standard eigenvector scores.
-Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering a version of the projection matrices of the Southern Women data where the proximities were designed to detect \textit{structural equivalence}.\footnote{In that application, the cell entry in each projection matrix was the \textit{proportion of matches} between each pair of people and groups.
-As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is Myrna/Myra's position, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
-This may be partly due to the discrepant position of ``core event'' E9 along the columns of the blocked matrix which is grouped with the second group of ``peripheral'' events \{E10, E11, E12, E13, E14\} in Figure~\ref{fig:bon-reord2}.
-
-This analysis points to a new way to interpret the second Bonacich dimension, namely, as a way to reveal secondary core-periphery partitions net of a main core-periphery partition in the two-mode network (if any exists).
-Furthermore, the similarities between persons and groups defined over the two-dimensional Bonacich space, shown in Figure~\ref{fig:bon-sim} (c), can thus be used to reveal \textit{structurally equivalent} clusters of persons and events similar to those obtained by \citet[Table 5]{everett2013dual}.
-This is confirmed by the coloring of the persons and groups in Figure~\ref{fig:bon-reord2}) which comes from a k-means cluster analysis using the first two Bonacich dimensions to define initial cluster centroids, revealing a three-cluster solution almost equivalent to that obtained by Everett and Borgatti, with the exception of Myrna/Myra's and event E9's assingment as discussed earlier.
-This suggests that the Bonacich approach to similarity clustering is distinct from CA in being useful for revealing structural equivalence, thus making the Bonacich dimensions a more faithful basis for a low-rank reconstruction of the original ``input data'' than CA if that is the goal of the analysis \citep{desposito2014comparison}.
-
-\section{Correspondence Analysis and Generalized Relational Similarity} \label{sec:cagensim}
-As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy.
-Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}, where objects' similarities are given by their pattern of connections to objects in the other mode.
-This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence \citep{everett1994regular}.
-In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.\footnote{\citet{lizardo2024two} shows the connection between Kovacs's idea of generalized relations similarity and the two-mode network projection.}
-
-An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}.
-In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong to, and the similarity of groups is a function of the people who belong to them, making the similarity of persons and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}.
-Thus,
-
-\begin{itemize}
-    \item People are \textit{similar} if they belong to \textit{similar} groups.
-    \item Groups are \textit{similar} if they share \textit{similar} members.
-\end{itemize}
-
-This definition consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
-To accomplish this, \citet[540, eq.
-2 and eq.
-3]{jeh2002simrank} propose that we define the similarity of each pair of people $S(p, p')$  as given by:
-
-\begin{equation}
-    S(p, p') = \frac{\alpha}{C^R(1)_pC^R(1)_{p'}}
-    \sum_{i = 1}^{C^R(1)_p} \sum_{j = 1}^{C^R(1)_{p'}} 
-    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
-    \label{eq:simrank1}
-\end{equation}
-
-Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to, and $\alpha$ is a free parameter obeying the restriction: $0 > \alpha < 1$.
-By construction, $S(p, p) = 1$, for all $p$.
-Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of each unordered pair of groups they both belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.
-
-Likewise, for groups, the SimRank similarities are given by:
-
-\begin{equation}
-    S(g, g') = \frac{\alpha}{C^R(1)_gC^R(1)_{g'}}
-    \sum_{i = 1}^{C^R(1)_g} \sum_{j = 1}^{C^R(1)_{g'}} 
-    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
-    \label{eq:simrank2}
-\end{equation}
-
-Where $p(i)_{i \in N(g)}$ is the $i^{th}$ person in the set of members of group $g$, and $p(j)_{j \in N(g')}$ is the $j^{th}$ person in the set of members of group $g'$, and $S(g, g) = 1$.
-Thus, the SimRank similarity between two groups is a function of the sum of the similarities of each pair of people who belong to both groups, weighted by the reciprocal of the products of the number of members of each group multiplied by $\alpha$.
-
-\begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
-    \centering
-        \includegraphics[width=0.9\textwidth]{Plots/simrank-v-ca.png}
-            \label{fig:sim-v-ca}
-    \caption{Simrank versus CA comparison.
-Panels (a) and (c) show the final SimRank similarity matrices for persons and groups (respectively) with rows and columns ordered according to the first CA dimension scores.
-Panel (b) shows the Pearson correlation between the scores corresponding to the second eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis) for persons, panel (d) shows the same correlation for groups.}
-    \label{fig:simrank}
-\end{figure}
-
-In two-mode networks, SimRank scores for each pair of nodes across the two sets can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
-Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
-We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}, which is confirmed here for the Southern Women data.
-
-Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
-The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
-Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections \citep{mealy2019interpreting, van2021correspondence}.
-Beyond that, the SimRank approach can be thought of as yet another way of obtaining one-mode projections from an original two-mode network, in line with Everett and Borgatti's ``dual projection'' approach \citep{everett2013dual}.
-
-Figure~\ref{fig:sr-reord1} shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices---with $\alpha$ set to $0.8$---for the row and column nodes.
-Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling displays a block-diagonal structure separating the two communities of persons and groups as that revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
-Figure~\ref{fig:simrank-v-ca}(a) and Figure~\ref{fig:simrank-v-ca}(b) show an eigenvector plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
-Looking at the plots from left to right, we can see that the partitioning of the node sets on the most informative eigenvector of the SimRank similarity matrix is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
-Figures~\ref{fig:simrank-v-ca}(c) and (d) show a regression plot of the first CA dimension scores for persons and groups (respectively) on the x-axis and the SimRank scores extracted from the main non-trivial eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
-As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that these first eigenvectors capture the same (GSR) information across the two approaches.
-
-\section{Discussion and Concluding Remarks} \label{sec:disc}
-This paper reconsiders the role of CA in the analysis of two-mode network data.
-We began with an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to simple CA of the affiliation matrix.
-Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}.
-I showed that just like the reflective scores that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums.
-This exercise clarifies the links between CA and the dual two-mode eigenvector centrality more coherently and systematically, an issue that was left open and somewhat unclear in Bonacich's \citeyearpar{bonacich1991simultaneous} classic paper.
-Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.
-
-But CA does more than reveal a latent dual ordering of nodes in the two-mode network.
-In addition to doing this, CA can be shown to reveal \textit{groupings} of nodes based on some conception of the similarity of their connections to nodes in the other mode.
-This is evident once we use the scores of the first CA dimension to re-order the rows and columns of the original affiliation matrix.
-When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data.
-This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events \citep{everett2013dual}.
-
-This is another way in which CA and the Bonacich dual centrality approach systematically differ; one is geared to uncovering blocks of actors with similar connections to events in the other mode (and vice versa), while the other reveals blocks of actors who are the most active and who attend the largest events.
-Both are, of course, legitimate ways of analyzing the structure of a two-mode network.
-Still, they differ in terms of the structural patterns that they are sensitive to, with the CA approach closer to a community partition where nodes that are surprisingly similar end up in the same clusters---where ``surprising'' means similarity based on deviations from a suitable null model based on random mixing given their first-order degrees.
-
-But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to?
-Here, we show that CA recovers a type of generalized relational similarity (GRS).
-That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar.
-Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, we saw that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach.
-Thus, we can clarify that two-mode network CA reveals latent groupings of nodes based on a GRS criterion.
-
-Overall, the preceding has shown that CA can be upgraded from a method designed to generate joint plots and visualization of two-mode data to one that can be seen as more ``central'' to the usual social-network-analytic tasks, like ordering the nodes in the two sets according to some substantively meaningful rank---centrality analysis---or finding sets of similar actors in the network (subgroup or community detection).
-At the very least, it seems like the Bonacich style dual centrality based on the eigenvector decomposition of the raw affiliation matrix should not be the default ``reflective'' centrality approach unless the analyst has the explicit analytic goal of exploring center-periphery partitioning in the network.
-
-A better approach, similar to the one exemplified here, would be to present a comparison of the reflective scores obtained by CA and Bonacich side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering.
-Of course, suppose the analyst is more interested in such a ``subgroup'' analytic approach.
-In that case, the Bonacich dual centrality approach is less helpful (even if multiple dimensions are considered), and CA should be the first line of attack.
-
 %% Loading bibliography style file]
@@
 % Loading bibliography database
-\bibliography{ca.bib}
+\bibliography{ca}
 \end{document}

### 2025-03-30 main-R1.tex
Updates from Overleaf

@@
 Yet, more recently, this line of interpretation has been controverted by  \citep{desposito2014comparison} who have noticed that whatever the CA similarities are designed to do, they are a less-than-optimal representation of structural equivalence (and thus presumably the original input data), given the specific weighting by the degree of nodes in each mode that is built into their construction.
+As such, the exact relationship that the CA representation bears to the original affiliation matrix data, as well as the similarities and contrasts between correspondence analysis, centrality analysis, and both to traditional positional analysis in social networks, remains to be specified and clarified.
 
@@
 Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks.
-Hence, the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
+Hence, they referred to their approach as a way to extract the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix.
 As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.
@@
 
-Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes).
-On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to).
-In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity).
-On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size).
-While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).
+Note that for people, the even-numbered reflection $C^R_p(2)$ assigns scores based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes).
+On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns scores based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to).
+In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns scores based on a formal feature of the \textit{people} who belong to them (in this case, their activity).
+On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns scores based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size).
+While these are distinct metrics in principle, in practice, the ordering of the nodes in each mode ends up being identical across even and odd scores after their rank ordering ``freezes'' past a given number of iterations (proportional to the network size).\footnote{This means that at least in the initial iterations, there is a formal resemblance between the reflection scores on the odd iterations and Bonacich's dual centrality idea, namely, that central people are those who belong to central groups, where the centrality of groups is defined by the centrality of people who belong to them (and vice versa).
+And, indeed, we find that the absolute value of the Pearson product-moment correlation between the odd reflection scores and the Bonacich eigenvector centrality is fairly high for both persons ($r = 0.80$) in the third iteration but then settles to a lower value after the equilibrium is reached past the 20th iteration for the Southern Women data ($r = 0.54$).
+The even iterations for persons, on the other hand, begin with a relatively low correlation with the Bonacich centrality at the second iteration ($r = 0.30$) but then settle into the same equilibrium correlation with the Bonacich centrality as the odd iterations after the 20th reflection.
+I thank an anonymous reviewer for pointing out this pattern.}
 
@@
 
-Equation~\ref{eq:Rq_p} says that the reflective score of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
-Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective score is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.
+Equation~\ref{eq:Rq_p} says that the reflective score of a person \textit{p} at iteration \textit{q} is the sum of the reflective scores the groups they belong to at the $q-1$ iteration ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$.
+Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective score is the sum of the reflective scores of their members at the $q-1$ iteration $C^R_p(q-1)$, divided by the number of group members $C^R_g(1)$.
 
@@
     \centering
-    \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
-    \caption{Southern Women data matrix with rows and columns ordered as in the original DGG table.}
+    \includegraphics[width=0.65\textwidth]{southern-women.jpg}
+    \caption{Southern Women affiliation data matrix with rows and columns ordered as in the original DGG table.}
     \label{fig:sw}
@@
 Despite its undisputed pedigree as a methodological touchstone in studies of duality and dual constitution in two-mode networks, a key source of confusion in the mini-literature that exists around the Southern Women data concerns the order of presentation of the rows and columns.
-The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous.
+The reason for this is that while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous.
 It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.
@@
 Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering (the average of) the average number of memberships of its members (and so forth).
-Notably, the equilibrium reflective scores recover and ordering of groups columns that re-appears across many articles that have analyzed these data with events 1-6 separated from events 10-15 by events 7-9 \citep[see e.g.,][]{doreian2004generalized, kovacs2010generalized, lizardo2024two, everett2013dual}.
+Notably, the equilibrium reflective scores recover and the ordering of groups columns that re-appears across many articles that have analyzed these data with events 1-6 separated from events 10-15 by events 7-9 \citep[see e.g.,][]{doreian2004generalized, kovacs2010generalized, lizardo2024two, everett2013dual}.
 
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
+        \includegraphics[width=1.0\textwidth]{p-reflections.png}
             \caption{Reflective score trajectories of persons.}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-reflections.png}
+        \includegraphics[width=1.0\textwidth]{g-reflections.png}
             \caption{Reflective score trajectories of groups.}
@@
 \begin{equation}
-   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
+   C^B_p(q)' = \frac{C^B_p(q)}{||C^B_p(q)||_2}
    \label{eq:BRq_pn}
@@
 \begin{equation}
-   C^B_g(q) = \frac{C^B_g(q)}{||C^B_g(q)||_2}
+   C^B_g(q)' = \frac{C^B_g(q)}{||C^B_g(q)||_2}
    \label{eq:BRq_gn}
@@
 
-Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}.
+Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected, rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}.
 Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations.
@@
 \begin{equation}
-    \left(\mathbf{AA}^T\right)C^B_p = \lambda C^B_p
+    \left(\mathbf{AA}^T\right)C^B_p = \lambda^2 C^B_p
     \label{eq:bon3}
@@
 \begin{equation}
-    \left(\mathbf{A}^T\mathbf{A}\right)C^B_g = \lambda C^B_g
+    \left(\mathbf{A}^T\mathbf{A}\right)C^B_g = \lambda^2 C^B_g
     \label{eq:bon4}
@@
 
-This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode ``Breiger'' \citeyearpar{breiger1974duality} projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.
+This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode ``Breiger'' \citeyearpar{breiger1974duality} projection matrices corresponding to the first (largest) eigenvalue (which is equivalent to the singular value $\lambda$ from equations~\ref{eq:bon1} and~\ref{eq:bon2} squared).
+This linkage qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.
 
@@
 
-To see this, consider the $|P| \times |P|$ matrix $\mathbf{D}p$, containing the ``first order'' reflective scores of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
-In the same way, consider the $|G| \times |G|$ matrix $\mathbf{D}g$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
-Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose.\footnote{Recall that a matrix is row stochastic if its rows sum to one} For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
-
-\begin{equation}
-    \mathbf{P}_{pg} = \mathbf{D}p^{-1}\mathbf{A}
-    \label{eq:ca1}
-\end{equation}
-
-We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the row-stochastic matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
-
-\begin{equation}
-    \mathbf{P}_{gp} = \mathbf{D}_G^{-1}\mathbf{A}^T
-    \label{eq:ca2}
-\end{equation}
-
-Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step---due to \citet{breiger1974duality}---of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
-
-For the people, the degree-weighted projection is:
-
-\begin{equation}
-\mathbf{P}_{pp} = \mathbf{P}_{pg}\mathbf{P}_{gp}
-    \label{eq:ca3}
-\end{equation}
-
-And for the groups:
-
-\begin{equation}
-    \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
-    \label{eq:ca4}
-\end{equation}
-
-Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
-Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), positive definite, but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
-Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row person (group) and, following any Markovian sequence of $person-group-person'-group'$ hops, will reach the column person (group), and where the probability of jumping from any person to a group they belong is given by the inverse of the number of groups they belong to (and similarly for groups) \citep[240]{deng2009generalized}.
-Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
-The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).
-Figure~\ref{fig:dwproj} shows the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices for the Southern Women Data.
-
-\begin{figure}[ht!]
+\begin{figure}
     \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-     \begin{subfigure}[b]{0.55\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-norm.png}
+     \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=1.0\textwidth]{p-norm.png}
             \caption{People.}
@@
     \end{subfigure} \\
-     \begin{subfigure}[b]{0.55\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-norm.png}
+     \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=1.0\textwidth]{g-norm.png}
             \caption{Groups.}
@@
 
+To see this, consider the $|P| \times |P|$ matrix $\mathbf{D}p$, containing the ``first order'' reflective scores of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell.
+In the same way, consider the $|G| \times |G|$ matrix $\mathbf{D}g$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell.
+Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose.\footnote{Recall that a matrix is row stochastic if its rows sum to one} For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:
+
+\begin{equation}
+    \mathbf{P}_{pg} = \mathbf{D}p^{-1}\mathbf{A}
+    \label{eq:ca1}
+\end{equation}
+
+We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:
+
+\begin{equation}
+    \mathbf{P}_{gp} = \mathbf{D}_G^{-1}\mathbf{A}^T
+    \label{eq:ca2}
+\end{equation}
+
+Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step---due to \citet{breiger1974duality}---of calculating the degree-weighted \textit{projections} for people and groups from these matrices.
+
+For the people, the degree-weighted projection is:
+
+\begin{equation}
+\mathbf{P}_{pp} = \mathbf{P}_{pg}\mathbf{P}_{gp}
+    \label{eq:ca3}
+\end{equation}
+
+And for the groups:
+
+\begin{equation}
+    \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
+    \label{eq:ca4}
+\end{equation}
+
+Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups.
+Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), positive definite, but not symmetric, as can be seen in Figure~\ref{fig:dwproj}.
+Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row person (group) and, following any Markovian sequence of $person-group-person'-group'$ hops, will reach the column person (group), and where the probability of jumping from any person to a group they belong is given by the inverse of the number of groups they belong to (and similarly for groups) \citep[240]{deng2009generalized}.
+Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix).
+The reason for the asymmetry is that, depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).
+Figure~\ref{fig:dwproj} shows the $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ matrices for the Southern Women Data.
+
 It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective scores can be obtained as a solution of the eigensystem:
@@
     \mathbf{P}_{pp}C^R_p = \lambda_2C^R_p
-    %\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
     \label{eq:dam1}
@@
     \mathbf{P}_{gg}C^R_g = \lambda_2C^R_g
-    %\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
     \label{eq:dam2}
@@
 
-With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices, which are the same as the CA scores and the limit HH reflective scores \citep{van2021correspondence}.\footnote{As we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.
-Note also that the degree normalized projection matrices have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix, with values between zero and one \citep{van2021correspondence}.} These scores will also be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups, respectively) obtained from a simple CA of the original affiliation matrix via SVD \citep[398, eq.
+With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices, which are the same as the CA scores and the limit HH reflective scores \citep{van2021correspondence}.\footnote{As we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.} These scores will also be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups, respectively) obtained from a simple CA of the original affiliation matrix via SVD \citep[398, eq.
 9.17]{fouss2016algorithms}.
 This exact correspondence can be verified in Figure~\ref{fig:ca-ref-corr}, which shows a scatterplot and associated linear regression line of the standardized scores computed via the iterative method of reflections ($q = 26$) on the y-axis against those obtained from the second eigenvector of the degree-weighted projection matrices for persons and groups on the x-axis ($r = 1.0$).\footnote{Formally, the eigendecomposition of the degree-weighted projection (for people) is given by the solution to the equation $\mathbf{P}_{pp} = \mathbf{UVU}^{-1}$, where $\mathbf{U}$ is a $|P| \times |P|$ matrix whose columns are the eigenvectors of $\mathbf{P}_{pp}$ and $\mathbf{V}$ is a $|P| \times |P|$ matrix whose diagonals are the eigenvalues of $\mathbf{P}_{pp}$ (with zeros in every other cell).
-Columns two through $|P|$ of matrix $\mathbf{U}$ contain the standard CA scores for the people along the complete set of $|P| - 1$ dimensions; the same goes for $\mathbf{P}_{gg}$ in the case of groups.}
+Columns two through $|G| - 1$ of matrix $\mathbf{U}$ contain the standard CA scores for the people along the complete set of $|G| - 1$ dimensions; the same goes for $\mathbf{P}_{gg}$ in the case of groups.
+Note that both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ have the same set of non-zero eigenvalues, with cardinality equal to the rank of the original affiliation matrix $\mathbf{A}$ minus one or $|G| - 1$; thus only the first $|G| - 2$ eigenvectors of both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are informative in the Southern Women data, since $|G| < |P|$.}
 
 Because $\mathbf{P}_{pp} = \mathbf{D}p^{-1}\mathbf{A}\mathbf{D}g^{-1}\mathbf{A}^T$ and $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}.
-Both extract individual and group scores as eigenvectors of the one-mode projection of the original affiliation matrix: $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups.
+Both extract individual and group scores as eigenvectors of the one-mode projection of the original affiliation matrix, $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups.
 The key difference is that in CA, we pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/p-ca-ref-corr.png}
+        \includegraphics[width=1.0\textwidth]{p-ca-ref-corr.png}
             \caption{}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/g-ca-ref-corr.png}
+        \includegraphics[width=1.0\textwidth]{g-ca-ref-corr.png}
             \caption{}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/ca-reord1.png}
+        \includegraphics[width=1.0\textwidth]{ca-reord1.png}
             \caption{First CA Dimension}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
+        \includegraphics[width=1.0\textwidth]{bon-reord1.png}
             \caption{First Bonacich Dimension}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
+        \includegraphics[width=1.0\textwidth]{bon-reord2.png}
             \caption{Second Bonacich Dimension}
@@
      \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=1.0\textwidth]{Plots/SR-reord1.png}
-            \caption{Second SimRank Eigenvector}
+        \includegraphics[width=1.0\textwidth]{SR-reord1.png}
+            \caption{Main SimRank Eigenvector}
             \label{fig:sr-reord1}
@@
 206]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic \citet{breiger1974duality} projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods.
-In fact, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is substantively identical to that shown in~\citet[206, Table 2]{everett2013dual}, who previously extracted a core-periphery partitioning of the same data.
+Accordingly, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is substantively identical to that shown in~\citet[206, Table 2]{everett2013dual}, who previously extracted a core-periphery partitioning of the same data using a similar approach.
 
@@
     \centering
-        \includegraphics[width=0.8\textwidth]{Plots/ca-eigvec.png}            
+        \includegraphics[width=0.8\textwidth]{ca-eigvec.png}            
         \label{fig:eigvec}
@@
     \centering
-        \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
+        \includegraphics[width=1.0\textwidth]{ca-corr-plot.png}
     \caption{The top panels show the degree-weighted one-mode projection heatmap plots for (a) persons and (b) groups in the Southern Women Data with rows and columns re-ordered according to the first CA dimension.
@@
 We know from Figure~\ref{fig:bon-reord1} that event E8, along with E7, is one of the most popularly attended events and thus contributes little to differentiate between the main partition of people and events on the first correspondence plot in panel (c), hence its location near the plot's origin.
-The pattern of attendance of these six women is one characterized by affiliation with these popular events combined with sparser attendance at the events that most serve to differentiate the two primary clusters.
+The pattern of attendance of these six women is one characterized by affiliation with these popular events, combined with sparser attendance at the events that most serve to differentiate the two primary clusters.
 In his original paper, \citet[164]{bonacich1991simultaneous} noted that this third dimension ``corresponded closely to the centrality analysis pattern,'' an observation likely driven by the affinity between the actors who load highly on this dimension and two of the most central events by raw popularity, \{E8, E9\}.
@@
 Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes \citep{desposito2014comparison}.
-In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix---or a Markov transition matrix under a different interpretation---between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.
+In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix---or a Markov transition matrix under a different interpretation \citep[2]{mealy2019interpreting}---between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.
 
@@
 Persons and groups high in eigenvector centrality can be thought of as being the most similar to most others, condensing the information in the Breiger projection matrix into a single ordinal rank.
-Also, just like with CA and the degree-weighted projections, higher-order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode forms of similarity between pairs of people and pairs of groups, net of the main axis dominated by overall person activity and group size.
-
-\begin{figure}[ht!]
+Also, just like with CA and the degree-weighted projections, higher-order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should encode forms of similarity between pairs of people and pairs of groups, net of the main axis dominated by overall person activity and group size.
+
+\begin{figure}
     \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
     \centering
-        \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
-            \caption{Bonacich correspondence plot.}
-            \label{fig:bon-corr}
+        \includegraphics[width=1.0\textwidth]{bon-corr-plot.png}
     \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel.
-The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis, points are colored according a k-means three-cluster solution.}
+The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the projection matrices on the vertical axis, points are colored according to a k-means three-cluster solution.}
     \label{fig:bon-sim}
@@
 
-While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice \citep[see][for discussion in the one-mode network case]{iacobucci2017eigenvector} the exact meaning of the mysterious second eigenvector remains unclear in the two-mode case.
+While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice \citep[see][for discussion in the one-mode network case]{iacobucci2017eigenvector}, the exact meaning of the mysterious second eigenvector remains unclear in the two-mode case.
 We can use the same ``a posteriori blockmodeling'' approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue \citep{wasserman1987stochastic}.
@@
 
-We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension diplays a \textit{dual triangular} structure, revealing two secondary core and peripheries not evident in the blocking according to either the first CA or standard eigenvector scores.
+We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension displays a \textit{dual triangular} structure, revealing two secondary cores and peripheries not evident in the blocking according to either the first CA or standard eigenvector scores.
 Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering a version of the projection matrices of the Southern Women data where the proximities were designed to detect \textit{structural equivalence}.\footnote{In that application, the cell entry in each projection matrix was the \textit{proportion of matches} between each pair of people and groups.
-As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is Myrna/Myra's position, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
-This may be partly due to the discrepant position of ``core event'' E9 along the columns of the blocked matrix which is grouped with the second group of ``peripheral'' events \{E10, E11, E12, E13, E14\} in Figure~\ref{fig:bon-reord2}.
+As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' \citet[208]{everett2013dual} noted at the time that this was a partition ``different from all of the other published blockings for these data.'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is Myrna/Myra's position, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.
+This may be partly due to the discrepant position of ``core event'' E9 along the columns of the blocked matrix, which is grouped with the second group of ``peripheral'' events \{E10, E11, E12, E13, E14\} in Figure~\ref{fig:bon-reord2}.
 
@@
 Furthermore, the similarities between persons and groups defined over the two-dimensional Bonacich space, shown in Figure~\ref{fig:bon-sim} (c), can thus be used to reveal \textit{structurally equivalent} clusters of persons and events similar to those obtained by \citet[Table 5]{everett2013dual}.
-This is confirmed by the coloring of the persons and groups in Figure~\ref{fig:bon-reord2}) which comes from a k-means cluster analysis using the first two Bonacich dimensions to define initial cluster centroids, revealing a three-cluster solution almost equivalent to that obtained by Everett and Borgatti, with the exception of Myrna/Myra's and event E9's assingment as discussed earlier.
+This is confirmed by the coloring of the persons and groups in Figure~\ref{fig:bon-sim}(c), which comes from a k-means cluster analysis using the first two Bonacich dimensions to define initial cluster centroids, revealing a three-cluster solution almost equivalent to that obtained by Everett and Borgatti, with the exception of Myrna/Myra's and event E9's assingment as discussed earlier.
 This suggests that the Bonacich approach to similarity clustering is distinct from CA in being useful for revealing structural equivalence, thus making the Bonacich dimensions a more faithful basis for a low-rank reconstruction of the original ``input data'' than CA if that is the goal of the analysis \citep{desposito2014comparison}.
+
+\begin{figure}
+    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
+    \centering
+     \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=1.0\textwidth]{p-SimRank.png}
+            \caption{People.}
+            \label{fig:p-SRproj}
+    \end{subfigure} \\
+     \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=1.0\textwidth]{g-SimRank.png}
+            \caption{Groups.}
+            \label{fig:g-SRproj}
+    \end{subfigure}
+    \caption{SimRank projection matrices of the Southern Women data.}
+    \label{fig:SRproj}
+\end{figure}
 
@@
 This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence \citep{everett1994regular}.
-In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.\footnote{\citet{lizardo2024two} shows the connection between Kovacs's idea of generalized relations similarity and the two-mode network projection.}
+In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.
+
+In this section, I show that the contrast between the CA and Bonacich similarities can be understood along the same lines, with the Bonacich decomposition being more in line with a first-order similarity (structural equivalence) partition of nodes in the network (person and groups are similar to the extent that they connect to the \textit{same} groups and persons respectively), and the CA clustering producing similarities more congruent with a GRS criterion (thus closer to regular equivalence), where persons (groups) are similar to the extent that they connect to groups (persons) that are themselves similar.
 
@@
 
-This definition consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
+This definition is consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}.
 To accomplish this, \citet[540, eq.
@@
 
-\begin{figure}
-    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
+In two-mode networks, SimRank scores for each pair of nodes across the two sets can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
+Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
+We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}, which is confirmed here for the Southern Women data.
+
+\begin{figure}[ht!]
     \centering
-        \includegraphics[width=0.9\textwidth]{Plots/simrank-v-ca.png}
-            \label{fig:sim-v-ca}
+        \includegraphics[width=0.75\textwidth]{simrank-v-ca.png}
     \caption{Simrank versus CA comparison.
@@
 Panel (b) shows the Pearson correlation between the scores corresponding to the second eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis) for persons, panel (d) shows the same correlation for groups.}
-    \label{fig:simrank}
+    \label{fig:sim-v-ca}
 \end{figure}
 
-In two-mode networks, SimRank scores for each pair of nodes across the two sets can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values.
-Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset.
-We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}, which is confirmed here for the Southern Women data.
-
-Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode.
+Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on information from nodes in the other mode, averaged by the degree of nodes in the focal mode.
 The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}.
-Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections \citep{mealy2019interpreting, van2021correspondence}.
+Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections noted earlier.
 Beyond that, the SimRank approach can be thought of as yet another way of obtaining one-mode projections from an original two-mode network, in line with Everett and Borgatti's ``dual projection'' approach \citep{everett2013dual}.
-
-Figure~\ref{fig:sr-reord1} shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices---with $\alpha$ set to $0.8$---for the row and column nodes.
-Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling displays a block-diagonal structure separating the two communities of persons and groups as that revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
-Figure~\ref{fig:simrank-v-ca}(a) and Figure~\ref{fig:simrank-v-ca}(b) show an eigenvector plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
+The SimRank projection matrices for the Southern Women data---obtained after successful convergence with $\alpha$ set to $0.8$---are shown in Figure~\ref{fig:SRproj}.
+
+I compare the results obtained from the SimRank analysis to those obtained via CA in three different ways.
+First, Figure~\ref{fig:sr-reord1} shows the original affiliation matrix, this time with rows and columns re-arranged according to the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank projection matrices for persons and column groups.
+Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling displays a block-diagonal structure separating the same two blocks of persons and groups as that revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.
+
+Second, Figure~\ref{fig:sim-v-ca}(a) and Figure~\ref{fig:sim-v-ca}(b) show an eigenvector plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis.
 Looking at the plots from left to right, we can see that the partitioning of the node sets on the most informative eigenvector of the SimRank similarity matrix is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.
-Figures~\ref{fig:simrank-v-ca}(c) and (d) show a regression plot of the first CA dimension scores for persons and groups (respectively) on the x-axis and the SimRank scores extracted from the main non-trivial eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
-As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that these first eigenvectors capture the same (GSR) information across the two approaches.
+
+Finally, Figures~\ref{fig:sim-v-ca}(c) and (d) show a scatter plot and associated regression line with the first CA dimension scores for persons and groups (respectively) on the x-axis and the scores extracted from the main non-trivial eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis.
+As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that they capture substantively identical similarity information for both persons and groups.
 
@@
 This paper reconsiders the role of CA in the analysis of two-mode network data.
-We began with an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to simple CA of the affiliation matrix.
-Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}.
-I showed that just like the reflective scores that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums.
+After considering the various ways CA has been dealt with in the two-mode network analysis literature, I examined an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to a simple CA of the affiliation matrix.
+Working backward from this, I also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual scores in two-mode data due to \citet{bonacich1991simultaneous}.
+I showed that just like the reflective scores that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode.
+At the same time, the Bonacich approach works with unweighted (but normalized) sums.
 This exercise clarifies the links between CA and the dual two-mode eigenvector centrality more coherently and systematically, an issue that was left open and somewhat unclear in Bonacich's \citeyearpar{bonacich1991simultaneous} classic paper.
-Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.
-
-But CA does more than reveal a latent dual ordering of nodes in the two-mode network.
-In addition to doing this, CA can be shown to reveal \textit{groupings} of nodes based on some conception of the similarity of their connections to nodes in the other mode.
+
+As \citet{bonacich1991simultaneous} suspected in his original paper, CA can be shown to reveal a \textit{clustering} of nodes based on the similarity of their connections to nodes in the other mode.
 This is evident once we use the scores of the first CA dimension to re-order the rows and columns of the original affiliation matrix.
 When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data.
-This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events \citep{everett2013dual}.
+This partition is substantively distinct from that which would emerge if we use the same a posterior blocking approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events \citep{everett2013dual}.
 
@@
 Both are, of course, legitimate ways of analyzing the structure of a two-mode network.
-Still, they differ in terms of the structural patterns that they are sensitive to, with the CA approach closer to a community partition where nodes that are surprisingly similar end up in the same clusters---where ``surprising'' means similarity based on deviations from a suitable null model based on random mixing given their first-order degrees.
-
-But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to?
-Here, we show that CA recovers a type of generalized relational similarity (GRS).
+Still, they differ in terms of the structural patterns that they are sensitive to, with the CA approach closer to a blocking where nodes that are surprisingly similar end up in the same clusters---where ``surprising'' means similarity based on deviations from a suitable null model based on random mixing given their first-order degrees.
+
+Nevertheless, the contrast between the Bonacich approach as one geared toward ``centrality'' analysis and CA as one geared toward ``positional analysis,'' is much too simple.
+After all, as many have noted before, the Bonacich-style eigendecomposition of a proximity matrix yields dimensions beyond the first eigenvector, which are sensitive to mesolevel structure in the network, including that linked to group and community partitioning \citep{iacobucci2017eigenvector}.
+I show that this is indeed the case in the two-mode case, with the second eigenvector of the Bonacich decomposition revealing a secondary dual core-periphery partition, and with a clustering that incorporates that dimension, revealing structurally equivalent blocks of persons and events \citep{everett2013dual}.
+
+This left open the question of what kind of similarity the clustering based on the CA of the two-mode network is sensitive to.
+I showed that, in contrast to the higher-order Bonacich analysis, CA recovers a type of generalized relational similarity (GRS).
 That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar.
-Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, we saw that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach.
-Thus, we can clarify that two-mode network CA reveals latent groupings of nodes based on a GRS criterion.
-
-Overall, the preceding has shown that CA can be upgraded from a method designed to generate joint plots and visualization of two-mode data to one that can be seen as more ``central'' to the usual social-network-analytic tasks, like ordering the nodes in the two sets according to some substantively meaningful rank---centrality analysis---or finding sets of similar actors in the network (subgroup or community detection).
-At the very least, it seems like the Bonacich style dual centrality based on the eigenvector decomposition of the raw affiliation matrix should not be the default ``reflective'' centrality approach unless the analyst has the explicit analytic goal of exploring center-periphery partitioning in the network.
-
-A better approach, similar to the one exemplified here, would be to present a comparison of the reflective scores obtained by CA and Bonacich side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering.
-Of course, suppose the analyst is more interested in such a ``subgroup'' analytic approach.
-In that case, the Bonacich dual centrality approach is less helpful (even if multiple dimensions are considered), and CA should be the first line of attack.
+Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, I showed that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach.
+Thus, we can clarify that a two-mode network CA reveals latent groupings of nodes based on a GRS criterion (although in the Southern Women data, both he structural equivalence and GRS partitionings will tend to agree pretty closely).
+
+Overall, the preceding has shown that it may be time to reconsider the place of CA in the pantheon of two-mode network analysis methods.
+As we saw at the outset, CA has either been thought of as a ``data reduction'' method designed to provide low-dimensional visualizations of the original affiliation matrix, or contrasted individually with methods---such as eigenvector centrality analysis---designed to produce scores for ranking the two sets of nodes in the affiliation network.
+This paper shows that neither of these CA considerations is quite on track.
+
+Minimally, the previous analysis shows that CA can be upgraded from being a method geared exclusively to visualization to one that can be seen as more ``central'' to the usual social-network-analytic tasks, like finding blocks of structurally similar actors in the two-mode network, thus bringing the use of CA within the SNA community closer to the how it is used in link analysis, information science, and network science, where it is seen primarily as a spectral clustering method \citep{van2021correspondence, yen2010link, mealy2019interpreting}.
+At the very least, it seems like the Bonacich style dual centrality based on the eigenvector decomposition of the raw affiliation matrix should not be the default ``reflective'' centrality approach---out of the many options available \citep{yang2022comparing}---unless the analyst has the explicit analytic goal of exploring center-periphery partitioning in the network.
+
+A better approach, similar to the one exemplified here, would be to present a comparison of the blocking produced via the reflective scores obtained by CA and the Bonacich approach side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering.
+Of course, suppose the analyst is interested in a blocking or clustering approach.
+In that case, the contrast should move to the---possibly distinct, possibly overlapping---similarity clustering that can be obtained via CA or the traditional Bonacich approach, which, as we have seen, are keyed to analytically distinct conceptions of the ways nodes can be similar in a two-mode network.
+In that respect, once the commonalities and differences in the underlying mathematical machinery are clarified, both CA and Bonacich centrality analysis are equally worthy weapons in the arsenal of the ``dual projection'' \citep{everett2013dual} approach to the analysis of two-mode networks.
 

### 2025-03-30 Reviews and Response/response-memo.tex
Updates from Overleaf

--- 2025-01-31 Reviews and Response/response-memo.tex
@@
 \usepackage{graphicx}
+\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
 \begin{document}
 
-\textbf{Editors}
+\section*{Editors}
 
-\textsc{Reviewer 1}
+Thanks to the editors for the opportunity to revise and resubmit the previous version of the manuscript.
+I have attended to all of yours and the reviewers's recommendations for revision, and I believe the result is a stronger, clearer, and potentially more impactful manuscript.
+Here I note my responses to the editors' suggestions, more detailed responses to the reviewer suggestions follow below.
 
-\underline{State key points and conclusions at the beginning}:
+\subsection*{Major Points}
+\begin{itemize}
+    \item Regarding errors in the Southern Women data matrix, these are now fixed, and the data analyzed are the canonical data set. See response to Reviewer 1 below.
+    \item I also addressed R1's insightful point about the differences between odd and even iterations with regard to their correlation with the Bonacich centrality; see response to the reviewer below.
+    \item Regarding R2's call for a more effective statement of the novelty of the argument and key contributions, see various responses to the reviewer's suggestions below, which I took seriously and became the basis of a thorough rewrite of the manuscript.
+    \item With regard to broadening the coverage of the work, I have added a section on previous work as background (section 2).
+This version of the paper now incorporates Everett and Borgatti's 2013 piece throughout, and reframes CA (and Bonacich centrality analysis) as options within the ``dual projection'' framework they outline in that article, while making new connections between higher order axes in the Bonacich decomposition and their proposed approach to finding sets of structurally equivalent actors in two-mode networks.
+    \item With regard to strengthening the linkages between the paper's argument and Breiger's duality idea, see the response to R2's suggestions below. 
+\end{itemize}
 
-\textbf{Reviewer}: \textit{In terms of its overall organization, your paper is (probably, necessarily) complicated.
-For me, the discussion/conclusion (Section 5) is much more clear and powerfully written than is the abstract and Introduction (Section 1).
-As you think about improving the paper, I hope you will find ways to assert the main Section 5 points crisply at the beginning of the paper.}
+\subsection*{Minor Points}
+\begin{itemize}
+    \item - \underline{Editors}: \textit{p.
+4 of ``the sum of the q-1 centralities'' is a little confusing.
+If we correctly understand your point, q-1 refers to the iteration rather than how many centralities are being summed.
+Perhaps this could be that could be reworded.''} \textbf{Response}: The relevant passage has been reworded to improve clarity and accuracy.
+    \item -  \underline{Editors}: \textit{Notation in eq.
+19 and 20 seems slightly awkward - the left-hand side should somehow show that those are the transformed values, not the same as on the right-hand side (which currently is written the same way).} \textbf{Response}: I have added a single apostrophe (prime) symbol to indicate that the left hand side quantity is a transformed value.
+    \item - \underline{Editors}: \textit{For eq.
+21 and 22, it might be helpful to note how the eigenanalysis of the two product matrices relates to the SVD of the original.} \textbf{Response}: I have added some wording noting the links between eq.
+21 and 22 and eq.
+9 and 10.
+\end{itemize}
 
-\textbf{Response}:
+\newpage
+\section*{Reviewer 1}
 
-\underline{Add distinction between Bonacich and CA to the highlights}:
+\noindent\textbf{State key points and conclusions at the beginning}\newline
+\textit{Response}: Thanks for this comment.
+Given the newly added material, I streamlined the introduction in this version of the paper.
+However, the introductory paragraphs have been rewritten to highlight the key points and conclusions, which are also summarized in the rewritten abstract and highlights.
+\newline
 
-\textbf{Reviewer}: \textit{(For example, I would like to see a ``Highlights" bullet point that says that Bonacich centrality is useful for revealing the position of the nodes in a center-periphery pattern, to the extent one exists, whereas CA is useful for revealing the presence of different groups or communities in the data, to the extent that the dual network is not homogeneous.)}
+\noindent\textbf{Add distinction between Bonacich and CA to the highlights}\newline
+\textit{Response}: Thanks for this suggestion. I have now included a relevant highlight bullet point noting the relevant analytic contrast of Bonacich versus CA. \newline
 
-\textbf{Response}: Thanks for this suggestion. I have now included a relevant highlight bulletpoint noting the relevant analytic contrast of Bonacich versus CA.
-
-\underline{Data errors in the Southern Women two-mode matrix}:
-
-\textbf{Reviewer}: \textit{Importantly, there are 4 errors in the matrix (``Southern Women data") that you analyze.
-For this dataset, you (in the note to Table 1) cite Doreian et al (2004, Table 4).
-Among the attendees at Event E4, you show Frances, Eleanor, and Ruth.
-However they did not attend this event.
-You also show Nora as absent from Event E8; however, Nora attended E8.
-These four errors are discrepant in comparison to your cited source (Doreian et al, 2004, Table 4), and they are discrepant as well in comparison to the original source (Davis, Gardener, and Gardener, 1941, p.
-141, which has events in the same order as in your Table 1, and a slightly different order of women; this table is reproduced in Freeman, 2003, p.
-42).}
-
-\begin{figure}[h!]
-    \includegraphics[width=1.0\textwidth]{Reviews and Response/sw-original.png}
-\end{figure}
-
-\textbf{Response}: Thanks for calling attention to this---rather embarrassing---error; it's definitely a cautionary tale against manually copying data.
-I have corrected the data, and as you can see in the analysis file included as supplementary material, I now extract the Southern Women data directly from the \texttt{networkdata} package.
-My table now matches those in the references you mentioned exactly.
+\noindent\textbf{Data errors in the Southern Women two-mode matrix}\newline
+\textit{Response}: Thanks for calling attention to this---rather embarrassing---error; it's definitely a cautionary tale against manually copying data.
+I have corrected the data, and as you can see in the analysis file included as supplementary material, I now extract the Southern Women data directly from the \texttt{networkdata} package.\footnote{\texttt{https://www.rdocumentation.org/packages/networkdata/versions/0.1.6}} My table now matches those in the references you mentioned exactly.
 Note, however, that while you are correct that Frances, Eleanor, and Ruth did not attend E4, Nora did not attend E8, at least according to the published tables in Doreian et al (2004), Kovacs (2010), and the Figure 1 from Freeman (2003), shown above; in these papers, as you can see, Nora's event set is \{E6, E7, E9, E10, E11, E12, E13, E14\}.
@@
 2a).
-In any case, to forestall any confusion, and given that the in the \texttt{networkdata} package, the \texttt{igraph} object containing the Southern Women data come ordered using the Freeman (2003) ordering, I'm sticking to the non-chronological event labeling.
-I now include a brief discussion in the paper clarifying this issue.
+In any case, to forestall any confusion, and given that, in the \texttt{networkdata} package, the columns of the biadjacency matrix corresponding to the \texttt{southern\_women} object containing the Southern Women data come ordered using the Freeman (2003) ordering, I'm sticking to the non-chronological event labeling.
+I now include a brief discussion in the paper clarifying this issue.\newline
 
-\underline{Appropriateness of the term ``reflective centrality''}:
+\begin{figure}
+    \includegraphics[width=1.0\textwidth]{Reviews and Response/sw-original.png}
+\end{figure}
 
-\textbf{Reviewer}: \textit{I have some resistance to your term "reflective centrality" (Section 2 title) to characterize the scores of the respective modes - precisely because, as you importantly show, in the limit the reflective scores are first dimensions of a CA, and (as you also importantly emphasize) the first CA dimension is about finding groups, not about ordering nodes.
-Hidalgo and Hausmann (2009) do not use the term ``centrality" for the reflective scores they define and compute.
-I wonder whether ``reflective scores" might be a better term; I'm not sure, and simply ask you to think about it.}
+\noindent\textbf{Appropriateness of the term ``reflective centrality''}\newline
+\textit{Response}: Thanks for this comment, which I thought was very helpful.
+The term centrality introduces a level of conceptual imprecision that is best avoided.
+I liked your suggestion of using the more neutral term ``scores'' instead, which I implemented throughout the paper.\newline
 
-\textbf{Response}:
-
-\underline{Differences between odd and even iterations}:
-
-\textbf{Reviewer}: \textit{What complicates things, as you know, is the difference between odd and even iterations of the reflective algorithm (see your comments in the first paragraph on p.
+\noindent\textbf{Differences between odd and even iterations}\newline
+\underline{Reviewer}: \textit{What complicates things, as you know, is the difference between odd and even iterations of the reflective algorithm (see your comments in the first paragraph on p.
 4, as well as comments of H \& H).
@@
 
-\textbf{Response}:
+\noindent\textit{Response}: Thanks for this very insightful observation.
+I think that indeed there is a link between how the even and odd iterations are defined mathematically and their different patterns of correlations with the Bonacich centrality.
+Indeed, the early definitions of the odd iterations are closest to Bonacich's dual centrality idea, whereby people are central if they belong to central groups (and the centrality of groups is a function of the centrality of people who belong to them), which explains why we see a stronger correlation between the third reflection and Bonacich centrality which attenuates to the equilibrium value after higher reflections, and a lower correlation between the second iteration and the same quantity, which increases to the same value as the two rankings converge.
+This is now noted and explained in footnote 9 of the main text, along with an acknowledgment.
+Note, however, that I get a different value than you for the equilibrium correlation (see attached replication materials).
+\newline
 
-\underline{Missed connection between Bonacich and CA}:
+\noindent\textbf{Missed connection between Bonacich and CA regarding ``groupiness''}\newline
+\textit{Response}: Thanks very much for raising this point, which made me think about and consider issues that I missed in the first version of the paper.
+You are right that the second dimension of the Bonacich centrality reveals a dimension of ``groupiness'' in the data; however, this dimension is not quite the same as that of CA.
+In fact, it became clear to me that what it reveals are continuous approximations to structural equivalence in the data, which makes sense if, indeed, the CA scores are continuous approximations to a regular equivalence style criterion (or generalized relational similarity).
+We can check this claim by comparing the blocking obtained from the second Bonacich dimension and that produced by Everett and Borgatti (2013, Table 5) using a proportion-matching criterion for structural equivalence, and see that they agree quite closely, a discussion that I now include in the paper.
+The groupings in the second Bonacich dimension end up being secondary core-periphery partitions, net of the main one.
+This reinforces my previous claim, which was poorly argued for in the previous version of the paper, that the Bonacich decomposition comes closer to reproducing the original ``input data'' than the CA decomposition.
+\newline
 
-\textbf{Reviewer}: \textit{One of the high points of your paper - bringing great clarity, and of great usefulness to practitioners - is your Fig.
-2, showing how the first (person and group) dimensions of the CA reveal ``groupiness" (my poor word choice), whereas the dual Bonacich centrality scores reveal the extent to which a center-periphery pattern obtains in the data.
-A point you miss here (however, I apologize if you do say this somewhere) is that the second dimension of Bonacich centrality (that is, the second dimension of the SVD of an affiliation matrix like Table 1) looks very much like the first CA dimension, emphasizing communities rather than center-periphery.
-This comment, like that of the previous paragraph, I think helps you to show how highly entangled are the concepts of reflective scores and centrality, and (therefore) how important it is to recognize when each is appropriate.}
+\noindent\textbf{Reversing column ordering on some figures}\newline
+\textit{Response}: Thanks for this suggestion. I have re-plotted the figures using the suggested column ordering. \newline
 
-\textbf{Response}:
-
-\underline{Reversing column ordering on some figures}:
-
-\textbf{Reviewer}: \textit{(In Fig.
-2, both parts (a) and (b), nothing substantive or mathematical would change if you keep the row ordering you show, but reverse the column ordering.
-However, reversing the column ordering would help the reader to see, in panel (a), the groups, and, in panel (b), the center and the periphery.
-A similar comment applies to several of your other figures.)}
-
-\textbf{Response}: Thanks for this suggestion. I have re-plotted the figures using the suggested column ordering.
-
-\underline{Confusing equations}:
-
-\textbf{Reviewer}: \textit{What confused me about eqs.
+\noindent\textbf{Confusing equations}\newline
+\underline{Reviewer}: \textit{What confused me about eqs.
 (23) and (24) is that I don't think the (respective) expressions in parentheses are symmetric matrices, and so I don't see how to take an eigenvector.
@@
 
-\textbf{Response}: Thanks for this observation.
+\noindent \textit{Response}: Thanks for this observation.
 You are correct that I did not adequately explain where the eigenvectors corresponding to the CA (and HH) scores come from.
@@
 126 notes).
-The relevant section of the paper has been rewritten to clarify these issues.
+The relevant section of the paper has been rewritten to clarify these issues.\newline
 
-\underline{Confusing Figures}:
+\noindent\textbf{Confusing labeling of figures across multiple panels}\newline
+\textit{Response}: I have re-rendered the figures so that the labeling is clearer. \newline
 
-\textbf{Reviewer}: \textit{Figs 3, 4, 5, and 6 each have multiple panels (referred to as a, b, c, etc), but, in the current version of the paper, it is difficult or impossible to figure out which part of the figure has which label (a, b, c, …).
-Hopefully the formatting here can be improved.}
-
-\textbf{Response}:
-
-\underline{Missing Freeman citation}
-
-\textbf{Response}:
+\noindent\textbf{Missing Freeman citation}\newline
+\textit{Response}: Freeman citation is now added and discussed in the paper. \newline
 
 \newpage
-\textsc{Reviewer 2}
+\section*{Reviewer 2}
 
-\underline{Insufficient engagement with the duality idea}
+\textbf{Insufficient engagement with Breiger's duality idea}\newline
+\textit{Response}: Thanks for this comment, which proved pivotal in my revision efforts for this draft.
+In this version of the paper, I have endeavored to make the linkages between CA, Bonacich centrality analysis, and duality clearer in the text.
+I have also incorporated a more in-depth discussion of Everett and Borgatti's dual projection approach, and re-framed both CA and centrality analysis as two distinct but related options within that wider framework, which also helps to make the link to the duality notion crisper.
+\newline
 
-\textbf{Reviewer}: \textit{In my opinion, though, the paper does not take Breiger's argument in due consideration, even more as the piece is aimed at contributing to a special issue on ``Duality in social networks".
-The latter principle seems to remain in the background, while it should explicitly inform most of the analyses presented in this work.}
+\noindent\textbf{More extensive introduction and background}\newline
+\textit{Response}: Thanks also for this comment, which informed most of the new material that has been added to this version of the paper, particularly the new section 2.
+I believe that with this background in hand, the paper's arguments and contribution can be more clearly stated and appreciated.
+\newline
 
-\textbf{Response}:
+\noindent\textbf{More even-handed consideration of the visualization aspects of the CA/MCA}\newline
+\textit{Response}: Thanks for this comment, which made me realize that perhaps my previous comments regarding the focus on visualization on previous uses of CA and MCA for two-mode network analysis were coming off as overly dismissive.
+I have tried to rewrite the relevant sections in this draft to remove these connotations.
+My key point was more in reference to Borgatti and Everett's original 1997 piece, which seemed to think of CA primarily as a visualization tool, but was ignoring its usefulness for other network-analytic tasks such as clustering and ``a posteriori blockmodeling'' (in Wasserman and Anderson's terms).
+I hope that these points come out clearer in this version of the paper.
+\newline
 
-\underline{More extensive introduction and background}
+\noindent\textbf{Concerns over the interpretation of CA biplot distances}\newline
+\textit{Response}: Thanks for this comment.
+In this version of the paper I provide a more in-depth discussion of what the inter-point distances in the usual CA correspondence plot get at.
+The rewritten section now ties the correspondence plot distances directly to the information in the degree-adjusted projection matrices, which leads to a more straightforward interpretation of the CA correspondence plot as a low-rank representation of the degree-adjusted proximity matrix, which makes the contrast to the Bonacich ``correspondence plot'' also clearer.
+\newline
 
-\textbf{Reviewer}: \textit{I suggest extending the introduction, enriching the discussion of the background as for CA in the context of two-mode networks.
-Given the importance of the technique and its variants (like MCA), it is surprising that the authors devote such a limited room to the relevant framework.
-Consider that the discussion of CA in the context of two-mode networks seems not comprehensive as it misses the use of MCA in the same context \ldots}
+\noindent\textbf{Insufficient discussion of positional equivalence and other blockmodeling approaches}\newline
+\textit{Response}: Thanks for this comment, which I took to heart in rewriting key portions of the paper.
+As you can see, I contrasted different conceptions of positional equivalence (e.g., structural versus regular) central to my contrast between what CA does versus what the Bonacich decomposition does, also drawing on some of the literature that you suggested I look at (particularly D'Esposito et al's 2014 piece the contrast between CA and MCA distances), and previous work by Everett and Borgatti.
+While I did not engage in a full-on review of previous work on positional analyses of two-mode networks for reasons of relevance and space, the emphasis on the usefulness of CA for positional analysis (e.g., aka, ``spectral clustering'' in other fields) is evident throughout.
+Note that in the different flavors of blockmodeling, CA (or even Bonacich centrality) is closer to indirect or ``a posteriori'' blockmodeling (e.g., data-driven rather than model-driven) so the emphasis is different from optimization-based methods like Doreian, Batagelj, and Ferligoj's direct blockmodeling approach.
+Hopefully, what is new in the current approach relative to this previous work can be more clearly appreciated in this draft.
+\newline
 
-\textbf{Response}:
+\noindent\textbf{Insufficient discussion of other CA-based analytic strategies}\newline
+\textit{Response}: Thanks for the reference to D’Ambrosio et al's work, which I read with great interest.
+In the end, when crafting the new background section, I realized that I couldn't cover all previous uses of CA and related techniques for two-mode network analysis.
+As a result, I decided to just focus on previous work that dealt with the simple-CA of the affiliation matrix, since that's what the core discussion of the paper is relevant for.
+However, I do note that this previous work focuses on MCA and related techniques (which your reference led me to) in various footnotes.
+\newline
 
-\underline{More consideration of the visualization aspects of the CA/MCA}
+\noindent\textbf{Better highlights section}\newline
+\textit{Response}: Thanks for this comment, which along with a similar observation by the other review informed by revision of the highlights section. \newline
 
-\textbf{Reviewer}: \textit{Further, as the introduction states, there is an attempt at seemingly diminishing the power of visualization as a key feature of CA, as if it were a secondary concern in this domain.
-I think the authors should first consider carefully the actual utility of data visualization via CA/MCA and then also mind that, although their work aims to move ``beyond the focus on data visualization" (p.
-2), they in fact make use of visual representation quite heavily in this paper.
-This seems to me rather misleading.}
-
-\textbf{Response}:
-
-\underline{Concerns over the interpretation of CA biplot distances}
-
-\textbf{Reviewer}: \textit{As regards CA and its ``conventional" - so to speak - interpretation which the authors criticize and want to go beyond, I disagree on the idea that using CA in a way different than the authors' would be less useful/informative or, even worse, ``off the mark".
-Here I specifically refer to a passage on p.
-12 \ldots While I would agree that one can interpret the distances between units - and, even more aptly, between the units and the axes origin - on a CA plot in terms of distances from an independence condition, I would not dismiss the importance of evaluating, in the above fashion, how ``people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups".
-This has clear substantive value for interpretation of the data thanks to use of CA.}
-
-\textbf{Response}:
-
-\underline{Insufficient discussion of positional equivalence}
-
-\textbf{Reviewer}: \textit{As for the paper's argument, one may ask what's new in the current proposal and what's the point in it, beyond already known things.
-This issue appears to affect, for example, the opportunity to analyse the forms of structural or regular equivalence in two-mode networks.
-This theme is paid very little attention, though.
-And yet, it is very relevant to the paper - as the authors seem at least to know, as they mention it (one time only, on p.
-14).
-In the course of the paper there are some passages where the point would be relevant, as is the case with the bottom of p.
-10, when the authors recall the works of Doreian, Batagelj and Ferligoj (2004), Kovács (2010) and Lizardo (2024), and assert that their own results are comparable to the latter's \ldots This is clearly a matter of positional equivalence, which deserves more room in the paper, in my opinion.}
-
-\textbf{Response}:
-
-\underline{Insufficient discussion of blockmodeling approaches}
-
-\textbf{Reviewer}: \textit{In addition, I would like to stress that the article sets forth an argument that, in the beginning, revolves around centrality, but ends with a more detailed discussion about how to assess relational similarity in two-mode networks.
-Here, there seems to be a lack of consideration regarding other relevant approaches.
-Blockmodeling, for instance, is not duly considered, although it is well known as a sort of gold standard in the analysis of equivalences - particularly generalized equivalences (I suggest seeing the book by Doreian, Batagelj, and Ferligoj, 2005, in addition to the Social Networks article by the same authors that is cited in the paper) \ldots I would like the authors to pay more attention to this absence in their discussion.
-More explicitly, what does the authors' proposal add to the analysis of structural/regular/generalized equivalences as it is already treated via blockmodeling (and also via MCA)?}
-
-\textbf{Response}:
-
-\underline{Other analytic strategies}
-
-\textbf{Reviewer}: \textit{I also think about other analytic strategies like Cluster Correspondence Analysis (Cluster CA) applied to two-mode networks, for which I may suggest reading and referencing relevant works (such as D'Ambrosio, Serino and Ragozini, 2021).}
-
-\textbf{Response}:
-
-\underline{Better highlights section}
-
-\textbf{Reviewer}: \textit{- The Highglights section should be improved in the sense that it looks now like a sort of abstract, while it needs to appear as a bullet-like summary of the paper that will make its argument better understood at the first glance}
-
-\textbf{Response}:
-
-\underline{Unclear figure}
-
-\textbf{Reviewer}: \textit{- Figure 3 is not clear, in particular it should have a proper layout where the content of the current insets would be shown separately (now it is a bit confusing)}
-
-\textbf{Response}:
+\noindent\textbf{Unclear figure}\newline
+\textit{Response}: I have redone the old Figure 3 to remove any remaining confusion. \newline
 

### 2025-06-02 analysis-noproj.qmd
oof

---
title: "CA of Music Genre Stereotypes"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  output: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
   library(conflicted)
   library(dplyr)
   library(FactoMineR)
   library(ggplot2)
   library(ggrepel)
   library(ggcorrplot)
   library(ggridges)
   library(haven)
   library(here)
   library(paletteer)
   cols <- paletteer_dynamic("cartography::multi.pal", 20)
   conflicts_prefer(dplyr::filter)
   conflicts_prefer(dplyr::select)
```

```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
   head(dat)
```

```{r Creating individual matrices}
      r <- c("Classical", "Opera", "Jazz", "Broadway", "Easy", "Big_Band", "Oldies", "Country", "Bluegrass", "Folk", "Gospel", "Latin", "Rap", "Blues", "Reggae", "Pop", "Rock", "AltRock", "Dance", "Metal")
       c <- c("Female", "Male", "White", "Black", "Hispanic", "Asian", "College", "No_College", "Young", "Mid_Age", "Old", "Lower_Class", "Working_Class", "Middle_Class", "Upper_Class")
   den <- rowSums(dat[,2:301])/300
   bad.den <- which(den > .9 | den < 0.1)
   dat2 <- dat[-c(bad.den), 2:301]
   ind.list <- split(dat2, seq(nrow(dat2))) # splitting each row of data frame into list objects
   ind.mat <- lapply(ind.list, function(x) {matrix(as.numeric(x), 20, 15, byrow = TRUE)}) # creating individual two mode network of genre by labels
   ind.mat.t <- lapply(ind.list, function(x) {matrix(as.numeric(x), 20, 15, byrow = TRUE)}) # creating individual two mode network of genre by labels
   ind.mat.t <- lapply(ind.mat.t, function(x) {t(x)})
   ind.mat <- lapply(ind.mat, function(x) {rownames(x) <- r; x}) # naming individual rows
   ind.mat <- lapply(ind.mat, function(x) {colnames(x) <- c; x}) # naming individual columns
   ind.mat.t <- lapply(ind.mat.t, function(x) {rownames(x) <- c; x}) # naming individual transposed rows
   ind.mat.t <- lapply(ind.mat.t, function(x) {colnames(x) <- r; x}) # naming individual transposed columns
   names(ind.mat) <- dat$id_char[-c(bad.den)]
   names(ind.mat.t) <- dat$id_char[-c(bad.den)]
   row.stack <- matrix(unlist(ind.mat), ncol = 15, byrow = TRUE) # genre stacked matrix
   col.stack <- matrix(unlist(ind.mat.t), ncol = 20, byrow = TRUE) # label stacked matrix
   colnames(row.stack) <- c # naming genre stacked matrix columns
   colnames(col.stack) <- r # naming label stacked matrix columns
   rownames(row.stack) <- rep(r, nrow(dat2)) # naming genre stacked matrix rows
   rownames(col.stack) <- rep(c, nrow(dat2)) # naming label stacked matrix rows
   zero.rows.r <- which(rowSums(row.stack) == 0) # which genre stacked matrix rows sum to zero?
   zero.rows.c <- which(rowSums(col.stack) == 0) # which label stacked matrix rows sum to zero?
```

```{r Data Examples (Two Mode)}
   png(file = here("Plots", "data-ex1.png"), width=700, height=700)
      corrplot(ind.mat[["id_65"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex2.png"), width=700, height=700)
      corrplot(ind.mat[["id_236"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex3.png"), width=700, height=700)
      corrplot(ind.mat[["id_1114"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex4.png"), width=700, height=700)
      corrplot(ind.mat[["id_2002"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()

   ex.stack <- rbind(ind.mat[["id_65"]], ind.mat[["id_236"]])
   ex.stack.t <- rbind(t(ind.mat[["id_1114"]]), t(ind.mat[["id_2002"]]))
   png(file = here("Plots", "data-ex-stack.png"), width=700, height=700)
      corrplot(ex.stack, is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 0.75, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex-stack-t.png"), width=700, height=700)
      corrplot(ex.stack.t, is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 0.75, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
```

```{r CA}
   #genres
   row.stack.d <- data.frame(row.stack)
   row.stack.d$genre <- rep(r, nrow(dat2))
   row.stack.d$id <- rep(dat$id_char[-bad.den], each = 20)
   CA.r <- CA(row.stack.d, graph = FALSE, row.sup = zero.rows.r, quali.sup = c(16, 17))
   #labels
   col.stack.d <- data.frame(col.stack)
   col.stack.d$label <- rep(c, nrow(dat2))
   col.stack.d$id <- rep(dat$id_char[-bad.den], each = 15)
   CA.c <- CA(col.stack.d, graph = FALSE, row.sup = zero.rows.c, quali.sup = c(21, 22))
```

```{r Plot Data}
     make.plot.dat <- function(x) { 
         w <-  data.frame(x) %>% 
         mutate(col.obj = gsub("\\..*","",rownames(x))) 
         rownames(w) <- 1:nrow(x)
         return(w)
         }
      col.lab.dat <-  data.frame(CA.r$col$coord) #genre column scores
      col.gen.dat <-  data.frame(CA.c$col$coord) #label column scores
      sup.gen.dat <- CA.r$quali.sup$coord[1:20, ] 
      sup.lab.dat <- CA.c$quali.sup$coord[1:15, ] 
      sup.gen.dat <- data.frame(sup.gen.dat, 
                                col.obj = gsub("genre.", "", rownames(sup.gen.dat)))
      rownames(sup.gen.dat) <- sup.gen.dat$col.obj
      sup.lab.dat <- data.frame(sup.lab.dat, 
                                col.obj = gsub("label.", "", rownames(sup.lab.dat)))
      rownames(sup.lab.dat) <- sup.lab.dat$col.obj
      row.gen.dat <- make.plot.dat(CA.r$row$coord) #genre row scores
      row.lab.dat <- make.plot.dat(CA.c$row$coord) #label row scores
```

```{r Column Score Similarity Plots}
   h <- 625
   w <- 750
   source(here("make.sim.plot.R"))
   #genres
   png(file = here("Plots", "ca-genre-col12.png"), width=w, height=h)
   make.sim.plot(col.gen.dat, 1, 2, "red", l = 1)
   dev.off()
   png(file = here("Plots", "ca-genre-col13.png"), width=w, height=h)
   make.sim.plot(col.gen.dat, 1, 3, "red", l = 1)
   dev.off()
   png(file = here("Plots", "ca-genre-col14.png"), width=w, height=h)
   make.sim.plot(col.gen.dat, 1, 4, "red", l = 1)
   dev.off()
   # labels
   png(file = here("Plots", "ca-label-col12.png"), width=w, height=h)
   make.sim.plot(col.lab.dat, 1, 2, "blue", l = 1)
   dev.off()
   png(file = here("Plots", "ca-label-col13.png"), width=w, height=h)
   make.sim.plot(col.lab.dat, 1, 3, "blue", l = 1)
   dev.off()
   png(file = here("Plots", "ca-label-col14.png"), width=w, height=h)
   make.sim.plot(col.lab.dat, 1, 4, "blue", l = 1)
   dev.off()
```

```{r Supplementary Score Similarity Plots}
   h <- 625
   w <- 750
   source(here("make.sim.plot.R"))
   #genres
   png(file = here("Plots", "ca-genre-sup12.png"), width=w, height=h)
   make.sim.plot(sup.gen.dat, 1, 2, "red", l = .6)
   dev.off()
   png(file = here("Plots", "ca-genre-sup13.png"), width=w, height=h)
   make.sim.plot(sup.gen.dat, 2, 3, "red", l = .6)
   dev.off()
   png(file = here("Plots", "ca-genre-sup14.png"), width=w, height=h)
   make.sim.plot(sup.gen.dat, 3, 4, "red", l = .6)
   dev.off()
   # labels
   png(file = here("Plots", "ca-label-sup12.png"), width=w, height=h)
   make.sim.plot(sup.lab.dat, 1, 2, "blue", l = .6)
   dev.off()
   png(file = here("Plots", "ca-label-sup13.png"), width=w, height=h)
   make.sim.plot(sup.lab.dat, 2, 3, "blue", l = .6)
   dev.off()
   png(file = here("Plots", "ca-label-sup14.png"), width=w, height=h)
   make.sim.plot(sup.lab.dat, 3, 4, "blue", l = .6)
   dev.off()
```

```{r}
   d <- filter(row.gen.dat, col.obj %in% 
                  c("Broadway", "Folk"))
   p <- ggplot(data = d, aes(Dim.1, Dim.2, 
                             group = col.obj, 
                             fill = col.obj,
                             color = col.obj))
   p <- p + geom_vline(xintercept = 0, color = "gray")
   p <- p + geom_hline(yintercept = 0, color = "gray")
   #p <- p + stat_density_2d(geom = "polygon", aes(alpha=..level..^2))
   p <- p + geom_jitter(aes(shape = col.obj), alpha = 0.15, size = 2)
   p
```

```{r Comparing Genre Scores to Cloud Centroids}
   source(here("make.fuzz.plot.R"))
   h <- 625
   w <- 750
   # genre plots
   png(file = here("Plots", "genre-proto12.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, 1, 2, c("Folk", "Rap", "Classical"))
   dev.off()
   png(file = here("Plots", "genre-proto13.png"), width=w, height=h)
      make.fuzz.plot(plot.dat.g, plot.dat3, 1, 3, c("AltRock", "Opera", "Gospel"))
   dev.off()
   png(file = here("Plots", "genre-proto14.png"), width=w, height=h)
      make.fuzz.plot(plot.dat.g, plot.dat3, 1, 4, c("Reggae", "Easy", "Classical"))
   dev.off()
   # label plots
   png(file = here("Plots", "label-proto12-1.png"), width=w, height=h)
      make.fuzz.plot(plot.dat.l, plot.dat4, 1, 2, c("Asian", "Female", "College"), lim = 1)
   dev.off()  
   png(file = here("Plots", "label-proto12-2.png"), width=w, height=h)
      make.fuzz.plot(plot.dat.l, plot.dat4, 1, 2, c("Hispanic", "Black", "No_College"), lim = 1)
   dev.off()  
   png(file = here("Plots", "label-proto13.png"), width=w, height=h)
      make.fuzz.plot(plot.dat.l, plot.dat4, 1, 3, c("Upper_Class", "Male", "No_College", "Asian"), lim = 1)
   dev.off() 
   png(file = here("Plots", "label-proto14.png"), width=w, height=h)
      make.fuzz.plot(plot.dat.l, plot.dat4, 1, 4, c("Lower_Class", "White", "College"), lim = 1)
   dev.off()
```

```{r Genre Ridge Plots}
      make.ridge.dat <- function(col.dat, ca.rows) {
        b <- col.dat %>% 
         rename(Prot.1 = Dim.1, Prot.2 = Dim.2, Prot.3 = Dim.3) %>% 
         mutate(col.obj = rownames(col.dat)) 
      x <-  data.frame(rbind(ca.rows)) %>% 
         mutate(col.obj = gsub("\\..*","",rownames(ca.rows))) %>% 
         select(-c("Dim.4", "Dim.5")) %>% 
         right_join(b) 
      return(x)
      }
      make.ridge.plot <- function(x, d1, d2, lim = 1) {
      p <- ggplot(data = x, aes(x = x[, d1], 
                                y = reorder(col.obj, x[, d2]),
                                group = col.obj, fill = col.obj, color = col.obj))
      p <- p + geom_density_ridges(alpha = .2, point_alpha = .2, jittered_points = TRUE)
      p <- p + guides(fill = "none", color = "none")
      p <- p + geom_vline(xintercept = 0, color = "black", linewidth = 1.25)
      p <- p + scale_color_manual(values = cols)
      p <- p + theme(axis.text = element_text(size = 20), 
                     axis.title = element_text(size = 24))
      p <- p + labs(x = paste("Dim.", d1))
      p <- p + theme_ridges() + xlim(-lim, lim) + labs(y = "") 
      return(p)
      }
   # genres
   ridge.dat.g <- make.ridge.dat(plot.dat1, CA.r$row$coord)
   png(file = here("Plots", "genre-ridge-dim1.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.g, 1, 5)
   dev.off()
   png(file = here("Plots", "genre-ridge-dim2.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.g, 2, 6)
   dev.off()
   png(file = here("Plots", "genre-ridge-dim3.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.g, 3, 7)
   dev.off()
   #labels
   ridge.dat.l <- make.ridge.dat(plot.dat2, CA.c$row$coord)
   png(file = here("Plots", "label-ridge-dim1.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.l, 1, 5)
   dev.off()
   png(file = here("Plots", "label-ridge-dim2.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.l, 2, 6)
   dev.off()
   png(file = here("Plots", "label-ridge-dim3.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.l, 3, 7)
   dev.off()
```
